---
title: "Default Priors for Binomial GLMs: BF Calibration"
shorttitle: "BF Calibration for Binomial GLMs"
author:
  - name: Klaus Oberauer
    orcid: 0000-0001-8815-6189
    affiliations:
      - name: University of Zurich
        department: Department of Psychology
  - name: Joscha Dutli
    affiliations:
      - name: University of Zurich
        department: Department of Psychology
  - name: Philipp Musfeld
    affiliations:
      - name: University of Zurich
        department: Department of Psychology
  - name: Gidon T. Frischkorn
    orcid: 0000-0002-5055-9764
    email: gidon.frischkorn@unilu.ch
    corresponding: true
    affiliations:
      - name: University of Lucerne
        department: Faculty of Health Sciences and Medicine
author-note:
  disclosures:
    study-registration: "This study was not preregistered."
    data-sharing: "Simulation code and analysis scripts are available at [repository link]."
    conflict-of-interest: "The authors declare no conflicts of interest."
    financial-support: "This research was supported by a grant from the Swiss National Science Foundation (SNSF)."
    authorship-agreements: |
      Klaus Oberauer: Conceptualization, Methodology, Writing -- Original Draft.
      Joscha Dutli: Methodology, Software, Formal Analysis.
      Philipp Musfeld: Methodology, Software, Formal Analysis.
      Gidon T. Frischkorn: Conceptualization, Methodology, Software, Formal Analysis, Visualization, Writing -- Original Draft.
format:
  apaquarto-docx: default
  apaquarto-pdf:
    keep-tex: true
  apaquarto-html: default
execute:
  echo: false
  warning: false
  message: false
bibliography: references.bib
csl: apa.csl
---

```{r}
#| label: setup
library(dplyr)
library(tidyr)
library(ggplot2)
library(viridis)
library(here)

source(here("R", "bf_helpers.R"))

summaries <- readRDS(here("output", "bf_calibration_summaries.rds"))

posteriors <- tryCatch(
  readRDS(here("output", "bf_calibration_posteriors.rds")),
  error = function(e) NULL
)
posteriors_available <- !is.null(posteriors)

re_summaries <- tryCatch(
  readRDS(here("output", "bf_calibration_re_summaries.rds")),
  error = function(e) NULL
)
re_results_available <- !is.null(re_summaries)

re_posteriors <- tryCatch(
  readRDS(here("output", "bf_calibration_re_posteriors.rds")),
  error = function(e) NULL
)
re_posteriors_available <- !is.null(re_posteriors)

# Factor labels
dist_labels <- c(cauchy = "Cauchy", normal = "Normal", logistic = "Logistic")

base_theme <- function() {
  theme_minimal(base_size = 12) +
    theme(
      panel.grid.minor = element_blank(),
      strip.text       = element_text(face = "bold"),
      legend.position  = "bottom"
    )
}
```

## Introduction

The prior predictive analysis established that matched priors — Logistic(0, 0.75)
for the logit link and Normal(0, 0.75) for the probit link — produce well-behaved
distributions on the probability scale, and that effect priors with
$\text{sd}_{b_1} = 0.25$ (the standard deviation of the prior on the effect
coefficient $b_1$) concentrate prior mass on substantively plausible effect sizes.
However, good prior predictive behaviour is a necessary condition for sound
Bayesian inference, not a sufficient one. We also need to know how these priors
perform when used for hypothesis testing with Bayes Factors.

Prior specification directly affects discovery rates. A prior that is too wide
spreads probability mass across implausible effect sizes, diluting the evidence
for real effects and reducing the power of the Bayes Factor to detect them. A
prior that is too narrow may spuriously favour or disfavour the null. This tension
is well documented for normal-data models: @schonbrodt2018 introduced Bayes Factor
Design Analysis (BFDA) as a formal framework for evaluating power and error rates
under a given prior, and @stefan2019 extended it to informed priors. These tools,
however, have been developed for simple designs with closed-form BF solutions
(e.g., the JZS prior of @rouder2009). For binomial GLMs — where BFs must be
computed numerically via bridge sampling [@gronau2017; @gronau2020] or the
Savage–Dickey density ratio — no systematic calibration exists. The present
simulation fills this gap by applying the BFDA logic to the recommended priors
from the prior predictive analysis.

We address four questions:

1. **Do the recommended priors provide adequate power** to detect effects across
   realistic sample sizes and numbers of trials?
2. **Do they protect against false evidence** — that is, do they correctly support
   the null hypothesis when there is no effect?
3. **How sensitive is the Bayes Factor to the prior family?** Does it matter whether
   researchers use a Normal, Logistic, or Cauchy prior, or is the prior scale the
   key decision?
4. **Does the matched-prior principle carry over to inference?** The prior
   predictive analysis showed that matched intercept priors produce better-behaved
   distributions on the probability scale — but does using a mismatched intercept
   prior actually harm Bayes Factor calibration?

Finally, we extend the analysis to within-subjects designs with correlated random
effects, asking whether the choice of random-effects hyperprior — which the prior
predictive analysis showed affects the distribution of subject-level probabilities —
also affects inferential performance.

### Simulation Design

We simulated data across a range of realistic experimental scenarios. The true
value of the effect coefficient $b_1$ (the regression weight for the experimental
condition in $g(p_i) = b_0 + b_1 x_i$) ranged from $b_1 = 0$ (no effect) to
$b_1 = 0.50$ on the logit scale,
covering the span from a null effect through a large effect in cognitive
experiments (on the probability scale with a typical 50% baseline, $b_1 = 0.50$
corresponds to roughly 24 percentage points between conditions). Sample sizes
ranged from 30 to 100 subjects, and each subject contributed either 20 or 50
trials. We crossed these data-generating factors with three effect prior families
(Normal, Logistic, Cauchy) at two scale values ($\text{sd}_{b_1} = 0.25$ and
$0.50$), and two intercept priors (matched Logistic and misfit Normal, both at
$\text{sd}_{b_0} = 0.75$). This produced 360 conditions with 100 replications
each (see `scripts/03_BFCalibration.R`).

A second simulation (`scripts/03b_BFCalibration_RE.R`) specifically investigated
the random-effects hyperprior. It used a within-subjects design with correlated
random intercepts and slopes, holding the fixed-effects priors at the recommended
matched configuration and varying only the hyperprior on the random-effects SDs
across three candidates: the brms default Student-*t*(3, 0, 2.5), Gamma(2, 4),
and Exponential(4). This produced 90 conditions with 100 replications each. We
return to these results in @sec-re-hyperprior.

### How We Computed Bayes Factors

All models used the logit link (the probit link is expected to produce analogous
results via the matched-prior principle established in the prior predictive
analysis). To ensure that results generalize across realistic baseline accuracies,
we did not fix the true intercept: in each replication, $b_0$ was sampled from
$\text{Uniform}(0.4, 0.9)$ on the probability scale — the range spanning
two-alternative to four-alternative forced choice — and transformed to the logit
scale. Each simulated dataset included subject-level random intercepts
($u_i \sim \mathcal{N}(0, 0.25)$) to mimic between-subject heterogeneity.

The analysis model mirrored this structure: a binomial GLMM with one random
intercept per subject, fitted with brms [@burkner2017], with the intercept prior
SD fixed at 0.75 and an $\text{Exponential}(4)$ hyperprior on the random-intercept
SD. Bayes Factors were computed via the Savage–Dickey density ratio
[@wagenmakers2018; @ly2016], which compares how plausible a zero effect was before
versus after seeing the data:
$\text{BF}_{10} = \pi(b_1 = 0) / p(b_1 = 0 \mid \text{data})$.
When the posterior concentrates away from zero (i.e., the data suggest an effect),
$\text{BF}_{10}$ exceeds 1 and favours H1; when the posterior piles up at zero
(i.e., the data are consistent with no effect), $\text{BF}_{10}$ falls below 1
and favours H0. Note that the prior density at zero — $\pi(b_1 = 0)$ — enters the
numerator directly. This has a critical consequence: different prior families with
the same scale parameter place different densities at zero, which creates
systematic differences in specificity (the ability to correctly support H0). We
return to this point in the specificity analysis below.

To quantify uncertainty in detection rates given the finite number of replications,
we applied Bayesian conjugate updating (see
@sec-appendix-uncertainty for details). All error
bars in the figures below show 95% credible intervals from this procedure.

## How Much Data Do Bayes Factors Need?

The most fundamental question for study planning is: given a real effect, how
many participants and trials are needed for the Bayes Factor to detect it? We
examine power — the probability that BF10 exceeds 3, providing at least
moderate evidence for the alternative — across the full range of sample sizes,
trial counts, and prior specifications.

### Power by sample size and effect size

```{r}
#| label: fig-power-n
#| fig-cap: "Power to detect effects increases with sample size and effect
#|   magnitude. With 50 trials per subject and the recommended effect prior
#|   (sd = 0.25), power exceeds 80% at N = 30 for effects of b1 >= 0.20. The
#|   tighter prior (sd = 0.25, solid lines) yields slightly higher power than
#|   the wider prior (sd = 0.50, dashed), but the difference is small compared
#|   to the effect of sample size and true effect. Prior family (color) has
#|   minimal impact on power. Error bars: 95% credible intervals."
#| fig-height: 6
#| fig-width: 10

if (posteriors_available) {
  matched <- posteriors |>
    filter(
      dist_b0 == "logistic",
      true_b1 > 0,
      threshold == 3
    ) |>
    mutate(
      dist_b1  = factor(dist_b1, levels = c("cauchy", "normal", "logistic"),
                        labels = c("Cauchy", "Normal", "Logistic")),
      sd_b1    = factor(paste0("sd = ", sd_b1))
    )

  ggplot(matched, aes(x = n_subjects, y = power_post_mean,
                      color = dist_b1, linetype = sd_b1)) +
    geom_line(linewidth = 0.8) +
    geom_point(size = 2) +
    geom_errorbar(aes(ymin = power_ci_lower, ymax = power_ci_upper),
                  width = 3, linewidth = 0.4, alpha = 0.6) +
    geom_hline(yintercept = 0.80, linetype = "dotted", color = "grey40") +
    scale_x_continuous("Number of subjects", breaks = c(30, 60, 100)) +
    scale_y_continuous("P(BF10 > 3)", limits = c(NA, 1)) +
    scale_color_viridis_d("Prior family", option = "D", end = 0.85) +
    scale_linetype_manual("Prior SD",
      values = c("sd = 0.25" = "solid", "sd = 0.5" = "dashed")) +
    facet_grid(n_trials ~ true_b1,
               labeller = labeller(
                 true_b1  = function(x) paste0("b1 = ", x),
                 n_trials = function(x) paste0(x, " trials")
               )) +
    labs(title = "Power: P(BF10 > 3) by sample size and true effect",
         subtitle = "Matched intercept prior (Logistic, sd = 0.75); dotted line = 80% power") +
    base_theme()
} else {
  matched <- summaries |>
    filter(dist_b0 == "logistic", true_b1 > 0) |>
    mutate(
      dist_b1  = factor(dist_b1, levels = c("cauchy", "normal", "logistic"),
                        labels = c("Cauchy", "Normal", "Logistic")),
      sd_b1    = factor(paste0("sd = ", sd_b1))
    )

  ggplot(matched, aes(x = n_subjects, y = P_BF10_gt3,
                      color = dist_b1, linetype = sd_b1)) +
    geom_line(linewidth = 0.8) +
    geom_point(size = 2) +
    geom_hline(yintercept = 0.80, linetype = "dotted", color = "grey40") +
    scale_x_continuous("Number of subjects", breaks = c(30, 60, 100)) +
    scale_y_continuous("P(BF10 > 3)", limits = c(NA, 1)) +
    scale_color_viridis_d("Prior family", option = "D", end = 0.85) +
    scale_linetype_manual("Prior SD",
      values = c("sd = 0.25" = "solid", "sd = 0.5" = "dashed")) +
    facet_grid(n_trials ~ true_b1,
               labeller = labeller(
                 true_b1  = function(x) paste0("b1 = ", x),
                 n_trials = function(x) paste0(x, " trials")
               )) +
    labs(title = "Power: P(BF10 > 3) by sample size and true effect",
         subtitle = "Matched intercept prior (Logistic, sd = 0.75); dotted line = 80% power") +
    base_theme()
}
```

For larger effects ($b_1 \geq 0.20$), @fig-power-n shows that power reaches
or exceeds 80% even at the smallest sample size ($n = 30$) when subjects
complete 50 trials. The tighter prior ($\text{sd} = 0.25$, solid lines) yields
slightly higher power than the wider alternative ($\text{sd} = 0.50$, dashed),
because concentrating the prior near plausible effect sizes makes the
Savage--Dickey ratio more sensitive to genuine shifts away from zero. But this
advantage is modest; the true effect magnitude and the amount of data drive
power far more than the prior specification does.

Two additional patterns stand out. The three prior families (Normal, Logistic,
Cauchy) produce nearly overlapping power curves, foreshadowing a conclusion we
return to below: the choice of prior *family* matters far less than the choice
of prior *scale*. And the difference between 20 and 50 trials per subject is
most visible at smaller effects, where 20 trials may not suffice. Increasing
trials per subject can be at least as effective as recruiting more
participants.

### The role of trials per subject

Power also depends strongly on how many trials each subject completes. @fig-power-trials
isolates this factor for a mid-range sample size ($n = 60$).

```{r}
#| label: fig-power-trials
#| fig-cap: "Increasing from 20 to 50 trials per subject substantially
#|   improves power, particularly for smaller effects. Shown for N = 60
#|   subjects with matched Logistic intercept prior. Error bars: 95%
#|   credible intervals."
#| fig-height: 5
#| fig-width: 10

if (posteriors_available) {
  posteriors |>
    filter(
      dist_b0 == "logistic",
      true_b1 > 0,
      threshold == 3,
      n_subjects == 60
    ) |>
    mutate(
      dist_b1  = factor(dist_b1, levels = c("cauchy", "normal", "logistic"),
                        labels = c("Cauchy", "Normal", "Logistic")),
      sd_b1    = factor(paste0("sd = ", sd_b1))
    ) |>
    ggplot(aes(x = n_trials, y = power_post_mean,
               color = dist_b1, linetype = sd_b1)) +
    geom_line(linewidth = 0.8) +
    geom_point(size = 2) +
    geom_errorbar(aes(ymin = power_ci_lower, ymax = power_ci_upper),
                  width = 2, linewidth = 0.4, alpha = 0.6) +
    geom_hline(yintercept = 0.80, linetype = "dotted", color = "grey40") +
    scale_x_continuous("Trials per subject", breaks = c(20, 50)) +
    scale_y_continuous("P(BF10 > 3)", limits = c(NA, 1)) +
    scale_color_viridis_d("Prior family", option = "D", end = 0.85) +
    scale_linetype_manual("Prior SD",
      values = c("sd = 0.25" = "solid", "sd = 0.5" = "dashed")) +
    facet_wrap(~ true_b1,
               labeller = labeller(
                 true_b1 = function(x) paste0("b1 = ", x)
               )) +
    labs(title = "Power by trials per subject (n_subjects = 60, matched intercept prior)") +
    base_theme()
} else {
  matched |>
    filter(n_subjects == 60) |>
    ggplot(aes(x = n_trials, y = P_BF10_gt3,
               color = dist_b1, linetype = sd_b1)) +
    geom_line(linewidth = 0.8) +
    geom_point(size = 2) +
    geom_hline(yintercept = 0.80, linetype = "dotted", color = "grey40") +
    scale_x_continuous("Trials per subject", breaks = c(20, 50)) +
    scale_y_continuous("P(BF10 > 3)", limits = c(NA, 1)) +
    scale_color_viridis_d("Prior family", option = "D", end = 0.85) +
    scale_linetype_manual("Prior SD",
      values = c("sd = 0.25" = "solid", "sd = 0.5" = "dashed")) +
    facet_wrap(~ true_b1,
               labeller = labeller(
                 true_b1 = function(x) paste0("b1 = ", x)
               )) +
    labs(title = "Power by trials per subject (n_subjects = 60, matched intercept prior)") +
    base_theme()
}
```

Increasing trials from 20 to 50 per subject substantially improves power across
all prior configurations. More trials per subject reduce within-subject
variability, which sharpens the posterior for $b_1$ and makes the Bayes Factor
more decisive. For study design, this means collecting more trials from each
participant (e.g., by extending the session) can be as effective as recruiting
additional participants, and is often easier to arrange.

## Do These Priors Produce False Evidence?

Good power is only valuable if the Bayes Factor does not also produce spurious
evidence for effects that do not exist. We therefore examine specificity: when
the null is true ($b_1 = 0$), how often does the BF correctly support H0 (i.e.,
BF01 > 3)?

```{r}
#| label: fig-specificity
#| fig-cap: "When no effect exists, the Bayes Factor correctly supports the
#|   null in the large majority of replications. Wider priors (sd = 0.50,
#|   dashed) yield substantially higher specificity because they place less
#|   density at exactly zero, making the Savage-Dickey ratio more conservative.
#|   At sd = 0.25 (solid lines), the Normal prior shows notably lower
#|   specificity than the Logistic and Cauchy, especially at small sample
#|   sizes — a consequence of its higher density at exactly zero (see text).
#|   Error bars: 95% credible intervals."
#| fig-height: 5
#| fig-width: 10

if (posteriors_available) {
  posteriors |>
    filter(
      true_b1 == 0,
      dist_b0 == "logistic",
      threshold == 3
    ) |>
    mutate(
      dist_b1 = factor(dist_b1, levels = c("cauchy", "normal", "logistic"),
                       labels = c("Cauchy", "Normal", "Logistic")),
      sd_b1   = factor(paste0("sd = ", sd_b1))
    ) |>
    ggplot(aes(x = n_subjects, y = spec_post_mean,
               color = dist_b1, linetype = sd_b1)) +
    geom_line(linewidth = 0.8) +
    geom_point(size = 2) +
    geom_errorbar(aes(ymin = spec_ci_lower, ymax = spec_ci_upper),
                  width = 3, linewidth = 0.4, alpha = 0.6) +
    scale_x_continuous("Number of subjects", breaks = c(30, 60, 100)) +
    scale_y_continuous("P(BF01 > 3)", limits = c(0, 1)) +
    scale_color_viridis_d("Prior family", option = "D", end = 0.85) +
    scale_linetype_manual("Prior SD",
      values = c("sd = 0.25" = "solid", "sd = 0.5" = "dashed")) +
    facet_wrap(~ n_trials,
               labeller = labeller(
                 n_trials = function(x) paste0(x, " trials")
               )) +
    labs(title = "Specificity under H0: P(BF01 > 3)",
         subtitle = "Matched intercept prior (Logistic, sd = 0.75)") +
    base_theme()
} else {
  summaries |>
    filter(true_b1 == 0, dist_b0 == "logistic") |>
    mutate(
      dist_b1 = factor(dist_b1, levels = c("cauchy", "normal", "logistic"),
                       labels = c("Cauchy", "Normal", "Logistic")),
      sd_b1   = factor(paste0("sd = ", sd_b1))
    ) |>
    ggplot(aes(x = n_subjects, y = P_BF01_gt3,
               color = dist_b1, linetype = sd_b1)) +
    geom_line(linewidth = 0.8) +
    geom_point(size = 2) +
    scale_x_continuous("Number of subjects", breaks = c(30, 60, 100)) +
    scale_y_continuous("P(BF01 > 3)", limits = c(0, 1)) +
    scale_color_viridis_d("Prior family", option = "D", end = 0.85) +
    scale_linetype_manual("Prior SD",
      values = c("sd = 0.25" = "solid", "sd = 0.5" = "dashed")) +
    facet_wrap(~ n_trials,
               labeller = labeller(
                 n_trials = function(x) paste0(x, " trials")
               )) +
    labs(title = "Specificity under H0: P(BF01 > 3)",
         subtitle = "Matched intercept prior (Logistic, sd = 0.75)") +
    base_theme()
}
```

Specificity is high across most configurations, but at $\text{sd}_{b_1} = 0.25$
(solid lines), the Normal prior yields substantially lower specificity than the
Logistic and Cauchy, particularly at small sample sizes (@fig-specificity). At
$n = 30$ with 20 trials, the Normal prior correctly supports H0 in only 41% of
null-effect replications, compared to 69% for the Logistic and 76% for the
Cauchy.

The reason is that the three families differ in their density at exactly zero
when the scale parameter is equated. At $\text{sd} = 0.25$:
$\text{Normal}(0, 0.25)$ places density 1.60 at zero, $\text{Cauchy}(0, 0.25)$
places 1.27, and $\text{Logistic}(0, 0.25)$ places 1.00. A higher prior density
at zero makes the numerator of the Savage--Dickey ratio
$\pi(b_1 = 0) / p(b_1 = 0 \mid \text{data})$ larger, so the posterior must pile
up more at zero before $\text{BF}_{01}$ crosses the threshold, reducing
specificity. The wider prior ($\text{sd} = 0.50$, dashed lines) lowers the
prior density at zero for all families (Normal: 0.80, Cauchy: 0.64, Logistic:
0.50), which is why specificity improves and between-family differences shrink
at the wider scale.

At the recommended $\text{sd}_{b_1} = 0.25$, then, the Logistic prior has the
best specificity (lowest density at zero), the Cauchy is intermediate, and the
Normal has the lowest, though all three converge to adequate specificity at
larger sample sizes ($n \geq 60$ with 50 trials). False positive rates
($\text{BF}_{10} > 3$ under H0) remain below 5% across all conditions.

Given that the choice of prior family affects specificity through the
Savage–Dickey mechanism (though not power), we next ask directly: how similar are
the three prior families in their overall BF performance?

## Does the Choice of Prior Family Matter?

A researcher choosing a prior for the effect coefficient faces three common
options — Normal, Logistic, and Cauchy — that differ in their tail behavior.
The prior predictive analysis showed that the Cauchy's heavy tails assign
implausible probability to extreme effects. But does this translate into
meaningful differences in Bayes Factor performance? @fig-prior-sensitivity
addresses this question by holding the prior scale constant at
$\text{sd}_{b_1} = 0.25$ and comparing power across families.

```{r}
#| label: fig-prior-sensitivity
#| fig-cap: "At the same prior scale (sd = 0.25), the three prior families
#|   produce nearly identical power curves. This holds across all sample sizes,
#|   trial counts, and effect sizes examined, indicating that the choice of
#|   prior family is far less consequential than the choice of prior scale.
#|   Error bars: 95% credible intervals."
#| fig-height: 6
#| fig-width: 10

if (posteriors_available) {
  posteriors |>
    filter(
      sd_b1 == 0.25,
      dist_b0 == "logistic",
      true_b1 > 0,
      threshold == 3
    ) |>
    mutate(
      dist_b1 = factor(dist_b1, levels = c("cauchy", "normal", "logistic"),
                       labels = c("Cauchy", "Normal", "Logistic"))
    ) |>
    ggplot(aes(x = true_b1, y = power_post_mean, color = dist_b1)) +
    geom_line(linewidth = 0.8) +
    geom_point(size = 2) +
    geom_errorbar(aes(ymin = power_ci_lower, ymax = power_ci_upper),
                  width = 0.02, linewidth = 0.4, alpha = 0.6) +
    geom_hline(yintercept = 0.80, linetype = "dotted", color = "grey40") +
    scale_x_continuous("True effect b1") +
    scale_y_continuous("P(BF10 > 3)", limits = c(NA, 1)) +
    scale_color_viridis_d("Prior family", option = "D", end = 0.85) +
    facet_grid(n_trials ~ n_subjects,
               labeller = labeller(
                 n_subjects = function(x) paste0(x, " subj"),
                 n_trials   = function(x) paste0(x, " trials")
               )) +
    labs(title = "Sensitivity to prior family (sd_b1 = 0.25, matched intercept prior)") +
    base_theme()
} else {
  summaries |>
    filter(sd_b1 == 0.25, dist_b0 == "logistic", true_b1 > 0) |>
    mutate(
      dist_b1 = factor(dist_b1, levels = c("cauchy", "normal", "logistic"),
                       labels = c("Cauchy", "Normal", "Logistic"))
    ) |>
    ggplot(aes(x = true_b1, y = P_BF10_gt3, color = dist_b1)) +
    geom_line(linewidth = 0.8) +
    geom_point(size = 2) +
    geom_hline(yintercept = 0.80, linetype = "dotted", color = "grey40") +
    scale_x_continuous("True effect b1") +
    scale_y_continuous("P(BF10 > 3)", limits = c(NA, 1)) +
    scale_color_viridis_d("Prior family", option = "D", end = 0.85) +
    facet_grid(n_trials ~ n_subjects,
               labeller = labeller(
                 n_subjects = function(x) paste0(x, " subj"),
                 n_trials   = function(x) paste0(x, " trials")
               )) +
    labs(title = "Sensitivity to prior family (sd_b1 = 0.25, matched intercept prior)") +
    base_theme()
}
```

For power, the three families are nearly interchangeable: at the same prior
scale, they produce overlapping power curves across all sample sizes and effect
sizes. Specificity is more differentiated. As shown in @fig-specificity, the
Normal prior's higher density at zero reduces its specificity relative to the
Logistic and Cauchy at $\text{sd}_{b_1} = 0.25$, especially at small $n$.
Taken together, these results give a practical edge to the Logistic prior: it
has the best specificity profile (lowest density at zero), well-behaved prior
predictive distributions, and identical power to the alternatives. The Cauchy
produces acceptable power but assigns implausible probability to extreme
effects with no compensating advantage. We therefore recommend the **Logistic
effect prior** at $\text{sd}_{b_1} = 0.25$ as the default for Bayes Factor
testing in binomial GLMs.

A similar question arises for the intercept prior: does the matched-prior
principle — which the prior predictive analysis showed produces cleaner
distributions on the probability scale — also matter for Bayes Factor calibration?

## Does the Intercept Prior Affect Bayes Factor Calibration?

The prior predictive analysis established the matched-prior principle: placing
a Logistic prior on the logit-scale intercept produces a uniform distribution on
the probability scale, whereas a Normal prior (misfit) concentrates mass
differently. But does this distinction matter for Bayes Factor calibration?
@fig-misfit-bf compares power under the matched (Logistic) and misfit (Normal)
intercept priors, holding the effect prior constant.

```{r}
#| label: fig-misfit-bf
#| fig-cap: "Using a mismatched intercept prior (Normal instead of the matched
#|   Logistic) has negligible impact on power. The two intercept priors produce
#|   overlapping power curves across all effect sizes, sample sizes, and trial
#|   counts. This confirms that the intercept prior primarily regularizes the
#|   baseline probability and does not alter inference about the effect.
#|   Error bars: 95% credible intervals."
#| fig-height: 6
#| fig-width: 10

if (posteriors_available) {
  posteriors |>
    filter(
      dist_b1 == "logistic",
      sd_b1   == 0.25,
      true_b1 > 0,
      threshold == 3
    ) |>
    mutate(
      dist_b0 = factor(dist_b0, levels = c("logistic", "normal"),
                       labels = c("Logistic (matched)", "Normal (misfit)"))
    ) |>
    ggplot(aes(x = true_b1, y = power_post_mean, color = dist_b0)) +
    geom_line(linewidth = 0.8) +
    geom_point(size = 2) +
    geom_errorbar(aes(ymin = power_ci_lower, ymax = power_ci_upper),
                  width = 0.02, linewidth = 0.4, alpha = 0.6) +
    geom_hline(yintercept = 0.80, linetype = "dotted", color = "grey40") +
    scale_x_continuous("True effect b1") +
    scale_y_continuous("P(BF10 > 3)", limits = c(NA, 1)) +
    scale_color_viridis_d("Intercept prior", option = "D", end = 0.7) +
    facet_grid(n_trials ~ n_subjects,
               labeller = labeller(
                 n_subjects = function(x) paste0(x, " subjects"),
                 n_trials   = function(x) paste0(x, " trials")
               )) +
    labs(title = "Effect of intercept prior misfit on power (logit link, sd_b1 = 0.25)") +
    base_theme()
} else {
  summaries |>
    filter(dist_b1 == "logistic", sd_b1 == 0.25, true_b1 > 0) |>
    mutate(
      dist_b0 = factor(dist_b0, levels = c("logistic", "normal"),
                       labels = c("Logistic (matched)", "Normal (misfit)"))
    ) |>
    ggplot(aes(x = true_b1, y = P_BF10_gt3, color = dist_b0)) +
    geom_line(linewidth = 0.8) +
    geom_point(size = 2) +
    geom_hline(yintercept = 0.80, linetype = "dotted", color = "grey40") +
    scale_x_continuous("True effect b1") +
    scale_y_continuous("P(BF10 > 3)", limits = c(NA, 1)) +
    scale_color_viridis_d("Intercept prior", option = "D", end = 0.7) +
    facet_grid(n_trials ~ n_subjects,
               labeller = labeller(
                 n_subjects = function(x) paste0(x, " subjects"),
                 n_trials   = function(x) paste0(x, " trials")
               )) +
    labs(title = "Effect of intercept prior misfit on power (logit link, sd_b1 = 0.25)") +
    base_theme()
}
```

The matched and misfit intercept priors produce essentially overlapping power
curves. The intercept prior regularizes the baseline probability, keeping it
away from 0 and 1, but has no meaningful effect on inference about $b_1$ as
long as the prior SD is reasonable (here, $\text{sd}_{b_0} = 0.75$). Even a
"misfit" Normal intercept prior on the logit scale does not compromise BF
calibration.

## What Happens When All Three Outcomes Are Considered?

So far we have examined power and specificity separately — asking how often the
Bayes Factor correctly supports H1 when effects exist, and how often it supports
H0 when they do not. But each replication actually falls into one of three
categories: the BF may favor H1 (BF10 > 3), favor H0 (BF01 > 3), or remain
inconclusive (neither threshold crossed). Decomposing outcomes this way reveals
where the "undecided" fraction concentrates and what conditions demand the most
attention to sample size.

```{r}
#| label: fig-three-category
#| fig-cap: "When effects are large or absent, the Bayes Factor is decisive in
#|   the majority of replications. The inconclusive fraction (middle category)
#|   is largest at small effects with few trials — the conditions where sample
#|   size recommendations matter most. Under the null, most replications
#|   correctly support H0, with a small fraction remaining undecided.
#|   Bars show Dirichlet posterior means for the recommended Normal effect
#|   prior (sd = 0.25) with matched Logistic intercept prior."
#| fig-height: 7
#| fig-width: 10
#| eval: !expr posteriors_available

posteriors |>
  filter(
    dist_b0 == "logistic",
    dist_b1 == "normal",
    sd_b1   == 0.25,
    threshold == 3
  ) |>
  select(true_b1, n_subjects, n_trials, h1_mean, inc_mean, h0_mean) |>
  pivot_longer(cols = c(h1_mean, inc_mean, h0_mean),
               names_to = "outcome", values_to = "proportion") |>
  mutate(
    outcome = factor(outcome,
      levels = c("h1_mean", "inc_mean", "h0_mean"),
      labels = c("H1 detected (BF10 > 3)",
                  "Inconclusive",
                  "H0 detected (BF01 > 3)")),
    true_b1_label = factor(paste0("b1 = ", true_b1))
  ) |>
  ggplot(aes(x = factor(n_subjects), y = proportion, fill = outcome)) +
  geom_col(position = "stack", width = 0.7) +
  facet_grid(n_trials ~ true_b1_label,
             labeller = labeller(
               n_trials = function(x) paste0(x, " trials")
             )) +
  scale_x_discrete("Number of subjects") +
  scale_y_continuous("Proportion of replications", limits = c(0, 1)) +
  scale_fill_viridis_d("BF outcome", option = "D", end = 0.85) +
  labs(title = "Three-category BF outcome decomposition (BF threshold = 3)",
       subtitle = "Normal effect prior (sd = 0.25), matched Logistic intercept prior") +
  base_theme()
```

Under the null ($b_1 = 0$), the large majority of replications correctly favor
H0 (@fig-three-category). As the true effect grows, the H1-detected fraction
dominates. The inconclusive fraction is largest at small effects ($b_1 = 0.05$
and $0.10$) with 20 trials, where the BF neither confirms nor rules out the
effect in a substantial share of replications. This is where collecting more
data --- more trials per subject or more subjects --- makes the biggest
practical difference. At 50 trials, the inconclusive fraction shrinks
considerably even for small effects.

The following sections address the role of the random-effects hyperprior, which
operates through a different mechanism than the fixed-effects priors examined so
far.

## Does the Random-Effects Hyperprior Matter? {#sec-re-hyperprior}

The analyses above used a single random-effects hyperprior
($\text{Exponential}(4)$) and focused exclusively on the fixed-effects priors.
But the prior predictive analysis showed that the choice of hyperprior affects
the distribution of subject-level probabilities — the brms default
Student-*t*(3, 0, 2.5) places a heavy right tail on between-subject SDs,
pushing individual participants toward floor and ceiling. Does this also affect
Bayes Factor performance?

To answer this, we ran a second simulation with a within-subjects design
(each subject in both conditions, with correlated random intercepts and slopes)
while holding the fixed-effects priors at the recommended matched configuration.
The only design factor was the hyperprior on the random-effects SDs: the brms
default Student-*t*(3, 0, 2.5), Gamma(2, 4), and Exponential(4).

```{r}
#| label: setup-re
#| eval: !expr re_results_available

re_plot_data <- re_summaries |>
  mutate(
    sd_prior_re = factor(sd_prior_re,
                         levels = c("default", "gamma", "exponential"),
                         labels = c("Student-t(3,0,2.5)", "Gamma(2,4)",
                                    "Exponential(4)")),
    n_trials_label = paste0(n_trials, " trials")
  )
```

```{r}
#| label: setup-re-posteriors
#| eval: !expr re_posteriors_available

re_post_data <- re_posteriors |>
  mutate(
    sd_prior_re = factor(sd_prior_re,
                         levels = c("default", "gamma", "exponential"),
                         labels = c("Student-t(3,0,2.5)", "Gamma(2,4)",
                                    "Exponential(4)")),
    n_trials_label = paste0(n_trials, " trials")
  )
```

### Power is unaffected by the RE hyperprior

```{r}
#| label: fig-re-power
#| fig-cap: "The choice of random-effects hyperprior has no meaningful impact
#|   on power. All three hyperpriors produce nearly identical power curves
#|   across all effect sizes and sample sizes. This confirms that Bayes Factor
#|   power is driven by the fixed-effects priors and the amount of data, not
#|   by the RE hyperprior. Within-subjects design with correlated RE; matched
#|   fixed-effects priors. Error bars: 95% credible intervals."
#| fig-height: 6
#| fig-width: 10
#| eval: !expr re_posteriors_available || re_results_available

if (re_posteriors_available) {
  re_post_data |>
    filter(true_b1 > 0, threshold == 3) |>
    ggplot(aes(x = n_subjects, y = power_post_mean, color = sd_prior_re,
               linetype = sd_prior_re)) +
    geom_line(linewidth = 0.8) +
    geom_point(size = 2) +
    geom_errorbar(aes(ymin = power_ci_lower, ymax = power_ci_upper),
                  width = 3, linewidth = 0.4, alpha = 0.6) +
    geom_hline(yintercept = 0.80, linetype = "dotted", color = "grey40") +
    scale_x_continuous("Number of subjects", breaks = c(30, 60, 100)) +
    scale_y_continuous("P(BF10 > 3)", limits = c(NA, 1)) +
    scale_color_viridis_d("RE hyperprior", option = "D", end = 0.85) +
    scale_linetype_manual("RE hyperprior",
      values = c("Student-t(3,0,2.5)" = "solid",
                 "Gamma(2,4)"         = "dashed",
                 "Exponential(4)"     = "dotdash")) +
    facet_grid(n_trials_label ~ true_b1,
               labeller = labeller(
                 true_b1 = function(x) paste0("b1 = ", x)
               )) +
    labs(title = "Power by RE hyperprior (within-subjects, matched fixed-effects priors)",
         subtitle = "Dotted line = 80% power criterion") +
    base_theme()
} else {
  re_plot_data |>
    filter(true_b1 > 0) |>
    ggplot(aes(x = n_subjects, y = P_BF10_gt3, color = sd_prior_re,
               linetype = sd_prior_re)) +
    geom_line(linewidth = 0.8) +
    geom_point(size = 2) +
    geom_hline(yintercept = 0.80, linetype = "dotted", color = "grey40") +
    scale_x_continuous("Number of subjects", breaks = c(30, 60, 100)) +
    scale_y_continuous("P(BF10 > 3)", limits = c(NA, 1)) +
    scale_color_viridis_d("RE hyperprior", option = "D", end = 0.85) +
    scale_linetype_manual("RE hyperprior",
      values = c("Student-t(3,0,2.5)" = "solid",
                 "Gamma(2,4)"         = "dashed",
                 "Exponential(4)"     = "dotdash")) +
    facet_grid(n_trials_label ~ true_b1,
               labeller = labeller(
                 true_b1 = function(x) paste0("b1 = ", x)
               )) +
    labs(title = "Power by RE hyperprior (within-subjects, matched fixed-effects priors)",
         subtitle = "Dotted line = 80% power criterion") +
    base_theme()
}
```

The RE hyperprior has no meaningful impact on power (@fig-re-power). All three
hyperpriors produce nearly identical detection rates across effect sizes and
sample sizes. This is expected: the Bayes Factor for $b_1$ is driven by the
fixed-effects priors and the data, not by how between-subject variability is
modeled.

### Specificity is similar across RE hyperpriors

Does the RE hyperprior affect specificity — the ability to correctly support the
null when there is no effect? @fig-re-specificity addresses this question.

```{r}
#| label: fig-re-specificity
#| fig-cap: "All three RE hyperpriors produce broadly similar specificity,
#|   with widely overlapping credible intervals across all conditions. At
#|   N = 30 with 20 trials, specificity is low (~43%) regardless of
#|   hyperprior choice. At larger samples, the three hyperpriors remain
#|   closely matched. Error bars: 95% credible intervals."
#| fig-height: 5
#| fig-width: 10
#| eval: !expr re_posteriors_available || re_results_available

if (re_posteriors_available) {
  re_post_data |>
    filter(true_b1 == 0, threshold == 3) |>
    ggplot(aes(x = n_subjects, y = spec_post_mean, color = sd_prior_re,
               linetype = sd_prior_re)) +
    geom_line(linewidth = 0.8) +
    geom_point(size = 2) +
    geom_errorbar(aes(ymin = spec_ci_lower, ymax = spec_ci_upper),
                  width = 3, linewidth = 0.4, alpha = 0.6) +
    scale_x_continuous("Number of subjects", breaks = c(30, 60, 100)) +
    scale_y_continuous("P(BF01 > 3)", limits = c(0, 1)) +
    scale_color_viridis_d("RE hyperprior", option = "D", end = 0.85) +
    scale_linetype_manual("RE hyperprior",
      values = c("Student-t(3,0,2.5)" = "solid",
                 "Gamma(2,4)"         = "dashed",
                 "Exponential(4)"     = "dotdash")) +
    facet_wrap(~ n_trials_label) +
    labs(title = "Specificity under H0 by RE hyperprior") +
    base_theme()
} else {
  re_plot_data |>
    filter(true_b1 == 0) |>
    ggplot(aes(x = n_subjects, y = P_BF01_gt3, color = sd_prior_re,
               linetype = sd_prior_re)) +
    geom_line(linewidth = 0.8) +
    geom_point(size = 2) +
    scale_x_continuous("Number of subjects", breaks = c(30, 60, 100)) +
    scale_y_continuous("P(BF01 > 3)", limits = c(0, 1)) +
    scale_color_viridis_d("RE hyperprior", option = "D", end = 0.85) +
    scale_linetype_manual("RE hyperprior",
      values = c("Student-t(3,0,2.5)" = "solid",
                 "Gamma(2,4)"         = "dashed",
                 "Exponential(4)"     = "dotdash")) +
    facet_wrap(~ n_trials_label) +
    labs(title = "Specificity under H0 by RE hyperprior") +
    base_theme()
}
```

The three RE hyperpriors produce broadly similar specificity across all
conditions, with widely overlapping 95% credible intervals
(@fig-re-specificity). At $n = 30$ with only 20 trials per subject, all three
yield similarly low specificity (Student-*t*: 0.44 [0.35, 0.54]; Gamma: 0.43
[0.34, 0.53]; Exponential: 0.42 [0.33, 0.52]); the low values reflect the
limited data, not the hyperprior. At larger sample sizes, the three remain
closely matched. Unlike the fixed-effects prior family (which affects
specificity through the Savage--Dickey density at zero), the RE hyperprior does
not meaningfully influence inferential performance for
the fixed effect. The recommendation for Exponential(4) as a default hyperprior
therefore rests on its prior predictive properties — it concentrates
between-subject variability in a substantively plausible range and avoids the
heavy right tail of the Student-*t* that can push individual probabilities
toward floor and ceiling (see the prior predictive analysis) — rather than on
BF calibration differences.

@tbl-re-summary provides the corresponding minimum sample sizes for 80% power,
confirming that all three hyperpriors require the same sample sizes to reach
adequate power.

```{r}
#| label: tbl-re-summary
#| tbl-cap: "Minimum sample size for 80% power (50 trials per subject) is
#|   identical across RE hyperpriors, confirming that the hyperprior does not
#|   affect power. Within-subjects design with correlated RE;
#|   matched fixed-effects priors."
#| eval: !expr re_results_available

re_plot_data |>
  filter(n_trials == 50, true_b1 > 0) |>
  group_by(sd_prior_re, true_b1) |>
  summarise(
    min_n_80pct = {
      n_meeting <- n_subjects[P_BF10_gt3 >= 0.80]
      if (length(n_meeting) > 0) min(n_meeting) else NA_integer_
    },
    .groups = "drop"
  ) |>
  pivot_wider(names_from = true_b1, values_from = min_n_80pct,
              names_prefix = "b1 = ") |>
  rename("RE hyperprior" = sd_prior_re) |>
  knitr::kable(caption = "Minimum n_subjects for 80% power by RE hyperprior (n_trials = 50)")
```

## Putting It All Together

Across all conditions examined, Bayes Factor calibration is primarily driven by
two factors: the scale of the effect prior and the amount of data (both sample
size and trials per subject). The choice of prior *family* has negligible impact
on power, but does affect specificity at the recommended scale of
$\text{sd}_{b_1} = 0.25$: the Normal prior's higher density at zero reduces its
specificity relative to the Logistic and Cauchy, particularly at small $n$. The
intercept prior (matched vs. misfit) makes no practical difference. The
random-effects hyperprior has no systematic effect on either power or specificity
— the three hyperpriors examined produce broadly similar results across all
conditions, though the Exponential(4) is preferred on prior predictive grounds.

These results confirm and extend the prior predictive analysis. The matched
priors recommended there — Logistic(0, 0.75) for the intercept and a Logistic
effect prior at $\text{sd}_{b_1} = 0.25$ — not only produce substantively
plausible prior predictive distributions, but also yield well-calibrated Bayes
Factor inference with the best specificity profile among the families examined.
Priors that are too wide degrade power, while the Normal effect prior at the
recommended scale incurs a specificity cost at small samples that favours the
Logistic alternative.

### Scale matters most

The choice that matters most is the *scale* of the effect prior, not the family,
the intercept prior, or the random-effects hyperprior. This parallels findings
from the BFDA literature [@schonbrodt2018; @stefan2019], where informed priors at
appropriate scales consistently outperform uninformed defaults. What the present
work adds is establishing *what* that appropriate scale is for binomial models
and *why*, connecting the prior scale to interpretable quantities on the
probability scale via the matched prior principle.

### A template for other model classes

The workflow here --- (1) derive matched priors analytically, (2) characterise
the prior--observation mapping via prior predictive simulation, (3) calibrate BF
performance via BFDA --- is not specific to binomial models. Any model class
with a nonlinear link function faces the same challenge: ordinal models
(cumulative logit/probit), count models (log link), beta regression (logit link
on the mean), and distributional regression models [@klein2015; @burkner2021].
In each case, the probability integral transform identifies the matched prior
family, prior predictive simulation reveals the implied effect sizes, and BFDA
quantifies inferential properties. The same template can establish principled
defaults for Bayes Factor testing beyond the normal model, where the
Rouder--Wagenmakers framework [@rouder2009; @rouder2012] does not currently
reach.

## Sample Size Guidance

For practical study planning, the following tables summarize the minimum sample
sizes needed to achieve 80% power across different prior specifications and
effect sizes, assuming 50 trials per subject and the recommended matched
intercept prior.

```{r}
#| label: tbl-summary
#| tbl-cap: "Minimum number of subjects for 80% power (P(BF10 > 3) >= 0.80)
#|   at 50 trials per subject, by prior family, prior SD, and true effect size.
#|   Between-subjects design with matched Logistic intercept prior."

summaries |>
  filter(
    n_trials == 50,
    dist_b0 == "logistic",
    true_b1 > 0
  ) |>
  group_by(dist_b1, sd_b1, true_b1) |>
  summarise(
    min_n_80pct = {
      n_meeting <- n_subjects[P_BF10_gt3 >= 0.80]
      if (length(n_meeting) > 0) min(n_meeting) else NA_integer_
    },
    .groups = "drop"
  ) |>
  mutate(
    dist_b1 = factor(dist_b1, levels = c("cauchy", "normal", "logistic"),
                     labels = c("Cauchy", "Normal", "Logistic"))
  ) |>
  pivot_wider(names_from = true_b1, values_from = min_n_80pct,
              names_prefix = "b1 = ") |>
  knitr::kable(caption = "Minimum n_subjects for 80% power (n_trials = 50, matched intercept prior)")
```

@tbl-summary-ci provides more detail for the recommended configuration (Normal
effect prior, $\text{sd}_{b_1} = 0.25$), showing the posterior mean and 95%
credible interval for power at each combination of effect size and sample size.

```{r}
#| label: tbl-summary-ci
#| tbl-cap: "Power estimates with uncertainty for the recommended prior
#|   configuration (Normal effect prior, sd = 0.25, 50 trials per subject,
#|   matched Logistic intercept prior). Values show posterior mean [95%
#|   credible interval] for P(BF10 > 3)."
#| eval: !expr posteriors_available

posteriors |>
  filter(
    n_trials == 50,
    dist_b0 == "logistic",
    dist_b1 == "normal",
    sd_b1   == 0.25,
    true_b1 > 0,
    threshold == 3
  ) |>
  mutate(
    power_summary = sprintf("%.2f [%.2f, %.2f]",
                            power_post_mean, power_ci_lower, power_ci_upper)
  ) |>
  select(true_b1, n_subjects, power_summary) |>
  pivot_wider(names_from = true_b1, values_from = power_summary,
              names_prefix = "b1 = ") |>
  rename("n_subjects" = n_subjects) |>
  knitr::kable(caption = "Power posterior mean [95% CI] (Normal prior, sd = 0.25, n_trials = 50)")
```

## Recommendations for Practice

The most important decision is the **scale of the effect prior**. An effect
prior with $\text{sd}_{b_1} = 0.25$ (under $\pm 1$ sum-to-zero coding) balances
power and specificity for typical cognitive psychology experiments. At this
scale, the prior concentrates mass on effects up to roughly 20 percentage points
on the probability scale --- the range most commonly observed in cognitive
experiments [@oberauer2019] --- while still allowing larger effects to receive
non-trivial prior probability. The prior predictive analysis confirmed that
$\text{sd}_{b_1} = 0.25$ yields a 90th percentile of
$|\delta p| \approx 0.20$.

This parallels a core insight from BFDA [@schonbrodt2018]: prior specification
is a design decision that directly affects what evidence a study can produce.
For normal-data models, the JZS prior [@rouder2009] provides a principled scale
through the Cauchy distribution on standardised effect sizes. For binomial GLMs,
the link function makes standardisation less natural, and the matched prior
principle offers an alternative anchor grounded in the probability integral
transform.

The **choice of prior family matters far less than the choice of scale** for
power, but it does affect specificity at $\text{sd}_{b_1} = 0.25$. All three
families produce nearly identical power curves. For specificity, the families
differ in their density at exactly zero (Normal: 1.60, Cauchy: 1.27, Logistic:
1.00 at scale $= 0.25$), which directly influences the Savage--Dickey ratio
under the null. We recommend the **Logistic effect prior** as the default: it
has the best specificity profile, well-behaved prior predictive distributions,
and identical power to the alternatives. The Cauchy, despite acceptable power,
assigns implausible probability to extreme effects, consistent with concerns
raised by @ghosh2018 about Cauchy priors under separation.

The **intercept prior** has minimal impact on Bayes Factor calibration. We
recommend the matched intercept prior (Logistic(0, 0.75) for logit, Normal(0,
0.75) for probit) for its better prior predictive properties, but a mismatched
intercept prior does not noticeably degrade inference. The intercept prior's
main role is to regularize the baseline probability away from floor and ceiling,
which benefits model fitting but does not substantially affect the Bayes Factor
for the effect.

For **sample size planning**, the number of trials per subject is at least as
important as the number of subjects. With 50 trials and
$\text{sd}_{b_1} = 0.25$, the 80% power criterion is met at $n = 30$ for
effects of $b_1 \geq 0.20$ on the logit scale across all three prior families
(Logistic: 0.83, Normal: 0.91, Cauchy: 0.89). At
$\text{sd}_{b_1} = 0.50$, the Logistic prior just misses the 80% threshold
(0.79), while the Normal and Cauchy remain above it — a minor trade-off of the
Logistic prior's specificity advantage. Smaller effects
($b_1 = 0.05$–$0.10$, corresponding to 2–5 percentage points) require larger
samples. Reducing from 50 to 20 trials per subject substantially lowers power
for these smaller effects, making trial count a critical design variable.

The **random-effects hyperprior** has no systematic impact on BF calibration.
All three hyperpriors examined --- the brms default Student-*t*(3, 0, 2.5),
Gamma(2, 4), and Exponential(4) --- produce similar power and specificity with
widely overlapping credible intervals. We recommend Exponential(4) as a default
on prior predictive grounds: it concentrates between-subject variability in a
plausible range and avoids the heavy right tail of the Student-*t* that can push
individual probabilities toward floor and ceiling (see the prior predictive
analysis). This choice does not compromise BF performance.

## Generalising to Other Model Classes

The binomial GLM is one member of a larger family of models where link functions
mediate between the parameter scale (where priors are specified) and the
observation scale (where scientific quantities live). The workflow presented here
— analytical matched priors, systematic prior predictive simulation, and
large-scale BF calibration — transfers directly to these other model classes.

**Ordinal regression.** Cumulative logit and probit models use the same link
functions as binomial GLMs, so the matched prior principle applies directly to
the threshold parameters. A Logistic prior on logit-scale thresholds
produces uniform implied category probabilities — a natural non-informative
starting point. Effect priors require the same prior predictive calibration
demonstrated here.

**Count data models.** Poisson and negative binomial regression use a log link.
The matched prior principle suggests an Exponential prior on the log-scale
intercept (since $\exp(X) \sim \text{Exponential}(1)$ when
$X \sim \text{Gumbel}$). Prior predictive simulation can reveal what prior
scales on log-scale coefficients imply for rate ratios.

**Distributional regression.** In GAMLSS [@klein2015] and brms distributional
models [@burkner2021], *all* parameters of the response distribution
(mean, variance, shape) are linked to predictors through link functions.
Each parameter therefore faces its own prior–link interaction. The matched
prior principle and the simulation approach developed here provide a
systematic way to address prior specification for each parameter.

**Beta regression.** For proportional data, beta regression models the mean via
a logit link. The matched prior principle applies identically to the intercept,
and prior predictive simulation can calibrate the implied distribution of
predicted proportions.

In each case, the key steps are: (1) identify the link function and the
corresponding matched prior family via the probability integral transform;
(2) use prior predictive simulation to map prior scales to interpretable
quantities on the observation scale; and (3) if Bayes Factor testing is
intended, run a calibration simulation to verify adequate power and
specificity. We provide all simulation code to facilitate such extensions.

## Appendix: Quantifying Uncertainty in Detection Rates {#sec-appendix-uncertainty}

Each simulation condition yields $n_{\text{valid}}$ replications (out of 100),
each producing a BF10 value. Rather than reporting detection rates as
point-estimate proportions, we apply Bayesian conjugate updating to obtain
full posterior distributions that reflect the sampling uncertainty from the
finite number of replications.

### Two-category model (Beta-Binomial conjugacy)

For a given BF threshold $\tau$ (e.g., $\tau = 3$ or $\tau = 10$), each
replication is a Bernoulli trial: "success" if BF10 $> \tau$ (for power) or
BF01 $> \tau$ (for specificity). With $k$ successes out of $n$ valid
replications and a $\text{Beta}(\alpha_0, \beta_0)$ prior, the posterior is:

$$\theta \mid k, n \sim \text{Beta}(\alpha_0 + k,\; \beta_0 + n - k)$$

We use $\alpha_0 = \beta_0 = 3$, a symmetric prior centered at 0.5 with 95%
of its mass in $[0.12,\, 0.88]$. This mildly regularises the estimate away
from extreme values while being dominated by the data: the effective prior
sample size is 6, negligible relative to the 100 replications per condition.
For comparison, the maximum shrinkage relative to the MLE is $\leq 0.03$
(at extreme rates near 0 or 1), and results are virtually identical under
the uniform $\text{Beta}(1, 1)$ or Jeffreys $\text{Beta}(0.5, 0.5)$ priors.

### Three-category model (Dirichlet-Multinomial conjugacy)

Each replication falls into exactly one of three categories: (1) H1 detected
(BF10 $> \tau$), (2) inconclusive ($1/\tau \leq \text{BF10} \leq \tau$), or
(3) H0 detected (BF10 $< 1/\tau$). The Dirichlet-Multinomial conjugate
update gives:

$$(\theta_{\text{H1}},\, \theta_{\text{inc}},\, \theta_{\text{H0}}) \mid \mathbf{k} \sim \text{Dirichlet}(\boldsymbol{\alpha}_0 + \mathbf{k})$$

with $\boldsymbol{\alpha}_0 = (1, 3, 1)$. This prior encodes the expectation
that most BFs will be inconclusive (prior mean 60%), with lower prior mass on
decisive evidence for either hypothesis (20% each). The effective prior sample
size is 5, again negligible relative to the data. The marginal posterior for
each category is Beta:
$\theta_j \mid \mathbf{k} \sim \text{Beta}(\alpha_{0,j} + k_j,\; \sum_{m \neq j}(\alpha_{0,m} + k_m))$.

### Sensitivity to Dirichlet prior choice

To confirm that the Dirichlet prior specification is inconsequential given
$n = 100$ replications per condition, we compare the $\text{Dirichlet}(1, 3, 1)$
used in the main analyses to the symmetric $\text{Dirichlet}(1, 1, 1)$ (uniform
on the simplex).

```{r}
#| label: fig-dirichlet-sensitivity
#| fig-cap: "The choice of Dirichlet prior has negligible impact on posterior
#|   means. Points compare the Dirichlet(1,3,1) and Dirichlet(1,1,1) priors
#|   for each simulation condition; all points fall near the identity line,
#|   confirming that 100 replications dominate the prior. Dashed line: identity."
#| fig-height: 5
#| fig-width: 10
#| eval: !expr posteriors_available

# Recompute posteriors under Dirichlet(1,1,1) for comparison
results_dir <- here("output", "Simulation_BFCalibration")
if (dir.exists(results_dir)) {
  cond_files <- list.files(results_dir, pattern = "^BFCalib_Cond-", full.names = TRUE)
} else {
  # Fall back to RE results for sensitivity check
  results_dir <- here("output", "Simulation_BFCalibration_RE")
  cond_files <- list.files(results_dir, pattern = "^BFCalib_RE_Cond-", full.names = TRUE)
}

sensitivity_rows <- lapply(cond_files, function(f) {
  cond_data <- readRDS(f)
  bf10_vec  <- cond_data$results[, "BF10"]
  counts    <- compute_bf_counts(bf10_vec, threshold = 3)

  dir_131 <- dirichlet_posterior_summary(
    counts["n_h1"], counts["n_h0"], counts["n_inconclusive"],
    alpha0 = c(1, 3, 1)
  )
  dir_111 <- dirichlet_posterior_summary(
    counts["n_h1"], counts["n_h0"], counts["n_inconclusive"],
    alpha0 = c(1, 1, 1)
  )

  data.frame(
    h1_mean_131  = dir_131["h1_mean"],
    h1_mean_111  = dir_111["h1_mean"],
    inc_mean_131 = dir_131["inc_mean"],
    inc_mean_111 = dir_111["inc_mean"],
    h0_mean_131  = dir_131["h0_mean"],
    h0_mean_111  = dir_111["h0_mean"]
  )
})

sensitivity <- bind_rows(sensitivity_rows)

p1 <- ggplot(sensitivity, aes(x = h1_mean_111, y = h1_mean_131)) +
  geom_abline(slope = 1, intercept = 0, linetype = "dashed", color = "grey50") +
  geom_point(alpha = 0.5, size = 1.5) +
  scale_x_continuous("Dirichlet(1,1,1)", limits = c(0, 1)) +
  scale_y_continuous("Dirichlet(1,3,1)", limits = c(0, 1)) +
  labs(title = "H1 detected") +
  base_theme()

p2 <- ggplot(sensitivity, aes(x = inc_mean_111, y = inc_mean_131)) +
  geom_abline(slope = 1, intercept = 0, linetype = "dashed", color = "grey50") +
  geom_point(alpha = 0.5, size = 1.5) +
  scale_x_continuous("Dirichlet(1,1,1)", limits = c(0, 1)) +
  scale_y_continuous("Dirichlet(1,3,1)", limits = c(0, 1)) +
  labs(title = "Inconclusive") +
  base_theme()

p3 <- ggplot(sensitivity, aes(x = h0_mean_111, y = h0_mean_131)) +
  geom_abline(slope = 1, intercept = 0, linetype = "dashed", color = "grey50") +
  geom_point(alpha = 0.5, size = 1.5) +
  scale_x_continuous("Dirichlet(1,1,1)", limits = c(0, 1)) +
  scale_y_continuous("Dirichlet(1,3,1)", limits = c(0, 1)) +
  labs(title = "H0 detected") +
  base_theme()

gridExtra::grid.arrange(p1, p2, p3, ncol = 3,
  top = "Dirichlet prior sensitivity: posterior means by category (BF > 3)")
```

```{r}
#| label: dirichlet-sensitivity-summary
#| eval: !expr posteriors_available

max_diff <- sensitivity |>
  summarise(
    max_h1  = max(abs(h1_mean_131 - h1_mean_111)),
    max_inc = max(abs(inc_mean_131 - inc_mean_111)),
    max_h0  = max(abs(h0_mean_131 - h0_mean_111))
  )
```

`r if (posteriors_available) paste0("The maximum absolute difference in posterior means between the two priors across all conditions is ", sprintf("%.3f", max(max_diff$max_h1, max_diff$max_inc, max_diff$max_h0)), " — confirming that the Dirichlet prior choice is negligible relative to the data.")`

## References

::: {#refs}
:::
