---
title: "Default Priors for Binomial GLMs: BF Calibration"
author:
  - name: Klaus Oberauer
    affiliation: University of Zurich
  - name: Joscha Dutli
    affiliation: University of Zurich
  - name: Philipp Musfeld
    affiliation: University of Zurich
  - name: Gidon Frischkorn
    affiliation: University of Lucerne
date: today
format:
  html:
    toc: true
    toc-depth: 3
    code-fold: true
    embed-resources: true
  pdf:
    toc: false
    toc-depth: 3
  docx:
    number-sections: false
execute:
  echo: false
  warning: false
  message: false
bibliography: references.bib
csl: apa.csl
---

```{r}
#| label: setup
library(dplyr)
library(tidyr)
library(ggplot2)
library(viridis)
library(here)

summaries <- readRDS(here("output", "bf_calibration_summaries.rds"))

re_summaries <- tryCatch(
  readRDS(here("output", "bf_calibration_re_summaries.rds")),
  error = function(e) NULL
)
re_results_available <- !is.null(re_summaries)

# Factor labels
dist_labels <- c(cauchy = "Cauchy", normal = "Normal", logistic = "Logistic")

base_theme <- function() {
  theme_minimal(base_size = 12) +
    theme(
      panel.grid.minor = element_blank(),
      strip.text       = element_text(face = "bold"),
      legend.position  = "bottom"
    )
}
```

## Introduction

Goal 2 examines how the choice of prior specification affects the behavior of
Bayes factors (BFs) in binomial GLMs. Specifically, we ask:

1. **Power:** Given a true effect of size $b_1$, how often does the BF exceed
   the evidence threshold (BF10 > 3 or BF10 > 10)?
2. **Specificity:** When the null is true ($b_1 = 0$), how often does the BF
   correctly support the null (BF01 > 3)?
3. **Sensitivity to prior:** How much do results change across the three prior
   families (Normal, Logistic, Cauchy)?
4. **Matched vs. misfit priors:** Does using the matched intercept prior (from
   Goal 1) matter for BF calibration?

### Design

The simulation crosses the following factors
(see `scripts/03_BFCalibration.R`):

- **Intercept prior family** (`dist_b0`): Logistic (matched for logit),
  Normal (misfit for logit)
- **Effect prior family** (`dist_b1`): Normal, Logistic, Cauchy
- **Effect prior SD** (`sd_b1`): 0.25, 0.50
- **True effect size** (`true_b1`): 0.00, 0.05, 0.10, 0.20, 0.50
- **Number of subjects** (`n_subjects`): 30, 60, 100
- **Trials per subject** (`n_trials`): 20, 50

This yields **360 conditions** with **100 replications** each.

A second simulation (`scripts/03b_BFCalibration_RE.R`) specifically investigates
the effect of the **random-effects hyperprior** on BF calibration. It uses a
structurally different data-generating model — a **within-subjects** design with
**correlated random intercepts and slopes** — and fixes the fixed-effects priors
at the matched configuration from Script 03:

- **RE hyperprior** (`sd_prior_re`): Default Student-*t*(3, 0, 2.5), Gamma(2, 4),
  Exponential(4) — this is the sole design factor
- **True effect size** (`true_b1`): 0.00, 0.05, 0.10, 0.20, 0.50
- **Number of subjects** (`n_subjects`): 30, 60, 100
- **Trials per subject** (`n_trials`): 20, 50
- Fixed: Logistic intercept prior ($\text{sd}_{b_0} = 0.75$), Normal effect
  prior ($\text{sd}_{b_1} = 0.25$); true intercept sampled per replication
  from $\text{Uniform}(0.4, 0.9)$ on the probability scale (same as Script 03)

This yields **90 conditions** with **100 replications** each.

### Model

All models use the **logit link** (the probit link is expected to produce
analogous results via the matched-prior principle established in Goal 1).
The true intercept $b_0$ is **not** fixed: in each replication it is sampled
from $\text{Uniform}(0.4,\, 0.9)$ on the probability scale and then
transformed to the logit scale. This covers the range of realistic baseline
accuracies in cognitive experiments (two-alternative to four-alternative
forced choice).

Each simulated dataset includes **subject-level random intercepts**
$u_i \sim \mathcal{N}(0, 0.25)$ to mimic between-subject heterogeneity.
The analysis model mirrors this structure: a binomial GLMM with one
random intercept per subject, fitted with brms [@burkner2017].

- Intercept prior SD fixed at 0.75 (recommended by Goal 1).
- Hyperprior on the random-intercept SD: $\text{Exponential}(4)$.

Bayes factors are computed via the **Savage–Dickey density ratio**
[@wagenmakers2018]: $\text{BF}_{10} = \pi(b_1 = 0) \,/\, p(b_1 = 0 \mid \text{data})$,
where the prior density is evaluated analytically and the posterior density
is estimated by kernel density estimation (KDE with Silverman's bandwidth).

---

## Results: Power (H1 true)

### Power by sample size and effect size

```{r}
#| label: fig-power-n
#| fig-cap: "Power (P(BF10 > 3)) as a function of the number of subjects and
#|   the true effect size, for each effect prior family and SD.
#|   Intercept prior: Logistic with sd_b0 = 0.75."
#| fig-height: 6
#| fig-width: 10

matched <- summaries |>
  filter(
    dist_b0 == "logistic",
    true_b1 > 0
  ) |>
  mutate(
    dist_b1  = factor(dist_b1, levels = c("cauchy", "normal", "logistic"),
                      labels = c("Cauchy", "Normal", "Logistic")),
    sd_b1    = factor(paste0("sd = ", sd_b1))
  )

ggplot(matched, aes(x = n_subjects, y = P_BF10_gt3,
                    color = dist_b1, linetype = sd_b1)) +
  geom_line(linewidth = 0.8) +
  geom_point(size = 2) +
  geom_hline(yintercept = 0.80, linetype = "dotted", color = "grey40") +
  scale_x_continuous("Number of subjects", breaks = c(30, 60, 100)) +
  scale_y_continuous("P(BF10 > 3)", limits = c(NA, 1)) +
  scale_color_viridis_d("Prior family", option = "D", end = 0.85) +
  scale_linetype_manual("Prior SD",
    values = c("sd = 0.25" = "solid", "sd = 0.5" = "dashed")) +
  facet_grid(n_trials ~ true_b1,
             labeller = labeller(
               true_b1  = function(x) paste0("b1 = ", x),
               n_trials = function(x) paste0(x, " trials")
             )) +
  labs(title = "Power: P(BF10 > 3) by sample size and true effect",
       subtitle = "Matched intercept prior (Logistic, sd = 0.75); dotted line = 80% power") +
  base_theme()
```

@fig-power-n shows that power increases with both sample size and true
effect size, as expected. The tighter prior ($\text{sd} = 0.25$) yields
slightly higher power than $\text{sd} = 0.50$ on average, because the
narrower prior concentrates more density near the true effect and less at
extreme values. However, this difference is practically negligible: power is
at or near ceiling for both effect sizes with 50 trials and $\text{sd} = 0.25$,
exceeding 97% even at the smallest sample size ($n = 30$). Differences between
the three prior families are small relative to the effect of prior SD. The
more informative variation appears at 20 trials per subject, where power is
lower and sample-size effects become more visible.

### Power by n_trials

```{r}
#| label: fig-power-trials
#| fig-cap: "Power as a function of trials per subject, for n_subjects = 60
#|   and each prior specification (matched Logistic intercept prior)."
#| fig-height: 5
#| fig-width: 10

matched |>
  filter(n_subjects == 60) |>
  ggplot(aes(x = n_trials, y = P_BF10_gt3,
             color = dist_b1, linetype = sd_b1)) +
  geom_line(linewidth = 0.8) +
  geom_point(size = 2) +
  geom_hline(yintercept = 0.80, linetype = "dotted", color = "grey40") +
  scale_x_continuous("Trials per subject", breaks = c(20, 50)) +
  scale_y_continuous("P(BF10 > 3)", limits = c(NA, 1)) +
  scale_color_viridis_d("Prior family", option = "D", end = 0.85) +
  scale_linetype_manual("Prior SD",
    values = c("sd = 0.25" = "solid", "sd = 0.5" = "dashed")) +
  facet_wrap(~ true_b1,
             labeller = labeller(
               true_b1 = function(x) paste0("b1 = ", x)
             )) +
  labs(title = "Power by trials per subject (n_subjects = 60, matched intercept prior)") +
  base_theme()
```

Increasing the number of trials per subject from 20 to 50 substantially
improves power for all prior configurations. This gain is expected because
more trials per subject reduce within-subject uncertainty, which in turn
sharpens the posterior for the fixed-effect parameter $b_1$.

---

## Results: Specificity Under H0

```{r}
#| label: fig-specificity
#| fig-cap: "Specificity (P(BF01 > 3)) under the null (true_b1 = 0) for each
#|   prior specification and sample size (matched Logistic intercept prior)."
#| fig-height: 5
#| fig-width: 10

summaries |>
  filter(
    true_b1 == 0,
    dist_b0 == "logistic"
  ) |>
  mutate(
    dist_b1 = factor(dist_b1, levels = c("cauchy", "normal", "logistic"),
                     labels = c("Cauchy", "Normal", "Logistic")),
    sd_b1   = factor(paste0("sd = ", sd_b1))
  ) |>
  ggplot(aes(x = n_subjects, y = P_BF01_gt3,
             color = dist_b1, linetype = sd_b1)) +
  geom_line(linewidth = 0.8) +
  geom_point(size = 2) +
  scale_x_continuous("Number of subjects", breaks = c(30, 60, 100)) +
  scale_y_continuous("P(BF01 > 3)", limits = c(0, 1)) +
  scale_color_viridis_d("Prior family", option = "D", end = 0.85) +
  scale_linetype_manual("Prior SD",
    values = c("sd = 0.25" = "solid", "sd = 0.5" = "dashed")) +
  facet_wrap(~ n_trials,
             labeller = labeller(
               n_trials = function(x) paste0(x, " trials")
             )) +
  labs(title = "Specificity under H0: P(BF01 > 3)",
       subtitle = "Matched intercept prior (Logistic, sd = 0.75)") +
  base_theme()
```

Specificity is generally high across all prior configurations: when the null
is true ($b_1 = 0$), the BF correctly supports H0 (BF01 > 3) in the large
majority of replications. The wider prior ($\text{sd} = 0.50$) tends to
yield slightly higher specificity because it places less prior density at
exactly zero, making the Savage–Dickey ratio more conservative. Prior family
has minimal impact on specificity.

---

## Results: Sensitivity to Prior Family

```{r}
#| label: fig-prior-sensitivity
#| fig-cap: "Comparison of power across prior families at sd_b1 = 0.25,
#|   showing the impact of prior family choice on BF calibration."
#| fig-height: 6
#| fig-width: 10

summaries |>
  filter(
    sd_b1 == 0.25,
    dist_b0 == "logistic",
    true_b1 > 0
  ) |>
  mutate(
    dist_b1 = factor(dist_b1, levels = c("cauchy", "normal", "logistic"),
                     labels = c("Cauchy", "Normal", "Logistic"))
  ) |>
  ggplot(aes(x = true_b1, y = P_BF10_gt3, color = dist_b1)) +
  geom_line(linewidth = 0.8) +
  geom_point(size = 2) +
  geom_hline(yintercept = 0.80, linetype = "dotted", color = "grey40") +
  scale_x_continuous("True effect b1") +
  scale_y_continuous("P(BF10 > 3)", limits = c(NA, 1)) +
  scale_color_viridis_d("Prior family", option = "D", end = 0.85) +
  facet_grid(n_trials ~ n_subjects,
             labeller = labeller(
               n_subjects = function(x) paste0(x, " subj"),
               n_trials   = function(x) paste0(x, " trials")
             )) +
  labs(title = "Sensitivity to prior family (sd_b1 = 0.25, matched intercept prior)") +
  base_theme()
```

Holding $\text{sd}_{b_1} = 0.25$ constant, the three prior families
produce nearly identical power curves across all sample sizes and effect
sizes. This indicates that the choice of prior family (Normal, Logistic, or
Cauchy) is far less consequential than the choice of prior scale for BF
calibration in binomial GLMs.

---

## Results: Matched vs. Misfit Intercept Prior

```{r}
#| label: fig-misfit-bf
#| fig-cap: "Effect of intercept prior misfit on power: matched (Logistic)
#|   vs. misfit (Normal) intercept prior, with Logistic effect prior at
#|   sd_b1 = 0.25."
#| fig-height: 6
#| fig-width: 10

summaries |>
  filter(
    dist_b1 == "logistic",
    sd_b1   == 0.25,
    true_b1 > 0
  ) |>
  mutate(
    dist_b0 = factor(dist_b0, levels = c("logistic", "normal"),
                     labels = c("Logistic (matched)", "Normal (misfit)"))
  ) |>
  ggplot(aes(x = true_b1, y = P_BF10_gt3, color = dist_b0)) +
  geom_line(linewidth = 0.8) +
  geom_point(size = 2) +
  geom_hline(yintercept = 0.80, linetype = "dotted", color = "grey40") +
  scale_x_continuous("True effect b1") +
  scale_y_continuous("P(BF10 > 3)", limits = c(NA, 1)) +
  scale_color_viridis_d("Intercept prior", option = "D", end = 0.7) +
  facet_grid(n_trials ~ n_subjects,
             labeller = labeller(
               n_subjects = function(x) paste0(x, " subjects"),
               n_trials   = function(x) paste0(x, " trials")
             )) +
  labs(title = "Effect of intercept prior misfit on power (logit link, sd_b1 = 0.25)") +
  base_theme()
```

Using a mismatched intercept prior (Normal instead of the matched Logistic)
has only a small impact on power. This confirms that the intercept prior
primarily serves to regularise the baseline probability and does not
meaningfully alter inference about the effect parameter $b_1$, as long as
its SD is reasonable (here, $\text{sd}_{b_0} = 0.75$).

---

## Results: Effect of Random-Effects Hyperprior

```{r}
#| label: setup-re
#| eval: !expr re_results_available

re_plot_data <- re_summaries |>
  mutate(
    sd_prior_re = factor(sd_prior_re,
                         levels = c("default", "gamma", "exponential"),
                         labels = c("Student-t(3,0,2.5)", "Gamma(2,4)",
                                    "Exponential(4)")),
    n_trials_label = paste0(n_trials, " trials")
  )
```

Script `03b_BFCalibration_RE.R` varies the prior on the random-effects SD while
keeping the fixed-effects priors at the recommended matched configuration
(Logistic intercept at $\text{sd}_{b_0} = 0.75$; Normal effect prior at
$\text{sd}_{b_1} = 0.25$). The three candidate hyperpriors for the random-intercept
and random-slope SDs are:

- **Default**: $\text{Student-}t(3, 0, 2.5)$ (brms default before version 2.20)
- **Gamma**: $\text{Gamma}(2, 4)$
- **Exponential**: $\text{Exponential}(4)$ (used throughout Script 03)

Unlike Script 03 (between-subjects, random intercepts only), this simulation uses
a **within-subjects** design in which each subject appears in both conditions,
with **correlated random intercepts and slopes** ($u_i \sim \mathcal{N}(0, 0.25)$,
$v_i \sim \mathcal{N}(0, 0.15)$). The true effect size range is extended to
$b_1 = 0.75$.

### Power by RE Hyperprior

```{r}
#| label: fig-re-power
#| fig-cap: "Power (P(BF10 > 3)) by number of subjects for each RE hyperprior,
#|   with matched fixed-effects priors (Logistic intercept sd = 0.75,
#|   Normal effect sd = 0.25). Within-subjects design with correlated RE."
#| fig-height: 6
#| fig-width: 10
#| eval: !expr re_results_available

re_plot_data |>
  filter(true_b1 > 0) |>
  ggplot(aes(x = n_subjects, y = P_BF10_gt3, color = sd_prior_re,
             linetype = sd_prior_re)) +
  geom_line(linewidth = 0.8) +
  geom_point(size = 2) +
  geom_hline(yintercept = 0.80, linetype = "dotted", color = "grey40") +
  scale_x_continuous("Number of subjects", breaks = c(30, 60, 100)) +
  scale_y_continuous("P(BF10 > 3)", limits = c(NA, 1)) +
  scale_color_viridis_d("RE hyperprior", option = "D", end = 0.85) +
  scale_linetype_manual("RE hyperprior",
    values = c("Student-t(3,0,2.5)" = "solid",
               "Gamma(2,4)"         = "dashed",
               "Exponential(4)"     = "dotdash")) +
  facet_grid(n_trials_label ~ true_b1,
             labeller = labeller(
               true_b1 = function(x) paste0("b1 = ", x)
             )) +
  labs(title = "Power by RE hyperprior (within-subjects, matched fixed-effects priors)",
       subtitle = "Dotted line = 80% power criterion") +
  base_theme()
```

### Specificity by RE Hyperprior

```{r}
#| label: fig-re-specificity
#| fig-cap: "Specificity (P(BF01 > 3)) under the null (true_b1 = 0) for each
#|   RE hyperprior, by number of subjects and trials per subject."
#| fig-height: 5
#| fig-width: 10
#| eval: !expr re_results_available

re_plot_data |>
  filter(true_b1 == 0) |>
  ggplot(aes(x = n_subjects, y = P_BF01_gt3, color = sd_prior_re,
             linetype = sd_prior_re)) +
  geom_line(linewidth = 0.8) +
  geom_point(size = 2) +
  scale_x_continuous("Number of subjects", breaks = c(30, 60, 100)) +
  scale_y_continuous("P(BF01 > 3)", limits = c(0, 1)) +
  scale_color_viridis_d("RE hyperprior", option = "D", end = 0.85) +
  scale_linetype_manual("RE hyperprior",
    values = c("Student-t(3,0,2.5)" = "solid",
               "Gamma(2,4)"         = "dashed",
               "Exponential(4)"     = "dotdash")) +
  facet_wrap(~ n_trials_label) +
  labs(title = "Specificity under H0 by RE hyperprior") +
  base_theme()
```

### Minimum Sample Size by RE Hyperprior

```{r}
#| label: tbl-re-summary
#| tbl-cap: "Power at 80% criterion: minimum n_subjects required to achieve
#|   P(BF10 > 3) >= 0.80, by RE hyperprior and true effect size (n_trials = 50).
#|   Within-subjects design with correlated RE; matched fixed-effects priors."
#| eval: !expr re_results_available

re_plot_data |>
  filter(n_trials == 50, true_b1 > 0) |>
  group_by(sd_prior_re, true_b1) |>
  summarise(
    min_n_80pct = {
      n_meeting <- n_subjects[P_BF10_gt3 >= 0.80]
      if (length(n_meeting) > 0) min(n_meeting) else NA_integer_
    },
    .groups = "drop"
  ) |>
  pivot_wider(names_from = true_b1, values_from = min_n_80pct,
              names_prefix = "b1 = ") |>
  rename("RE hyperprior" = sd_prior_re) |>
  knitr::kable(caption = "Minimum n_subjects for 80% power by RE hyperprior (n_trials = 50)")
```

---

## Summary Table

```{r}
#| label: tbl-summary
#| tbl-cap: "Power at 80% criterion (Script 03, between-subjects, random intercepts
#|   only): minimum n_subjects required to achieve P(BF10 > 3) >= 0.80, for each
#|   prior specification and true effect size (n_trials = 50, matched intercept
#|   prior)."

summaries |>
  filter(
    n_trials == 50,
    dist_b0 == "logistic",
    true_b1 > 0
  ) |>
  group_by(dist_b1, sd_b1, true_b1) |>
  summarise(
    min_n_80pct = {
      n_meeting <- n_subjects[P_BF10_gt3 >= 0.80]
      if (length(n_meeting) > 0) min(n_meeting) else NA_integer_
    },
    .groups = "drop"
  ) |>
  mutate(
    dist_b1 = factor(dist_b1, levels = c("cauchy", "normal", "logistic"),
                     labels = c("Cauchy", "Normal", "Logistic"))
  ) |>
  pivot_wider(names_from = true_b1, values_from = min_n_80pct,
              names_prefix = "b1 = ") |>
  knitr::kable(caption = "Minimum n_subjects for 80% power (n_trials = 50, matched intercept prior)")
```

---

## Recommendations

Based on the BF calibration results:

1. **Prior scale matters more than prior family.** The Normal and Logistic
   prior families produce very similar BF calibration at the same
   $\text{sd}_{b_1}$, and even the heavier-tailed Cauchy prior differs only
   modestly. The most impactful choice is the prior SD.

2. **Recommended default:** $\text{sd}_{b_1} = 0.25$ with a Normal or
   Logistic effect prior provides a good balance of power and specificity
   for typical cognitive psychology effect sizes ($b_1 \approx 0.05$–$0.20$
   on the logit scale, corresponding to roughly 2–10 percentage points
   between conditions).

3. **Intercept prior misfit has minimal impact** on BF calibration when
   $\text{sd}_{b_0} = 0.75$, confirming that the main priority is matching
   the effect prior scale to the expected range of effects.

4. **Sample size guidance:** With 50 trials per condition and
   $\text{sd}_{b_1} = 0.25$, the 80% power criterion is already met at
   $n = 30$ subjects for the larger effects examined here ($b_1 \geq 0.20$).
   On the probability scale with a typical baseline ($p_0 = 0.50$), $b_1 = 0.50$
   on the logit scale corresponds to approximately 24 percentage points between
   conditions, while $b_1 = 0.20$ corresponds to approximately 10 percentage
   points. Smaller, more typical effects ($b_1 = 0.05$–$0.10$; roughly 2–5 pp)
   are expected to require larger samples and more trials to achieve adequate power.
   Power is more informatively differentiated by number of trials:
   reducing from 50 to 20 trials per subject substantially lowers power for
   smaller effects.

5. **RE hyperprior has minimal impact on power but affects specificity at small samples.**
   All three hyperpriors produce identical power (at or near ceiling) across the effect
   sizes examined. However, specificity under H0 shows meaningful differences at small
   samples with few trials: at $n = 30$ and 20 trials per subject, the default
   Student-*t*(3, 0, 2.5) yields notably lower specificity (P(BF01 > 3) $\approx$ 0.45)
   compared to $\text{Exponential}(4)$ ($\approx$ 0.59). These differences attenuate at
   50 trials per subject, where the three hyperpriors converge to P(BF01 > 3) $\approx$
   0.72–0.81 at $n = 30$. The $\text{Exponential}(4)$ prior offers a reasonable default
   with good specificity properties at small $n$.

---

## References

::: {#refs}
:::
