---
title: "Default Priors for Binomial GLMs: BF Calibration"
author:
  - name: Klaus Oberauer
    affiliation: University of Zurich
  - name: Joscha Dutli
    affiliation: University of Zurich
  - name: Philipp Musfeld
    affiliation: University of Zurich
  - name: Gidon Frischkorn
    affiliation: University of Lucerne
date: today
format:
  html:
    toc: true
    toc-depth: 3
    code-fold: true
    embed-resources: true
  pdf:
    toc: false
    toc-depth: 3
  docx:
    number-sections: false
execute:
  echo: false
  warning: false
  message: false
bibliography: references.bib
csl: apa.csl
---

```{r}
#| label: setup
library(dplyr)
library(tidyr)
library(ggplot2)
library(viridis)
library(here)

source(here("R", "bf_helpers.R"))

summaries <- readRDS(here("output", "bf_calibration_summaries.rds"))

posteriors <- tryCatch(
  readRDS(here("output", "bf_calibration_posteriors.rds")),
  error = function(e) NULL
)
posteriors_available <- !is.null(posteriors)

re_summaries <- tryCatch(
  readRDS(here("output", "bf_calibration_re_summaries.rds")),
  error = function(e) NULL
)
re_results_available <- !is.null(re_summaries)

re_posteriors <- tryCatch(
  readRDS(here("output", "bf_calibration_re_posteriors.rds")),
  error = function(e) NULL
)
re_posteriors_available <- !is.null(re_posteriors)

# Factor labels
dist_labels <- c(cauchy = "Cauchy", normal = "Normal", logistic = "Logistic")

base_theme <- function() {
  theme_minimal(base_size = 12) +
    theme(
      panel.grid.minor = element_blank(),
      strip.text       = element_text(face = "bold"),
      legend.position  = "bottom"
    )
}
```

## Introduction

The prior predictive analysis established that matched priors — Logistic(0, 0.75)
for the logit link and Normal(0, 0.75) for the probit link — produce well-behaved
distributions on the probability scale, and that effect priors at
$\text{sd}_{b_1} = 0.25$ concentrate prior mass on substantively plausible effect
sizes. However, good prior predictive behaviour is a necessary condition for sound
Bayesian inference, not a sufficient one. We also need to know how these priors
perform when used for hypothesis testing with Bayes Factors.

Prior specification directly affects discovery rates. A prior that is too wide
spreads probability mass across implausible effect sizes, diluting the evidence
for real effects and reducing the power of the Bayes Factor to detect them. A
prior that is too narrow may spuriously favour or disfavour the null. This tension
is well documented for normal-data models: @schonbrodt2018 introduced Bayes Factor
Design Analysis (BFDA) as a formal framework for evaluating power and error rates
under a given prior, and @stefan2019 extended it to informed priors. These tools,
however, have been developed for simple designs with closed-form BF solutions
(e.g., the JZS prior of @rouder2009). For binomial GLMs — where BFs must be
computed numerically via bridge sampling [@gronau2017; @gronau2020] or the
Savage–Dickey density ratio — no systematic calibration exists. The present
simulation fills this gap by applying the BFDA logic to the recommended priors
from the prior predictive analysis.

We address four questions:

1. **Do the recommended priors provide adequate power** to detect effects across
   realistic sample sizes and numbers of trials?
2. **Do they protect against false evidence** — that is, do they correctly support
   the null hypothesis when there is no effect?
3. **How sensitive is the Bayes Factor to the prior family?** Does it matter whether
   researchers use a Normal, Logistic, or Cauchy prior, or is the prior scale the
   key decision?
4. **Does the matched-prior principle carry over to inference?** The prior
   predictive analysis showed that matched intercept priors produce better-behaved
   distributions on the probability scale — but does using a mismatched intercept
   prior actually harm Bayes Factor calibration?

Finally, we extend the analysis to within-subjects designs with correlated random
effects, asking whether the choice of random-effects hyperprior — which the prior
predictive analysis showed affects the distribution of subject-level probabilities —
also affects inferential performance.

### Simulation Design

We simulated data across a range of realistic experimental scenarios. The true
effect size ranged from $b_1 = 0$ (no effect) to $b_1 = 0.50$ on the logit scale,
covering the span from a null effect through a large effect in cognitive
experiments (on the probability scale with a typical 50% baseline, $b_1 = 0.50$
corresponds to roughly 24 percentage points between conditions). Sample sizes
ranged from 30 to 100 subjects, and each subject contributed either 20 or 50
trials. We crossed these data-generating factors with three effect prior families
(Normal, Logistic, Cauchy) at two scale values ($\text{sd}_{b_1} = 0.25$ and
$0.50$), and two intercept priors (matched Logistic and misfit Normal, both at
$\text{sd}_{b_0} = 0.75$). This produced 360 conditions with 100 replications
each (see `scripts/03_BFCalibration.R`).

A second simulation (`scripts/03b_BFCalibration_RE.R`) specifically investigated
the random-effects hyperprior. It used a within-subjects design with correlated
random intercepts and slopes, holding the fixed-effects priors at the recommended
matched configuration and varying only the hyperprior on the random-effects SDs
across three candidates: the brms default Student-*t*(3, 0, 2.5), Gamma(2, 4),
and Exponential(4). This produced 90 conditions with 100 replications each. We
return to these results in @sec-re-hyperprior.

### How We Computed Bayes Factors

All models used the logit link (the probit link is expected to produce analogous
results via the matched-prior principle established in the prior predictive
analysis). To ensure that results generalize across realistic baseline accuracies,
we did not fix the true intercept: in each replication, $b_0$ was sampled from
$\text{Uniform}(0.4, 0.9)$ on the probability scale — the range spanning
two-alternative to four-alternative forced choice — and transformed to the logit
scale. Each simulated dataset included subject-level random intercepts
($u_i \sim \mathcal{N}(0, 0.25)$) to mimic between-subject heterogeneity.

The analysis model mirrored this structure: a binomial GLMM with one random
intercept per subject, fitted with brms [@burkner2017], with the intercept prior
SD fixed at 0.75 and an $\text{Exponential}(4)$ hyperprior on the random-intercept
SD. Bayes Factors were computed via the Savage–Dickey density ratio
[@wagenmakers2018; @ly2016], which compares how plausible a zero effect was before
versus after seeing the data:
$\text{BF}_{10} = \pi(b_1 = 0) / p(b_1 = 0 \mid \text{data})$.
When the posterior concentrates away from zero (i.e., the data suggest an effect),
$\text{BF}_{10}$ exceeds 1 and favours H1; when the posterior piles up at zero
(i.e., the data are consistent with no effect), $\text{BF}_{10}$ falls below 1
and favours H0. Note that the prior density at zero — $\pi(b_1 = 0)$ — enters the
numerator directly. This has a critical consequence: different prior families with
the same scale parameter place different densities at zero, which creates
systematic differences in specificity (the ability to correctly support H0). We
return to this point in the specificity analysis below.

To quantify uncertainty in detection rates given the finite number of replications,
we applied Bayesian conjugate updating (see
@sec-appendix-uncertainty for details). All error
bars in the figures below show 95% credible intervals from this procedure.


## How Much Data Do Bayes Factors Need?

The most fundamental question for study planning is: given a real effect, how
many participants and trials are needed for the Bayes Factor to detect it? We
examine power — the probability that BF10 exceeds 3, providing at least
moderate evidence for the alternative — across the full range of sample sizes,
trial counts, and prior specifications.

### Power by sample size and effect size

```{r}
#| label: fig-power-n
#| fig-cap: "Power to detect effects increases with sample size and effect
#|   magnitude. With 50 trials per subject and the recommended effect prior
#|   (sd = 0.25), power exceeds 80% at N = 30 for effects of b1 >= 0.20. The
#|   tighter prior (sd = 0.25, solid lines) yields slightly higher power than
#|   the wider prior (sd = 0.50, dashed), but the difference is small compared
#|   to the effect of sample size and true effect. Prior family (color) has
#|   minimal impact on power. Error bars: 95% credible intervals."
#| fig-height: 6
#| fig-width: 10

if (posteriors_available) {
  matched <- posteriors |>
    filter(
      dist_b0 == "logistic",
      true_b1 > 0,
      threshold == 3
    ) |>
    mutate(
      dist_b1  = factor(dist_b1, levels = c("cauchy", "normal", "logistic"),
                        labels = c("Cauchy", "Normal", "Logistic")),
      sd_b1    = factor(paste0("sd = ", sd_b1))
    )

  ggplot(matched, aes(x = n_subjects, y = power_post_mean,
                      color = dist_b1, linetype = sd_b1)) +
    geom_line(linewidth = 0.8) +
    geom_point(size = 2) +
    geom_errorbar(aes(ymin = power_ci_lower, ymax = power_ci_upper),
                  width = 3, linewidth = 0.4, alpha = 0.6) +
    geom_hline(yintercept = 0.80, linetype = "dotted", color = "grey40") +
    scale_x_continuous("Number of subjects", breaks = c(30, 60, 100)) +
    scale_y_continuous("P(BF10 > 3)", limits = c(NA, 1)) +
    scale_color_viridis_d("Prior family", option = "D", end = 0.85) +
    scale_linetype_manual("Prior SD",
      values = c("sd = 0.25" = "solid", "sd = 0.5" = "dashed")) +
    facet_grid(n_trials ~ true_b1,
               labeller = labeller(
                 true_b1  = function(x) paste0("b1 = ", x),
                 n_trials = function(x) paste0(x, " trials")
               )) +
    labs(title = "Power: P(BF10 > 3) by sample size and true effect",
         subtitle = "Matched intercept prior (Logistic, sd = 0.75); dotted line = 80% power") +
    base_theme()
} else {
  matched <- summaries |>
    filter(dist_b0 == "logistic", true_b1 > 0) |>
    mutate(
      dist_b1  = factor(dist_b1, levels = c("cauchy", "normal", "logistic"),
                        labels = c("Cauchy", "Normal", "Logistic")),
      sd_b1    = factor(paste0("sd = ", sd_b1))
    )

  ggplot(matched, aes(x = n_subjects, y = P_BF10_gt3,
                      color = dist_b1, linetype = sd_b1)) +
    geom_line(linewidth = 0.8) +
    geom_point(size = 2) +
    geom_hline(yintercept = 0.80, linetype = "dotted", color = "grey40") +
    scale_x_continuous("Number of subjects", breaks = c(30, 60, 100)) +
    scale_y_continuous("P(BF10 > 3)", limits = c(NA, 1)) +
    scale_color_viridis_d("Prior family", option = "D", end = 0.85) +
    scale_linetype_manual("Prior SD",
      values = c("sd = 0.25" = "solid", "sd = 0.5" = "dashed")) +
    facet_grid(n_trials ~ true_b1,
               labeller = labeller(
                 true_b1  = function(x) paste0("b1 = ", x),
                 n_trials = function(x) paste0(x, " trials")
               )) +
    labs(title = "Power: P(BF10 > 3) by sample size and true effect",
         subtitle = "Matched intercept prior (Logistic, sd = 0.75); dotted line = 80% power") +
    base_theme()
}
```

@fig-power-n shows a clear and encouraging pattern. For the larger effects
examined here ($b_1 \geq 0.20$), power reaches or exceeds 80% even at the
smallest sample size ($n = 30$) when subjects complete 50 trials. The tighter
prior ($\text{sd} = 0.25$, solid lines) yields slightly higher power than the
wider alternative ($\text{sd} = 0.50$, dashed), because concentrating the prior
near plausible effect sizes makes the Savage–Dickey ratio more sensitive to
genuine shifts away from zero. However, this advantage is modest — the dominant
drivers of power are the true effect magnitude and the amount of data, not the
prior specification.

Two additional patterns are worth noting. First, the three prior families (Normal,
Logistic, Cauchy) produce nearly overlapping power curves, foreshadowing a key
conclusion: the choice of prior *family* matters far less than the choice of
prior *scale*. Second, the difference between 20 and 50 trials per subject is
particularly informative at smaller effects, where 20 trials may not suffice.
This suggests that increasing trials per subject can be at least as effective as
recruiting more participants.

### The role of trials per subject

Power also depends strongly on how many trials each subject completes. @fig-power-trials
isolates this factor for a mid-range sample size ($n = 60$).

```{r}
#| label: fig-power-trials
#| fig-cap: "Increasing from 20 to 50 trials per subject substantially
#|   improves power, particularly for smaller effects. Shown for N = 60
#|   subjects with matched Logistic intercept prior. Error bars: 95%
#|   credible intervals."
#| fig-height: 5
#| fig-width: 10

if (posteriors_available) {
  posteriors |>
    filter(
      dist_b0 == "logistic",
      true_b1 > 0,
      threshold == 3,
      n_subjects == 60
    ) |>
    mutate(
      dist_b1  = factor(dist_b1, levels = c("cauchy", "normal", "logistic"),
                        labels = c("Cauchy", "Normal", "Logistic")),
      sd_b1    = factor(paste0("sd = ", sd_b1))
    ) |>
    ggplot(aes(x = n_trials, y = power_post_mean,
               color = dist_b1, linetype = sd_b1)) +
    geom_line(linewidth = 0.8) +
    geom_point(size = 2) +
    geom_errorbar(aes(ymin = power_ci_lower, ymax = power_ci_upper),
                  width = 2, linewidth = 0.4, alpha = 0.6) +
    geom_hline(yintercept = 0.80, linetype = "dotted", color = "grey40") +
    scale_x_continuous("Trials per subject", breaks = c(20, 50)) +
    scale_y_continuous("P(BF10 > 3)", limits = c(NA, 1)) +
    scale_color_viridis_d("Prior family", option = "D", end = 0.85) +
    scale_linetype_manual("Prior SD",
      values = c("sd = 0.25" = "solid", "sd = 0.5" = "dashed")) +
    facet_wrap(~ true_b1,
               labeller = labeller(
                 true_b1 = function(x) paste0("b1 = ", x)
               )) +
    labs(title = "Power by trials per subject (n_subjects = 60, matched intercept prior)") +
    base_theme()
} else {
  matched |>
    filter(n_subjects == 60) |>
    ggplot(aes(x = n_trials, y = P_BF10_gt3,
               color = dist_b1, linetype = sd_b1)) +
    geom_line(linewidth = 0.8) +
    geom_point(size = 2) +
    geom_hline(yintercept = 0.80, linetype = "dotted", color = "grey40") +
    scale_x_continuous("Trials per subject", breaks = c(20, 50)) +
    scale_y_continuous("P(BF10 > 3)", limits = c(NA, 1)) +
    scale_color_viridis_d("Prior family", option = "D", end = 0.85) +
    scale_linetype_manual("Prior SD",
      values = c("sd = 0.25" = "solid", "sd = 0.5" = "dashed")) +
    facet_wrap(~ true_b1,
               labeller = labeller(
                 true_b1 = function(x) paste0("b1 = ", x)
               )) +
    labs(title = "Power by trials per subject (n_subjects = 60, matched intercept prior)") +
    base_theme()
}
```

Increasing trials from 20 to 50 per subject substantially improves power across
all prior configurations. The mechanism is straightforward: more trials per
subject reduce within-subject variability, which sharpens the posterior for the
fixed-effect parameter $b_1$ and makes the Bayes Factor more decisive. For
researchers designing studies, this underscores a practical point — collecting
more trials from each participant (e.g., by extending the experimental session)
can be as effective as recruiting additional participants, and is often easier
to implement.


## Do These Priors Produce False Evidence?

Good power is only valuable if the Bayes Factor does not also produce spurious
evidence for effects that do not exist. We therefore examine specificity: when
the null is true ($b_1 = 0$), how often does the BF correctly support H0 (i.e.,
BF01 > 3)?

```{r}
#| label: fig-specificity
#| fig-cap: "When no effect exists, the Bayes Factor correctly supports the
#|   null in the large majority of replications. Wider priors (sd = 0.50,
#|   dashed) yield substantially higher specificity because they place less
#|   density at exactly zero, making the Savage-Dickey ratio more conservative.
#|   At sd = 0.25 (solid lines), the Normal prior shows notably lower
#|   specificity than the Logistic and Cauchy, especially at small sample
#|   sizes — a consequence of its higher density at exactly zero (see text).
#|   Error bars: 95% credible intervals."
#| fig-height: 5
#| fig-width: 10

if (posteriors_available) {
  posteriors |>
    filter(
      true_b1 == 0,
      dist_b0 == "logistic",
      threshold == 3
    ) |>
    mutate(
      dist_b1 = factor(dist_b1, levels = c("cauchy", "normal", "logistic"),
                       labels = c("Cauchy", "Normal", "Logistic")),
      sd_b1   = factor(paste0("sd = ", sd_b1))
    ) |>
    ggplot(aes(x = n_subjects, y = spec_post_mean,
               color = dist_b1, linetype = sd_b1)) +
    geom_line(linewidth = 0.8) +
    geom_point(size = 2) +
    geom_errorbar(aes(ymin = spec_ci_lower, ymax = spec_ci_upper),
                  width = 3, linewidth = 0.4, alpha = 0.6) +
    scale_x_continuous("Number of subjects", breaks = c(30, 60, 100)) +
    scale_y_continuous("P(BF01 > 3)", limits = c(0, 1)) +
    scale_color_viridis_d("Prior family", option = "D", end = 0.85) +
    scale_linetype_manual("Prior SD",
      values = c("sd = 0.25" = "solid", "sd = 0.5" = "dashed")) +
    facet_wrap(~ n_trials,
               labeller = labeller(
                 n_trials = function(x) paste0(x, " trials")
               )) +
    labs(title = "Specificity under H0: P(BF01 > 3)",
         subtitle = "Matched intercept prior (Logistic, sd = 0.75)") +
    base_theme()
} else {
  summaries |>
    filter(true_b1 == 0, dist_b0 == "logistic") |>
    mutate(
      dist_b1 = factor(dist_b1, levels = c("cauchy", "normal", "logistic"),
                       labels = c("Cauchy", "Normal", "Logistic")),
      sd_b1   = factor(paste0("sd = ", sd_b1))
    ) |>
    ggplot(aes(x = n_subjects, y = P_BF01_gt3,
               color = dist_b1, linetype = sd_b1)) +
    geom_line(linewidth = 0.8) +
    geom_point(size = 2) +
    scale_x_continuous("Number of subjects", breaks = c(30, 60, 100)) +
    scale_y_continuous("P(BF01 > 3)", limits = c(0, 1)) +
    scale_color_viridis_d("Prior family", option = "D", end = 0.85) +
    scale_linetype_manual("Prior SD",
      values = c("sd = 0.25" = "solid", "sd = 0.5" = "dashed")) +
    facet_wrap(~ n_trials,
               labeller = labeller(
                 n_trials = function(x) paste0(x, " trials")
               )) +
    labs(title = "Specificity under H0: P(BF01 > 3)",
         subtitle = "Matched intercept prior (Logistic, sd = 0.75)") +
    base_theme()
}
```

@fig-specificity shows that specificity is high across most prior configurations,
but with a notable pattern: at $\text{sd}_{b_1} = 0.25$ (solid lines), the
Normal prior yields substantially lower specificity than the Logistic and Cauchy,
particularly at small sample sizes. At $n = 30$ with 20 trials, the Normal prior
correctly supports H0 in only 41% of null-effect replications, compared to 69%
for the Logistic and 76% for the Cauchy.

This pattern is a direct consequence of the Savage–Dickey density ratio mechanism
and the fact that the three families differ in their density at exactly zero when
the scale parameter is equated. At $\text{sd} = 0.25$: $\text{Normal}(0, 0.25)$
places density 1.60 at zero, $\text{Cauchy}(0, 0.25)$ places 1.27, and
$\text{Logistic}(0, 0.25)$ places 1.00. A higher prior density at zero makes the
numerator of the Savage–Dickey ratio
$\pi(b_1 = 0) / p(b_1 = 0 \mid \text{data})$ larger, so a greater posterior
pile-up at zero is needed for $\text{BF}_{01}$ to exceed the threshold —
reducing specificity. The wider prior ($\text{sd} = 0.50$, dashed lines) lowers
the prior density at zero for all families (Normal: 0.80, Cauchy: 0.64,
Logistic: 0.50), which is why specificity improves and the between-family
differences shrink at the wider scale.

The key take-home is that at the recommended scale of $\text{sd}_{b_1} = 0.25$,
the Logistic prior offers the best specificity (due to its lowest density at
zero), the Cauchy is intermediate, and the Normal yields the lowest specificity —
though all three families converge to adequate specificity at larger sample sizes
($n \geq 60$ with 50 trials). The recommended priors do not produce misleading
evidence: false positive rates ($\text{BF}_{10} > 3$ under H0) remain below 5%
across all conditions.

Given that the choice of prior family affects specificity through the
Savage–Dickey mechanism (though not power), we next ask directly: how similar are
the three prior families in their overall BF performance?


## Does the Choice of Prior Family Matter?

A researcher choosing a prior for the effect coefficient faces three common
options — Normal, Logistic, and Cauchy — that differ in their tail behavior.
The prior predictive analysis showed that the Cauchy's heavy tails assign
implausible probability to extreme effects. But does this translate into
meaningful differences in Bayes Factor performance? @fig-prior-sensitivity
addresses this question by holding the prior scale constant at
$\text{sd}_{b_1} = 0.25$ and comparing power across families.

```{r}
#| label: fig-prior-sensitivity
#| fig-cap: "At the same prior scale (sd = 0.25), the three prior families
#|   produce nearly identical power curves. This holds across all sample sizes,
#|   trial counts, and effect sizes examined, indicating that the choice of
#|   prior family is far less consequential than the choice of prior scale.
#|   Error bars: 95% credible intervals."
#| fig-height: 6
#| fig-width: 10

if (posteriors_available) {
  posteriors |>
    filter(
      sd_b1 == 0.25,
      dist_b0 == "logistic",
      true_b1 > 0,
      threshold == 3
    ) |>
    mutate(
      dist_b1 = factor(dist_b1, levels = c("cauchy", "normal", "logistic"),
                       labels = c("Cauchy", "Normal", "Logistic"))
    ) |>
    ggplot(aes(x = true_b1, y = power_post_mean, color = dist_b1)) +
    geom_line(linewidth = 0.8) +
    geom_point(size = 2) +
    geom_errorbar(aes(ymin = power_ci_lower, ymax = power_ci_upper),
                  width = 0.02, linewidth = 0.4, alpha = 0.6) +
    geom_hline(yintercept = 0.80, linetype = "dotted", color = "grey40") +
    scale_x_continuous("True effect b1") +
    scale_y_continuous("P(BF10 > 3)", limits = c(NA, 1)) +
    scale_color_viridis_d("Prior family", option = "D", end = 0.85) +
    facet_grid(n_trials ~ n_subjects,
               labeller = labeller(
                 n_subjects = function(x) paste0(x, " subj"),
                 n_trials   = function(x) paste0(x, " trials")
               )) +
    labs(title = "Sensitivity to prior family (sd_b1 = 0.25, matched intercept prior)") +
    base_theme()
} else {
  summaries |>
    filter(sd_b1 == 0.25, dist_b0 == "logistic", true_b1 > 0) |>
    mutate(
      dist_b1 = factor(dist_b1, levels = c("cauchy", "normal", "logistic"),
                       labels = c("Cauchy", "Normal", "Logistic"))
    ) |>
    ggplot(aes(x = true_b1, y = P_BF10_gt3, color = dist_b1)) +
    geom_line(linewidth = 0.8) +
    geom_point(size = 2) +
    geom_hline(yintercept = 0.80, linetype = "dotted", color = "grey40") +
    scale_x_continuous("True effect b1") +
    scale_y_continuous("P(BF10 > 3)", limits = c(NA, 1)) +
    scale_color_viridis_d("Prior family", option = "D", end = 0.85) +
    facet_grid(n_trials ~ n_subjects,
               labeller = labeller(
                 n_subjects = function(x) paste0(x, " subj"),
                 n_trials   = function(x) paste0(x, " trials")
               )) +
    labs(title = "Sensitivity to prior family (sd_b1 = 0.25, matched intercept prior)") +
    base_theme()
}
```

The answer is clear for power: at the same prior scale, the three families
produce nearly identical power curves across all sample sizes and effect sizes.
For *specificity*, however, the picture is more nuanced — as shown in
@fig-specificity, the Normal prior's higher density at zero reduces its
specificity relative to the Logistic and Cauchy at $\text{sd}_{b_1} = 0.25$,
especially at small $n$. Taken together, these results give a practical edge to
the Logistic prior, which combines the best specificity profile (lowest density
at zero among the three families), well-behaved prior predictive distributions,
and identical power to the alternatives. The Cauchy, while producing acceptable
power, assigns implausibly heavy-tailed prior predictive distributions and
offers no compensating advantage. We therefore recommend the **Logistic effect
prior** at $\text{sd}_{b_1} = 0.25$ as the default choice for Bayes Factor
testing in binomial GLMs.

A similar question arises for the intercept prior: does the matched-prior
principle — which the prior predictive analysis showed produces cleaner
distributions on the probability scale — also matter for Bayes Factor calibration?


## Does the Intercept Prior Affect Bayes Factor Calibration?

The prior predictive analysis established the matched-prior principle: placing
a Logistic prior on the logit-scale intercept produces a uniform distribution on
the probability scale, whereas a Normal prior (misfit) concentrates mass
differently. But does this distinction matter for Bayes Factor calibration?
@fig-misfit-bf compares power under the matched (Logistic) and misfit (Normal)
intercept priors, holding the effect prior constant.

```{r}
#| label: fig-misfit-bf
#| fig-cap: "Using a mismatched intercept prior (Normal instead of the matched
#|   Logistic) has negligible impact on power. The two intercept priors produce
#|   overlapping power curves across all effect sizes, sample sizes, and trial
#|   counts. This confirms that the intercept prior primarily regularizes the
#|   baseline probability and does not alter inference about the effect.
#|   Error bars: 95% credible intervals."
#| fig-height: 6
#| fig-width: 10

if (posteriors_available) {
  posteriors |>
    filter(
      dist_b1 == "logistic",
      sd_b1   == 0.25,
      true_b1 > 0,
      threshold == 3
    ) |>
    mutate(
      dist_b0 = factor(dist_b0, levels = c("logistic", "normal"),
                       labels = c("Logistic (matched)", "Normal (misfit)"))
    ) |>
    ggplot(aes(x = true_b1, y = power_post_mean, color = dist_b0)) +
    geom_line(linewidth = 0.8) +
    geom_point(size = 2) +
    geom_errorbar(aes(ymin = power_ci_lower, ymax = power_ci_upper),
                  width = 0.02, linewidth = 0.4, alpha = 0.6) +
    geom_hline(yintercept = 0.80, linetype = "dotted", color = "grey40") +
    scale_x_continuous("True effect b1") +
    scale_y_continuous("P(BF10 > 3)", limits = c(NA, 1)) +
    scale_color_viridis_d("Intercept prior", option = "D", end = 0.7) +
    facet_grid(n_trials ~ n_subjects,
               labeller = labeller(
                 n_subjects = function(x) paste0(x, " subjects"),
                 n_trials   = function(x) paste0(x, " trials")
               )) +
    labs(title = "Effect of intercept prior misfit on power (logit link, sd_b1 = 0.25)") +
    base_theme()
} else {
  summaries |>
    filter(dist_b1 == "logistic", sd_b1 == 0.25, true_b1 > 0) |>
    mutate(
      dist_b0 = factor(dist_b0, levels = c("logistic", "normal"),
                       labels = c("Logistic (matched)", "Normal (misfit)"))
    ) |>
    ggplot(aes(x = true_b1, y = P_BF10_gt3, color = dist_b0)) +
    geom_line(linewidth = 0.8) +
    geom_point(size = 2) +
    geom_hline(yintercept = 0.80, linetype = "dotted", color = "grey40") +
    scale_x_continuous("True effect b1") +
    scale_y_continuous("P(BF10 > 3)", limits = c(NA, 1)) +
    scale_color_viridis_d("Intercept prior", option = "D", end = 0.7) +
    facet_grid(n_trials ~ n_subjects,
               labeller = labeller(
                 n_subjects = function(x) paste0(x, " subjects"),
                 n_trials   = function(x) paste0(x, " trials")
               )) +
    labs(title = "Effect of intercept prior misfit on power (logit link, sd_b1 = 0.25)") +
    base_theme()
}
```

The answer is reassuring: the matched and misfit intercept priors produce
essentially overlapping power curves. The intercept prior regularizes the
baseline probability — keeping it away from 0 and 1 — but has no meaningful
effect on inference about the experimental effect $b_1$, as long as the prior
SD is reasonable (here, $\text{sd}_{b_0} = 0.75$). This is good news for
researchers who may not use the exact matched prior: even a "misfit" Normal
intercept prior on the logit scale does not compromise BF calibration.


## What Happens When All Three Outcomes Are Considered?

So far we have examined power and specificity separately — asking how often the
Bayes Factor correctly supports H1 when effects exist, and how often it supports
H0 when they do not. But each replication actually falls into one of three
categories: the BF may favor H1 (BF10 > 3), favor H0 (BF01 > 3), or remain
inconclusive (neither threshold crossed). Decomposing outcomes this way reveals
where the "undecided" fraction concentrates and what conditions demand the most
attention to sample size.

```{r}
#| label: fig-three-category
#| fig-cap: "When effects are large or absent, the Bayes Factor is decisive in
#|   the majority of replications. The inconclusive fraction (middle category)
#|   is largest at small effects with few trials — the conditions where sample
#|   size recommendations matter most. Under the null, most replications
#|   correctly support H0, with a small fraction remaining undecided.
#|   Bars show Dirichlet posterior means for the recommended Normal effect
#|   prior (sd = 0.25) with matched Logistic intercept prior."
#| fig-height: 7
#| fig-width: 10
#| eval: !expr posteriors_available

posteriors |>
  filter(
    dist_b0 == "logistic",
    dist_b1 == "normal",
    sd_b1   == 0.25,
    threshold == 3
  ) |>
  select(true_b1, n_subjects, n_trials, h1_mean, inc_mean, h0_mean) |>
  pivot_longer(cols = c(h1_mean, inc_mean, h0_mean),
               names_to = "outcome", values_to = "proportion") |>
  mutate(
    outcome = factor(outcome,
      levels = c("h1_mean", "inc_mean", "h0_mean"),
      labels = c("H1 detected (BF10 > 3)",
                  "Inconclusive",
                  "H0 detected (BF01 > 3)")),
    true_b1_label = factor(paste0("b1 = ", true_b1))
  ) |>
  ggplot(aes(x = factor(n_subjects), y = proportion, fill = outcome)) +
  geom_col(position = "stack", width = 0.7) +
  facet_grid(n_trials ~ true_b1_label,
             labeller = labeller(
               n_trials = function(x) paste0(x, " trials")
             )) +
  scale_x_discrete("Number of subjects") +
  scale_y_continuous("Proportion of replications", limits = c(0, 1)) +
  scale_fill_viridis_d("BF outcome", option = "D", end = 0.85) +
  labs(title = "Three-category BF outcome decomposition (BF threshold = 3)",
       subtitle = "Normal effect prior (sd = 0.25), matched Logistic intercept prior") +
  base_theme()
```

@fig-three-category reveals the full picture. Under the null ($b_1 = 0$), the
large majority of replications correctly favor H0, with only a small fraction
remaining undecided. As the true effect grows, the H1-detected fraction
dominates. The most informative pattern appears at small effects ($b_1 = 0.05$
and $0.10$) with 20 trials: here, a substantial proportion of replications
remain inconclusive, meaning the BF neither confirms nor rules out the effect.
This is precisely the regime where collecting more data — either more trials per
subject or more subjects — makes the biggest practical difference. At 50 trials,
the inconclusive fraction shrinks considerably even for small effects.

The following sections address the role of the random-effects hyperprior, which
operates through a different mechanism than the fixed-effects priors examined so
far.


## Does the Random-Effects Hyperprior Matter? {#sec-re-hyperprior}

The analyses above used a single random-effects hyperprior
($\text{Exponential}(4)$) and focused exclusively on the fixed-effects priors.
But the prior predictive analysis showed that the choice of hyperprior affects
the distribution of subject-level probabilities — the brms default
Student-*t*(3, 0, 2.5) places a heavy right tail on between-subject SDs,
pushing individual participants toward floor and ceiling. Does this also affect
Bayes Factor performance?

To answer this, we ran a second simulation with a within-subjects design
(each subject in both conditions, with correlated random intercepts and slopes)
while holding the fixed-effects priors at the recommended matched configuration.
The only design factor was the hyperprior on the random-effects SDs: the brms
default Student-*t*(3, 0, 2.5), Gamma(2, 4), and Exponential(4).

```{r}
#| label: setup-re
#| eval: !expr re_results_available

re_plot_data <- re_summaries |>
  mutate(
    sd_prior_re = factor(sd_prior_re,
                         levels = c("default", "gamma", "exponential"),
                         labels = c("Student-t(3,0,2.5)", "Gamma(2,4)",
                                    "Exponential(4)")),
    n_trials_label = paste0(n_trials, " trials")
  )
```

```{r}
#| label: setup-re-posteriors
#| eval: !expr re_posteriors_available

re_post_data <- re_posteriors |>
  mutate(
    sd_prior_re = factor(sd_prior_re,
                         levels = c("default", "gamma", "exponential"),
                         labels = c("Student-t(3,0,2.5)", "Gamma(2,4)",
                                    "Exponential(4)")),
    n_trials_label = paste0(n_trials, " trials")
  )
```

### Power is unaffected by the RE hyperprior

```{r}
#| label: fig-re-power
#| fig-cap: "The choice of random-effects hyperprior has no meaningful impact
#|   on power. All three hyperpriors produce nearly identical power curves
#|   across all effect sizes and sample sizes. This confirms that Bayes Factor
#|   power is driven by the fixed-effects priors and the amount of data, not
#|   by the RE hyperprior. Within-subjects design with correlated RE; matched
#|   fixed-effects priors. Error bars: 95% credible intervals."
#| fig-height: 6
#| fig-width: 10
#| eval: !expr re_posteriors_available || re_results_available

if (re_posteriors_available) {
  re_post_data |>
    filter(true_b1 > 0, threshold == 3) |>
    ggplot(aes(x = n_subjects, y = power_post_mean, color = sd_prior_re,
               linetype = sd_prior_re)) +
    geom_line(linewidth = 0.8) +
    geom_point(size = 2) +
    geom_errorbar(aes(ymin = power_ci_lower, ymax = power_ci_upper),
                  width = 3, linewidth = 0.4, alpha = 0.6) +
    geom_hline(yintercept = 0.80, linetype = "dotted", color = "grey40") +
    scale_x_continuous("Number of subjects", breaks = c(30, 60, 100)) +
    scale_y_continuous("P(BF10 > 3)", limits = c(NA, 1)) +
    scale_color_viridis_d("RE hyperprior", option = "D", end = 0.85) +
    scale_linetype_manual("RE hyperprior",
      values = c("Student-t(3,0,2.5)" = "solid",
                 "Gamma(2,4)"         = "dashed",
                 "Exponential(4)"     = "dotdash")) +
    facet_grid(n_trials_label ~ true_b1,
               labeller = labeller(
                 true_b1 = function(x) paste0("b1 = ", x)
               )) +
    labs(title = "Power by RE hyperprior (within-subjects, matched fixed-effects priors)",
         subtitle = "Dotted line = 80% power criterion") +
    base_theme()
} else {
  re_plot_data |>
    filter(true_b1 > 0) |>
    ggplot(aes(x = n_subjects, y = P_BF10_gt3, color = sd_prior_re,
               linetype = sd_prior_re)) +
    geom_line(linewidth = 0.8) +
    geom_point(size = 2) +
    geom_hline(yintercept = 0.80, linetype = "dotted", color = "grey40") +
    scale_x_continuous("Number of subjects", breaks = c(30, 60, 100)) +
    scale_y_continuous("P(BF10 > 3)", limits = c(NA, 1)) +
    scale_color_viridis_d("RE hyperprior", option = "D", end = 0.85) +
    scale_linetype_manual("RE hyperprior",
      values = c("Student-t(3,0,2.5)" = "solid",
                 "Gamma(2,4)"         = "dashed",
                 "Exponential(4)"     = "dotdash")) +
    facet_grid(n_trials_label ~ true_b1,
               labeller = labeller(
                 true_b1 = function(x) paste0("b1 = ", x)
               )) +
    labs(title = "Power by RE hyperprior (within-subjects, matched fixed-effects priors)",
         subtitle = "Dotted line = 80% power criterion") +
    base_theme()
}
```

@fig-re-power confirms that the RE hyperprior has no meaningful impact on power.
All three hyperpriors produce nearly identical detection rates across effect sizes
and sample sizes. This makes sense: the Bayes Factor for the fixed effect $b_1$
is primarily driven by the fixed-effects priors and the data, not by how between-
subject variability is modeled. For power, researchers can choose any reasonable
hyperprior without concern.

### Specificity is similar across RE hyperpriors

Does the RE hyperprior affect specificity — the ability to correctly support the
null when there is no effect? @fig-re-specificity addresses this question.

```{r}
#| label: fig-re-specificity
#| fig-cap: "All three RE hyperpriors produce broadly similar specificity,
#|   with widely overlapping credible intervals across all conditions. At
#|   N = 30 with 20 trials, specificity is low (~43%) regardless of
#|   hyperprior choice. At larger samples, the three hyperpriors remain
#|   closely matched. Error bars: 95% credible intervals."
#| fig-height: 5
#| fig-width: 10
#| eval: !expr re_posteriors_available || re_results_available

if (re_posteriors_available) {
  re_post_data |>
    filter(true_b1 == 0, threshold == 3) |>
    ggplot(aes(x = n_subjects, y = spec_post_mean, color = sd_prior_re,
               linetype = sd_prior_re)) +
    geom_line(linewidth = 0.8) +
    geom_point(size = 2) +
    geom_errorbar(aes(ymin = spec_ci_lower, ymax = spec_ci_upper),
                  width = 3, linewidth = 0.4, alpha = 0.6) +
    scale_x_continuous("Number of subjects", breaks = c(30, 60, 100)) +
    scale_y_continuous("P(BF01 > 3)", limits = c(0, 1)) +
    scale_color_viridis_d("RE hyperprior", option = "D", end = 0.85) +
    scale_linetype_manual("RE hyperprior",
      values = c("Student-t(3,0,2.5)" = "solid",
                 "Gamma(2,4)"         = "dashed",
                 "Exponential(4)"     = "dotdash")) +
    facet_wrap(~ n_trials_label) +
    labs(title = "Specificity under H0 by RE hyperprior") +
    base_theme()
} else {
  re_plot_data |>
    filter(true_b1 == 0) |>
    ggplot(aes(x = n_subjects, y = P_BF01_gt3, color = sd_prior_re,
               linetype = sd_prior_re)) +
    geom_line(linewidth = 0.8) +
    geom_point(size = 2) +
    scale_x_continuous("Number of subjects", breaks = c(30, 60, 100)) +
    scale_y_continuous("P(BF01 > 3)", limits = c(0, 1)) +
    scale_color_viridis_d("RE hyperprior", option = "D", end = 0.85) +
    scale_linetype_manual("RE hyperprior",
      values = c("Student-t(3,0,2.5)" = "solid",
                 "Gamma(2,4)"         = "dashed",
                 "Exponential(4)"     = "dotdash")) +
    facet_wrap(~ n_trials_label) +
    labs(title = "Specificity under H0 by RE hyperprior") +
    base_theme()
}
```

@fig-re-specificity shows that the three RE hyperpriors produce broadly similar
specificity across all conditions, with widely overlapping 95% credible
intervals. At $n = 30$ with only 20 trials per subject, all three hyperpriors
yield similarly low specificity (Student-*t*: 0.44 [0.35, 0.54]; Gamma: 0.43
[0.34, 0.53]; Exponential: 0.42 [0.33, 0.52]) — the low values here reflect
the limited data rather than the hyperprior choice. At larger sample sizes, the
three hyperpriors remain closely matched, with no systematic advantage for any
one candidate. These results indicate that, unlike the fixed-effects prior
family (which affects specificity through the Savage–Dickey density at zero),
the RE hyperprior does not meaningfully influence inferential performance for
the fixed effect. The recommendation for Exponential(4) as a default hyperprior
therefore rests on its prior predictive properties — it concentrates
between-subject variability in a substantively plausible range and avoids the
heavy right tail of the Student-*t* that can push individual probabilities
toward floor and ceiling (see the prior predictive analysis) — rather than on
BF calibration differences.

@tbl-re-summary provides the corresponding minimum sample sizes for 80% power,
confirming that all three hyperpriors require the same sample sizes to reach
adequate power.

```{r}
#| label: tbl-re-summary
#| tbl-cap: "Minimum sample size for 80% power (50 trials per subject) is
#|   identical across RE hyperpriors, confirming that the hyperprior does not
#|   affect power. Within-subjects design with correlated RE;
#|   matched fixed-effects priors."
#| eval: !expr re_results_available

re_plot_data |>
  filter(n_trials == 50, true_b1 > 0) |>
  group_by(sd_prior_re, true_b1) |>
  summarise(
    min_n_80pct = {
      n_meeting <- n_subjects[P_BF10_gt3 >= 0.80]
      if (length(n_meeting) > 0) min(n_meeting) else NA_integer_
    },
    .groups = "drop"
  ) |>
  pivot_wider(names_from = true_b1, values_from = min_n_80pct,
              names_prefix = "b1 = ") |>
  rename("RE hyperprior" = sd_prior_re) |>
  knitr::kable(caption = "Minimum n_subjects for 80% power by RE hyperprior (n_trials = 50)")
```


## Putting It All Together

Across all conditions examined, Bayes Factor calibration is primarily driven by
two factors: the scale of the effect prior and the amount of data (both sample
size and trials per subject). The choice of prior *family* has negligible impact
on power, but does affect specificity at the recommended scale of
$\text{sd}_{b_1} = 0.25$: the Normal prior's higher density at zero reduces its
specificity relative to the Logistic and Cauchy, particularly at small $n$. The
intercept prior (matched vs. misfit) makes no practical difference. The
random-effects hyperprior has no systematic effect on either power or specificity
— the three hyperpriors examined produce broadly similar results across all
conditions, though the Exponential(4) is preferred on prior predictive grounds.

These results confirm and extend the prior predictive analysis. The matched
priors recommended there — Logistic(0, 0.75) for the intercept and a Logistic
effect prior at $\text{sd}_{b_1} = 0.25$ — not only produce substantively
plausible prior predictive distributions, but also yield well-calibrated Bayes
Factor inference with the best specificity profile among the families examined.
Priors that are too wide degrade power, while the Normal effect prior at the
recommended scale incurs a specificity cost at small samples that favours the
Logistic alternative.

### Practical implication: scale matters most

For researchers choosing priors for Bayes Factor testing in binomial GLMs, these
results deliver a reassuringly simple message. The choice that matters most is
the *scale* of the effect prior — not the family, not the intercept prior, and
not the random-effects hyperprior. This parallels findings from the broader BFDA
literature [@schonbrodt2018; @stefan2019], where informed priors at appropriate
scales consistently outperform uninformed defaults. The key contribution of the
present work is to establish *what* that appropriate scale is for binomial models
and *why* — connecting the prior scale to interpretable quantities on the
probability scale via the matched prior principle.

### A template for other model classes

The workflow demonstrated here — (1) derive matched priors analytically,
(2) characterise the prior–observation mapping via prior predictive simulation,
(3) calibrate BF performance via large-scale BFDA — is not specific to binomial
models. Any model class with a nonlinear link function faces the same challenge:
ordinal models (cumulative logit/probit), count models (log link), beta
regression (logit link on the mean), and distributional regression models more
broadly [@klein2015; @burkner2021]. In each case, the probability integral
transform identifies the matched prior family, prior predictive simulation
reveals the implied effect sizes, and BFDA quantifies inferential properties.
We encourage researchers and methodologists to apply this template to
establish principled defaults for Bayes Factor testing beyond the normal model
— an area where the existing Rouder–Wagenmakers framework
[@rouder2009; @rouder2012] currently does not reach.


## Sample Size Guidance

For practical study planning, the following tables summarize the minimum sample
sizes needed to achieve 80% power across different prior specifications and
effect sizes, assuming 50 trials per subject and the recommended matched
intercept prior.

```{r}
#| label: tbl-summary
#| tbl-cap: "Minimum number of subjects for 80% power (P(BF10 > 3) >= 0.80)
#|   at 50 trials per subject, by prior family, prior SD, and true effect size.
#|   Between-subjects design with matched Logistic intercept prior."

summaries |>
  filter(
    n_trials == 50,
    dist_b0 == "logistic",
    true_b1 > 0
  ) |>
  group_by(dist_b1, sd_b1, true_b1) |>
  summarise(
    min_n_80pct = {
      n_meeting <- n_subjects[P_BF10_gt3 >= 0.80]
      if (length(n_meeting) > 0) min(n_meeting) else NA_integer_
    },
    .groups = "drop"
  ) |>
  mutate(
    dist_b1 = factor(dist_b1, levels = c("cauchy", "normal", "logistic"),
                     labels = c("Cauchy", "Normal", "Logistic"))
  ) |>
  pivot_wider(names_from = true_b1, values_from = min_n_80pct,
              names_prefix = "b1 = ") |>
  knitr::kable(caption = "Minimum n_subjects for 80% power (n_trials = 50, matched intercept prior)")
```

@tbl-summary-ci provides more detail for the recommended configuration (Normal
effect prior, $\text{sd}_{b_1} = 0.25$), showing the posterior mean and 95%
credible interval for power at each combination of effect size and sample size.

```{r}
#| label: tbl-summary-ci
#| tbl-cap: "Power estimates with uncertainty for the recommended prior
#|   configuration (Normal effect prior, sd = 0.25, 50 trials per subject,
#|   matched Logistic intercept prior). Values show posterior mean [95%
#|   credible interval] for P(BF10 > 3)."
#| eval: !expr posteriors_available

posteriors |>
  filter(
    n_trials == 50,
    dist_b0 == "logistic",
    dist_b1 == "normal",
    sd_b1   == 0.25,
    true_b1 > 0,
    threshold == 3
  ) |>
  mutate(
    power_summary = sprintf("%.2f [%.2f, %.2f]",
                            power_post_mean, power_ci_lower, power_ci_upper)
  ) |>
  select(true_b1, n_subjects, power_summary) |>
  pivot_wider(names_from = true_b1, values_from = power_summary,
              names_prefix = "b1 = ") |>
  rename("n_subjects" = n_subjects) |>
  knitr::kable(caption = "Power posterior mean [95% CI] (Normal prior, sd = 0.25, n_trials = 50)")
```


## Recommendations for Practice

The most important decision when specifying priors for Bayes Factor testing in
binomial GLMs is the **scale of the effect prior**. An effect prior with
$\text{sd}_{b_1} = 0.25$ (under ±1 sum-to-zero coding) provides a good balance
between power and specificity for typical cognitive psychology experiments. At
this scale, the prior concentrates mass on effects up to roughly 20 percentage
points on the probability scale — the range most commonly observed in cognitive
experiments [@oberauer2019] — while still allowing larger effects to receive
non-trivial prior probability. Combined with the prior predictive analysis, which
showed that $\text{sd}_{b_1} = 0.25$ yields a 90th percentile of
$|\delta p| \approx 0.20$, this represents a principled default that is both
substantively informed and inferentially well-calibrated.

This parallels the core insight from BFDA [@schonbrodt2018]: prior specification
is not merely a philosophical choice but a design decision that directly affects
what evidence a study can produce. For normal-data models, the JZS prior
[@rouder2009] provides a principled scale through the Cauchy distribution on
standardised effect sizes. For binomial GLMs, the link function makes
standardisation less natural, and the matched prior principle offers an
alternative anchor — grounded in the probability integral transform rather
than standardised effect sizes.

The **choice of prior family matters far less than the choice of scale** for
power, but does affect specificity at $\text{sd}_{b_1} = 0.25$. All three
families produce nearly identical power curves. For specificity, however, the
families differ in their density at exactly zero — Normal: 1.60, Cauchy: 1.27,
Logistic: 1.00 at scale $= 0.25$ — which directly influences the Savage–Dickey
ratio under the null. The Logistic prior, with the lowest density at zero, yields
the best specificity. We therefore recommend the **Logistic effect prior** as the
default: it offers the best specificity profile, well-behaved prior predictive
distributions, and identical power to the alternatives. The Cauchy, despite
acceptable power, assigns implausible probability to extreme effects — consistent
with the concerns raised by @ghosh2018 about Cauchy priors producing
extraordinarily large posterior samples under separation — and is not recommended
as a default for hypothesis testing in this context.

Similarly, the **intercept prior** has minimal impact on Bayes Factor
calibration. While we recommend the matched intercept prior (Logistic(0, 0.75)
for logit, Normal(0, 0.75) for probit) based on its superior prior predictive
properties, using a mismatched intercept prior does not noticeably degrade
inferential performance. The main role of the intercept prior is to regularize
the baseline probability away from floor and ceiling, which benefits model
fitting but does not substantially affect the Bayes Factor for the effect.

For **sample size planning**, the number of trials per subject is at least as
important as the number of subjects. With 50 trials and
$\text{sd}_{b_1} = 0.25$, the 80% power criterion is met at $n = 30$ for
effects of $b_1 \geq 0.20$ on the logit scale across all three prior families
(Logistic: 0.83, Normal: 0.91, Cauchy: 0.89). At
$\text{sd}_{b_1} = 0.50$, the Logistic prior just misses the 80% threshold
(0.79), while the Normal and Cauchy remain above it — a minor trade-off of the
Logistic prior's specificity advantage. Smaller effects
($b_1 = 0.05$–$0.10$, corresponding to 2–5 percentage points) require larger
samples. Reducing from 50 to 20 trials per subject substantially lowers power
for these smaller effects, making trial count a critical design variable.

Finally, the **random-effects hyperprior** has no systematic impact on BF
calibration. All three hyperpriors examined — the brms default Student-*t*(3, 0,
2.5), Gamma(2, 4), and Exponential(4) — produce similar power and specificity
profiles with widely overlapping credible intervals. We nonetheless recommend
Exponential(4) as a default hyperprior for the random-effects SDs on prior
predictive grounds: it concentrates between-subject variability in a
substantively plausible range and avoids the heavy right tail of the Student-*t*
that can push individual probabilities toward floor and ceiling (see the prior
predictive analysis). This choice does not compromise BF performance.


## Generalising to Other Model Classes

The binomial GLM is one member of a larger family of models where link functions
mediate between the parameter scale (where priors are specified) and the
observation scale (where scientific quantities live). The workflow presented here
— analytical matched priors, systematic prior predictive simulation, and
large-scale BF calibration — transfers directly to these other model classes.

**Ordinal regression.** Cumulative logit and probit models use the same link
functions as binomial GLMs, so the matched prior principle applies directly to
the threshold parameters. A Logistic prior on logit-scale thresholds
produces uniform implied category probabilities — a natural non-informative
starting point. Effect priors require the same prior predictive calibration
demonstrated here.

**Count data models.** Poisson and negative binomial regression use a log link.
The matched prior principle suggests an Exponential prior on the log-scale
intercept (since $\exp(X) \sim \text{Exponential}(1)$ when
$X \sim \text{Gumbel}$). Prior predictive simulation can reveal what prior
scales on log-scale coefficients imply for rate ratios.

**Distributional regression.** In GAMLSS [@klein2015] and brms distributional
models [@burkner2021], *all* parameters of the response distribution
(mean, variance, shape) are linked to predictors through link functions.
Each parameter therefore faces its own prior–link interaction. The matched
prior principle and the simulation approach developed here provide a
systematic way to address prior specification for each parameter.

**Beta regression.** For proportional data, beta regression models the mean via
a logit link. The matched prior principle applies identically to the intercept,
and prior predictive simulation can calibrate the implied distribution of
predicted proportions.

In each case, the key steps are: (1) identify the link function and the
corresponding matched prior family via the probability integral transform;
(2) use prior predictive simulation to map prior scales to interpretable
quantities on the observation scale; and (3) if Bayes Factor testing is
intended, run a calibration simulation to verify adequate power and
specificity. We provide all simulation code to facilitate such extensions.


## Appendix: Quantifying Uncertainty in Detection Rates {#sec-appendix-uncertainty}


Each simulation condition yields $n_{\text{valid}}$ replications (out of 100),
each producing a BF10 value. Rather than reporting detection rates as
point-estimate proportions, we apply Bayesian conjugate updating to obtain
full posterior distributions that reflect the sampling uncertainty from the
finite number of replications.

### Two-category model (Beta-Binomial conjugacy)

For a given BF threshold $\tau$ (e.g., $\tau = 3$ or $\tau = 10$), each
replication is a Bernoulli trial: "success" if BF10 $> \tau$ (for power) or
BF01 $> \tau$ (for specificity). With $k$ successes out of $n$ valid
replications and a $\text{Beta}(\alpha_0, \beta_0)$ prior, the posterior is:

$$\theta \mid k, n \sim \text{Beta}(\alpha_0 + k,\; \beta_0 + n - k)$$

We use $\alpha_0 = \beta_0 = 3$, a symmetric prior centered at 0.5 with 95%
of its mass in $[0.12,\, 0.88]$. This mildly regularises the estimate away
from extreme values while being dominated by the data: the effective prior
sample size is 6, negligible relative to the 100 replications per condition.
For comparison, the maximum shrinkage relative to the MLE is $\leq 0.03$
(at extreme rates near 0 or 1), and results are virtually identical under
the uniform $\text{Beta}(1, 1)$ or Jeffreys $\text{Beta}(0.5, 0.5)$ priors.

### Three-category model (Dirichlet-Multinomial conjugacy)

Each replication falls into exactly one of three categories: (1) H1 detected
(BF10 $> \tau$), (2) inconclusive ($1/\tau \leq \text{BF10} \leq \tau$), or
(3) H0 detected (BF10 $< 1/\tau$). The Dirichlet-Multinomial conjugate
update gives:

$$(\theta_{\text{H1}},\, \theta_{\text{inc}},\, \theta_{\text{H0}}) \mid \mathbf{k} \sim \text{Dirichlet}(\boldsymbol{\alpha}_0 + \mathbf{k})$$

with $\boldsymbol{\alpha}_0 = (1, 3, 1)$. This prior encodes the expectation
that most BFs will be inconclusive (prior mean 60%), with lower prior mass on
decisive evidence for either hypothesis (20% each). The effective prior sample
size is 5, again negligible relative to the data. The marginal posterior for
each category is Beta:
$\theta_j \mid \mathbf{k} \sim \text{Beta}(\alpha_{0,j} + k_j,\; \sum_{m \neq j}(\alpha_{0,m} + k_m))$.

### Sensitivity to Dirichlet prior choice

To confirm that the Dirichlet prior specification is inconsequential given
$n = 100$ replications per condition, we compare the $\text{Dirichlet}(1, 3, 1)$
used in the main analyses to the symmetric $\text{Dirichlet}(1, 1, 1)$ (uniform
on the simplex).

```{r}
#| label: fig-dirichlet-sensitivity
#| fig-cap: "The choice of Dirichlet prior has negligible impact on posterior
#|   means. Points compare the Dirichlet(1,3,1) and Dirichlet(1,1,1) priors
#|   for each simulation condition; all points fall near the identity line,
#|   confirming that 100 replications dominate the prior. Dashed line: identity."
#| fig-height: 5
#| fig-width: 10
#| eval: !expr posteriors_available

# Recompute posteriors under Dirichlet(1,1,1) for comparison
results_dir <- here("output", "Simulation_BFCalibration")
if (dir.exists(results_dir)) {
  cond_files <- list.files(results_dir, pattern = "^BFCalib_Cond-", full.names = TRUE)
} else {
  # Fall back to RE results for sensitivity check
  results_dir <- here("output", "Simulation_BFCalibration_RE")
  cond_files <- list.files(results_dir, pattern = "^BFCalib_RE_Cond-", full.names = TRUE)
}

sensitivity_rows <- lapply(cond_files, function(f) {
  cond_data <- readRDS(f)
  bf10_vec  <- cond_data$results[, "BF10"]
  counts    <- compute_bf_counts(bf10_vec, threshold = 3)

  dir_131 <- dirichlet_posterior_summary(
    counts["n_h1"], counts["n_h0"], counts["n_inconclusive"],
    alpha0 = c(1, 3, 1)
  )
  dir_111 <- dirichlet_posterior_summary(
    counts["n_h1"], counts["n_h0"], counts["n_inconclusive"],
    alpha0 = c(1, 1, 1)
  )

  data.frame(
    h1_mean_131  = dir_131["h1_mean"],
    h1_mean_111  = dir_111["h1_mean"],
    inc_mean_131 = dir_131["inc_mean"],
    inc_mean_111 = dir_111["inc_mean"],
    h0_mean_131  = dir_131["h0_mean"],
    h0_mean_111  = dir_111["h0_mean"]
  )
})

sensitivity <- bind_rows(sensitivity_rows)

p1 <- ggplot(sensitivity, aes(x = h1_mean_111, y = h1_mean_131)) +
  geom_abline(slope = 1, intercept = 0, linetype = "dashed", color = "grey50") +
  geom_point(alpha = 0.5, size = 1.5) +
  scale_x_continuous("Dirichlet(1,1,1)", limits = c(0, 1)) +
  scale_y_continuous("Dirichlet(1,3,1)", limits = c(0, 1)) +
  labs(title = "H1 detected") +
  base_theme()

p2 <- ggplot(sensitivity, aes(x = inc_mean_111, y = inc_mean_131)) +
  geom_abline(slope = 1, intercept = 0, linetype = "dashed", color = "grey50") +
  geom_point(alpha = 0.5, size = 1.5) +
  scale_x_continuous("Dirichlet(1,1,1)", limits = c(0, 1)) +
  scale_y_continuous("Dirichlet(1,3,1)", limits = c(0, 1)) +
  labs(title = "Inconclusive") +
  base_theme()

p3 <- ggplot(sensitivity, aes(x = h0_mean_111, y = h0_mean_131)) +
  geom_abline(slope = 1, intercept = 0, linetype = "dashed", color = "grey50") +
  geom_point(alpha = 0.5, size = 1.5) +
  scale_x_continuous("Dirichlet(1,1,1)", limits = c(0, 1)) +
  scale_y_continuous("Dirichlet(1,3,1)", limits = c(0, 1)) +
  labs(title = "H0 detected") +
  base_theme()

gridExtra::grid.arrange(p1, p2, p3, ncol = 3,
  top = "Dirichlet prior sensitivity: posterior means by category (BF > 3)")
```

```{r}
#| label: dirichlet-sensitivity-summary
#| eval: !expr posteriors_available

max_diff <- sensitivity |>
  summarise(
    max_h1  = max(abs(h1_mean_131 - h1_mean_111)),
    max_inc = max(abs(inc_mean_131 - inc_mean_111)),
    max_h0  = max(abs(h0_mean_131 - h0_mean_111))
  )
```

`r if (posteriors_available) paste0("The maximum absolute difference in posterior means between the two priors across all conditions is ", sprintf("%.3f", max(max_diff$max_h1, max_diff$max_inc, max_diff$max_h0)), " — confirming that the Dirichlet prior choice is negligible relative to the data.")`


## References

::: {#refs}
:::
