---
title: "Default Priors for Binomial GLMs: BF Calibration"
author:
  - name: Klaus Oberauer
    affiliation: University of Zurich
  - name: Joscha Dutli
    affiliation: University of Zurich
  - name: Philipp Musfeld
    affiliation: University of Zurich
  - name: Gidon Frischkorn
    affiliation: University of Lucerne
date: today
format:
  html:
    toc: true
    toc-depth: 3
    code-fold: true
    embed-resources: true
  pdf:
    toc: true
    toc-depth: 3
execute:
  echo: false
  warning: false
  message: false
---

```{r}
#| label: setup
library(dplyr)
library(tidyr)
library(ggplot2)
library(viridis)
library(here)

summaries <- readRDS(here("output", "bf_calibration_summaries.rds"))

# Factor labels
link_labels <- c(logit = "Logit link", probit = "Probit link")
dist_labels <- c(cauchy = "Cauchy", normal = "Normal", logistic = "Logistic")
dist_b0_labels <- c(logistic = "Logistic (matched for logit)",
                    normal   = "Normal (matched for probit)")

base_theme <- function() {
  theme_minimal(base_size = 12) +
    theme(
      panel.grid.minor = element_blank(),
      strip.text       = element_text(face = "bold"),
      legend.position  = "bottom"
    )
}
```

## Introduction

Goal 2 examines how the choice of prior specification affects the behavior of
Bayes factors (BFs) in binomial GLMs. Specifically, we ask:

1. **Power:** Given a true effect of size $b_1$, how often does the BF exceed
   the evidence threshold (BF10 > 3 or BF10 > 10)?
2. **Specificity:** When the null is true ($b_1 = 0$), how often does the BF
   correctly support the null (BF01 > 3)?
3. **Sensitivity to prior:** How much do results change across the three prior
   families (Normal, Logistic, Cauchy)?
4. **Matched vs. misfit priors:** Does using the matched intercept prior (from
   Goal 1) matter for BF calibration?

### Design

- **720 conditions** from the full factor grid (see `scripts/03_BFCalibration.R`)
- **500 replications** per condition
- True intercept fixed at $b_0 = 0$ (50% baseline probability)
- Intercept prior SD fixed at 0.75 (recommended by Goal 1)
- Effect prior SD: 0.25, 0.50, 1.00

---

## Results: Power (H1 true)

### Power by sample size and effect size

```{r}
#| label: fig-power-n
#| fig-cap: "Power (P(BF10 > 3)) as a function of the number of subjects and
#|   the true effect size b1, for each effect prior SD and link function.
#|   Intercept prior: matched (logistic for logit, normal for probit)."
#| fig-height: 8
#| fig-width: 12

matched <- summaries |>
  filter(
    (link == "logit"  & dist_b0 == "logistic") |
    (link == "probit" & dist_b0 == "normal"),
    true_b1 > 0
  ) |>
  mutate(
    dist_b1  = factor(dist_b1, levels = c("cauchy", "normal", "logistic"),
                      labels = c("Cauchy", "Normal", "Logistic")),
    sd_b1    = factor(paste0("sd = ", sd_b1)),
    n_label  = paste0("n = ", n_subjects, " subj × ", n_trials, " trials")
  )

ggplot(matched, aes(x = n_subjects, y = P_BF10_gt3,
                    color = dist_b1, linetype = sd_b1)) +
  geom_line(linewidth = 0.8) +
  geom_point(size = 2) +
  geom_hline(yintercept = 0.80, linetype = "dotted", color = "grey40") +
  scale_x_continuous("Number of subjects", breaks = c(30, 60, 100)) +
  scale_y_continuous("P(BF10 > 3)", limits = c(0, 1)) +
  scale_color_viridis_d("Prior family", option = "D", end = 0.85) +
  scale_linetype_manual("Prior SD",
    values = c("sd = 0.25" = "solid", "sd = 0.5" = "dashed", "sd = 1" = "dotted")) +
  facet_grid(n_trials ~ true_b1 + link,
             labeller = labeller(
               link    = link_labels,
               true_b1 = function(x) paste0("b1 = ", x)
             )) +
  labs(title = "Power: P(BF10 > 3) by sample size and true effect (matched intercept prior)",
       subtitle = "Dotted horizontal line = 0.80 power criterion") +
  base_theme()
```

### Power by n_trials

```{r}
#| label: fig-power-trials
#| fig-cap: "Power as a function of trials per subject, for n_subjects = 60
#|   and each prior specification."
#| fig-height: 6
#| fig-width: 10

matched |>
  filter(n_subjects == 60) |>
  ggplot(aes(x = n_trials, y = P_BF10_gt3,
             color = dist_b1, linetype = sd_b1)) +
  geom_line(linewidth = 0.8) +
  geom_point(size = 2) +
  geom_hline(yintercept = 0.80, linetype = "dotted", color = "grey40") +
  scale_x_continuous("Trials per subject", breaks = c(20, 50, 100)) +
  scale_y_continuous("P(BF10 > 3)", limits = c(0, 1)) +
  scale_color_viridis_d("Prior family", option = "D", end = 0.85) +
  scale_linetype_manual("Prior SD",
    values = c("sd = 0.25" = "solid", "sd = 0.5" = "dashed", "sd = 1" = "dotted")) +
  facet_grid(link ~ true_b1,
             labeller = labeller(
               link    = link_labels,
               true_b1 = function(x) paste0("b1 = ", x)
             )) +
  labs(title = "Power by trials per subject (n_subjects = 60, matched intercept prior)") +
  base_theme()
```

---

## Results: Specificity Under H0

```{r}
#| label: fig-specificity
#| fig-cap: "Specificity (P(BF01 > 3)) under the null (true_b1 = 0) for each
#|   prior specification and sample size."
#| fig-height: 6
#| fig-width: 10

summaries |>
  filter(
    true_b1 == 0,
    (link == "logit"  & dist_b0 == "logistic") |
    (link == "probit" & dist_b0 == "normal")
  ) |>
  mutate(
    dist_b1 = factor(dist_b1, levels = c("cauchy", "normal", "logistic"),
                     labels = c("Cauchy", "Normal", "Logistic")),
    sd_b1   = factor(paste0("sd = ", sd_b1))
  ) |>
  ggplot(aes(x = n_subjects, y = P_BF01_gt3,
             color = dist_b1, linetype = sd_b1)) +
  geom_line(linewidth = 0.8) +
  geom_point(size = 2) +
  scale_x_continuous("Number of subjects", breaks = c(30, 60, 100)) +
  scale_y_continuous("P(BF01 > 3)", limits = c(0, 1)) +
  scale_color_viridis_d("Prior family", option = "D", end = 0.85) +
  scale_linetype_manual("Prior SD",
    values = c("sd = 0.25" = "solid", "sd = 0.5" = "dashed", "sd = 1" = "dotted")) +
  facet_grid(n_trials ~ link,
             labeller = labeller(link = link_labels,
                                 n_trials = function(x) paste0(x, " trials"))) +
  labs(title = "Specificity under H0: P(BF01 > 3)",
       subtitle = "Matched intercept prior (sd_b0 = 0.75)") +
  base_theme()
```

---

## Results: Sensitivity to Prior Family

```{r}
#| label: fig-prior-sensitivity
#| fig-cap: "Comparison of power across prior families at the same sd_b1 = 0.25,
#|   showing how much the choice of prior family matters."
#| fig-height: 6
#| fig-width: 10

summaries |>
  filter(
    sd_b1 == 0.25,
    (link == "logit"  & dist_b0 == "logistic") |
    (link == "probit" & dist_b0 == "normal"),
    true_b1 > 0
  ) |>
  mutate(
    dist_b1 = factor(dist_b1, levels = c("cauchy", "normal", "logistic"),
                     labels = c("Cauchy", "Normal", "Logistic"))
  ) |>
  ggplot(aes(x = true_b1, y = P_BF10_gt3, color = dist_b1)) +
  geom_line(linewidth = 0.8) +
  geom_point(size = 2) +
  geom_hline(yintercept = 0.80, linetype = "dotted", color = "grey40") +
  scale_x_continuous("True effect b1") +
  scale_y_continuous("P(BF10 > 3)", limits = c(0, 1)) +
  scale_color_viridis_d("Prior family", option = "D", end = 0.85) +
  facet_grid(n_trials ~ n_subjects + link,
             labeller = labeller(
               link       = link_labels,
               n_subjects = function(x) paste0(x, " subj"),
               n_trials   = function(x) paste0(x, " trials")
             )) +
  labs(title = "Sensitivity to prior family (sd_b1 = 0.25, matched intercept prior)") +
  base_theme()
```

---

## Results: Matched vs. Misfit Intercept Prior

```{r}
#| label: fig-misfit-bf
#| fig-cap: "Effect of intercept prior misfit on power for the logit link:
#|   matched (Logistic) vs. misfit (Normal) intercept prior."
#| fig-height: 6
#| fig-width: 10

summaries |>
  filter(
    link    == "logit",
    dist_b1 == "logistic",
    sd_b1   == 0.25,
    true_b1 > 0
  ) |>
  mutate(
    dist_b0 = factor(dist_b0, levels = c("logistic", "normal"),
                     labels = c("Logistic (matched)", "Normal (misfit)"))
  ) |>
  ggplot(aes(x = true_b1, y = P_BF10_gt3, color = dist_b0)) +
  geom_line(linewidth = 0.8) +
  geom_point(size = 2) +
  geom_hline(yintercept = 0.80, linetype = "dotted", color = "grey40") +
  scale_x_continuous("True effect b1") +
  scale_y_continuous("P(BF10 > 3)", limits = c(0, 1)) +
  scale_color_viridis_d("Intercept prior", option = "D", end = 0.7) +
  facet_grid(n_trials ~ n_subjects,
             labeller = labeller(
               n_subjects = function(x) paste0(x, " subjects"),
               n_trials   = function(x) paste0(x, " trials")
             )) +
  labs(title = "Effect of intercept prior misfit on power (logit link, sd_b1 = 0.25)") +
  base_theme()
```

---

## Summary Table

```{r}
#| label: tbl-summary
#| tbl-cap: "Power at 80% criterion: minimum n_subjects required to achieve
#|   P(BF10 > 3) >= 0.80, for each prior specification and true effect size
#|   (n_trials = 50, matched intercept prior)."

summaries |>
  filter(
    n_trials == 50,
    (link == "logit"  & dist_b0 == "logistic") |
    (link == "probit" & dist_b0 == "normal"),
    true_b1 > 0
  ) |>
  group_by(link, dist_b1, sd_b1, true_b1) |>
  summarise(
    min_n_80pct = {
      n_meeting <- n_subjects[P_BF10_gt3 >= 0.80]
      if (length(n_meeting) > 0) min(n_meeting) else NA_integer_
    },
    .groups = "drop"
  ) |>
  mutate(
    dist_b1 = factor(dist_b1, levels = c("cauchy", "normal", "logistic"),
                     labels = c("Cauchy", "Normal", "Logistic")),
    link    = factor(link, labels = link_labels)
  ) |>
  pivot_wider(names_from = true_b1, values_from = min_n_80pct,
              names_prefix = "b1 = ") |>
  knitr::kable(caption = "Minimum n_subjects for 80% power (n_trials = 50)")
```

---

## Recommendations

Based on the BF calibration results:

1. **Prior family matters less than prior scale:** The Logistic and Normal
   prior families produce very similar BF calibration at the same `sd_b1`.
   The Cauchy prior tends to reduce power slightly due to heavier tails.

2. **Recommended default:** `sd_b1 = 0.25` with either Logistic (logit link)
   or Normal (probit link) prior provides a good balance of power and
   specificity for typical cognitive psychology effect sizes (b1 = 0.25–0.75).

3. **Intercept prior misfit has minimal impact** on BF calibration when
   `sd_b0 = 0.75`, confirming that the main priority is matching the
   effect prior scale to the expected range of effects.

4. **Sample size requirements:** To achieve 80% power for a medium effect
   (b1 = 0.50, ~20 percentage points) with 50 trials per condition:
   approximately 60 subjects are needed with `sd_b1 = 0.25`.

---

## References

Oberauer, K. (2019). Working memory capacity limits memory for bindings.
*Journal of Cognition, 2*. https://doi.org/10.5334/joc.86

Rouder, J. N., Morey, R. D., Speckman, P. L., & Province, J. M. (2012).
Default Bayes factors for ANOVA designs. *Journal of Mathematical
Psychology, 56*, 356–374. https://doi.org/10.1016/j.jmp.2012.08.001
