---
title: "Supplementary Materials: A Three-Step Workflow for Prior Specification in GLMs"
shorttitle: "Supplement — Prior Specification Workflow"
author:
  - name: Gidon T. Frischkorn
    affiliations:
      - name: University of Zurich
      - name: University of Lucerne
  - name: Joscha Dutli
    affiliations:
      - name: University of Zurich
  - name: Philipp Musfeld
    affiliations:
      - name: University of Amsterdam
      - name: University of Zurich
  - name: Klaus Oberauer
    affiliations:
      - name: University of Zurich
format:
  apaquarto-docx: default
  apaquarto-html: default
  apaquarto-pdf:
    documentmode: man
bibliography:
  - references.bib
csl: apa.csl
---

```{r}
#| label: setup
#| include: false

library(dplyr)
library(tidyr)
library(ggplot2)
library(ggridges)
library(viridis)
library(scales)
library(here)
library(knitr)

source(here("R", "link_functions.R"))
source(here("R", "prior_sampling.R"))
source(here("R", "plotting.R"))

# Goal 1: Prior predictive results
summaries_pp  <- readRDS(here("output", "prior_predictive_summaries.rds"))
draws_long    <- readRDS(here("output", "prior_predictive_draws.rds"))

# RE prior predictive results
re_pp_summaries <- tryCatch(
  readRDS(here("output", "prior_predictive_re_summaries.rds")),
  error = function(e) NULL
)
re_pp_draws <- tryCatch(
  readRDS(here("output", "prior_predictive_re_draws.rds")),
  error = function(e) NULL
)
re_pp_available <- !is.null(re_pp_summaries) && !is.null(re_pp_draws)
```

## Full Simulation Design

### Prior Predictive Simulation (Step 2)

The prior predictive simulation crossed the following factors:

- **Link function**: logit, probit (2 levels)
- **Intercept prior family** ($b_0$): Normal, Logistic, Cauchy (3 levels)
- **Intercept prior SD** ($\text{sd}_{b_0}$): 0.50, 0.75, 1.00, 1.50 (4 levels)
- **Effect prior family** ($b_1$): Normal, Logistic, Cauchy (3 levels)
- **Effect prior SD** ($\text{sd}_{b_1}$): 0.10, 0.15, 0.20, 0.25, 0.30, 0.40, 0.50, 0.75, 1.00, 2.00 (10 levels)

This yields $2 \times 3 \times 4 \times 3 \times 10 = 720$ conditions. For each condition, 10 replications of 10,000 draws were performed using the SimDesign package.

### Bayes Factor Calibration (Step 3)

The BF calibration crossed:

- **Intercept prior family**: Logistic, Normal (2 levels; both at $\text{sd}_{b_0} = 0.75$)
- **Effect prior family**: Normal, Logistic, Cauchy (3 levels)
- **Effect prior SD**: 0.25, 0.50 (2 levels)
- **True effect** ($b_1$): 0, 0.05, 0.10, 0.20, 0.50 (5 levels)
- **Number of subjects**: 30, 60, 100 (3 levels)
- **Trials per subject**: 20, 50 (2 levels)

This yields $2 \times 3 \times 2 \times 5 \times 3 \times 2 = 360$ conditions with 100 replications each.

A separate random-effects simulation held fixed-effects priors at the recommended matched configuration and varied only the hyperprior: Student-*t*(3, 0, 2.5), Gamma(2, 4), and Exponential(4), yielding 90 conditions with 100 replications each.

### Density Formulas for Analytical Overlays

For the logit link with a $\text{Logistic}(0, s)$ prior on $b_0$: $f(p) = f_{\text{Logis}}(\text{logit}(p), 0, s) / [p(1-p)]$. At $s = 1$, the logistic density and the Jacobian $p(1-p)$ cancel, giving $f(p) = 1$. For the probit link with a $\text{Normal}(0, \sigma)$ prior: $f(p) = \phi(\text{probit}(p)/\sigma) / [\sigma \cdot \phi(\text{probit}(p))]$, where $\phi$ denotes the standard normal density. At $\sigma = 1$, the normal densities cancel, again yielding $f(p) = 1$.


## Supplementary Figures

```{r}
#| label: fig-S1-floor-ceiling
#| fig-cap: "Proportion of prior predictive mass in the floor/ceiling region
#|   (p < 0.05 or p > 0.95) as a function of intercept prior SD and family.
#|   The matched prior at $sd_{b_0} = 1.0$ places exactly 10% in this region
#|   (2 x 0.05 = 0.10), confirming the uniform result. At $sd_{b_0} = 0.75$,
#|   floor/ceiling mass drops below 10% for both matched priors."
#| fig-height: 6
#| fig-width: 10

plot_floor_ceiling_heatmap(summaries_pp)
```

```{r}
#| label: fig-S2-re-delta-p
#| fig-cap: "Effect size heterogeneity by hyperprior family and scale for the
#|   intercept + slope model. At matched scale, all families produce similar
#|   effect heterogeneity. Dotted lines mark +/-0.10, +/-0.20, +/-0.30."
#| fig-height: 8
#| fig-width: 12
#| eval: !expr re_pp_available

re_pp_draws |>
  filter(re_structure == "Intercept + slope") |>
  group_by(hp_family, hp_mean) |>
  slice_sample(n = 10000) |>
  ungroup() |>
  mutate(hp_mean_label = paste0("mu == ", hp_mean)) |>
  ggplot(aes(x = delta_p, y = hp_family,
             fill = hp_family, colour = hp_family)) +
  geom_density_ridges(stat = "binline", bins = 80,
                      alpha = 0.5, scale = 0.9) +
  geom_vline(xintercept = 0, linewidth = 0.5, colour = "grey20") +
  geom_vline(xintercept = c(-0.30, -0.20, -0.10, 0.10, 0.20, 0.30),
             linetype = "dotted", linewidth = 0.5, colour = "grey40") +
  facet_wrap(~ hp_mean_label, ncol = 2,
             labeller = label_parsed) +
  scale_fill_viridis_d(option = "D", end = 0.85, guide = "none") +
  scale_colour_viridis_d(option = "D", end = 0.85, guide = "none") +
  scale_x_continuous(
    expression(delta * italic(p)),
    limits = c(-1, 1), breaks = seq(-1, 1, 0.25)
  ) +
  scale_y_discrete("Family") +
  prior_theme()
```
