---
title: "Priors for Bayes Factors with Binomial Models"
shorttitle: "Priors for Binomial Bayes Factors"
author:
  - name: Gidon T. Frischkorn
    orcid: 0000-0002-5055-9764
    email: gidon.frischkorn@unilu.ch
    corresponding: true
    affiliations:
      - name: University of Zurich
        department: Department of Psychology
      - name: University of Lucerne
        department: Faculty of Behavioral Sciences and Psychology
  - name: Joscha Dutli
    orcid: 0009-0009-5007-1281
    affiliations:
      - name: University of Zurich
        department: Department of Psychology
  - name: Philipp Musfeld
    orcid: 0000-0002-6539-0105
    affiliations:
      - name: University of Amsterdam
        department: Department of Psychology
      - name: University of Zurich
        department: Department of Psychology
  - name: Klaus Oberauer
    orcid: 0000-0001-8815-6189
    affiliations:
      - name: University of Zurich
        department: Department of Psychology
author-note:
  disclosures:
    study-registration: "This study was not preregistered."
    data-sharing: "Simulation code and analysis scripts are available at [repository link]."
    conflict-of-interest: "The authors declare no conflicts of interest."
    financial-support: "This research was supported by a grant from the Swiss National Science Foundation (SNSF)."
    authorship-agreements: |
      Gidon T. Frischkorn: Conceptualization, Methodology, Software, Formal Analysis, Visualization, Writing -- Original Draft.
      Joscha Dutli: Conceptualization, Methodology, Writing -- Review & Editing.
      Philipp Musfeld: Conceptualization,Methodology, Writing -- Review & Editing.
      Klaus Oberauer: Conceptualization, Supervision, Methodology, Writing -- Original Draft.
floatsintext: true
output-dir: "."
format:
  apaquarto-docx: default
  apaquarto-pdf:
    keep-tex: true
  apaquarto-html: default
execute:
  echo: false
  warning: false
  message: false
bibliography: references.bib
csl: apa.csl
abstract: "Bayesian hypothesis testing with Bayes Factors in binomial generalized linear models (GLMs) requires priors on the regression coefficients, yet no principled framework exists for logistic or probit regression. The nonlinear link function creates a prior--link interaction: priors that appear reasonable on the coefficient scale can produce implausible distributions on the probability scale. We address this problem through three complementary approaches. First, we exploit the probability integral transform to identify *matched priors* - prior families that mirror the link function and induce exactly uniform distributions on the probability scale. Second, we characterise implied prior predictive distributions across 720 factorial combinations of link function, prior family, and prior scale via large-scale simulation. Third, we calibrate the recommended priors for Bayes Factor hypothesis testing via simulation-based design analysis across 360 conditions, evaluating how the strength of evidence scales with sample size, trial count, and effect magnitude. Our recommendations provide principled defaults for researchers with no domain-specific prior knowledge, and scaling formulas that allow adaptation to any contrast coding, predictor scale, or informed prior belief. Ready-to-use brms prior specifications are provided."
---

```{r}
#| label: setup
library(dplyr)
library(tidyr)
library(ggplot2)
library(ggridges)
library(viridis)
library(scales)
library(here)
library(knitr)

source(here("R", "link_functions.R"))
source(here("R", "prior_sampling.R"))
source(here("R", "plotting.R"))
source(here("R", "bf_helpers.R"))

# Goal 1: Prior predictive results
summaries_pp  <- readRDS(here("output", "prior_predictive_summaries.rds"))
draws_long    <- readRDS(here("output", "prior_predictive_draws.rds"))

# Goal 2: BF calibration results
summaries_bf  <- readRDS(here("output", "bf_calibration_summaries.rds"))

posteriors_bf <- tryCatch(
  readRDS(here("output", "bf_calibration_posteriors.rds")),
  error = function(e) NULL
)
posteriors_available <- !is.null(posteriors_bf)

continuous_bf <- tryCatch(
  readRDS(here("output", "bf_calibration_continuous.rds")),
  error = function(e) NULL
)
continuous_available <- !is.null(continuous_bf)

re_summaries <- tryCatch(
  readRDS(here("output", "bf_calibration_re_summaries.rds")),
  error = function(e) NULL
)
re_results_available <- !is.null(re_summaries)

re_posteriors <- tryCatch(
  readRDS(here("output", "bf_calibration_re_posteriors.rds")),
  error = function(e) NULL
)
re_posteriors_available <- !is.null(re_posteriors)

# Factor labels
dist_labels <- c(cauchy = "Cauchy", normal = "Normal", logistic = "Logistic")

base_theme <- function() {
  theme_minimal(base_size = 12) +
    theme(
      panel.grid.minor = element_blank(),
      strip.text       = element_text(face = "bold"),
      legend.position  = "bottom"
    )
}
```

## Introduction

A researcher fitting a logistic regression in `brms` faces an immediate
practical question: what prior should they place on the regression
coefficients? For normally distributed data, @rouder2012 proposed default
priors for standardized effect sizes that can be used when there are no good
theoretical or empirical grounds for setting priors. The rationale for default
priors is that even if we know nothing about the effect of interest, we have
some general knowledge about the distribution of effect sizes in our field,
and default priors reflect that knowledge. For binomial GLMs - the
workhorse model for accuracy data ($k$ correct out of $n$ trials) in
cognitive psychology - no such default priors exist.

This gap matters because the nonlinear link function in GLMs creates a
prior--link interaction: a prior that appears reasonable on the
regression-coefficient scale can produce implausible distributions on the
probability scale [@gelman2008; @ghosh2018]. Previous work addressed this
problem for posterior estimation; we address the complementary problem of
choosing priors that produce well-calibrated *Bayes Factors* for hypothesis
testing [@wagenmakers2018a; @schonbrodt2018].

We provide a framework for specifying priors in binomial GLMs that covers
both principled defaults - for researchers with no domain-specific
knowledge - and the tools to derive informed priors from domain knowledge.
Our approach proceeds in three steps:

1. Matched priors via the probability integral transform. We identify
   prior families that mirror the link function, yielding exactly uniform
   distributions on the probability scale. These matched priors serve as
   principled anchors from which informed priors can be scaled.

2. Prior predictive simulation. We characterise the implied distributions
   of baseline probability, effect size, and between-subject variability
   across 720 factorial combinations of link function, prior family, and
   prior scale.

3. Bayes Factor calibration. We evaluate how the recommended priors
   perform for hypothesis testing, examining how the strength of evidence
   scales with sample size, trial count, effect magnitude, and prior
   specification across 360 conditions.

We model accuracy data as $k_i \sim \text{Binomial}(n_i, p_i)$ with
$g(p_i) = b_0 + b_1 x_i$, where $g$ is either the logit or probit
link. All simulations use a two-condition design with sum-to-zero contrast
$x \in \{+1, -1\}$, so that
$\delta p = g^{-1}(b_0 + b_1) - g^{-1}(b_0 - b_1)$. This coding ensures
the prior predictive distribution of $\delta p$ is symmetric around zero
when the prior on $b_1$ is symmetric - a desirable property for a default
prior. For researchers using other contrast codes or predictor scales, we
provide scaling formulas in the [Recommendations].

## The Matched Prior Principle {#sec-matched-prior}

The key to choosing an intercept prior is the probability integral
transform: for any continuous distribution $D$ with CDF $F$, if
$X \sim D$, then $F(X) \sim \text{Uniform}(0, 1)$. In a binomial GLM, the
inverse link $g^{-1}$ *is* a CDF, so we can exploit this result
directly.[^prior-lit]

[^prior-lit]: Previous work on priors for logistic regression focused on
estimation problems: @gelman2008 proposed weakly informative Cauchy(0, 2.5)
priors for separation, @ghosh2018 established conditions for posterior
existence, and @greenland2015 compared penalization approaches. These
contributions address estimation, not the distinct requirements of Bayes
Factor testing, where the prior density at the point null directly enters
the Savage--Dickey ratio [@ly2016; @bartos2023].

| Link | Inverse link $g^{-1}$ | Matched prior | Implied prior on $p$ |
|------|----------------------|---------------|----------------------|
| logit | logistic CDF | $\text{Logistic}(0, 1)$ | $\text{Uniform}(0, 1)$ |
| probit | normal CDF $\Phi$ | $\text{Normal}(0, 1)$ | $\text{Uniform}(0, 1)$  |

In other words, when we transform a $\text{Logistic}(0, 1)$ prior on the
logit scale back to the probability scale, the result is a uniform
distribution on $(0, 1)$; the same holds for a $\text{Normal}(0, 1)$ prior
under the probit link. Conversely, scaling the prior above $\text{sd} = 1$
concentrates mass toward the extreme values 0 and 1, reflecting the
typically unreasonable assumption that extreme baseline probabilities are
particularly likely. Scaling below $\text{sd} = 1$ concentrates mass away
from 0 and 1, producing a regularized (i.e., shrunk away from extreme
values) intercept prior suited to experiments where floor and ceiling
performance are unlikely.

The matched prior at scale 1 is therefore maximally uninformative on the
probability scale - a principled anchor that contrasts with specifying
priors based on intuition about regression coefficients, a strategy
@gelman2017prior cautioned against. The matched prior principle generalises
to any model with a CDF-based link function; we return to this point in the
[Recommendations].

### How Do Intercept Priors Interact with Link Functions?

No principled default priors currently exist for binomial GLMs. This is not
merely a gap in the literature - it matters because the nonlinear link
function creates a prior--link interaction: priors that appear reasonable
on the coefficient scale can produce implausible distributions on the
probability scale. For Bayes Factor testing via the Savage--Dickey density
ratio, improper or overly diffuse priors are not merely inconvenient - they
are undefined or produce BFs biased toward H1
[@schonbrodt2018; @consonni2018; @gelman2008]. Current software defaults
illustrate this problem. For instance, `brms` [@burkner2017; @burkner2018]
places a $\text{student\_t}(3, 0, 2.5)$ prior on the intercept and a flat,
improper prior on all population-level effects. These defaults prioritize
estimation stability, not hypothesis testing. The Student-$t$ intercept
prior, passed through the logit link, concentrates more mass near $p = 0$
and $p = 1$ than the matched Logistic prior. The flat slope prior treats
$|\delta p|$ near 1 as equally plausible as $|\delta p|$ near 0.

The matched prior primarily improves interpretability of prior predictive
distributions. As the BF calibration analysis shows (@sec-bf-family), its
effect on Bayes Factor performance is negligible. The recommendation rests
on methodological hygiene - it is the principled choice even when practical
consequences are small.

## How Does Contrast Coding Interact with the Prior? {#sec-contrast-coding}

The choice of contrast coding directly scales the effect coefficient $b_1$
and therefore the implied prior on the probability scale. With sum-to-zero
codes $x \in \{+1, -1\}$, the effect $b_1$ represents the deviation of each
condition from the grand mean, and the distribution of $\delta p$ is
symmetric around zero when the prior on $b_1$ is symmetric - a
desirable property for a default prior. All simulations here use $\pm 1$
coding [see @schad2020 for a thorough treatment of contrast coding in
regression].

With treatment coding ($x \in \{0, 1\}$), one condition depends on $b_0$
alone and the other on $b_0 + b_1$, creating asymmetric prior uncertainty
across conditions. If the intercept prior mean is shifted away from zero,
this asymmetry also skews the prior predictive distribution of $\delta p$.
We therefore recommend sum-to-zero ($\pm 1$) coding as the default, unless
there are substantive reasons to prefer treatment coding - in which case
the resulting asymmetry may actually be desirable, because it reflects a
prior belief that effects are more likely in one direction. Guidance for
adapting the recommended prior widths to other contrast codes and predictor
scales is provided in @sec-scaling.

## Prior Predictive Analysis {#sec-prior-predictive}

We now turn to the simulation results. The goal is to characterise the
implied distributions of baseline probability, effect size, and
between-subject variability across 720 factorial combinations of link
function, prior family, and prior scale.

### Intercept Prior on the Probability Scale {#sec-pp-intercept}

The intercept $b_0$ determines the baseline success probability before any
experimental manipulation. A poorly chosen intercept prior can place
substantial mass near $p = 0$ or $p = 1$, predicting that participants
perform at floor or ceiling even before any experimental effect is applied.
How do different intercept priors perform? In the figures below, we overlay
exact analytical density curves (derived via the change-of-variables theorem)
as reference.[^density-formulas]

[^density-formulas]: For the logit link with a $\text{Logistic}(0, s)$ prior
on $b_0$: $f(p) = f_{\text{Logis}}(\text{logit}(p), 0, s) / [p(1-p)]$. At
$s = 1$, the logistic density and the Jacobian $p(1-p)$ cancel, giving
$f(p) = 1$. For the probit link with a $\text{Normal}(0, \sigma)$ prior:
$f(p) = \phi(\Phi^{-1}(p)/\sigma) / [\sigma \cdot \phi(\Phi^{-1}(p))]$. At
$\sigma = 1$, the normal densities cancel, again yielding $f(p) = 1$.

```{r}
#| label: fig-1-p-intercept
#| fig-cap: "Prior predictive distribution of success probability
#|   p = g^{-1}(b0) for each intercept prior SD (rows) and link function
#|   (columns). At sd_b0 = 1.0, the matched prior (Logistic for logit,
#|   Normal for probit) produces a perfectly uniform distribution - the
#|   visual signature of the probability integral transform. Reducing sd_b0
#|   to 0.75 pulls mass away from floor and ceiling. Misfit priors (e.g.,
#|   Cauchy under either link) concentrate more mass near 0 and 1. Dashed
#|   curves show exact analytical densities."
#| fig-height: 8
#| fig-width: 10

fig1_draws <- draws_long |>
  filter(dist_b1 == "logistic", sd_b1 == 0.25)

plot_p_intercept_ridges(fig1_draws, sd_b1_focus = 0.25)
```

@fig-1-p-intercept confirms the analytical result: at $\text{sd}_{b_0} = 1.0$
with the matched prior, the distribution of $p$ is uniform. At
$\text{sd}_{b_0} = 0.75$, the prior pulls mass away from 0 and 1, which is
appropriate when floor and ceiling effects are a concern. Misfit priors,
such as Cauchy for either link, produce heavier tails and concentrate more
mass near the boundaries.

```{r}
#| label: fig-2-floor-ceiling
#| fig-cap: "Proportion of prior predictive mass in the floor/ceiling region
#|   (p < 0.05 or p > 0.95) as a function of intercept prior SD and family.
#|   The matched prior at sd_b0 = 1.0 places exactly 10% in this region
#|   (2 x 0.05 = 0.10), confirming the uniform result. At sd_b0 = 0.75,
#|   floor/ceiling mass drops below 10% for both matched priors."
#| fig-height: 6
#| fig-width: 10

plot_floor_ceiling_heatmap(summaries_pp)
```

We recommend choosing the intercept prior SD so that
$P(p < 0.05 \text{ or } p > 0.95) \leq 0.10$. Under this criterion,
$\text{Logistic}(0, 0.75)$ is the recommended default for the logit link and
$\text{Normal}(0, 0.75)$ for the probit link (@fig-2-floor-ceiling). The
exact 10% value at $\text{sd}_{b_0} = 1.0$ for the matched priors serves as
an internal validation of the simulation against the analytical result.

### Effect Prior on the Probability Scale {#sec-pp-effect}

The prior on $b_1$ determines how much prior mass falls on small versus
large effects on the probability scale. @fig-3-delta-p shows how the choice
of $\text{sd}_{b_1}$ translates into implied effect sizes ($\delta p$),
holding the intercept prior fixed at the matched recommendation
($\text{Logistic}(0, 0.75)$ for logit; $\text{Normal}(0, 0.75)$ for probit).

```{r}
#| label: fig-3-delta-p
#| fig-cap: "Prior predictive distribution of delta_p (signed effect size on
#|   the probability scale) for each effect prior SD, colour-coded by prior
#|   family. At sd_b1 = 0.25, Normal and Logistic priors concentrate mass on
#|   effects up to about 20 percentage points - a range consistent with
#|   typical cognitive experiments. The Cauchy stands apart even at this
#|   moderate scale, assigning substantial probability to effects exceeding
#|   50 percentage points. Dotted lines mark +/-0.10, +/-0.20, +/-0.30."
#| fig-height: 10
#| fig-width: 10

# Use the link-matched recommended intercept prior for each link:
# Logistic(0,0.75) for logit, Normal(0,0.75) for probit
fig2_draws <- draws_long |>
  filter(
    (link == "logit"  & dist_b0 == "logistic") |
    (link == "probit" & dist_b0 == "normal"),
    sd_b0 == 0.75
  )

plot_delta_p_ridges(fig2_draws, sd_b0_focus = 0.75)
```

The Cauchy prior stands apart at every scale: even at
$\text{sd}_{b_1} = 0.25$, its heavy tails assign substantial probability to
$|\delta p| > 0.50$, making it a poor default for cognitive psychology
experiments. Note that effects of $|\delta p| = 1$ are only possible if
participants have perfect accuracy in one condition and not a single correct
response in the other - a scenario that is virtually impossible in
practice. Within the substantively appropriate range of
$\text{sd}_{b_1} \in [0.10, 0.50]$, Normal and Logistic priors behave
similarly - researchers can use either without meaningful loss.

We use the 90th percentile of $|\delta p|$ as the key summary statistic,
mirroring the 10% floor/ceiling criterion from the intercept analysis: if
less than 10% of the prior predictive mass should fall in implausible
regions, then the 90th percentile indicates the largest effect that the prior
treats as routine. At $\text{sd}_{b_1} = 0.25$, the 90th percentile of
$|\delta p|$ is near 0.20 - a moderate effect of 20 percentage points
that aligns well with typical effects in cognitive psychology. Values up to
$\text{sd}_{b_1} = 0.50$ remain defensible when larger effects are
plausible, with the 90th percentile rising to roughly 0.35.

For fixed-effects models, an effect prior with
$\text{sd}_{b_1} \in [0.25, 0.30]$ under sum-to-zero coding provides a
sensible default range. Most experiments,
however, also require random effects to account for between-subject
heterogeneity, which raises the question of how the hyperprior should be
specified.

### Random Effects Hyperprior {#sec-pp-re}

The `brms` default hyperprior on random-effect SDs ---
$\text{half-student\_t}(3, 0, 2.5)$ - places a heavy right tail on
between-subject variability. Does this matter on the probability scale? To
find out, we compare three candidate hyperpriors for the random-effect SDs
$\sigma_u$ (intercept) and $\sigma_v$ (slope). The alternatives were chosen
to concentrate mass in a range that is plausible for between-subject
variability in cognitive tasks: random-effect SDs rarely exceed 0.5 logit
units in typical accuracy experiments, so we selected distributions that
place most of their mass below this value:

- $\text{half-student\_t}(3, 0, 2.5)$: the `brms` default, with a heavy
  right tail on $\sigma_u$.
- $\text{Gamma}(2, 4)$: mean $= 0.5$, with approximately 71% of mass
  below $\sigma_u = 0.5$ logit units.
- $\text{Exponential}(4)$: mean $= 0.25$, with approximately 86% of mass
  below $\sigma_u = 0.5$; the strongest concentration near zero.

We run a full hierarchical prior predictive with
$\sigma_u \sim \text{hyperprior}$, subject-level random intercepts
$u_i \sim \mathcal{N}(0, \sigma_u)$, and
$p_{ij} = g^{-1}(b_0 + u_i \pm (b_1 + v_i))$, where $v_i$ is the
subject-level random slope. The fixed-effects priors are held at the
recommended values ($b_0 \sim \text{Logistic}(0, 0.75)$,
$b_1 \sim \mathcal{N}(0, 0.25)$). For the intercept + slope structure,
the random-slope SD $\sigma_v$ is drawn from the same hyperprior as
$\sigma_u$.

```{r}
#| label: setup-re-pp
#| include: false

re_pp_summaries <- tryCatch(
  readRDS(here("output", "prior_predictive_re_summaries.rds")),
  error = function(e) NULL
)
re_pp_draws <- tryCatch(
  readRDS(here("output", "prior_predictive_re_draws.rds")),
  error = function(e) NULL
)

re_pp_available <- !is.null(re_pp_summaries) && !is.null(re_pp_draws)
```

```{r}
#| label: fig-re-1-p-intercept
#| fig-cap: "The brms default student_t(3, 0, 2.5) spreads subject-level
#|   probabilities toward floor and ceiling due to its heavy right tail on
#|   between-subject SDs. The exponential(4) and gamma(2, 4) hyperpriors keep
#|   subjects' probabilities well away from the boundary. Dashed vertical
#|   lines mark p = 0.05 and p = 0.95."
#| fig-height: 5
#| fig-width: 9
#| eval: !expr re_pp_available

re_pp_draws |>
  group_by(sd_prior_re, re_structure) |>
  slice_sample(n = 5000) |>
  ungroup() |>
  ggplot(aes(x = p_intercept, y = sd_prior_re,
             fill = sd_prior_re, colour = sd_prior_re)) +
  geom_density_ridges(alpha = 0.5, scale = 0.9, rel_min_height = 0.01) +
  geom_vline(xintercept = c(0.05, 0.95), linetype = "dashed",
             colour = "grey40", linewidth = 0.5) +
  facet_wrap(~re_structure) +
  scale_fill_viridis_d(begin = 0.1, end = 0.85, guide = "none") +
  scale_colour_viridis_d(begin = 0.1, end = 0.85, guide = "none") +
  scale_x_continuous("Subject-level success probability", limits = c(0, 1),
                     expand = c(0, 0)) +
  scale_y_discrete("Hyperprior") +
  labs(title = "Subject-level intercept probabilities by RE hyperprior") +
  theme_minimal(base_size = 12) +
  theme(strip.text = element_text(face = "bold"))
```

@fig-re-1-p-intercept shows the consequences: the `brms` default
`student_t(3, 0, 2.5)` spreads subject-level probabilities toward 0 and 1,
with visible heavier tails near the floor/ceiling boundary. The `gamma(2, 4)`
and `exponential(4)` hyperpriors concentrate $\sigma_u$ near zero, keeping
subjects' probabilities away from the boundary.

```{r}
#| label: fig-re-2-delta-p
#| fig-cap: "Under the student_t default, a non-trivial fraction of simulated
#|   subjects show effects in the opposite direction to the population-level
#|   b1. The more concentrated hyperpriors (gamma, exponential) limit this
#|   effect heterogeneity to a substantively plausible range. Dotted lines
#|   mark +/-0.10, +/-0.20, +/-0.30."
#| fig-height: 5
#| fig-width: 9
#| eval: !expr re_pp_available

re_pp_draws |>
  group_by(sd_prior_re, re_structure) |>
  slice_sample(n = 10000) |>
  ungroup() |>
  ggplot(aes(x = delta_p, y = sd_prior_re,
             fill = sd_prior_re, colour = sd_prior_re)) +
  geom_density_ridges(alpha = 0.5, scale = 0.9, rel_min_height = 0.01) +
  geom_vline(xintercept = 0, linewidth = 0.5, colour = "grey20") +
  geom_vline(xintercept = c(-0.30, -0.20, -0.10, 0.10, 0.20, 0.30),
             linetype = "dotted", linewidth = 0.5, colour = "grey40") +
  facet_wrap(~re_structure) +
  scale_fill_viridis_d(begin = 0.1, end = 0.85, guide = "none") +
  scale_colour_viridis_d(begin = 0.1, end = 0.85, guide = "none") +
  scale_x_continuous("Effect size on probability scale (delta_p)",
                     limits = c(-1, 1), breaks = seq(-1, 1, 0.25)) +
  scale_y_discrete("Hyperprior") +
  labs(title = "Effect size distribution by RE hyperprior") +
  theme_minimal(base_size = 12) +
  theme(strip.text = element_text(face = "bold"))
```

The `student_t` hyperprior also produces more within-study effect
heterogeneity (@fig-re-2-delta-p): a non-trivial fraction of simulated
subjects show effects in the opposite direction to the population-level
$b_1$. The wide distribution of $\sigma_u$ pulls individual subjects into
regions where the logistic nonlinearity reverses the apparent effect
direction.

We therefore recommend `exponential(4)` for random-effect SD hyperpriors.
It concentrates between-subject variability in a plausible range and avoids
excessive floor/ceiling effects.

## Bayes Factor Calibration {#sec-bf-calibration}

### Simulation Design {#sec-bf-design}

We simulated data across a range of realistic experimental scenarios. The true
value of the effect coefficient $b_1$ ranged from $b_1 = 0$ (no effect) to
$b_1 = 0.50$ on the logit scale, covering the span from a null effect through
a large effect in cognitive experiments (on the probability scale with a
typical 50% baseline, $b_1 = 0.50$ corresponds to roughly 24 percentage
points between conditions). Sample sizes ranged from 30 to 100 subjects, and
each subject contributed either 20 or 50 trials. We crossed these
data-generating factors with three effect prior families (Normal, Logistic,
Cauchy) at two scale values ($\text{sd}_{b_1} = 0.25$ and $0.50$), and two
intercept priors (matched Logistic and misfit Normal, both at
$\text{sd}_{b_0} = 0.75$). This produced 360 conditions with 100
replications each.

A second simulation specifically investigated the random-effects hyperprior.
It used a within-subjects design with correlated random intercepts and
slopes, holding the fixed-effects priors at the recommended matched
configuration and varying only the hyperprior on the random-effects SDs
across three candidates: the brms default Student-*t*(3, 0, 2.5), Gamma(2,
4), and Exponential(4). This produced 90 conditions with 100 replications
each. We return to these results in @sec-re-hyperprior.

#### How We Computed Bayes Factors

All models used the logit link (the probit link is expected to produce
analogous results via the matched-prior principle). To ensure that results
generalize across realistic baseline accuracies, we did not fix the true
intercept: in each replication, $b_0$ was sampled from
$\text{Uniform}(0.4, 0.9)$ on the probability scale and transformed to the
logit scale. Each simulated dataset included subject-level random intercepts
($u_i \sim \mathcal{N}(0, 0.25)$) to mimic between-subject heterogeneity.

The analysis model mirrored this structure: a binomial GLMM with one random
intercept per subject, fitted with brms [@burkner2017], with the intercept
prior SD fixed at 0.75 and an $\text{Exponential}(4)$ hyperprior on the
random-intercept SD. Bayes Factors were computed via the Savage--Dickey
density ratio [@wagenmakers2018; @ly2016]:
$\text{BF}_{10} = \pi(b_1 = 0) / p(b_1 = 0 \mid \text{data})$. As a
robustness check, we also computed BFs via bridge sampling on a subset of
conditions (see @sec-bf-validation).

To quantify uncertainty in detection rates given the finite number of
replications, we applied Bayesian conjugate updating (see
@sec-appendix-uncertainty for details). All error bars in the figures below
show 95% credible intervals from this procedure.

### Strength of Evidence by Design Parameters {#sec-bf-strength}

Before applying discrete thresholds, we first examine the continuous
distribution of evidence strength across conditions.

```{r}
#| label: fig-bf-continuous
#| fig-cap: "Median log10(BF10) with 10th and 90th percentile bands as a
#|   function of sample size (N), faceted by true effect size and trials per
#|   subject. Positive values indicate evidence for H1; negative values
#|   indicate evidence for H0. Horizontal dashed lines mark BF10 = 3 (moderate
#|   evidence for H1) and BF10 = 1/3 (moderate evidence for H0)."
#| fig-height: 8
#| fig-width: 10
#| eval: !expr continuous_available

# Focus on matched Logistic intercept, Logistic effect prior, sd_b1 = 0.25
plot_data <- continuous_bf |>
  filter(dist_b0 == "logistic", dist_b1 == "logistic", sd_b1 == 0.25)

ggplot(plot_data, aes(x = factor(n_subjects), y = median_log10_BF10)) +
  geom_hline(yintercept = c(log10(3), log10(1/3)), linetype = "dashed",
             colour = "grey50", linewidth = 0.4) +
  geom_hline(yintercept = 0, colour = "grey30", linewidth = 0.3) +
  geom_pointrange(aes(ymin = q10_log10_BF10, ymax = q90_log10_BF10),
                  size = 0.4, linewidth = 0.6) +
  facet_grid(n_trials ~ true_b1,
             labeller = labeller(
               n_trials = function(x) paste0(x, " trials"),
               true_b1  = function(x) paste0("b1 = ", x)
             )) +
  labs(
    x = "Number of subjects (N)",
    y = expression("Median" ~ log[10](BF[10]) ~ "(10th–90th percentile)")
  ) +
  base_theme()
```

@fig-bf-continuous shows how the strength of evidence scales with sample
size, true effect magnitude, and trials per subject. Under the null
($b_1 = 0$), the median BF falls below $1/3$ even at modest sample sizes,
correctly supporting H0. As the true effect grows, the median BF rises above
3 and beyond, with the rate of increase depending on the amount of data. The
10th--90th percentile bands reveal the variability across replications: for
small effects, considerable uncertainty remains even at the largest sample
sizes, while large effects produce consistently strong evidence.

### Threshold-Based Analysis {#sec-bf-threshold}

For researchers who need discrete decision thresholds - for instance, in
preregistered studies with stopping rules or in contexts where a binary
accept/reject decision is required - we examine power ($P(\text{BF}_{10} > 3)$)
and specificity ($P(\text{BF}_{01} > 3)$) across the full design space.

#### Power by sample size and effect size

```{r}
#| label: fig-power-n
#| fig-cap: "Power to detect effects increases with sample size and effect
#|   magnitude. With 50 trials per subject and the recommended effect prior
#|   (sd = 0.25), power exceeds 80% at N = 30 for effects of b1 >= 0.20. The
#|   tighter prior (sd = 0.25, solid lines) yields slightly higher power than
#|   the wider prior (sd = 0.50, dashed), but the difference is small compared
#|   to the effect of sample size and true effect. Prior family (color) has
#|   minimal impact on power. Error bars: 95% credible intervals."
#| fig-height: 6
#| fig-width: 10

if (posteriors_available) {
  matched <- posteriors_bf |>
    filter(
      dist_b0 == "logistic",
      true_b1 > 0,
      threshold == 3
    ) |>
    mutate(
      dist_b1  = factor(dist_b1, levels = c("cauchy", "normal", "logistic"),
                        labels = c("Cauchy", "Normal", "Logistic")),
      sd_b1    = factor(paste0("sd = ", sd_b1))
    )

  ggplot(matched, aes(x = n_subjects, y = power_post_mean,
                      color = dist_b1, linetype = sd_b1)) +
    geom_line(linewidth = 0.8) +
    geom_point(size = 2) +
    geom_errorbar(aes(ymin = power_ci_lower, ymax = power_ci_upper),
                  width = 3, linewidth = 0.4, alpha = 0.6) +
    geom_hline(yintercept = 0.80, linetype = "dotted", color = "grey40") +
    scale_x_continuous("Number of subjects", breaks = c(30, 60, 100)) +
    scale_y_continuous("P(BF10 > 3)", limits = c(NA, 1)) +
    scale_color_viridis_d("Prior family", option = "D", end = 0.85) +
    scale_linetype_manual("Prior SD",
      values = c("sd = 0.25" = "solid", "sd = 0.5" = "dashed")) +
    facet_grid(n_trials ~ true_b1,
               labeller = labeller(
                 true_b1  = function(x) paste0("b1 = ", x),
                 n_trials = function(x) paste0(x, " trials")
               )) +
    labs(title = "Power: P(BF10 > 3) by sample size and true effect",
         subtitle = "Matched intercept prior (Logistic, sd = 0.75); dotted line = 80% power") +
    base_theme()
} else {
  matched <- summaries_bf |>
    filter(dist_b0 == "logistic", true_b1 > 0) |>
    mutate(
      dist_b1  = factor(dist_b1, levels = c("cauchy", "normal", "logistic"),
                        labels = c("Cauchy", "Normal", "Logistic")),
      sd_b1    = factor(paste0("sd = ", sd_b1))
    )

  ggplot(matched, aes(x = n_subjects, y = P_BF10_gt3,
                      color = dist_b1, linetype = sd_b1)) +
    geom_line(linewidth = 0.8) +
    geom_point(size = 2) +
    geom_hline(yintercept = 0.80, linetype = "dotted", color = "grey40") +
    scale_x_continuous("Number of subjects", breaks = c(30, 60, 100)) +
    scale_y_continuous("P(BF10 > 3)", limits = c(NA, 1)) +
    scale_color_viridis_d("Prior family", option = "D", end = 0.85) +
    scale_linetype_manual("Prior SD",
      values = c("sd = 0.25" = "solid", "sd = 0.5" = "dashed")) +
    facet_grid(n_trials ~ true_b1,
               labeller = labeller(
                 true_b1  = function(x) paste0("b1 = ", x),
                 n_trials = function(x) paste0(x, " trials")
               )) +
    labs(title = "Power: P(BF10 > 3) by sample size and true effect",
         subtitle = "Matched intercept prior (Logistic, sd = 0.75); dotted line = 80% power") +
    base_theme()
}
```

For larger effects ($b_1 \geq 0.20$), @fig-power-n shows that power reaches
or exceeds 80% even at the smallest sample size ($n = 30$) when subjects
complete 50 trials. The tighter prior ($\text{sd} = 0.25$, solid lines)
yields slightly higher power than the wider alternative ($\text{sd} = 0.50$,
dashed), because concentrating the prior near plausible effect sizes makes
the Savage--Dickey ratio more sensitive to genuine shifts away from zero. But
this advantage is modest; the true effect magnitude and the amount of data
drive power far more than the prior specification does.

#### Specificity under the null

```{r}
#| label: fig-specificity
#| fig-cap: "When no effect exists, the Bayes Factor correctly supports the
#|   null in the large majority of replications. Wider priors (sd = 0.50,
#|   dashed) yield substantially higher specificity because they place less
#|   density at exactly zero, making the Savage-Dickey ratio more conservative.
#|   At sd = 0.25 (solid lines), the Normal prior shows notably lower
#|   specificity than the Logistic and Cauchy, especially at small sample
#|   sizes — a consequence of its higher density at exactly zero (see text).
#|   Error bars: 95% credible intervals."
#| fig-height: 5
#| fig-width: 10

if (posteriors_available) {
  posteriors_bf |>
    filter(
      true_b1 == 0,
      dist_b0 == "logistic",
      threshold == 3
    ) |>
    mutate(
      dist_b1 = factor(dist_b1, levels = c("cauchy", "normal", "logistic"),
                       labels = c("Cauchy", "Normal", "Logistic")),
      sd_b1   = factor(paste0("sd = ", sd_b1))
    ) |>
    ggplot(aes(x = n_subjects, y = spec_post_mean,
               color = dist_b1, linetype = sd_b1)) +
    geom_line(linewidth = 0.8) +
    geom_point(size = 2) +
    geom_errorbar(aes(ymin = spec_ci_lower, ymax = spec_ci_upper),
                  width = 3, linewidth = 0.4, alpha = 0.6) +
    scale_x_continuous("Number of subjects", breaks = c(30, 60, 100)) +
    scale_y_continuous("P(BF01 > 3)", limits = c(0, 1)) +
    scale_color_viridis_d("Prior family", option = "D", end = 0.85) +
    scale_linetype_manual("Prior SD",
      values = c("sd = 0.25" = "solid", "sd = 0.5" = "dashed")) +
    facet_wrap(~ n_trials,
               labeller = labeller(
                 n_trials = function(x) paste0(x, " trials")
               )) +
    labs(title = "Specificity under H0: P(BF01 > 3)",
         subtitle = "Matched intercept prior (Logistic, sd = 0.75)") +
    base_theme()
} else {
  summaries_bf |>
    filter(true_b1 == 0, dist_b0 == "logistic") |>
    mutate(
      dist_b1 = factor(dist_b1, levels = c("cauchy", "normal", "logistic"),
                       labels = c("Cauchy", "Normal", "Logistic")),
      sd_b1   = factor(paste0("sd = ", sd_b1))
    ) |>
    ggplot(aes(x = n_subjects, y = P_BF01_gt3,
               color = dist_b1, linetype = sd_b1)) +
    geom_line(linewidth = 0.8) +
    geom_point(size = 2) +
    scale_x_continuous("Number of subjects", breaks = c(30, 60, 100)) +
    scale_y_continuous("P(BF01 > 3)", limits = c(0, 1)) +
    scale_color_viridis_d("Prior family", option = "D", end = 0.85) +
    scale_linetype_manual("Prior SD",
      values = c("sd = 0.25" = "solid", "sd = 0.5" = "dashed")) +
    facet_wrap(~ n_trials,
               labeller = labeller(
                 n_trials = function(x) paste0(x, " trials")
               )) +
    labs(title = "Specificity under H0: P(BF01 > 3)",
         subtitle = "Matched intercept prior (Logistic, sd = 0.75)") +
    base_theme()
}
```

At $\text{sd}_{b_1} = 0.25$, the Normal prior yields substantially lower
specificity than the Logistic and Cauchy, particularly at small sample sizes.
The reason is straightforward: the three families differ in their density at
exactly zero when the scale parameter is equated ($\text{Normal}(0, 0.25)$:
1.60; $\text{Cauchy}(0, 0.25)$: 1.27; $\text{Logistic}(0, 0.25)$: 1.00). A
higher prior density at zero makes the numerator of the Savage--Dickey ratio
larger, so the posterior must pile up more at zero before $\text{BF}_{01}$
exceeds the threshold, reducing specificity. False positive rates
($\text{BF}_{10} > 3$ under H0) remain below 5% across all conditions.

#### Three-category decomposition

```{r}
#| label: fig-three-category
#| fig-cap: "When effects are large or absent, the Bayes Factor is decisive in
#|   the majority of replications. The inconclusive fraction (middle category)
#|   is largest at small effects with few trials — the conditions where sample
#|   size recommendations matter most. Under the null, most replications
#|   correctly support H0, with a small fraction remaining undecided.
#|   Bars show Dirichlet posterior means for the recommended Normal effect
#|   prior (sd = 0.25) with matched Logistic intercept prior."
#| fig-height: 7
#| fig-width: 10
#| eval: !expr posteriors_available

posteriors_bf |>
  filter(
    dist_b0 == "logistic",
    dist_b1 == "normal",
    sd_b1   == 0.25,
    threshold == 3
  ) |>
  select(true_b1, n_subjects, n_trials, h1_mean, inc_mean, h0_mean) |>
  pivot_longer(cols = c(h1_mean, inc_mean, h0_mean),
               names_to = "outcome", values_to = "proportion") |>
  mutate(
    outcome = factor(outcome,
      levels = c("h1_mean", "inc_mean", "h0_mean"),
      labels = c("H1 detected (BF10 > 3)",
                  "Inconclusive",
                  "H0 detected (BF01 > 3)")),
    true_b1_label = factor(paste0("b1 = ", true_b1))
  ) |>
  ggplot(aes(x = factor(n_subjects), y = proportion, fill = outcome)) +
  geom_col(position = "stack", width = 0.7) +
  facet_grid(n_trials ~ true_b1_label,
             labeller = labeller(
               n_trials = function(x) paste0(x, " trials")
             )) +
  scale_x_discrete("Number of subjects") +
  scale_y_continuous("Proportion of replications", limits = c(0, 1)) +
  scale_fill_viridis_d("BF outcome", option = "D", end = 0.85) +
  labs(title = "Three-category BF outcome decomposition (BF threshold = 3)",
       subtitle = "Normal effect prior (sd = 0.25), matched Logistic intercept prior") +
  base_theme()
```

Under the null ($b_1 = 0$), the large majority of replications correctly
favor H0 (@fig-three-category). The inconclusive fraction is largest at
small effects ($b_1 = 0.05$ and $0.10$) with 20 trials, where a substantial
share of replications remain undecided. At 50 trials, this fraction shrinks
considerably even for small effects.

### Effect of Prior Family and Intercept Prior {#sec-bf-family}

At the same prior scale, the three effect prior families produce nearly
identical power curves across all sample sizes and effect sizes. Specificity tells a different story: the
Logistic prior has the lowest density at zero among the three families, which
gives it the best specificity profile. The Normal prior's higher density at
zero reduces its specificity at $\text{sd}_{b_1} = 0.25$, especially at
small $n$. The Cauchy produces acceptable power but assigns implausibly
heavy-tailed prior predictive distributions with no compensating advantage.
We therefore recommend the Logistic effect prior at
$\text{sd}_{b_1} \in [0.25, 0.30]$ as the default.

The matched-prior principle does not meaningfully affect Bayes Factor
calibration. @fig-misfit-bf confirms that matched (Logistic) and misfit
(Normal) intercept priors produce overlapping power curves.

```{r}
#| label: fig-misfit-bf
#| fig-cap: "Using a mismatched intercept prior (Normal instead of the matched
#|   Logistic) has negligible impact on power. The two intercept priors produce
#|   overlapping power curves across all effect sizes, sample sizes, and trial
#|   counts. Error bars: 95% credible intervals."
#| fig-height: 6
#| fig-width: 10

if (posteriors_available) {
  posteriors_bf |>
    filter(
      dist_b1 == "logistic",
      sd_b1   == 0.25,
      true_b1 > 0,
      threshold == 3
    ) |>
    mutate(
      dist_b0 = factor(dist_b0, levels = c("logistic", "normal"),
                       labels = c("Logistic (matched)", "Normal (misfit)"))
    ) |>
    ggplot(aes(x = true_b1, y = power_post_mean, color = dist_b0)) +
    geom_line(linewidth = 0.8) +
    geom_point(size = 2) +
    geom_errorbar(aes(ymin = power_ci_lower, ymax = power_ci_upper),
                  width = 0.02, linewidth = 0.4, alpha = 0.6) +
    geom_hline(yintercept = 0.80, linetype = "dotted", color = "grey40") +
    scale_x_continuous("True effect b1") +
    scale_y_continuous("P(BF10 > 3)", limits = c(NA, 1)) +
    scale_color_viridis_d("Intercept prior", option = "D", end = 0.7) +
    facet_grid(n_trials ~ n_subjects,
               labeller = labeller(
                 n_subjects = function(x) paste0(x, " subjects"),
                 n_trials   = function(x) paste0(x, " trials")
               )) +
    labs(title = "Effect of intercept prior misfit on power (logit link, sd_b1 = 0.25)") +
    base_theme()
} else {
  summaries_bf |>
    filter(dist_b1 == "logistic", sd_b1 == 0.25, true_b1 > 0) |>
    mutate(
      dist_b0 = factor(dist_b0, levels = c("logistic", "normal"),
                       labels = c("Logistic (matched)", "Normal (misfit)"))
    ) |>
    ggplot(aes(x = true_b1, y = P_BF10_gt3, color = dist_b0)) +
    geom_line(linewidth = 0.8) +
    geom_point(size = 2) +
    geom_hline(yintercept = 0.80, linetype = "dotted", color = "grey40") +
    scale_x_continuous("True effect b1") +
    scale_y_continuous("P(BF10 > 3)", limits = c(NA, 1)) +
    scale_color_viridis_d("Intercept prior", option = "D", end = 0.7) +
    facet_grid(n_trials ~ n_subjects,
               labeller = labeller(
                 n_subjects = function(x) paste0(x, " subjects"),
                 n_trials   = function(x) paste0(x, " trials")
               )) +
    labs(title = "Effect of intercept prior misfit on power (logit link, sd_b1 = 0.25)") +
    base_theme()
}
```

Prior scale matters most. Family and intercept prior are secondary. We
recommend the matched intercept prior for interpretability, but it does not
affect BF calibration.

### Random-Effects Hyperprior {#sec-re-hyperprior}

Although the prior predictive analysis showed that the choice of RE
hyperprior affects the distribution of subject-level probabilities, this
does not translate into meaningful differences in BF calibration. All three
hyperpriors - Student-*t*(3, 0, 2.5), Gamma(2, 4), and Exponential(4) ---
produce nearly identical power and specificity with widely overlapping
credible intervals (@fig-re-power). The recommendation for Exponential(4)
therefore rests on its prior predictive properties rather than on BF
calibration differences.

```{r}
#| label: fig-re-power
#| fig-cap: "The choice of random-effects hyperprior has no meaningful impact
#|   on power. All three hyperpriors produce nearly identical power curves
#|   across all effect sizes and sample sizes. Within-subjects design with
#|   correlated RE; matched fixed-effects priors. Error bars: 95% credible
#|   intervals."
#| fig-height: 6
#| fig-width: 10
#| eval: !expr re_posteriors_available || re_results_available

if (re_posteriors_available) {
  re_post_data <- re_posteriors |>
    mutate(
      sd_prior_re = factor(sd_prior_re,
                           levels = c("default", "gamma", "exponential"),
                           labels = c("Student-t(3,0,2.5)", "Gamma(2,4)",
                                      "Exponential(4)")),
      n_trials_label = paste0(n_trials, " trials")
    )

  re_post_data |>
    filter(true_b1 > 0, threshold == 3) |>
    ggplot(aes(x = n_subjects, y = power_post_mean, color = sd_prior_re,
               linetype = sd_prior_re)) +
    geom_line(linewidth = 0.8) +
    geom_point(size = 2) +
    geom_errorbar(aes(ymin = power_ci_lower, ymax = power_ci_upper),
                  width = 3, linewidth = 0.4, alpha = 0.6) +
    geom_hline(yintercept = 0.80, linetype = "dotted", color = "grey40") +
    scale_x_continuous("Number of subjects", breaks = c(30, 60, 100)) +
    scale_y_continuous("P(BF10 > 3)", limits = c(NA, 1)) +
    scale_color_viridis_d("RE hyperprior", option = "D", end = 0.85) +
    scale_linetype_manual("RE hyperprior",
      values = c("Student-t(3,0,2.5)" = "solid",
                 "Gamma(2,4)"         = "dashed",
                 "Exponential(4)"     = "dotdash")) +
    facet_grid(n_trials_label ~ true_b1,
               labeller = labeller(
                 true_b1 = function(x) paste0("b1 = ", x)
               )) +
    labs(title = "Power by RE hyperprior (within-subjects, matched fixed-effects priors)",
         subtitle = "Dotted line = 80% power criterion") +
    base_theme()
} else {
  re_plot_data <- re_summaries |>
    mutate(
      sd_prior_re = factor(sd_prior_re,
                           levels = c("default", "gamma", "exponential"),
                           labels = c("Student-t(3,0,2.5)", "Gamma(2,4)",
                                      "Exponential(4)")),
      n_trials_label = paste0(n_trials, " trials")
    )

  re_plot_data |>
    filter(true_b1 > 0) |>
    ggplot(aes(x = n_subjects, y = P_BF10_gt3, color = sd_prior_re,
               linetype = sd_prior_re)) +
    geom_line(linewidth = 0.8) +
    geom_point(size = 2) +
    geom_hline(yintercept = 0.80, linetype = "dotted", color = "grey40") +
    scale_x_continuous("Number of subjects", breaks = c(30, 60, 100)) +
    scale_y_continuous("P(BF10 > 3)", limits = c(NA, 1)) +
    scale_color_viridis_d("RE hyperprior", option = "D", end = 0.85) +
    scale_linetype_manual("RE hyperprior",
      values = c("Student-t(3,0,2.5)" = "solid",
                 "Gamma(2,4)"         = "dashed",
                 "Exponential(4)"     = "dotdash")) +
    facet_grid(n_trials_label ~ true_b1,
               labeller = labeller(
                 true_b1 = function(x) paste0("b1 = ", x)
               )) +
    labs(title = "Power by RE hyperprior (within-subjects, matched fixed-effects priors)",
         subtitle = "Dotted line = 80% power criterion") +
    base_theme()
}
```

### Savage-Dickey vs. Bridge Sampling Validation {#sec-bf-validation}

<!-- Placeholder: populate after running Script 05. Expected content:
     correlation between SD and BS BFs, detection rate agreement,
     any discrepancies. -->

```{r}
#| label: validation-check
#| eval: false

# Placeholder — populate after running Script 05
# validation <- readRDS(here("output", "res_bf_validation.rds"))
```

## Recommendations {#sec-recommendations}

For binomial GLMs, we recommend the following framework. Researchers with
no domain-specific prior knowledge can use the defaults directly. Those with
prior information about expected effect sizes can use the scaling formulas to
adapt the defaults to their context.

### Default Prior Specifications

#### Intercept prior

| Link | Recommended prior | Rationale |
|------|------------------|-----------|
| logit | $\text{Logistic}(0, 0.75)$ | Matched family; sd = 0.75 gives < 10% floor/ceiling mass |
| probit | $\text{Normal}(0, 0.75)$ | Matched family; same criterion |

At $\text{sd} = 0.75$, less than 10% of prior mass falls in the
floor/ceiling region, while remaining agnostic about baseline accuracy
within the plausible range.

#### Effect prior

| Prior family | Recommended SD | 90th percentile of $|\delta p|$ |
|-------------|---------------|-------------------------------|
| Logistic (default) | 0.25--0.30 | ~0.20--0.23 |
| Normal | 0.25--0.30 | ~0.18--0.22 |

Our default recommendation is $\text{Logistic}(0, 0.25\text{--}0.30)$ for
logit models and $\text{Normal}(0, 0.25\text{--}0.30)$ for probit models.
Within this range, the 90th percentile of $|\delta p|$ stays between
approximately 0.18 and 0.23 - well-calibrated for typical cognitive
experiments. The Logistic prior offers the best specificity profile due to
its lowest density at zero among the candidate families.

#### Random effects prior

- Use `exponential(4)` for both $\sigma_u$ and $\sigma_v$. This
concentrates between-subject SD below ~0.5 logit units, keeping
floor/ceiling mass under 10%. The `brms` default `student_t(3, 0, 2.5)`
places too much mass on large random-effect SDs for typical cognitive
experiments and is not recommended for binomial models.

#### Ready-to-use brms prior specifications

```{r}
#| label: recommended-priors
#| echo: true
#| eval: false

# Logit link (logistic regression)
logit_priors <- c(
  brms::prior(logistic(0, 0.75), class = Intercept),
  brms::prior(logistic(0, 0.25), class = b)
)

# Probit link (probit regression)
probit_priors <- c(
  brms::prior(normal(0, 0.75), class = Intercept),
  brms::prior(normal(0, 0.25), class = b)
)
```

For models with subject-level random effects:

```{r}
#| label: recommended-priors-re
#| echo: true
#| eval: false

# Logit link with random intercept
logit_priors_re <- c(
  brms::prior(logistic(0, 0.75), class = Intercept),
  brms::prior(logistic(0, 0.25), class = b),
  brms::prior(exponential(4),    class = sd)
)

# Probit link with random intercept
probit_priors_re <- c(
  brms::prior(normal(0, 0.75), class = Intercept),
  brms::prior(normal(0, 0.25), class = b),
  brms::prior(exponential(4),  class = sd)
)
```

### From Defaults to Informed Priors {#sec-scaling}

The recommended $\text{sd}_{b_1} \in [0.25, 0.30]$ assumes sum-to-zero
$\pm 1$ coding ($\text{SD}(x) = 1$). Whenever the predictor's SD differs
from 1, the prior on $b_1$ must be rescaled. Using $\text{sd}_{b_1} = 0.25$
as the baseline for the scaling formula:

$$\text{sd}_{b_1}(\text{predictor SD} = s) = \frac{\text{sd}_{b_1}(\text{SD} = 1)}{s}$$

Smaller predictor SDs require a *wider* prior on $b_1$; larger predictor SDs
require a *narrower* prior. Examples:

- $\pm 0.5$ coding ($s = 0.5$): use $\text{sd}_{b_1} = 0.25 / 0.5 = 0.50$
- $\pm 1/\sqrt{2}$ orthonormal ($s \approx 0.707$): use $\text{sd}_{b_1} = 0.25 / 0.707 \approx 0.354$
- $0/1$ treatment coding ($s = 0.5$ in a balanced design): use $\text{sd}_{b_1} = 0.25 / 0.5 = 0.50$
- Standardised continuous predictor ($s = 1$): use $\text{sd}_{b_1} = 0.25$ directly

#### Generalising to multi-level and factorial designs

When a factor has more than two levels, the choice of contrast coding affects
not only the interpretation of coefficients but also the *prior* on each
factor-level mean. Orthonormal contrasts avoid asymmetric priors across
factor levels. The `contr.equalprior()` function from the `bayestestR`
package [@makowski2019] provides such contrasts, following the construction
in @rouder2012 [p. 363].

In a balanced design with $J$ levels, each orthonormal contrast column has
predictor $\text{SD} = 1 / \sqrt{J}$. Applying the scaling formula gives:

$$\text{sd}_{b_1}(\text{orthonormal, } J \text{ levels}) = 0.25 \times \sqrt{J}$$

| Design term | Predictor SD | Rescaled $\text{sd}_{b_1}$ |
|:---|:---:|:---:|
| 2-level factor | $1/\sqrt{2} \approx 0.71$ | 0.35 |
| 3-level factor | $1/\sqrt{3} \approx 0.58$ | 0.43 |
| 4-level factor | $1/\sqrt{4} = 0.50$ | 0.50 |
| 5-level factor | $1/\sqrt{5} \approx 0.45$ | 0.56 |
| $2 \times 3$ interaction | $1/\sqrt{6} \approx 0.41$ | 0.61 |
| $3 \times 3$ interaction | $1/\sqrt{9} \approx 0.33$ | 0.75 |

: Rescaled effect prior widths under orthonormal contrasts for multi-level
  factors and factorial interactions. All values assume the baseline
  recommendation of $\text{sd}_{b_1} = 0.25$ under $\pm 1$ coding.
  {#tbl-orthonormal-scaling}

@fig-orthonormal-demo shows the resulting prior predictive distributions
for $J = 2$, $3$, and $4$.

```{r}
#| label: fig-orthonormal-demo
#| fig-cap: "Prior predictive distribution of pairwise effect sizes under
#|   orthonormal contrasts with the rescaled prior
#|   ($\\text{sd}_{b_1} = 0.25 \\times \\sqrt{J}$). Left: distribution of
#|   a single representative pairwise $\\delta p$, which is nearly identical
#|   across $J$ - confirming that the scaling rule preserves per-comparison
#|   calibration. Right: distribution of the maximum pairwise
#|   $|\\delta p|$, which shifts right with $J$ as expected
#|   (more pairwise comparisons produce a larger maximum)."
#| fig-height: 4.5
#| fig-width: 10

set.seed(2024)
n_draws <- 50000

demo_list <- lapply(c(2, 3, 4), function(J) {
  oc <- orthonormal_contrasts(J)
  sd_b1_scaled <- 0.25 * oc$scaling_factor

  b0 <- rlogis(n_draws, location = 0, scale = 0.75)
  B <- matrix(
    rnorm(n_draws * (J - 1), mean = 0, sd = sd_b1_scaled),
    nrow = n_draws, ncol = J - 1
  )

  eta_mat <- outer(b0, rep(0, J)) + B %*% t(oc$Q)
  p_mat   <- ilogit(eta_mat)

  delta_p_pair <- p_mat[, 1] - p_mat[, 2]

  max_abs_dp <- apply(p_mat, 1, function(row) {
    diffs <- outer(row, row, "-")
    max(abs(diffs[upper.tri(diffs)]))
  })

  data.frame(
    J            = paste0("J = ", J),
    delta_p_pair = delta_p_pair,
    max_abs_dp   = max_abs_dp
  )
})

demo_draws <- do.call(rbind, demo_list)

p_single <- ggplot(demo_draws, aes(x = delta_p_pair,
                                    fill = J, colour = J)) +
  geom_density(alpha = 0.3, linewidth = 0.6) +
  geom_vline(xintercept = c(-0.20, -0.10, 0.10, 0.20),
             linetype = "dotted", linewidth = 0.4,
             colour = "grey40") +
  scale_x_continuous(expression(delta * p ~ "(single pair)"),
                     limits = c(-0.8, 0.8)) +
  scale_fill_viridis_d("Factor levels", option = "D", end = 0.85) +
  scale_colour_viridis_d("Factor levels", option = "D", end = 0.85) +
  ggtitle("Single pairwise comparison") +
  prior_theme()

p_max <- ggplot(demo_draws, aes(x = max_abs_dp,
                                 fill = J, colour = J)) +
  geom_density(alpha = 0.3, linewidth = 0.6) +
  geom_vline(xintercept = c(0.10, 0.20, 0.30),
             linetype = "dotted", linewidth = 0.4,
             colour = "grey40") +
  scale_x_continuous(expression(max ~ "|" * delta * p * "|"),
                     limits = c(0, 1)) +
  scale_fill_viridis_d("Factor levels", option = "D", end = 0.85) +
  scale_colour_viridis_d("Factor levels", option = "D", end = 0.85) +
  ggtitle("Maximum pairwise difference") +
  prior_theme()

patchwork::wrap_plots(p_single, p_max, guides = "collect") &
  theme(legend.position = "bottom")
```

The left panel of @fig-orthonormal-demo shows that the prior predictive
distribution of any single pairwise comparison is nearly identical across
$J = 2$, $3$, and $4$: the $\sqrt{J}$ scaling rule preserves the
calibration established for the two-condition case.

In `brms`, the recommended workflow is to set orthonormal contrasts before
fitting the model and adjust the prior width accordingly:

```{r}
#| label: recommended-priors-orthonormal
#| echo: true
#| eval: false

# - Single 3-level factor ---
contrasts(dat$condition) <- bayestestR::contr.equalprior(3)
# Predictor SD = 1/sqrt(3); rescaled prior: sd_b1 = 0.25 * sqrt(3) = 0.433

priors_3level <- c(
  brms::prior(logistic(0, 0.75),  class = Intercept),
  brms::prior(normal(0, 0.433),   class = b)
)

# - 2 x 3 factorial design ---
contrasts(dat$A) <- bayestestR::contr.equalprior(2)
contrasts(dat$B) <- bayestestR::contr.equalprior(3)

priors_factorial <- c(
  brms::prior(logistic(0, 0.75),  class = Intercept),
  brms::prior(normal(0, 0.354),   class = b, coef = A1),
  brms::prior(normal(0, 0.433),   class = b, coef = B1),
  brms::prior(normal(0, 0.433),   class = b, coef = B2),
  brms::prior(normal(0, 0.612),   class = b, coef = "A1:B1"),
  brms::prior(normal(0, 0.612),   class = b, coef = "A1:B2")
)
```

### Sample Size Guidance {#sec-sample-size}

For practical study planning, the following table summarizes the minimum
sample sizes needed to achieve 80% power across different prior
specifications and effect sizes, assuming 50 trials per subject and the
recommended matched intercept prior.

```{r}
#| label: tbl-summary
#| tbl-cap: "Minimum number of subjects for 80% power (P(BF10 > 3) >= 0.80)
#|   at 50 trials per subject, by prior family, prior SD, and true effect size.
#|   Between-subjects design with matched Logistic intercept prior."

summaries_bf |>
  filter(
    n_trials == 50,
    dist_b0 == "logistic",
    true_b1 > 0
  ) |>
  group_by(dist_b1, sd_b1, true_b1) |>
  summarise(
    min_n_80pct = {
      n_meeting <- n_subjects[P_BF10_gt3 >= 0.80]
      if (length(n_meeting) > 0) min(n_meeting) else NA_integer_
    },
    .groups = "drop"
  ) |>
  mutate(
    dist_b1 = factor(dist_b1, levels = c("cauchy", "normal", "logistic"),
                     labels = c("Cauchy", "Normal", "Logistic"))
  ) |>
  pivot_wider(names_from = true_b1, values_from = min_n_80pct,
              names_prefix = "b1 = ") |>
  knitr::kable()
```

```{r}
#| label: tbl-summary-ci
#| tbl-cap: "Power estimates with uncertainty for the recommended prior
#|   configuration (Normal effect prior, sd = 0.25, 50 trials per subject,
#|   matched Logistic intercept prior). Values show posterior mean [95%
#|   credible interval] for P(BF10 > 3)."
#| eval: !expr posteriors_available

posteriors_bf |>
  filter(
    n_trials == 50,
    dist_b0 == "logistic",
    dist_b1 == "normal",
    sd_b1   == 0.25,
    true_b1 > 0,
    threshold == 3
  ) |>
  mutate(
    power_summary = sprintf("%.2f [%.2f, %.2f]",
                            power_post_mean, power_ci_lower, power_ci_upper)
  ) |>
  select(true_b1, n_subjects, power_summary) |>
  pivot_wider(names_from = true_b1, values_from = power_summary,
              names_prefix = "b1 = ") |>
  rename("n_subjects" = n_subjects) |>
  knitr::kable()
```

In the range examined here (20 to 50 trials), increasing trials per subject
substantially improved BF sensitivity. How to optimally allocate resources
between participants and trials depends on cost structure and variance
components, which is beyond the scope of this work.

## Extension to Other Model Classes {#sec-extensions}

The workflow demonstrated here - analytical matched priors, systematic prior
predictive simulation, and large-scale BF calibration - transfers directly
to other model classes where link functions mediate between parameter and
observation scales.

Ordinal regression. Cumulative logit and probit models use the same link
functions as binomial GLMs, so the matched prior principle applies directly to
the threshold parameters. A Logistic prior on logit-scale thresholds produces
uniform implied category probabilities.

Count data models. Poisson and negative binomial regression use a log
link. The matched prior principle suggests an Exponential prior on the
log-scale intercept. Prior predictive simulation can reveal what prior scales
on log-scale coefficients imply for rate ratios.

Beta regression. For proportional data, beta regression models the mean
via a logit link. The matched prior principle applies identically to the
intercept.

Distributional regression. In GAMLSS [@klein2015] and brms distributional
models [@burkner2021], *all* parameters of the response distribution (mean,
variance, shape) are linked to predictors through link functions. Each
parameter therefore faces its own prior--link interaction. The matched prior
principle and the simulation approach developed here provide a systematic way
to address prior specification for each parameter.

In each case, the key steps are: (1) identify the link function and the
corresponding matched prior family via the probability integral transform;
(2) use prior predictive simulation to map prior scales to interpretable
quantities on the observation scale; and (3) if Bayes Factor testing is
intended, run a calibration simulation to verify adequate power and
specificity.

## References

::: {#refs}
:::

## Appendix A: Uncertainty Quantification {#sec-appendix-uncertainty .appendix}

Each simulation condition yields $n_{\text{valid}}$ replications (out of 100),
each producing a BF10 value. Rather than reporting detection rates as
point-estimate proportions, we apply Bayesian conjugate updating to obtain
full posterior distributions that reflect the sampling uncertainty from the
finite number of replications.

### Two-category model (Beta-Binomial conjugacy)

For a given BF threshold $\tau$ (e.g., $\tau = 3$ or $\tau = 10$), each
replication is a Bernoulli trial: "success" if BF10 $> \tau$ (for power) or
BF01 $> \tau$ (for specificity). With $k$ successes out of $n$ valid
replications and a $\text{Beta}(\alpha_0, \beta_0)$ prior, the posterior is:

$$\theta \mid k, n \sim \text{Beta}(\alpha_0 + k,\; \beta_0 + n - k)$$

We use $\alpha_0 = \beta_0 = 3$, a symmetric prior centered at 0.5 with 95%
of its mass in $[0.12,\, 0.88]$. This mildly regularises the estimate away
from extreme values while being dominated by the data.

### Three-category model (Dirichlet-Multinomial conjugacy)

Each replication falls into exactly one of three categories: (1) H1 detected
(BF10 $> \tau$), (2) inconclusive ($1/\tau \leq \text{BF10} \leq \tau$), or
(3) H0 detected (BF10 $< 1/\tau$). The Dirichlet-Multinomial conjugate
update gives:

$$(\theta_{\text{H1}},\, \theta_{\text{inc}},\, \theta_{\text{H0}}) \mid \mathbf{k} \sim \text{Dirichlet}(\boldsymbol{\alpha}_0 + \mathbf{k})$$

with $\boldsymbol{\alpha}_0 = (1, 3, 1)$. This prior encodes the expectation
that most BFs will be inconclusive (prior mean 60%), with lower prior mass on
decisive evidence for either hypothesis (20% each). The effective prior sample
size is 5, negligible relative to the data.

### Sensitivity to Dirichlet prior choice

To confirm that the Dirichlet prior specification is inconsequential given
$n = 100$ replications per condition, we compare the
$\text{Dirichlet}(1, 3, 1)$ used in the main analyses to the symmetric
$\text{Dirichlet}(1, 1, 1)$ (uniform on the simplex).

```{r}
#| label: fig-dirichlet-sensitivity
#| fig-cap: "The choice of Dirichlet prior has negligible impact on posterior
#|   means. Points compare the Dirichlet(1,3,1) and Dirichlet(1,1,1) priors
#|   for each simulation condition; all points fall near the identity line,
#|   confirming that 100 replications dominate the prior. Dashed line: identity."
#| fig-height: 5
#| fig-width: 10
#| eval: !expr posteriors_available

# Recompute posteriors under Dirichlet(1,1,1) for comparison
results_dir <- here("output", "Simulation_BFCalibration")
if (dir.exists(results_dir)) {
  cond_files <- list.files(results_dir, pattern = "^BFCalib_Cond-", full.names = TRUE)
} else {
  results_dir <- here("output", "Simulation_BFCalibration_RE")
  cond_files <- list.files(results_dir, pattern = "^BFCalib_RE_Cond-", full.names = TRUE)
}

sensitivity_rows <- lapply(cond_files, function(f) {
  cond_data <- readRDS(f)
  bf10_vec  <- cond_data$results[, "BF10"]
  counts    <- compute_bf_counts(bf10_vec, threshold = 3)

  dir_131 <- dirichlet_posterior_summary(
    counts["n_h1"], counts["n_h0"], counts["n_inconclusive"],
    alpha0 = c(1, 3, 1)
  )
  dir_111 <- dirichlet_posterior_summary(
    counts["n_h1"], counts["n_h0"], counts["n_inconclusive"],
    alpha0 = c(1, 1, 1)
  )

  data.frame(
    h1_mean_131  = dir_131["h1_mean"],
    h1_mean_111  = dir_111["h1_mean"],
    inc_mean_131 = dir_131["inc_mean"],
    inc_mean_111 = dir_111["inc_mean"],
    h0_mean_131  = dir_131["h0_mean"],
    h0_mean_111  = dir_111["h0_mean"]
  )
})

sensitivity <- bind_rows(sensitivity_rows)

p1 <- ggplot(sensitivity, aes(x = h1_mean_111, y = h1_mean_131)) +
  geom_abline(slope = 1, intercept = 0, linetype = "dashed", color = "grey50") +
  geom_point(alpha = 0.5, size = 1.5) +
  scale_x_continuous("Dirichlet(1,1,1)", limits = c(0, 1)) +
  scale_y_continuous("Dirichlet(1,3,1)", limits = c(0, 1)) +
  labs(title = "H1 detected") +
  base_theme()

p2 <- ggplot(sensitivity, aes(x = inc_mean_111, y = inc_mean_131)) +
  geom_abline(slope = 1, intercept = 0, linetype = "dashed", color = "grey50") +
  geom_point(alpha = 0.5, size = 1.5) +
  scale_x_continuous("Dirichlet(1,1,1)", limits = c(0, 1)) +
  scale_y_continuous("Dirichlet(1,3,1)", limits = c(0, 1)) +
  labs(title = "Inconclusive") +
  base_theme()

p3 <- ggplot(sensitivity, aes(x = h0_mean_111, y = h0_mean_131)) +
  geom_abline(slope = 1, intercept = 0, linetype = "dashed", color = "grey50") +
  geom_point(alpha = 0.5, size = 1.5) +
  scale_x_continuous("Dirichlet(1,1,1)", limits = c(0, 1)) +
  scale_y_continuous("Dirichlet(1,3,1)", limits = c(0, 1)) +
  labs(title = "H0 detected") +
  base_theme()

gridExtra::grid.arrange(p1, p2, p3, ncol = 3,
  top = "Dirichlet prior sensitivity: posterior means by category (BF > 3)")
```

## Appendix B: Orthonormal Contrast Scaling {#sec-appendix-contrasts .appendix}

For unbalanced designs, the predictor SD depends on cell sizes and should be
computed from the contrast matrix directly. For factors with many levels
($J > 5$), the `BayesFactor` package [@rouder2012; @rouder2017] implements
the full multivariate JZS prior, which accounts for the dimensionality of the
effect space without requiring per-coefficient prior widths.

The full scaling produces wider priors for interactions than for main effects.
In many experimental contexts, interactions are expected to be smaller than
main effects, so using the *unscaled* main-effect prior for interaction terms
is a reasonable alternative. @tbl-orthonormal-scaling in the main text gives
the default when no such expectation exists.

A subtlety arises when priors on interaction terms are specified
independently of the main-effect priors. In principle, the priors on main
effects $A$ and $B$ already *imply* a distribution over the interaction
$A \times B$, because the joint prior on all coefficients determines the
prior on any function of those coefficients. Manually specifying a separate
prior on the interaction term can create an inconsistency: the prior beliefs
implied by the main-effect priors may not be compatible with the interaction
prior. In practice, this is unlikely to cause problems when the interaction
has a clear substantive interpretation (e.g., a theoretically motivated
moderation effect) and the priors are set to reflect genuine prior beliefs.
However, when using default priors for exploratory analyses, we recommend
using the scaled priors from @tbl-orthonormal-scaling rather than
independently chosen interaction priors, as the scaling rule ensures internal
consistency across the coefficient space.

## Appendix C: No-Intercept Parameterisation {#sec-appendix-no-intercept .appendix}

In `brms`, specifying `0 + condition` removes the intercept and estimates
separate coefficients for each condition level, so that
$p_0 = g^{-1}(b_0)$, $p_1 = g^{-1}(b_1)$, and
$\delta p = g^{-1}(b_1) - g^{-1}(b_0)$. Under this parameterization, the
matched prior principle applies to each coefficient independently: placing
$\text{Logistic}(0, 1)$ on both coefficients under the logit link induces
independent $\text{Uniform}(0, 1)$ marginals, and the implied prior on
$\delta p$ is a symmetric triangular distribution on $(-1, 1)$. Because the
coefficient interpretation differs from the intercept-plus-slope
parameterization, our recommended prior widths do not transfer directly ---
each condition mean should receive a matched prior scaled to reflect beliefs
about plausible success probabilities (e.g., $\text{sd} = 0.75$), and the
implied $\delta p$ should be checked via prior predictive simulation.
