---
title: "A Three-Step Workflow for Prior Specification in GLMs: Default Recommendations for Binomial Models"
shorttitle: "Prior Specification Workflow for GLMs"
author:
  - name: Gidon T. Frischkorn
    orcid: 0000-0002-5055-9764
    email: gidon.frischkorn@psychologie.uzh.ch
    corresponding: true
    affiliations:
      - name: University of Zurich
        department: Department of Psychology
      - name: University of Lucerne
        department: Faculty of Behavioral Sciences and Psychology
  - name: Joscha Dutli
    orcid: 0009-0009-5007-1281
    affiliations:
      - name: University of Zurich
        department: Department of Psychology
  - name: Philipp Musfeld
    orcid: 0000-0002-6539-0105
    affiliations:
      - name: University of Amsterdam
        department: Department of Psychology
      - name: University of Zurich
        department: Department of Psychology
  - name: Klaus Oberauer
    orcid: 0000-0001-8815-6189
    affiliations:
      - name: University of Zurich
        department: Department of Psychology
author-note:
  disclosures:
    study-registration: "This study was not preregistered."
    data-sharing: "Simulation code and analysis scripts are available at [repository link]."
    conflict-of-interest: "The authors declare no conflicts of interest."
    financial-support: "This research was supported by a grant from the Swiss National Science Foundation (SNSF) awarded to Gidon T. Frischkorn (Grant No. 208828)."
    authorship-agreements: |
      Gidon T. Frischkorn: Conceptualization, Methodology, Software, Formal Analysis, Visualization, Writing -- Original Draft.
      Joscha Dutli: Conceptualization, Methodology, Writing -- Review & Editing.
      Philipp Musfeld: Conceptualization,Methodology, Writing -- Review & Editing.
      Klaus Oberauer: Conceptualization, Supervision, Methodology, Writing -- Original Draft.
floatsintext: true
output-dir: "."
format:
  apaquarto-docx: default
  apaquarto-html: default
  apaquarto-pdf:
    documentmode: man
bibliography:
  - references.bib
csl: apa.csl
abstract: "Bayesian hypothesis testing with Bayes Factors in generalized linear models (GLMs) requires priors on regression coefficients, yet no principled framework exists for choosing these priors when the model uses a nonlinear link function. The link function creates a prior--link interaction: priors that appear reasonable on the coefficient scale can imply implausible distributions on the observation scale. We propose a three-step workflow for prior specification that generalizes across GLM families: (1) identify *matched priors* via the probability integral transform -- prior families that mirror the link function and produce uniform distributions on the observation scale; (2) use prior predictive simulation to map prior scales to interpretable quantities; and (3) calibrate the recommended priors for Bayes Factor hypothesis testing. We demonstrate this workflow for binomial GLMs (logistic and probit regression), providing default prior recommendations and ready-to-use `brms` code. We show that the matched Logistic(0, 0.75) intercept prior and Logistic(0, 0.25--0.30) effect prior provide well-calibrated Bayes Factors across a wide range of experimental scenarios. Scaling formulas allow researchers to adapt these defaults to any contrast coding, predictor scale, or informed prior belief. The workflow transfers directly to ordinal, lognormal, count, beta, and distributional regression models."
---

```{r}
#| label: setup
#| include: false

library(dplyr)
library(tidyr)
library(ggplot2)
library(ggridges)
library(viridis)
library(scales)
library(here)
library(knitr)

source(here("R", "link_functions.R"))
source(here("R", "prior_sampling.R"))
source(here("R", "plotting.R"))
# Goal 1: Prior predictive results
summaries_pp  <- readRDS(here("output", "prior_predictive_summaries.rds"))
draws_long    <- readRDS(here("output", "prior_predictive_draws.rds"))

# Goal 2: BF calibration results
summaries_bf  <- readRDS(here("output", "bf_calibration_summaries.rds"))

re_summaries <- tryCatch(
  readRDS(here("output", "bf_calibration_re_summaries.rds")),
  error = function(e) NULL
)
re_results_available <- !is.null(re_summaries)

# Factor labels
dist_labels <- c(cauchy = "Cauchy", normal = "Normal", logistic = "Logistic")

# Use prior_theme() from R/plotting.R for all figures
```

## Introduction

In psychology and related fields, we often model data that reflect counts of successes out of trials -- for instance, accuracy in a cognitive task or endorsement rates in a survey. These data follow a binomial distribution and are best modeled with a binomial generalized linear model (GLM), where the success probability $p_i$ is related to a linear predictor via a link function $g$:

$$g(p_i) = b_0 + b_1 x_i$$

When testing whether an effect exists, Bayesian hypothesis testing via Bayes Factors [@wagenmakers2018a; @schonbrodt2018] compares a model including the effect ($H_1$) against one constraining it to zero ($H_0$). The Bayes Factor $BF_{10}$ depends critically on the prior for the effect parameter $b_1$. For models assuming normally distributed outcomes, @rouder2012 proposed principled default priors. For GLMs with nonlinear link functions, no such defaults exist.

This gap matters because the link function creates a prior--link interaction: a prior that appears reasonable on the regression-coefficient scale can produce implausible distributions on the probability scale [@gelman2008; @ghosh2018]. Priors recommended for posterior estimation [e.g., the weakly informative Cauchy(0, 2.5) of @gelman2008] are typically too broad for hypothesis testing, biasing Bayes Factors toward the null.[^elicitation]

[^elicitation]: Recent advances in prior elicitation have developed methods that translate expert knowledge into priors, either through structured queries about observable quantities [@hosack2024; @mikkola2023] or simulation-based optimization [@bockting2024]. These approaches are valuable when domain expertise is available, but they require expert input for each new application. Our contribution is complementary: we provide principled *default* priors that can be used off the shelf when no domain-specific knowledge is available -- analogous to what @rouder2012 provided for $t$-tests and ANOVA. Following the taxonomy of @mikkola2023, our workflow operates in the observable space, using the link function to bridge between prior distributions on the coefficient scale and their implications on the observation scale.

We propose a three-step workflow for prior specification in GLMs that addresses this problem. The workflow is general -- it applies whenever a link function mediates between parameter and observation scales -- but we demonstrate it fully for binomial models, providing default priors and ready-to-use code. We illustrate the approach with accuracy data from a hypothetical two-condition experiment using sum-to-zero contrast coding ($x \in \{+1, -1\}$), so that the effect on the probability scale is $\delta p = g^{-1}(b_0 + b_1) - g^{-1}(b_0 - b_1)$.

## The Three-Step Workflow {#sec-workflow}

Prior specification for GLMs with nonlinear link functions requires bridging the gap between the coefficient scale (where priors are set) and the observation scale (where effects are interpretable). Our workflow achieves this through three complementary steps:

**Step 1: Identify matched priors.** Using the probability integral transform, we identify prior families whose CDF matches the inverse link function. These matched priors produce exactly uniform distributions on the observation scale, serving as principled, maximally uninformative anchors from which to scale.

**Step 2: Prior predictive simulation.** We draw from the joint prior on all coefficients, transform through the inverse link, and examine the resulting distributions on the observation scale. This reveals what effect sizes and baseline rates the prior treats as plausible -- information that is not apparent from the prior on the coefficient scale alone.

**Step 3: Bayes Factor calibration.** We simulate data under known effects and evaluate how the strength of evidence varies across prior specifications. This verifies that the recommended priors yield well-calibrated Bayes Factors -- neither systematically favouring $H_0$ nor $H_1$.

Each step can be applied to any GLM family. We now demonstrate the full workflow for binomial models.

## Step 1: The Matched Prior Principle {#sec-matched-prior}

The key insight is that when the prior distribution on the coefficient scale belongs to the same family as the link function, it produces a flat (uninformative) prior on the probability scale. This follows from the probability integral transform: applying a distribution's own CDF to a random variable drawn from that distribution always yields a uniform result.

| Link | Inverse link $g^{-1}$ | Matched prior | Implied prior on $p$ |
|------|----------------------|---------------|----------------------|
| logit | logistic CDF | $\text{Logistic}(0, 1)$ | $\text{Uniform}(0, 1)$ |
| probit | normal CDF | $\text{Normal}(0, 1)$ | $\text{Uniform}(0, 1)$  |

Scaling the matched prior with $\text{sd} > 1$ concentrates mass toward the extreme values 0 and 1; scaling with $\text{sd} < 1$ concentrates mass away from the extremes, producing a regularized intercept prior suited to experiments where floor and ceiling performance are unlikely. The matched prior at scale 1 is therefore maximally uninformative on the probability scale -- a principled anchor. This principle generalises to any model with a CDF-based link function (see [Applying the Workflow Beyond Binomial Models]).

## Step 2: Prior Predictive Simulation {#sec-prior-predictive}

We simulated prior predictive distributions across a factorial design crossing two link functions (logit, probit), three prior families (Normal, Logistic, Cauchy), and multiple scale values for both intercept and effect priors (see Supplementary Materials for the full design). Concretely, we drew 10,000 samples per condition from the joint prior on $(b_0, b_1)$ and computed the implied distributions on the probability scale.

### Intercept Prior {#sec-pp-intercept}

```{r}
#| label: fig-1-p-intercept
#| fig-cap: "Prior predictive distribution of success probability
#|   p = $g^{-1}(b_0)$ for each intercept prior SD (rows) and link function
#|   (columns). At $sd_{b_0} = 1.0$, the matched prior (Logistic for logit,
#|   Normal for probit) produces a perfectly uniform distribution. Reducing
#|   $sd_{b_0}$ to 0.75 pulls mass away from floor and ceiling. Dashed
#|   curves show exact analytical densities."
#| fig-height: 8
#| fig-width: 10

fig1_draws <- draws_long |>
  filter(dist_b1 == "logistic", sd_b1 == 0.25)

plot_p_intercept_ridges(fig1_draws, sd_b1_focus = 0.25)
```

@fig-1-p-intercept confirms the analytical result: at $\text{sd}_{b_0} = 1.0$ with the matched prior, the distribution of $p$ is uniform. At $\text{sd}_{b_0} = 0.75$, mass is pulled away from 0 and 1, appropriate for experiments avoiding floor and ceiling effects. Misfit priors (e.g., Cauchy for either link) produce heavier tails. We recommend $\text{sd}_{b_0} = 0.75$ as the default, giving $\text{Logistic}(0, 0.75)$ for the logit link and $\text{Normal}(0, 0.75)$ for the probit link (see Supplementary Figure S1 for floor/ceiling mass quantification).

### Effect Prior {#sec-pp-effect}

The prior on $b_1$ determines how much prior mass falls on small versus large effects. Due to the nonlinear link, we must compute the implied prior on $\delta p$ by drawing from the joint prior on $(b_0, b_1)$ and transforming through the inverse link.

```{r}
#| label: fig-3-delta-p
#| fig-cap: "Prior predictive distribution of delta_p (signed effect size on
#|   the probability scale) for each effect prior SD (facets), colour-coded by
#|   prior family (ridges). Shown for the logit link with matched Logistic(0, 0.75)
#|   intercept prior. At sd_b1 = 0.25, Normal and Logistic priors concentrate
#|   mass on effects up to about 20 percentage points. The Cauchy assigns
#|   substantial probability to effects exceeding 50 percentage points even at
#|   this moderate scale. Dotted lines mark +/-0.10, +/-0.20, +/-0.30."
#| fig-height: 12
#| fig-width: 14

# Use matched Logistic intercept prior on logit link
fig2_draws <- draws_long |>
  filter(
    link == "logit",
    dist_b0 == "logistic",
    sd_b0 == 0.75
  )

plot_delta_p_ridges(fig2_draws, sd_b0_focus = 0.75)
```

The Cauchy prior stands apart at every scale: even at $\text{sd}_{b_1} = 0.25$, its heavy tails assign substantial probability to $|\delta p| > 0.50$, making it a poor default. We use the 90th percentile of $|\delta p|$ as the key summary: at $\text{sd}_{b_1} = 0.25$, the 90th percentile is near 0.20 -- a large effect by the standards of most psychological research. Values up to $\text{sd}_{b_1} = 0.50$ remain defensible when very large effects are plausible. We recommend $\text{sd}_{b_1} \in [0.25, 0.30]$ as the default range.

### Random Effects Hyperprior {#sec-pp-re}

```{r}
#| label: setup-re-pp
#| include: false

re_pp_summaries <- tryCatch(
  readRDS(here("output", "prior_predictive_re_summaries.rds")),
  error = function(e) NULL
)
re_pp_draws <- tryCatch(
  readRDS(here("output", "prior_predictive_re_draws.rds")),
  error = function(e) NULL
)

re_pp_available <- !is.null(re_pp_summaries) && !is.null(re_pp_draws)
```

For hierarchical models, the hyperprior on the random-effects SD $\sigma$ controls between-subject variability. We compared four hyperprior families -- Exponential, Gamma(shape = 2), Half-Normal, and Half-Student-t(df = 3) -- at five scale levels matched by their mean (see Supplementary Materials for details).

```{r}
#| label: fig-re-1-p-intercept
#| fig-cap: "Subject-level success probabilities by hyperprior family and scale
#|   (mean of the hyperprior). At matched scale, all four families produce
#|   nearly identical distributions; the primary driver is the scale parameter.
#|   Dashed vertical lines mark p = 0.05 and p = 0.95."
#| fig-height: 10
#| fig-width: 12
#| eval: !expr re_pp_available

re_pp_draws |>
  group_by(hp_family, hp_mean, re_structure) |>
  slice_sample(n = 5000) |>
  ungroup() |>
  mutate(hp_mean_label = paste0("mu == ", hp_mean)) |>
  ggplot(aes(x = p_intercept, y = hp_family,
             fill = hp_family, colour = hp_family)) +
  geom_density_ridges(stat = "binline", bins = 80,
                      alpha = 0.5, scale = 0.9) +
  geom_vline(xintercept = c(0.05, 0.95), linetype = "dashed",
             colour = "grey40", linewidth = 0.5) +
  facet_grid(hp_mean_label ~ re_structure,
             labeller = labeller(hp_mean_label = label_parsed)) +
  scale_fill_viridis_d(option = "D", end = 0.85, guide = "none") +
  scale_colour_viridis_d(option = "D", end = 0.85, guide = "none") +
  scale_x_continuous(
    expression("Success probability" ~ italic(p)),
    limits = c(0, 1), expand = expansion(mult = c(0.02, 0.02))
  ) +
  scale_y_discrete("Family") +
  prior_theme() +
  theme(strip.text.y = element_text(size = 10))
```

At matched scale, all four families produce nearly identical distributions (@fig-re-1-p-intercept). The primary driver is the scale, not the family. At a hyperprior mean of 0.25, subject-level probabilities are well concentrated away from floor and ceiling. We recommend $\text{Exponential}(4)$ (mean = 0.25) as the default for its simplicity. The `brms` default `student_t(3, 0, 2.5)` (mean $\approx$ 2.76) spreads subject-level probabilities too heavily toward 0 and 1 for typical experiments (see Supplementary Figure S2 for effect heterogeneity).


## Step 3: Bayes Factor Calibration {#sec-bf-calibration}

The prior predictive analysis established which priors produce plausible distributions. We now verify that these priors yield well-calibrated Bayes Factors. We simulated data with true effects ranging from $b_1 = 0$ to $b_1 = 0.50$, crossing three effect prior families (Normal, Logistic, Cauchy) at two scales ($\text{sd}_{b_1} = 0.25$ and $0.50$), two intercept priors (matched Logistic vs. misfit Normal), three sample sizes (30, 60, 100), and two trial counts (20, 50). Bayes Factors were computed via the Savage--Dickey density ratio [@wagenmakers2018; @ly2016]. We present the most data-sparse scenario ($N = 30$, 20 trials) where prior sensitivity is most visible.

```{r}
#| label: setup-bf-raw
#| include: false

# Load raw BF10 values from per-condition files for violin plots
bf_raw <- tryCatch({
  results_dir <- here("output", "Simulation_BFCalibration")
  cond_files <- list.files(results_dir, pattern = "^BFCalib_Cond-",
                           full.names = TRUE)
  raw_rows <- lapply(cond_files, function(f) {
    d <- readRDS(f)
    bf10_vec <- d$results[, "BF10"]
    cbind(
      as.data.frame(as.list(d$condition)),
      data.frame(BF10 = bf10_vec, log10_BF10 = log10(bf10_vec))
    )
  })
  bind_rows(raw_rows)
}, error = function(e) NULL)
bf_raw_available <- !is.null(bf_raw)

# Load raw RE BF10 values
bf_raw_re <- tryCatch({
  results_dir_re <- here("output", "Simulation_BFCalibration_RE")
  cond_files_re <- list.files(results_dir_re, pattern = "^BFCalib_RE_Cond-",
                              full.names = TRUE)
  raw_rows_re <- lapply(cond_files_re, function(f) {
    d <- readRDS(f)
    bf10_vec <- d$results[, "BF10"]
    cbind(
      as.data.frame(as.list(d$condition)),
      data.frame(BF10 = bf10_vec, log10_BF10 = log10(bf10_vec))
    )
  })
  bind_rows(raw_rows_re)
}, error = function(e) NULL)
bf_raw_re_available <- !is.null(bf_raw_re)
```

### Effect of Prior Family and Width

```{r}
#| label: fig-bf-violin-family
#| fig-cap: "Distribution of log10(BF10) by effect prior family and width for
#|   N = 30 subjects and 20 trials per subject. Rows: true effect size. Columns:
#|   prior SD. The choice of prior family has minimal impact; prior width shifts
#|   the distribution modestly. Black dots indicate medians. Dashed lines mark
#|   BF10 = 3 and BF10 = 1/3."
#| fig-height: 7
#| fig-width: 10
#| eval: !expr bf_raw_available

bf_raw |>
  filter(
    dist_b0 == "logistic",
    true_b1 %in% c(0, 0.10, 0.20),
    n_subjects == 30,
    n_trials == 20
  ) |>
  mutate(
    dist_b1 = factor(dist_b1, levels = c("cauchy", "normal", "logistic"),
                     labels = c("Cauchy", "Normal", "Logistic")),
    true_b1_label = factor(paste0("b[1] == ", true_b1)),
    sd_b1_label   = factor(paste0("sd[b[1]] == ", sd_b1))
  ) |>
  ggplot(aes(x = dist_b1, y = log10_BF10, fill = dist_b1)) +
  geom_hline(yintercept = c(log10(3), log10(1/3)), linetype = "dashed",
             colour = "grey50", linewidth = 0.5) +
  geom_hline(yintercept = 0, colour = "grey30", linewidth = 0.4) +
  geom_violin(alpha = 0.6, scale = "width", width = 0.8) +
  stat_summary(fun = median, geom = "point", size = 2, colour = "black") +
  facet_grid(true_b1_label ~ sd_b1_label, labeller = label_parsed) +
  scale_fill_viridis_d("Prior family", option = "D", end = 0.85) +
  scale_y_continuous(
    expression(BF[10]),
    breaks = log10(c(1/100, 1/10, 1/3, 1, 3, 10, 100)),
    labels = c("1/100", "1/10", "1/3", "1", "3", "10", "100")
  ) +
  coord_cartesian(ylim = c(-2.5, 2.5)) +
  labs(x = "Effect prior family") +
  prior_theme()
```

The choice of effect prior family has minimal impact on evidence strength (@fig-bf-violin-family). All three families produce nearly identical BF distributions. The prior width has a modest effect: narrower priors ($\text{sd} = 0.25$) yield slightly stronger evidence for $H_1$ when the effect is present, particularly at $b_1 = 0.2$, though the difference is small relative to variability across replications.

### Effect of Random-Effects Hyperprior

```{r}
#| label: fig-bf-violin-re
#| fig-cap: "Distribution of log10(BF10) by random-effects hyperprior for
#|   N = 30, 20 trials. All three hyperpriors produce nearly identical BF
#|   distributions. Black dots indicate medians. Dashed lines mark BF10 = 3
#|   and BF10 = 1/3."
#| fig-height: 5
#| fig-width: 8
#| eval: !expr bf_raw_re_available

bf_raw_re |>
  filter(
    true_b1 %in% c(0, 0.10, 0.20),
    n_subjects == 30,
    n_trials == 20
  ) |>
  mutate(
    sd_prior_re = factor(sd_prior_re,
                         levels = c("default", "gamma", "exponential"),
                         labels = c("Student-t(3,0,2.5)", "Gamma(2,4)",
                                    "Exponential(4)")),
    true_b1_label = factor(paste0("b[1] == ", true_b1))
  ) |>
  ggplot(aes(x = sd_prior_re, y = log10_BF10, fill = sd_prior_re)) +
  geom_hline(yintercept = c(log10(3), log10(1/3)), linetype = "dashed",
             colour = "grey50", linewidth = 0.5) +
  geom_hline(yintercept = 0, colour = "grey30", linewidth = 0.4) +
  geom_violin(alpha = 0.6, scale = "width", width = 0.7) +
  stat_summary(fun = median, geom = "point", size = 2, colour = "black") +
  facet_wrap(~ true_b1_label, nrow = 1, labeller = label_parsed) +
  scale_fill_viridis_d("RE hyperprior", option = "D", end = 0.85) +
  scale_y_continuous(
    expression(BF[10]),
    breaks = log10(c(1/100, 1/10, 1/3, 1, 3, 10, 100)),
    labels = c("1/100", "1/10", "1/3", "1", "3", "10", "100")
  ) +
  coord_cartesian(ylim = c(-2.5, 2.5)) +
  labs(x = "Random-effects hyperprior") +
  prior_theme() +
  theme(axis.text.x = element_text(angle = 30, hjust = 1))
```

The random-effects hyperprior has no meaningful impact on Bayes Factors for the fixed effect (@fig-bf-violin-re), even though the Student-*t*(3, 0, 2.5) is far wider than Exponential(4). This confirms that the hyperprior choice is secondary.

### Summary

Across all comparisons, the important choice is the scale of the effect prior, for which we recommend $\text{sd}_{b_1} \in [0.25, 0.30]$. Prior family, intercept prior family, and random-effect hyperprior have negligible impact. We recommend the matched Logistic intercept prior for interpretability and Exponential(4) for random effects, but these choices are not critical for BF performance. To test robustness, we recommend running a prior sensitivity analysis by computing BFs across a small set of plausible prior widths (e.g., $\text{sd}_{b_1} \in \{0.15, 0.25, 0.35\}$).


## Recommendations {#sec-recommendations}

### Default Prior Specifications

#### Intercept prior

| Link | Recommended prior | Rationale |
|------|------------------|-----------|
| logit | $\text{Logistic}(0, 0.75)$ | Matched family; sd = 0.75 gives < 10% floor/ceiling mass |
| probit | $\text{Normal}(0, 0.75)$ | Matched family; same criterion |

#### Effect prior

| Link | Recommended prior | Rationale |
|------|------------------|-----------|
| logit | $\text{Logistic}(0, 0.25\text{--}0.30)$ | Matched family; 90th percentile of $|\delta p| \approx 0.20\text{--}0.23$ |
| probit | $\text{Normal}(0, 0.25\text{--}0.30)$ | Matched family; 90th percentile of $|\delta p| \approx 0.18\text{--}0.22$ |

#### Random effects prior

The hyperprior on random-effect SDs should concentrate mass below approximately 0.5 logit units. Our recommended default is `exponential(4)` (mean = 0.25). For larger expected between-subject variability, `gamma(2, 4)` (mean = 0.50) is an alternative.

#### Ready-to-use brms prior specifications

```{r}
#| label: recommended-priors
#| echo: true
#| eval: false

# Logit link (logistic regression)
logit_priors <- c(
  brms::prior(logistic(0, 0.75), class = Intercept),
  brms::prior(logistic(0, 0.25), class = b)
)

# Probit link (probit regression)
probit_priors <- c(
  brms::prior(normal(0, 0.75), class = Intercept),
  brms::prior(normal(0, 0.25), class = b)
)
```

For models with subject-level random effects:

```{r}
#| label: recommended-priors-re
#| echo: true
#| eval: false

# Logit link with random intercept
logit_priors_re <- c(
  brms::prior(logistic(0, 0.75), class = Intercept),
  brms::prior(logistic(0, 0.25), class = b),
  brms::prior(exponential(4),    class = sd)
)
```

### From Defaults to Informed Priors {#sec-scaling}

The recommended $\text{sd}_{b_1} \in [0.25, 0.30]$ assumes $\text{SD}(x) = 1$ (sum-to-zero $\pm 1$ coding). When the predictor's SD differs from 1, rescale:

$$\text{sd}_{b_1}(\text{predictor SD} = s) = \frac{\text{sd}_{b_1}(\text{SD} = 1)}{s}$$

Examples: $\pm 0.5$ coding ($s = 0.5$): use $\text{sd}_{b_1} = 0.50$. Treatment coding ($s = 0.5$ in balanced designs): use $\text{sd}_{b_1} = 0.50$. Standardised continuous predictor ($s = 1$): use $\text{sd}_{b_1} = 0.25$ directly.

#### Multi-level factors with orthonormal contrasts

For factors with more than two levels, orthonormal contrasts [e.g., `contr.equalprior()` from the `bayestestR` package; @makowski2019; @rouder2012] equalize contrast column variances, ensuring a single prior width induces the same distribution on every pairwise comparison. In a balanced design with $J$ levels, each column has predictor $\text{SD} = 1 / \sqrt{J}$, giving:

$$\text{sd}_{b_1}(\text{orthonormal, } J \text{ levels}) = 0.25 \times \sqrt{J}$$

```{r}
#| label: fig-orthonormal-demo
#| fig-cap: "Prior predictive distribution of a single representative
#|   pairwise $\\delta p$ under orthonormal contrasts with the rescaled prior
#|   ($\\text{sd}_{b_1} = 0.25 \\times \\sqrt{J}$). The distribution is
#|   nearly identical across $J$ -- confirming that the scaling rule
#|   preserves per-comparison calibration."
#| fig-height: 4.5
#| fig-width: 6

set.seed(2024)
n_draws <- 50000

demo_list <- lapply(c(2, 3, 4), function(J) {
  oc <- orthonormal_contrasts(J)
  sd_b1_scaled <- 0.25 * oc$scaling_factor

  b0 <- rlogis(n_draws, location = 0, scale = 0.75)
  B <- matrix(
    rnorm(n_draws * (J - 1), mean = 0, sd = sd_b1_scaled),
    nrow = n_draws, ncol = J - 1
  )

  eta_mat <- outer(b0, rep(0, J)) + B %*% t(oc$Q)
  p_mat   <- ilogit(eta_mat)

  delta_p_pair <- p_mat[, 1] - p_mat[, 2]

  data.frame(
    J            = paste0("J = ", J),
    delta_p_pair = delta_p_pair
  )
})

demo_draws <- do.call(rbind, demo_list)

p_single <- ggplot(demo_draws, aes(x = delta_p_pair,
                                    y = J, fill = J)) +
  geom_density_ridges(alpha = 0.5, scale = 1.2) +
  geom_vline(xintercept = c(-0.20, -0.10, 0.10, 0.20),
             linetype = "dotted", linewidth = 0.4,
             colour = "grey40") +
  scale_x_continuous(expression(delta * p ~ "(single pair)"),
                     limits = c(-0.8, 0.8)) +
  scale_fill_viridis_d("Factor levels", option = "D", end = 0.85) +
  ggtitle("Single pairwise comparison") +
  prior_theme()

p_single
```


## Applying the Workflow Beyond Binomial Models {#sec-extensions}

The three-step workflow demonstrated above transfers directly to other model classes where link functions mediate between parameter and observation scales. We briefly sketch the application to four common model families and provide a more detailed walkthrough for ordinal regression.

### Ordinal regression

Cumulative logit and probit models use the same link functions as binomial GLMs, applied to cumulative probabilities across ordered response categories. Each threshold parameter $\tau_k$ determines the boundary between adjacent categories on the latent scale, and the inverse link converts these boundaries to cumulative probabilities.

**Step 1 (Matched priors).** The matched prior principle applies directly: a $\text{Logistic}(0, 1)$ prior on each threshold under the cumulative logit link produces a uniform distribution of cumulative probabilities at that threshold. Scaling to $\text{sd} < 1$ regularizes the thresholds toward the center of the latent scale, concentrating prior mass away from extreme category probabilities (where nearly all responses fall into one category). For the effect prior on predictors, the same matched families apply as in the binomial case.

**Step 2 (Prior predictive simulation).** The key difference from the binomial case is that the observation scale is now a vector of category probabilities rather than a single success probability. Prior predictive simulation should examine the implied distribution of category probabilities and, critically, of contrasts between conditions (e.g., shifts in the most likely category, or changes in cumulative probabilities). One can draw $b_0$, threshold spacings, and effect coefficients from the joint prior, compute category probabilities via the inverse link, and inspect whether the implied distributions match substantive expectations about response patterns.

**Step 3 (BF calibration).** Bayes Factor calibration follows the same template: simulate data with known effects, fit cumulative models with the candidate priors, and evaluate BF sensitivity. The Savage--Dickey density ratio applies to the effect coefficient just as in the binomial case, because the hypothesis $b_1 = 0$ has the same structure regardless of the outcome distribution.

### Other model classes

*Count data.* Poisson and negative binomial regression use a log link. The matched prior principle does not apply directly (the log link is not a CDF), but prior predictive simulation can reveal what prior scales on log-scale coefficients imply for rate ratios.

*Beta regression.* For proportional data modeled via a logit link on the mean, the matched prior principle applies identically to the intercept.

*Distributional regression.* In `brms` distributional models [@burkner2021], all parameters of the response distribution (mean, variance, shape) are linked to predictors through link functions. Each parameter faces its own prior--link interaction. The matched prior principle and simulation approach developed here provide a systematic way to address prior specification for each parameter.

### Lognormal regression for response times

Response times are among the most common continuous outcomes in experimental psychology. Because they are right-skewed and strictly positive, lognormal regression (`family = lognormal()` in `brms`, with an identity link on the log scale) is a natural modelling choice. Here, the matched prior principle from Step 1 does not apply: the identity link involves no nonlinear CDF, so there is no distribution family to "match." Nevertheless, a prior--transform interaction remains, because the exponential function maps the log-scale intercept to the median response time in milliseconds. For instance, a $\text{Normal}(6, 1)$ prior on the intercept implies a 95% range of median response times from $\exp(4) \approx 55$ ms to $\exp(8) \approx 2{,}981$ ms -- implausibly broad for most cognitive tasks. A tighter prior such as $\text{Normal}(6, 0.25)$ concentrates the median in the $\exp(5.5)$--$\exp(6.5) \approx 245$--$665$ ms range more typical of simple reaction-time paradigms. For effect coefficients, the mapping is multiplicative: a coefficient $b_1$ on the log scale implies a proportional change of $\exp(b_1)$ on the response-time scale, so a $\text{Normal}(0, 0.1)$ effect prior treats effects up to roughly 20% speedup or slowdown as plausible. Steps 2 and 3 of the workflow apply directly: prior predictive simulation reveals what ranges of absolute response times and proportional effects the prior treats as plausible, and Bayes Factor calibration verifies that the chosen scale yields well-behaved evidence. This illustrates that even when CDF-matching is unavailable, the simulation and calibration steps remain essential tools for principled prior specification [see also @hosack2024 for an expert-elicitation approach to lognormal GLMs].

## Discussion

We have proposed a three-step workflow for prior specification in GLMs with nonlinear link functions: (1) identify matched priors via the probability integral transform, (2) evaluate implied prior predictive distributions on the observation scale, and (3) calibrate priors for Bayes Factor hypothesis testing. Applying this workflow to binomial models, we recommend the matched $\text{Logistic}(0, 0.75)$ intercept prior and $\text{Logistic}(0, 0.25\text{--}0.30)$ effect prior for logistic regression (with analogous Normal priors for probit), and $\text{Exponential}(4)$ for random-effects hyperpriors. These defaults are well-calibrated across a wide range of experimental scenarios.

Several features of our results deserve emphasis. First, the matched prior principle provides a theoretically grounded anchor that is missing from current practice: by matching the prior family to the link function, the intercept prior becomes interpretable on the probability scale. Second, the prior predictive simulations revealed that the choice of prior *family* matters far less than the choice of prior *scale* -- a finding that held for both fixed effects and random-effects hyperpriors. Third, the BF calibration confirmed that the recommended priors produce well-behaved evidence across a range of effect sizes, without systematically biasing toward either hypothesis.

Our demonstration is limited to the binomial case, and the specific prior scale recommendations assume standard effect sizes in psychological experiments. Researchers in other fields may need to adjust the defaults based on domain-specific effect size distributions. The BF calibration used the Savage--Dickey density ratio; other BF estimation methods (e.g., bridge sampling for nested or non-nested comparisons) may show different prior sensitivities.[^elicitation-discussion]

[^elicitation-discussion]: When substantive domain expertise is available, expert elicitation methods [@bockting2024; @hosack2024; @mikkola2023] can produce more informative priors tailored to specific research questions. Our defaults are intended as principled starting points for researchers who lack the resources for full elicitation, or as reference points from which to build informed priors. The two approaches are complementary: the matched-prior framework identifies the appropriate prior *family*, while expert elicitation can refine the *scale*.

The workflow itself, however, is general. Any model with a CDF-based link function admits matched priors, and prior predictive simulation is universally applicable. We encourage the research community to apply this workflow to ordinal, count, and distributional regression models, building a library of principled default priors across the full range of GLM families used in psychological research.


## References

::: {#refs}
:::

## Appendix A: Orthonormal Contrast Scaling {#sec-appendix-contrasts .appendix}

| Design term | Predictor SD | Rescaled $\text{sd}_{b_1}$ |
|:---|:---:|:---:|
| 2-level factor | $1/\sqrt{2} \approx 0.71$ | 0.35 |
| 3-level factor | $1/\sqrt{3} \approx 0.58$ | 0.43 |
| 4-level factor | $1/\sqrt{4} = 0.50$ | 0.50 |
| 5-level factor | $1/\sqrt{5} \approx 0.45$ | 0.56 |
| $2 \times 3$ interaction | $1/\sqrt{6} \approx 0.41$ | 0.61 |
| $3 \times 3$ interaction | $1/\sqrt{9} \approx 0.33$ | 0.75 |

: Rescaled effect prior widths under orthonormal contrasts for multi-level
  factors and factorial interactions. All values assume the baseline
  recommendation of $\text{sd}_{b_1} = 0.25$ under $\pm 1$ coding.
  {#tbl-orthonormal-scaling}

For unbalanced designs, the predictor SD depends on cell sizes and should be computed from the contrast matrix directly. For factors with many levels ($J > 5$), the `BayesFactor` package [@rouder2012; @rouder2017] implements the full multivariate JZS prior, which accounts for the dimensionality of the effect space without requiring per-coefficient prior widths.

When using orthonormal contrasts in `brms`:

```{r}
#| label: recommended-priors-orthonormal
#| echo: true
#| eval: false

# - Single 3-level factor ---
contrasts(dat$condition) <- bayestestR::contr.equalprior(3)
# Predictor SD = 1/sqrt(3); rescaled prior: sd_b1 = 0.25 * sqrt(3) = 0.433

priors_3level <- c(
  brms::prior(logistic(0, 0.75),  class = Intercept),
  brms::prior(normal(0, 0.433),   class = b)
)

# - 2 x 3 factorial design ---
contrasts(dat$A) <- bayestestR::contr.equalprior(2)
contrasts(dat$B) <- bayestestR::contr.equalprior(3)

priors_factorial <- c(
  brms::prior(logistic(0, 0.75),  class = Intercept),
  brms::prior(normal(0, 0.354),   class = b, coef = A1),
  brms::prior(normal(0, 0.433),   class = b, coef = B1),
  brms::prior(normal(0, 0.433),   class = b, coef = B2),
  brms::prior(normal(0, 0.612),   class = b, coef = "A1:B1"),
  brms::prior(normal(0, 0.612),   class = b, coef = "A1:B2")
)
```

## Appendix B: No-Intercept Parameterisation {#sec-appendix-no-intercept .appendix}

In `brms`, specifying `0 + condition` removes the intercept and estimates separate coefficients for each condition level, so that $p_0 = g^{-1}(b_0)$, $p_1 = g^{-1}(b_1)$, and $\delta p = g^{-1}(b_1) - g^{-1}(b_0)$. Under this parameterization, the matched prior principle applies to each coefficient independently: placing $\text{Logistic}(0, 1)$ on both coefficients under the logit link induces independent $\text{Uniform}(0, 1)$ marginals, and the implied prior on $\delta p$ is a symmetric triangular distribution on $(-1, 1)$. Because the coefficient interpretation differs from the intercept-plus-slope parameterization, our recommended prior widths do not transfer directly -- each condition mean should receive a matched prior scaled to reflect beliefs
about plausible success probabilities (e.g., $\text{sd} = 0.75$), and the implied $\delta p$ should be checked via prior predictive simulation.

