---
title: "Principled Default Priors for Bayesian Hypothesis Testing in Binomial GLMs: A Matched-Prior Framework with Empirical Evaluation"
shorttitle: "Matched Priors for BF Testing in Binomial GLMs"
author:
  - name: Gidon T. Frischkorn
    orcid: 0000-0002-5055-9764
    email: gidon.frischkorn@psychologie.uzh.ch
    corresponding: true
    affiliations:
      - name: University of Zurich
        department: Department of Psychology
  - name: Joscha Dutli
    orcid: 0009-0009-5007-1281
    affiliations:
      - name: University of Zurich
        department: Department of Psychology
  - name: Philipp Musfeld
    orcid: 0000-0002-6539-0105
    affiliations:
      - name: University of Amsterdam
        department: Department of Psychology
      - name: University of Zurich
        department: Department of Psychology
  - name: Klaus Oberauer
    orcid: 0000-0001-8815-6189
    affiliations:
      - name: University of Zurich
        department: Department of Psychology
author-note:
  disclosures:
    study-registration: "This study was not preregistered."
    data-sharing: "Simulation code and analysis scripts are available at https://github.com/GidonFrischkorn/PriorsBinomial."
    conflict-of-interest: "The author(s) report having no conflicts of interest with respect to contents, authorship, or publication of this article."
    financial-support: "This research was supported by a grant from the Swiss National Science Foundation (SNSF) awarded to Gidon T. Frischkorn (Grant No. 208828)."
    authorship-agreements: |
      Gidon T. Frischkorn: Conceptualization, Methodology, Software, Formal Analysis, Visualization, Writing -- Original Draft, Writing -- Review & Editing.
      Joscha Dutli: Conceptualization, Methodology, Writing -- Review & Editing.
      Philipp Musfeld: Conceptualization,Methodology, Writing -- Review & Editing.
      Klaus Oberauer: Conceptualization, Supervision, Methodology, Writing -- Original Draft, Writing -- Review & Editing.
floatsintext: true
output-dir: "."
format:
  apaquarto-docx: default
  apaquarto-html: default
  apaquarto-pdf:
    documentmode: man
bibliography:
  - references.bib
keywords:
  - generalized linear models
  - prior specification
  - binomial models
  - Bayes Factor
  - Bayesian statistics
abstract: "Researchers using Bayesian hypothesis testing in binomial generalized linear models (GLMs) -- for instance, logistic regression for accuracy data -- need priors on regression coefficients, but no widely adopted default framework exists for Bayesian hypothesis testing in these models. The nonlinear link function creates a prior-link interaction: priors that appear reasonable on the coefficient scale can imply implausible distributions on the probability scale. We show that this problem has an elegant solution for CDF-based link functions: *matched priors* -- prior families that mirror the link function -- produce exactly uniform distributions on the probability scale, providing a maximally uninformative anchor. We embed this matched-prior principle in a three-step workflow and empirically evaluate it via simulation: (1) identify matched priors, (2) evaluate prior predictive distributions, (3) calibrate Bayes Factors including formal evaluation of consistency properties and comparison to the JZS/Cauchy default. For logistic regression, the matched Logistic(0, 0.75) intercept prior and Logistic(0, 0.25--0.30) effect prior provide well-calibrated Bayes Factors that exhibit information and Lindley consistency, and outperform the commonly used JZS/Cauchy default in both prior predictive plausibility and BF calibration for small effects, where the Cauchy prior frequently produces misleading evidence for $H_0$ when a genuine small effect is present. We further show that the workflow extends naturally to hierarchical models, and that the prior on the random-effects variance has negligible influence on Bayes factors for fixed effects. Ready-to-use `brms` code and scaling formulas for any contrast coding or predictor scale are provided."
abstract-public: "When psychologists use Bayesian analysis to test whether an experimental result is real --- for example, whether a memory training improved accuracy --- they must supply the software with a prior distribution that describes what effect sizes they consider plausible before seeing the data. For the most common type of outcome in laboratory experiments (how often participants respond correctly), no standard guidance has existed for choosing these distributions when using logistic regression. This matters because a poorly chosen prior can cause the software to report evidence against a genuine effect or to give unreliable conclusions when sample sizes are small. This paper provides clear default recommendations. The key idea is simple: because logistic regression uses a specific mathematical function to convert model parameters to probabilities, there is a natural family of distributions --- the logistic family --- that produces a completely neutral (uniform) starting assumption on the probability scale. We call these 'matched priors.' Starting from this neutral anchor, we show through extensive computer simulations which specific parameter values yield trustworthy results across the range of sample sizes and effect sizes typical in psychological research. The recommended defaults (Logistic(0, 0.75) for the baseline; Logistic(0, 0.25) for the effect) outperform the currently popular alternative for detecting small but real effects, while remaining reliable under the null. Ready-to-use code for the brms R package is provided so that researchers can apply the recommendations without technical expertise in Bayesian computation."
---

```{r}
#| label: setup
#| include: false

library(dplyr)
library(tidyr)
library(ggplot2)
library(ggridges)
library(viridis)
library(scales)
library(here)
library(knitr)

source(here("R", "link_functions.R"))
source(here("R", "prior_sampling.R"))
source(here("R", "plotting.R"))

# Goal 1: Prior predictive results
summaries_pp  <- readRDS(here("output", "prior_predictive_summaries.rds"))
draws_long    <- readRDS(here("output", "prior_predictive_draws.rds"))

# Goal 2: BF calibration results
summaries_bf  <- readRDS(here("output", "bf_calibration_summaries.rds"))

re_summaries <- tryCatch(
  readRDS(here("output", "bf_calibration_re_summaries.rds")),
  error = function(e) NULL
)
re_results_available <- !is.null(re_summaries)

# Goal 2 extended: information consistency results
consistency_summaries <- tryCatch(
  readRDS(here("output", "bf_calibration_consistency_summaries.rds")),
  error = function(e) NULL
)
consistency_continuous <- tryCatch(
  readRDS(here("output", "bf_calibration_consistency_continuous.rds")),
  error = function(e) NULL
)
consistency_available <- !is.null(consistency_summaries) && !is.null(consistency_continuous)

# Goal 2 extended: JZS comparison results
jzs_summaries <- tryCatch(
  readRDS(here("output", "bf_calibration_jzs_summaries.rds")),
  error = function(e) NULL
)
jzs_available <- !is.null(jzs_summaries)

# Factor labels
dist_labels <- c(cauchy = "Cauchy", normal = "Normal", logistic = "Logistic")

# Use prior_theme() from R/plotting.R for all figures
```

## Introduction

Consider a researcher studying working memory. She runs an experiment in which participants are tested on whether they can correctly identify previously shown items; her outcome is the number of correct responses out of a fixed number of trials. She wants to use Bayesian analysis to test whether a new encoding strategy improves accuracy. To do so, she must supply her software with a prior distribution — a mathematical statement of what effect sizes she considers plausible before seeing the data. No widely adopted guidance tells her what that prior should be when using logistic regression for this kind of count outcome. Use the wrong one, and her Bayes factor may systematically fail to detect genuine effects, or produce misleading evidence against them. This paper addresses that problem by deriving principled default priors from first principles, validating them through simulation, and providing ready-to-use code.

In psychology and related fields, we often have data that reflect counts of successes out of trials -- for instance, accuracy in a cognitive task or endorsement rates in a survey. These data follow a binomial distribution and are best modeled with a binomial generalized linear model (GLM):

$$k_i \sim \mathrm{Binomial}(n_i, p_i)$$

Here, the number of successes $k_i$ is the result of $n_i$ trials given the success probability $p_i$. To assess any effects on the number of successes, the success probability $p_i$ is related to a linear prediction term via a link function $g$ to accommodate the bounded nature of probabilities:

$$g(p_i) = b_0 + b_1 x_i$$

To translate the result of the linear prediction term back to the probability scale, the inverse of the link-function must be applied:

$$p_i = g^{-1}(b_0 + b_1 x_i)$$

For binomial models, the inverse of the link function $g^{-1}$ is either the logistic function (logistic regression), or the cumulative distribution function (CDF) of the standard normal distribution (probit regression).

In psychological experiments, participants typically complete multiple trials, so observations are nested within persons rather than independent. Between-participant differences in baseline success rates — some people are systematically faster, more accurate, or more conservative than others — are captured by a participant-level random intercept $u_j \sim \mathrm{Normal}(0, \sigma_u^2)$ in the linear predictor. Participants may further differ in the *magnitude* of the experimental effect they show, motivating an additional participant-level random slope $v_j \sim \mathrm{Normal}(0, \sigma_v^2)$ on the effect predictor $x_{ij}$; omitting random slopes when they are present in the data can substantially inflate false-positive rates in Bayesian hypothesis testing [@oberauer2022]. The present workflow addresses the hierarchical random-intercept model, as it introduces the key additional prior choice — the hyperprior on $\sigma_u$ — in the simplest setting. We present the formal hierarchical model specification in the [Hierarchical Binomial GLMs](#sec-hierarchical) section below and demonstrate the complete workflow for both the flat and the hierarchical random-intercept case.

When testing the hypothesis that an effect of interest exists, Bayesian hypothesis testing via Bayes Factors [@wagenmakers2018a; @schonbrodt2018; @heck2022] compares a model including the effect ($H_1$) against one constraining it to zero ($H_0$); the Bayes Factor $BF_{10}$ quantifies how much more likely the observed data are under $H_1$ than under $H_0$. When estimating the BF, we need to carefully think about the prior for the parameter of interest because the Bayes Factor critically depends on it. For models assuming normally distributed outcomes, @rouder2012 proposed principled default priors based on the JZS (Jeffreys--Zellner--Siow) framework.

Priors in Bayesian statistics can serve several distinct goals. Researchers may use them to *encode existing knowledge* --- prior research, meta-analytic summaries, or general knowledge about the distribution of effect sizes in a discipline --- so that this information informs the current analysis [@mikkola2024]. They may seek to *regularize* parameter estimates, keeping them within a plausible range and preventing overfitting, particularly when sample sizes are small [@gelman2008]. For parameters with bounded ranges --- such as probabilities --- the goal may be to *constrain* the model to logically valid values. And in settings with limited data, a well-chosen prior can serve to *stabilize* inference by supplying information where the data alone are uninformative, while being overwhelmed by the likelihood as sample size grows.

These goals are not always in harmony, and their relative importance depends on the inferential context. For *posterior estimation*, a weakly informative regularizing prior --- such as the Cauchy(0, 2.5) recommended by @gelman2008 --- is appropriate: it prevents extreme estimates while allowing the data to dominate at reasonable sample sizes. For *Bayes Factor hypothesis testing*, however, the prior on the effect parameter does more than regularize: it defines the alternative hypothesis $H_1$. Because the Bayes Factor $BF_{10}$ quantifies how much more likely the data are under $H_1$ than under $H_0$, and $H_1$ is characterized by the prior, the BF is highly sensitive to the prior choice. @hoijtink2021 demonstrated formally that null-hypothesis Bayes factors are highly sensitive to prior width: even modest changes in the prior scale can reverse the direction of evidence, making principled prior selection essential rather than optional [@schonbrodt2018]. Our matched-prior framework directly addresses this vulnerability for the binomial case. A prior that assigns substantial probability to implausibly large effects --- as the Cauchy default does --- biases BFs toward $H_0$, because those extreme effects repeatedly fail to materialize. Our contribution therefore focuses on two of the four prior goals: *encoding knowledge* about plausible effect sizes in behavioral research, and *constraining* the prior so that it implies interpretable distributions on the probability scale. Both goals are jointly addressed by the matched-prior framework developed below. For GLMs with nonlinear link functions, no such principled defaults have been widely adopted in psychology for Bayesian hypothesis testing.[^elicitation]

[^elicitation]: When domain expertise is available, structured elicitation methods [@hosack2026; @mikkola2024; @bockting2024; @stefan2021] can produce more targeted priors tailored to specific research questions. Our contribution is complementary: we provide principled *default* priors for use when such expertise is not available. Following the taxonomy of @mikkola2024, our workflow operates in the observable space, using the link function to bridge between prior distributions on the coefficient scale and their implications on the probability scale.

The difficulty for these models arises because the parameters of interest are estimated on the transformed, unbounded scale. Accordingly, priors must be expressed on the coefficient scale for model fitting, whereas our considerations regarding plausible effects --- and the constraints arising from the bounded nature of probability --- refer to the probability scale. This creates a prior-link interaction: a prior that appears reasonable on the regression-coefficient scale can produce implausible distributions on the probability scale [@gelman2008; @ghosh2018]. In addition, priors recommended for posterior estimation [e.g., the weakly informative Cauchy(0, 2.5) of @gelman2008] are typically too broad for hypothesis testing, biasing Bayes Factors toward the null.

We propose a three-step workflow for prior specification in GLMs that addresses this problem. The workflow is general; it applies whenever a link function mediates between parameter and probability scales. Here, we demonstrate it fully for binomial models, providing default priors and ready-to-use code for both the flat and the hierarchical random-intercept case.

We illustrate this approach with accuracy data from a hypothetical experiment with two conditions, in which we model the number of successes $k_i$ in $n_i$ trials as $k_i \sim \mathrm{Binomial}(n_i, p_i)$ with the linear model $g(p_i) = b_0 + b_1 x_i$, where $p_i$ is the success probability, $g$ is either the logit or probit link, and $x_i$ is a single discrete predictor with two levels. We code $x_i$ with a sum-to-zero contrast $x_i \in \{+1, -1\}$, so that the difference between conditions on the probability scale is $\delta p = g^{-1}(b_0 + b_1) - g^{-1}(b_0 - b_1)$, where $g^{-1}(\cdot)$ denotes the inverse link function that maps the linear predictor back to the probability scale. Before presenting the workflow and its empirical validation, we situate the matched-prior approach within existing frameworks for prior specification in GLMs.

## Related Work {#sec-related-work}

### The G-Prior Tradition: From Linear Models to GLMs

@liang2008 established mixtures of g-priors as a principled, analytically tractable framework for default priors in linear regression. In this approach, $\beta \sim \mathrm{Normal}\!\left(0,\, g \cdot (X'X)^{-1}/\phi\right)$, so the prior covariance scales with the data covariance matrix through the hyperparameter $g$. Rather than fixing $g$ — a choice that remains largely arbitrary — the mixture approach treats $g$ as random and marginalises over it. The key insight is that placing an Inverse-Gamma prior on $g$ (equivalently, a Gamma prior on the precision $1/g$) and integrating out $g$ yields Cauchy-distributed marginals for $\beta$: a Normal prior with Gamma-distributed variance is a scale mixture of normals that integrates to a Cauchy. The Zellner--Siow prior [@liang2008] formalises this by setting $g \sim \mathrm{Inverse\text{-}Gamma}\!\left(\tfrac{1}{2},\, \tfrac{n}{2}\right)$, giving multivariate Cauchy marginals for $\beta$ centred at zero. This Normal-variance-mixture construction provides the formal justification for the Cauchy default priors widely adopted in Bayesian hypothesis testing [e.g., @rouder2009]. Crucially, @rouder2012 showed that this framework achieves genuine design independence for linear ANOVA: by placing priors on *standardized* effects $\delta = \beta / \sigma$ — that is, setting $\beta \mid g, \sigma \sim \mathrm{Normal}(0,\, g\sigma^2 I)$ combined with a Jeffreys prior $\pi(\mu, \sigma^2) \propto 1/\sigma^2$ — the prior covariance reduces to $G = gI$, free of the Fisher information matrix altogether. Under sum-to-zero or orthonormal contrast coding, the resulting Bayes factor depends only on $R^2$ and is invariant to linear transformations of the outcome and to the specific design. 

The design independence achieved by @rouder2012 rests on two structural features of Gaussian linear models that do not carry over to binomial GLMs. First, linear models have a free scale parameter $\sigma^2$: by expressing effects on the standardized scale $\delta = \beta / \sigma$, both the prior and the resulting Bayes factor become invariant to measurement units of the outcome. Binomial GLMs have no separate variance parameter — the variance $p(1-p)$ is entirely determined by the mean — so there is no $\sigma$ to standardize against. Second, the identity link in linear models means a scale-invariant prior on $\beta$ translates directly into a scale-invariant distribution over outcome differences. With a nonlinear link function, the mapping from $\beta$ to the probability scale is distorted by the link's curvature in a way that depends on the baseline probability, and therefore on the design. 

To address this issue, @bove2011 and @liclyde2018 extended the g-Prior framework to GLMs, replacing $(X'X)$ with the Fisher information matrix evaluated at a reference point. For binomial GLMs, this yields approximate marginal likelihoods via Laplace approximation, enabling principled BF estimation. @sekulovski2025 provide a recent large-scale comparison of Bayesian variable selection methods for logistic regression, including g-prior-based approaches. @porwalraftery2022 provide a broader comparison of Bayesian and frequentist model selection methods, including g-prior-based approaches, across a range of regression settings.

However, the GLM extensions [@bove2011; @liclyde2018] substitute the Fisher information matrix $W = \mathrm{diag}\!\left(p_i(\hat{\beta}_0)[1 - p_i(\hat{\beta}_0)]\right)$, evaluated at a reference prediction, for $(X'X)$. This substitution reintroduces design dependence: the IRLS weight matrix changes with the predicted probabilities at $\hat{\beta}_0$, so the prior's implied scale on the probability scale changes with every dataset — no universal recommendation like "$\mathrm{Logistic}(0, 0.75)$" is possible. Crucially, even a hypothetical design-independent g-prior for GLMs would leave a second problem unresolved: it would not inform us what the prior implies on the probability scale. A prior that appears reasonable on the unbounded coefficient scale can assign substantial mass to probabilities near 0 or 1, implausible for most behavioral experiments [@gelman2008; @ghosh2018]. This *prior--link interaction* is precisely the problem the matched-prior framework addresses: by anchoring prior choice on the probability scale via the probability integral transform, a single scale recommendation applies across all datasets with the same link function, and prior predictive simulation provides a directly interpretable diagnostic. We provide a mathematical comparison in the supplementary materials. We also note that g-priors are not readily available in `brms` [@burkner2017; @burkner2018] or Stan, limiting their practical uptake in psychological research.

### Conditional Mean Priors and Prior Elicitation

@bedrick1996 introduced conditional mean priors for GLMs, in which the researcher specifies prior distributions on $P(Y=1 \mid x)$ at selected design points; the prior on $\beta$ is then implied by a change-of-variables argument. This approach is principled, directly interpretable on the probability scale, and has been extended in @christensen2010 and more recently by @hosack2026 for a broad family of GLMs. Simulation-based approaches to prior elicitation [@bockting2024] and comprehensive reviews of elicitation methodology [@mikkola2024] further enrich this tradition.

The matched-prior approach complements rather than competes with conditional mean priors. Conditional mean priors require the researcher to specify design points and probability distributions -- they are semi-informative and elicitation-based, suited to situations where genuine prior knowledge exists. The matched prior is a *default* requiring no elicitation, appropriate when such knowledge is unavailable. Additionally, conditional mean priors were developed primarily for parameter estimation rather than for Bayes factor hypothesis testing, which imposes the additional requirement of a proper prior on $\beta$ under $H_1$ with a point mass at zero under $H_0$. The matched-prior workflow is designed specifically for this hypothesis-testing context. The insight that the probability scale provides an interpretable basis for prior specification is shared by both approaches: while computational fitting requires priors to be expressed on the coefficient scale, the *motivation* for prior choice can -- and we argue *should* -- come from the probability scale.

## The Three-Step Workflow {#sec-workflow}

Prior specification for GLMs with nonlinear link functions requires bridging the gap between the coefficient scale (where priors are set) and the probability scale (where effects are interpretable). Our workflow instantiates the principled Bayesian workflow for cognitive science [@schad2021] in the specific context of prior selection for Bayes factor testing, and proceeds through three complementary steps:

**Step 1: Identify matched priors.** Using the probability integral transform, we identify prior families whose cumulative distribution function (CDF) matches the inverse link function. These matched priors produce exactly uniform distributions on the probability scale, serving as principled, maximally uninformative anchors from which the prior scale can be adjusted to reflect specific prior beliefs.

**Step 2: Prior predictive simulation.** Building on the matched priors, we draw samples from the joint prior on all coefficients while varying the prior scale, transform through the inverse link, and examine the resulting distributions on the probability scale. This reveals what effect sizes and baseline rates the prior treats as plausible -- information that is not apparent from the prior on the coefficient scale alone [@gelman2020].

**Step 3: Bayes Factor calibration.** Finally, we simulate data under known effects and evaluate how the strength of evidence varies across prior specifications. This verifies that the priors recommended based on the prior predictive simulations yield well-calibrated Bayes Factors -- neither systematically favouring $H_0$ nor $H_1$ -- and that they satisfy formal properties of principled Bayesian model comparison.

Each step can be applied to any GLM family. Here, we demonstrate the full workflow for binomial models.

## Step 1: The Matched Prior Principle {#sec-matched-prior}

The key mathematical result underlying this approach is the following: for any continuous cumulative distribution function (CDF) $F$ with location 0 and scale 1, if $B_0 \sim F(\cdot;\, 0, 1)$, then $p = F(B_0) \sim \mathrm{Uniform}(0, 1)$ [@casella2002]. This follows from the probability integral transform: applying a distribution's own CDF to a draw from that distribution maps the draw to its percentile rank within the distribution, and percentile ranks are, by definition, uniformly distributed. Consequently, a $\mathrm{Logistic}(0, 1)$ prior on $b_0$ combined with the logistic inverse link $g^{-1} = F_\mathrm{logistic}$ yields $p = F_\mathrm{logistic}(b_0) \sim \mathrm{Uniform}(0, 1)$; the same logic gives $\mathrm{Normal}(0, 1) \rightarrow \mathrm{Uniform}(0, 1)$ under the probit link.

Stated in practical terms: when the prior distribution on the coefficient scale is matched exactly to the link function --- same family and same parameters (location 0, scale 1) --- it produces a flat (uninformative) prior on the probability scale. For binomial GLMs that either use the logit or the probit link, the matched prior is the Logistic(0, 1) or Normal(0, 1), respectively:

| Link   | Inverse link $g^{-1}$ | Matched prior | Implied prior on $p$ |
|--------|-----------------------|---------------|----------------------|
| logit  | logistic CDF | $\mathrm{Logistic}(0, 1)$ | $\mathrm{Uniform}(0, 1)$ |
| probit | normal CDF | $\mathrm{Normal}(0, 1)$ | $\mathrm{Uniform}(0, 1)$  |

: Matched priors for binomial GLMs. By the probability integral transform, matching the prior family to the link function produces a uniform prior on the probability scale. {#tbl-matched-priors}

In other words, when we transform a $\mathrm{Logistic}(0, 1)$ prior on the coefficient scale back to the probability scale via the inverse logit function, the result is a uniform distribution on $(0, 1)$; the same holds for a $\mathrm{Normal}(0, 1)$ prior under the probit link. The matched prior at scale 1 is therefore maximally uninformative on the probability scale -- a principled anchor from which we can obtain desired priors through scaling. Conversely, scaling the matched prior with $\mathrm{sd} > 1$ concentrates probability mass toward the extreme values 0 and 1, reflecting a typically unreasonable assumption that extreme baseline probabilities are particularly likely. Scaling with $\mathrm{sd} < 1$ concentrates mass away from 0 and 1, producing a regularized (i.e., shrunk away from extreme values) intercept prior suited to experiments where floor and ceiling performance are unlikely. This principle generalises to any model with a CDF-based link function (see [Applying the Workflow Beyond Binomial Models](#sec-extensions)).

## Hierarchical Binomial GLMs {#sec-hierarchical}

Step 1 identifies matched priors for the fixed-effect coefficients $b_0$ and $b_1$. The simulation steps that follow (Steps 2 and 3) apply these priors within a hierarchical model, which introduces one further prior component: the hyperprior on the random-effects standard deviation $\sigma_u$. We therefore formalize the complete model specification here before proceeding.

Psychological experiments typically yield nested data: each participant contributes multiple trials, so observations are not independent across participants. We therefore extend the flat binomial GLM to a random-intercept model. Using the general link function $g$ introduced above, the complete hierarchical model is:

$$y_{ij} \sim \mathrm{Binomial}(n_{ij},\, p_{ij})$$

$$g(p_{ij}) = b_0 + u_j + b_1 x_{ij}$$

where $j = 1, \ldots, J$ indexes participants and $i = 1, \ldots, n_j$ indexes trials within participant $j$. The participant-level random intercept $u_j \sim \mathrm{Normal}(0, \sigma_u^2)$ captures between-participant variability in baseline success probability, and $\sigma_u$ is the random-effects standard deviation. The link functions are the same as in the flat model: $g(p) = \log(p/(1-p))$ for logistic regression and $g(p) = \Phi^{-1}(p)$ for probit regression, where $\Phi^{-1}$ is the quantile function of the standard normal distribution.

The complete prior specification is:

- **Fixed effects:** $b_0 \sim \mathrm{Logistic}(0, \mathrm{sd}_{b_0})$ (logit) or $\mathrm{Normal}(0, \mathrm{sd}_{b_0})$ (probit), matched to the link function; $b_1 \sim \mathrm{Logistic}(0, \mathrm{sd}_{b_1})$ or $\mathrm{Normal}(0, \mathrm{sd}_{b_1})$
- **Random effects:** $u_j \sim \mathrm{Normal}(0, \sigma_u^2)$ with hyperprior $\sigma_u \sim \mathrm{Exponential}(\lambda)$

Recommended scale values — $\mathrm{sd}_{b_0} \approx 0.75$, $\mathrm{sd}_{b_1} \in [0.25, 0.30]$, and $\lambda = 4$ — are derived from the three-step workflow in Steps 2 and 3 below.

The Bayes Factor for the fixed effect $b_1$ compares:

$$H_1:\; b_1 \neq 0 \quad \text{vs.} \quad H_0:\; b_1 = 0$$

Under both models, the random intercepts $u_j$ are present and integrated out in the marginal likelihoods used to compute the BF. Because $\sigma_u$ appears symmetrically in both marginal likelihoods, its hyperprior largely cancels from the ratio, leaving the BF dominated by the fixed-effect signal-to-noise ratio. This *nuisance-parameter robustness* means that the choice of hyperprior on $\sigma_u$ has little influence on BFs for the fixed effect $b_1$, provided a reasonable weakly informative prior is used.

This robustness is practically valuable: researchers do not need to fine-tune the hyperprior on $\sigma_u$ to obtain reliable BFs for the fixed effect. Step 3 confirms this empirically (see @fig-bf-violin-re): even when the hyperprior is dramatically misspecified — for example, using `student_t(3, 0, 2.5)`, which concentrates mass near $\sigma_u \approx 2.76$ on the logit scale — the BF for the fixed effect is virtually unchanged relative to the recommended `exponential(4)` hyperprior.

## Step 2: Prior Predictive Simulation {#sec-prior-predictive}

We simulated prior predictive distributions for Binomial GLMs across a factorial design crossing two link functions (logit, probit), three prior families for both intercept and effect coefficients (Normal, Logistic, and Cauchy), and a range of prior scale values. The Normal and Logistic families were included because each matches one of the two link functions; the Cauchy was included because @rouder2012 proposed Cauchy-distributed default priors for linear models, and one of us [@oberauer2019] has proposed them also for logistic models. The intercept prior SD ($\mathrm{sd}_{b_0}$) ranged from 0.50 to 1.50; the effect prior SD ($\mathrm{sd}_{b_1}$) ranged from 0.10 to 2.00, yielding 720 factorial combinations in total (see Supplementary Materials for the complete design table). For each condition, we drew 10,000 samples from the joint prior on $(b_0, b_1)$ and computed the implied distributions on the probability scale.[^software]

[^software]: All simulations were conducted in R 4.5.2 [@r2023] using the SimDesign package [version 2.25; @chalmers2020]. Bayesian models for Step 3 were fitted with brms [version 2.23.0; @burkner2017; @burkner2018] using CmdStan [version 2.38.0; @stan2023] via the cmdstanr interface [version 0.9.0; @gabry2024]. Marginal likelihoods for bridge sampling were estimated with the bridgesampling package [version 1.2.1; @gronau2017]. All simulation code and analysis scripts are available at <https://github.com/GidonFrischkorn/PriorsBinomial>.

### Intercept Prior {#sec-pp-intercept}

The intercept $b_0$ determines the baseline success probability before any experimental manipulation. A poorly chosen intercept prior can place substantial mass near $p = 0$ or $p = 1$, predicting that participants perform at floor or ceiling even before any experimental effect is applied.

```{r}
#| label: fig-1-p-intercept
#| fig-cap: "Prior predictive distribution of success probability
#|   p = $g^{-1}(b_0)$ for each intercept prior SD (rows) and link function
#|   (columns). At $\\mathrm{sd}_{b_0} = 1.0$, the matched prior (Logistic for logit,
#|   Normal for probit) produces a perfectly uniform distribution. Reducing
#|   $\\mathrm{sd}_{b_0}$ to 0.75 pulls mass away from floor and ceiling. Dashed
#|   curves show exact analytical densities."
#| fig-height: 10
#| fig-width: 10

fig1_draws <- draws_long |>
  filter(dist_b1 == "logistic", sd_b1 == 0.25)

plot_p_intercept_ridges(fig1_draws, sd_b1_focus = 0.25)
```

@fig-1-p-intercept confirms the analytical result from the previous step: at $\mathrm{sd}_{b_0} = 1.0$ with the matched prior, the distribution of $p$ is uniform. At $\mathrm{sd}_{b_0} = 0.75$, mass is pulled away from 0 and 1, appropriate for experiments avoiding floor and ceiling effects. Misfit priors (e.g., Cauchy for either link) produce heavier tails. We recommend $\mathrm{sd}_{b_0} = 0.75$ as the default, that is, $\mathrm{Logistic}(0, 0.75)$ for the logit link and $\mathrm{Normal}(0, 0.75)$ for the probit link (see Supplementary Figure S1 for a quantification of how much probability mass is assigned to the floor/ceiling regions by each prior specification).

### Effect Prior {#sec-pp-effect}

The prior on effect parameter $b_1$ determines how much prior mass falls on small versus large effects. Due to the nonlinear link, we must compute the implied prior on $\delta p$ by drawing from the joint prior on $(b_0, b_1)$ and transforming through the inverse link. Concretely, the implied prior on $\delta p$ is computed as $\delta p = g^{-1}(b_0 + b_1) - g^{-1}(b_0 - b_1)$. @fig-3-delta-p shows how the choice of $\mathrm{sd}_{b_1}$ translates into implied experimental effects in $\delta p$, holding the intercept prior fixed at the matched recommendation ($\mathrm{Logistic}(0, 0.75)$ for logit; $\mathrm{Normal}(0, 0.75)$ for probit).

```{r}
#| label: fig-3-delta-p
#| fig-cap: "Prior predictive distribution of $\\delta p$ (signed effect size on
#|   the probability scale) for each effect prior SD (facets), colour-coded by
#|   prior family (ridges). Shown for the logit link with matched Logistic(0, 0.75)
#|   intercept prior. At $\\mathrm{sd}_{b_1} = 0.25$, Normal and Logistic priors concentrate
#|   mass on effects up to about 20 percentage points. The Cauchy assigns
#|   substantial probability to effects exceeding 50 percentage points even at
#|   this moderate scale. Dotted lines mark +/-0.10, +/-0.20, +/-0.30."
#| fig-height: 12
#| fig-width: 14

fig2_draws <- draws_long |>
  filter(
    link == "logit",
    dist_b0 == "logistic",
    sd_b0 == 0.75
  )

plot_delta_p_ridges(fig2_draws, sd_b0_focus = 0.75)
```

The Cauchy prior stands apart at every scale: even at $\mathrm{sd}_{b_1} = 0.25$, its heavy tails assign substantial probability to $|\delta p| > 0.50$, making it a poor default. Note that effects of $|\delta p| \approx 1.0$ would require perfect accuracy in one condition and zero correct responses in the other -- a scenario that is virtually impossible in practice.

To quantify this more precisely, we computed the proportion of the prior predictive mass that falls within the plausible range $0.10 \leq |\delta p| \leq 0.90$ for an intercept-only model at $\mathrm{sd}_{b_1} = 0.25$. For the matched Logistic(0, 0.25), approximately 90% of the mass falls below $|\delta p| = 0.20$. In contrast, the $\mathrm{Cauchy}(0, \sqrt{2}/2)$ -- the JZS default from the BayesFactor package -- assigns substantial probability to extreme effects even on the probability scale, reflecting a poor fit between the prior's implicit assumptions and what is plausible in behavioral research. We return to this comparison in the BF calibration analysis below.

Within the substantively appropriate range, Normal and Logistic priors behave very similarly. We consider a prior *plausible* for behavioral research if the bulk of its prior predictive mass falls on effect sizes that could realistically be observed in the paradigm of interest; for typical accuracy and detection tasks, this means assigning the majority of prior probability to $|\delta p| < 0.30$. We use the 90th percentile of $|\delta p|$ as the key summary; 90% of the prior probability mass is assigned to effect sizes smaller than that value: at $\mathrm{sd}_{b_1} = 0.25$, the 90th percentile is near 0.20 -- a large effect by the standards of most psychological research. Values up to $\mathrm{sd}_{b_1} = 0.50$ remain defensible when very large effects are plausible. We recommend $\mathrm{sd}_{b_1} \in [0.25, 0.30]$ as the default range when continuous predictors are scaled to $\mathrm{SD} = 1$, and categorical predictors are sum-to-zero coded with $\pm 1$.

### Random Effects Hyperprior {#sec-pp-re}

```{r}
#| label: setup-re-pp
#| include: false

re_pp_summaries <- tryCatch(
  readRDS(here("output", "prior_predictive_re_summaries.rds")),
  error = function(e) NULL
)
re_pp_draws <- tryCatch(
  readRDS(here("output", "prior_predictive_re_draws.rds")),
  error = function(e) NULL
)

re_pp_available <- !is.null(re_pp_summaries) && !is.null(re_pp_draws)
```

For hierarchical models, the hyperprior on the random-effects SD $\sigma$ controls between-subject variability. Candidate hyperpriors should limit $\sigma$ to positive real values. Any principled hyperprior for a SD parameter should have three properties: (a) mode at zero, expressing a preference for parsimony; (b) monotonically decreasing density, so that larger variability receives progressively less prior support; and (c) a scale calibrated to domain knowledge about plausible between-subject variability [@gelman2006; @simpson2017]. Some distributions, like the Exponential, fulfill these properties naturally; others, such as the Student-*t* or Normal distribution, can be truncated to fulfill them (resulting in a half-Student-*t* or half-Normal distribution).

To assess the effects of distribution family and scale, we compared four hyperprior families -- Exponential, Gamma(shape = 2), Half-Normal, and Half-Student-*t*(df = 3) -- at five values of the hyperprior mean (0.10, 0.25, 0.50, 0.75, and 1.00; see Supplementary Materials for details). The Half-Student-*t* is included primarily as a reference point for comparison with the `brms` default `student_t(3, 0, 2.5)` (mean $\approx$ 2.76), which corresponds to the largest scale level. We included the Gamma because it is a widely used hyperprior family in Bayesian modelling practice, and because its positive mode (when shape > 1) provides a deliberate contrast to the three other families.

```{r}
#| label: fig-re-1-p-intercept
#| fig-cap: "Subject-level success probabilities by hyperprior family and hyperprior
#|   mean (rows: mean = 0.10, 0.25, 0.50, 0.75, 1.00; columns: random-effects
#|   structure). At each mean value, all four families produce nearly identical
#|   distributions; the primary driver is the scale parameter, not the family.
#|   Dashed vertical lines mark p = 0.05 and p = 0.95."
#| fig-height: 10
#| fig-width: 12
#| eval: !expr re_pp_available

re_pp_draws |>
  group_by(hp_family, hp_mean, re_structure) |>
  slice_sample(n = 5000) |>
  ungroup() |>
  mutate(hp_mean_label = paste0("mu == ", hp_mean)) |>
  ggplot(aes(x = p_intercept, y = hp_family,
             fill = hp_family, colour = hp_family)) +
  geom_density_ridges(stat = "binline", bins = 80,
                      alpha = 0.5, scale = 0.9) +
  geom_vline(xintercept = c(0.05, 0.95), linetype = "dashed",
             colour = "grey40", linewidth = 0.5) +
  facet_grid(hp_mean_label ~ re_structure,
             labeller = labeller(hp_mean_label = label_parsed)) +
  scale_fill_viridis_d(option = "D", end = 0.85, guide = "none") +
  scale_colour_viridis_d(option = "D", end = 0.85, guide = "none") +
  scale_x_continuous(
    expression("Success probability" ~ italic(p)),
    limits = c(0, 1), expand = expansion(mult = c(0.02, 0.02))
  ) +
  scale_y_discrete("Family") +
  prior_theme() +
  theme(strip.text.y = element_text(size = 10))
```

When matched by their mean value, all four families produce nearly identical prior predictive distributions for subject-level success probabilities (@fig-re-1-p-intercept). The primary driver is the scale, not the family. At a hyperprior mean of 0.25, subject-level probabilities are well concentrated away from floor and ceiling. In sum, we recommend $\mathrm{Exponential}(4)$ (mean = 0.25) as the default because it requires only one parameter. In `brms` [@burkner2017; @burkner2018] -- an R package for Bayesian multilevel modelling -- the default hyperprior `student_t(3, 0, 2.5)` (mean $\approx$ 2.76) spreads subject-level probabilities too heavily toward 0 and 1 for typical experiments (see Supplementary Figure S2 for effect heterogeneity).


## Step 3: Bayes Factor Calibration {#sec-bf-calibration}

The prior predictive analysis established which prior families and scales produce plausible distributions on the probability scale. We now verify that these priors also yield well-calibrated Bayes Factors -- that is, Bayes Factors that correctly favour $H_1$ when an effect exists and $H_0$ when it does not. Because the Bayes Factor reflects how much the data update our prior beliefs, a prior that places too much mass on implausibly large effects will be penalised even when a true effect exists, producing Bayes Factors that systematically favour $H_0$.

### Properties of Matched-Prior Bayes Factors {#sec-bf-properties}

Before presenting simulation results, we briefly characterise the theoretical properties of matched-prior Bayes factors against the desiderata for Bayesian model selection criteria [@bayarri2012]. @pfister2024 demonstrated through systematic comparison of BayesFactor and JASP implementations that BF estimates can diverge substantially across software packages, particularly for small effects and moderate sample sizes, raising concerns about reproducibility. The theoretical properties evaluated here — information and Lindley consistency — provide formal guarantees that matched-prior Bayes factors are well-behaved regardless of which compliant implementation is used.

**Information consistency.** A Bayesian test is information-consistent if the Bayes factor in favour of the true model grows without bound as evidence accumulates: $BF_{10} \to \infty$ as the effect size is fixed and $n \to \infty$, and similarly $BF_{10} \to \infty$ as the true effect size grows with fixed $n$. Fixed-scale normal (and logistic) priors are information-consistent: the prior density at zero is finite and fixed, while the posterior density at zero decreases to zero as evidence accumulates, driving $BF_{10} = f_{\mathrm{prior}}(0) / f_{\mathrm{post}}(0) \to \infty$. This contrasts with improper or overly diffuse priors, which can produce inconsistent behavior. We demonstrate this empirically below.

**Model selection consistency (Lindley consistency).** As $n \to \infty$ with a true nonzero effect fixed, the Bayes factor correctly identifies $H_1$ with probability approaching 1 (and analogously for $H_0$). This follows from standard Bayesian consistency results for proper priors with sufficient support at the true parameter value -- a condition satisfied by matched priors.

**Measurement invariance.** Under sum-to-zero $\pm 1$ contrast coding, the matched prior on $b_1$ implies a symmetric prior predictive distribution for $\delta p$: reversing the sign of the predictor reverses the sign of $b_1$ but leaves $|\delta p|$ unchanged. The recommended priors are therefore invariant to the direction of coding, so that the BF does not depend on which group is arbitrarily labelled as $+1$ or $-1$.

**Predictive matching.** The matched-prior workflow explicitly evaluates priors through their prior predictive distributions on the probability scale (Step 2), ensuring that the prior assigns the majority of its mass to scientifically plausible effect sizes. In the terminology of @bayarri2012, this constitutes a form of predictive matching: the prior is calibrated so that the prior predictive distribution reflects the researcher's actual uncertainty about plausible outcomes. The following simulation operationalizes all four criteria empirically.

### Simulation Design

We simulated data under the logit link with true effects ranging from $b_1 = 0$ (null) to $b_1 = 0.50$, crossing three effect prior families (Normal, Logistic, Cauchy) at two scales ($\mathrm{sd}_{b_1} = 0.25$ and $0.50$), two intercept prior families, three sample sizes (30, 60, 100), and two trial counts (20, 50). The two intercept prior families were a matched Logistic(0, 0.75) -- whose family mirrors the logit link -- and a misfit Normal(0, 0.75), which does not. A second simulation held fixed-effects priors at the recommended matched configuration and varied only the random-effects hyperprior across three candidates: Student-*t*(3, 0, 2.5), Gamma(2, 4), and Exponential(4). Bayes Factors were computed via the Savage-Dickey density ratio [@wagenmakers2018; @ly2016] -- the ratio of the prior density to the posterior density at $b_1 = 0$, which estimates how much the data have shifted belief away from the null.[^sdr]

[^sdr]: The Savage-Dickey ratio is valid when two conditions hold: (a) the prior on $b_1$ under $H_1$ is continuous (no point masses), and (b) the null hypothesis specifies $b_1 = 0$ as a nested point constraint within the full model [@ly2016]. Both conditions are satisfied here: the matched logistic and normal priors are continuous, and $H_0 : b_1 = 0$ is a point-mass restriction within the full model. When the posterior is far from zero, bridge sampling is preferred to avoid KDE underflow at the null, as used in the consistency analysis below. We present results for small to moderate effects ($b_1 \in \{0, 0.10, 0.20\}$) in the most data-sparse scenario ($N = 30$, 20 trials), where prior sensitivity is most visible; results for the full range of effect sizes and sample sizes are available in the Supplementary Materials.

```{r}
#| label: setup-bf-raw
#| include: false

bf_raw <- tryCatch({
  results_dir <- here("output", "Simulation_BFCalibration")
  cond_files <- list.files(results_dir, pattern = "^BFCalib_Cond-",
                           full.names = TRUE)
  raw_rows <- lapply(cond_files, function(f) {
    d <- readRDS(f)
    bf10_vec <- d$results[, "BF10"]
    cbind(
      as.data.frame(as.list(d$condition)),
      data.frame(BF10 = bf10_vec, log10_BF10 = log10(bf10_vec))
    )
  })
  bind_rows(raw_rows)
}, error = function(e) NULL)
bf_raw_available <- !is.null(bf_raw)

bf_raw_re <- tryCatch({
  results_dir_re <- here("output", "Simulation_BFCalibration_RE")
  cond_files_re <- list.files(results_dir_re, pattern = "^BFCalib_RE_Cond-",
                              full.names = TRUE)
  raw_rows_re <- lapply(cond_files_re, function(f) {
    d <- readRDS(f)
    bf10_vec <- d$results[, "BF10"]
    cbind(
      as.data.frame(as.list(d$condition)),
      data.frame(BF10 = bf10_vec, log10_BF10 = log10(bf10_vec))
    )
  })
  bind_rows(raw_rows_re)
}, error = function(e) NULL)
bf_raw_re_available <- !is.null(bf_raw_re)

# JZS and Consistency simulations use qs2 format; load from pre-processed flat files
# (generated by scripts/04c and 04d)
bf_raw_jzs <- tryCatch(
  readRDS(here("output", "bf_calibration_jzs_raw.rds")),
  error = function(e) NULL
)
bf_raw_jzs_available <- !is.null(bf_raw_jzs)

bf_raw_consistency <- tryCatch(
  readRDS(here("output", "bf_calibration_consistency_raw.rds")),
  error = function(e) NULL
)
bf_raw_consistency_available <- !is.null(bf_raw_consistency)
```

### Transparency and Openness

All simulation scripts, model-fitting code, and analysis code reported in this paper are publicly available at <https://github.com/GidonFrischkorn/PriorsBinomial>. No empirical data were collected; all results are based on computer simulations. This study was not preregistered. An earlier, more tutorial-oriented version of this manuscript is publicly available as a preprint [@frischkorn2026preprint]; that version was submitted to *Advances in Methods and Practices in Psychological Science* and not accepted. The present manuscript is a substantially revised and extended version produced in response to the recommendations from that review. Software versions used: R 4.5.2 [@r2023]; SimDesign 2.25 [@chalmers2020]; brms 2.23.0 [@burkner2017; @burkner2018]; CmdStan 2.38.0 [@stan2023] via cmdstanr 0.9.0 [@gabry2024]; bridgesampling 1.2.1 [@gronau2017]. A complete session information block is provided in the Supplementary Materials.

### Effect of Prior Family and Width

```{r}
#| label: fig-bf-violin-family
#| fig-cap: "Distribution of log10(BF10) by effect prior family and width for
#|   N = 30 subjects and 20 trials per subject. Rows (top to bottom): true effect
#|   b1 = 0 (null), b1 = 0.10, b1 = 0.20. Columns: prior SD (0.25 or 0.50).
#|   The choice of prior family has minimal impact; prior width shifts the
#|   distribution modestly. Black dots indicate medians. Dashed lines mark
#|   BF10 = 3 and BF10 = 1/3."
#| fig-height: 7
#| fig-width: 10
#| eval: !expr bf_raw_available

bf_raw |>
  filter(
    dist_b0 == "logistic",
    true_b1 %in% c(0, 0.10, 0.20),
    n_subjects == 30,
    n_trials == 20
  ) |>
  mutate(
    dist_b1 = factor(dist_b1, levels = c("cauchy", "normal", "logistic"),
                     labels = c("Cauchy", "Normal", "Logistic")),
    true_b1_label = factor(paste0("b[1] == ", true_b1)),
    sd_b1_label   = factor(paste0("sd[b[1]] == ", sd_b1))
  ) |>
  ggplot(aes(x = dist_b1, y = log10_BF10, fill = dist_b1)) +
  geom_hline(yintercept = c(log10(3), log10(1/3)), linetype = "dashed",
             colour = "grey50", linewidth = 0.5) +
  geom_hline(yintercept = 0, colour = "grey30", linewidth = 0.4) +
  geom_violin(alpha = 0.6, scale = "width", width = 0.8) +
  stat_summary(fun = median, geom = "point", size = 2, colour = "black") +
  facet_grid(true_b1_label ~ sd_b1_label, labeller = label_parsed) +
  scale_fill_viridis_d("Prior family", option = "D", end = 0.85) +
  scale_y_continuous(
    expression(BF[10]),
    breaks = log10(c(1/100, 1/10, 1/3, 1, 3, 10, 100)),
    labels = c("1/100", "1/10", "1/3", "1", "3", "10", "100")
  ) +
  coord_cartesian(ylim = c(-2.5, 2.5)) +
  labs(x = "Effect prior family") +
  prior_theme()
```

The choice of effect prior family has only very little impact on the BF (@fig-bf-violin-family). All three families produce nearly identical BF distributions. The prior width has a modest effect: narrower priors ($\mathrm{sd} = 0.25$) yield slightly stronger evidence for $H_1$ when the effect is present, particularly at $b_1 = 0.2$, though the difference is small relative to variability across replications.

### Comparison to the JZS/Cauchy Default

To assess whether the matched prior outperforms the JZS/Cauchy default, we ran a factorial simulation with two effect-prior conditions: the matched Logistic(0, 0.25) and the JZS Cauchy(0, $\sqrt{2}/2$) used in the `BayesFactor` package [@rouder2012]. These crossed four true effect sizes ($b_1 \in \{0, 0.10, 0.20, 0.50\}$), three sample sizes ($N \in \{30, 60, 100\}$), and two trial counts (20 or 50 per subject), yielding 48 conditions. Both prior conditions used the matched Logistic(0, 0.75) intercept prior and Exponential(4) random-effects hyperprior. We ran 100 replications per condition and estimated Bayes Factors via the Savage-Dickey density ratio.

```{r}
#| label: fig-bf-jzs-comparison
#| fig-cap: "Comparison of the matched Logistic(0, 0.25) prior against the
#|   JZS/Cauchy default (Cauchy(0, sqrt(2)/2), as used in the BayesFactor package)
#|   for N = 30, 20 trials. Rows show true effect b1 = 0 (null), 0.10, 0.20, 0.50.
#|   For small effects (b1 = 0.10), the matched prior yields substantially higher
#|   power and lower rates of false H0 support than the Cauchy default; the two
#|   priors converge for medium-to-large effects (b1 >= 0.20).
#|   Dashed lines mark BF10 = 3 and BF10 = 1/3."
#| fig-height: 8
#| fig-width: 8
#| eval: !expr bf_raw_jzs_available

bf_raw_jzs |>
  filter(n_subjects == 30, n_trials == 20) |>
  mutate(
    prior_label = case_when(
      dist_b1 == "logistic" ~ "Matched\nLogistic(0, 0.25)",
      dist_b1 == "cauchy"   ~ "JZS/Cauchy\nCauchy(0, sqrt(2)/2)"
    ),
    prior_label   = factor(prior_label,
                           levels = c("Matched\nLogistic(0, 0.25)",
                                      "JZS/Cauchy\nCauchy(0, sqrt(2)/2)")),
    true_b1_label = factor(paste0("b[1] == ", true_b1))
  ) |>
  ggplot(aes(x = prior_label, y = log10_BF10, fill = prior_label)) +
  geom_hline(yintercept = c(log10(3), log10(1/3)), linetype = "dashed",
             colour = "grey50", linewidth = 0.5) +
  geom_hline(yintercept = 0, colour = "grey30", linewidth = 0.4) +
  geom_violin(alpha = 0.6, scale = "width", width = 0.7) +
  stat_summary(fun = median, geom = "point", size = 2, colour = "black") +
  facet_wrap(~ true_b1_label, nrow = 2, labeller = label_parsed) +
  scale_fill_viridis_d("Prior", option = "D", end = 0.85) +
  scale_y_continuous(
    expression(BF[10]),
    breaks = log10(c(1/100, 1/10, 1/3, 1, 3, 10, 100)),
    labels = c("1/100", "1/10", "1/3", "1", "3", "10", "100")
  ) +
  coord_cartesian(ylim = c(-2.5, 2.5)) +
  labs(x = "Prior specification") +
  prior_theme() +
  theme(axis.text.x = element_text(size = 10))
```

@fig-bf-jzs-comparison compares the matched Logistic(0, 0.25) against the JZS/Cauchy default. The two priors diverge most strongly for small true effects. Under the true null ($b_1 = 0$), the Cauchy's heavier tails make it more skeptical and it correctly supports $H_0$ at a higher rate than the matched logistic (e.g., $P(\mathrm{BF}_{01} > 3) \approx 0.91$ vs.\ $0.60$ at $N = 30$, 20 trials). However, this skepticism comes at a cost when a small true effect is present: at $b_1 = 0.10$ and $N = 60$, the Cauchy achieves only 12% power ($P(\mathrm{BF}_{10} > 3)$) compared to 34% for the matched logistic, and 53% of Cauchy replications falsely return evidence for $H_0$ ($\mathrm{BF}_{01} > 3$) versus 20% for the matched logistic. The Cauchy default thus provides systematically misleading evidence when a small genuine effect is present. For medium and large effects ($b_1 \geq 0.20$), the two priors converge and both achieve high power. The matched prior therefore offers a genuine BF calibration advantage for small effects, over and above providing scientifically plausible prior predictive distributions on the probability scale; the JZS/Cauchy default assigns substantial prior mass to implausibly large effects (see Prior Predictive Simulation section above and Supplementary Materials for quantification).

### Information Consistency

Having established the BF calibration advantage of matched priors over the JZS default, we now verify whether they satisfy the formal consistency desiderata described above. To verify information and Lindley consistency empirically, we held the prior fixed at the recommended matched configuration (Logistic(0, 0.25) effect prior, Logistic(0, 0.75) intercept prior, Exponential(4) random-effects hyperprior) and varied the true effect size ($b_1 \in \{0, 0.05, 0.10, 0.20, 0.50\}$) and sample size ($N \in \{20, 30, 50, 100, 200\}$), with 20 trials per subject throughout. We ran 100 replications per condition. Bayes Factors were estimated via bridge sampling [@gronau2017], which remains reliable when the posterior is far from zero; the Savage-Dickey KDE at zero can underflow in such cases, producing inflated BF estimates.

```{r}
#| label: fig-bf-consistency-effect
#| fig-cap: "Information consistency: median $\\log_{10}(BF_{10})$ as a function
#|   of true effect size $b_1$ (20 trials per subject), colour-coded by sample
#|   size. Each line shows a monotonically increasing BF as the true effect grows,
#|   with larger samples yielding stronger evidence at every effect size. The null
#|   ($b_1 = 0$) anchors each line near or below the 1/3 threshold. Error bars
#|   show the interquartile range across replications. Dashed lines mark
#|   $BF_{10} = 3$ and $BF_{10} = 1/3$. BFs estimated via bridge sampling."
#| fig-height: 5
#| fig-width: 7
#| eval: !expr consistency_available

bf_breaks <- c(-3, -2, -1, 0, 1, 2, 3, 6, 9, 12, 15)
bf_labels <- expression(10^{-3}, 10^{-2}, 10^{-1}, 10^0,
                        10^1,   10^2,   10^3,  10^6,
                        10^9,   10^{12}, 10^{15})

dodge_effect <- position_dodge(width = 0.02)

consistency_continuous |>
  filter(!is.na(median_log10_BF10)) |>
  mutate(n_subjects = factor(n_subjects)) |>
  ggplot(aes(x = true_b1, y = median_log10_BF10,
             colour = n_subjects, group = n_subjects)) +
  geom_hline(yintercept = c(log10(3), log10(1/3)), linetype = "dashed",
             colour = "grey50", linewidth = 0.5) +
  geom_linerange(aes(ymin = q25_log10_BF10, ymax = q75_log10_BF10),
                 alpha = 0.4, linewidth = 0.6,
                 position = dodge_effect) +
  geom_line(linewidth = 0.8, position = dodge_effect) +
  geom_point(size = 2, position = dodge_effect) +
  scale_colour_viridis_d("N subjects", option = "D", end = 0.85) +
  scale_x_continuous(expression("True effect" ~ b[1])) +
  scale_y_continuous(
    expression(log[10](BF[10])),
    breaks = bf_breaks,
    labels = bf_labels
  ) +
  coord_cartesian(ylim = c(-3, 15)) +
  prior_theme()
```

```{r}
#| label: fig-bf-consistency-n
#| fig-cap: "Lindley consistency: median $\\log_{10}(BF_{10})$ as a function of
#|   sample size (20 trials per subject), colour-coded by true effect size $b_1$.
#|   For $b_1 \\geq 0.10$, the BF grows monotonically with $N$; for $b_1 = 0$
#|   (bottom line), the BF decreases monotonically, correctly accumulating evidence
#|   for $H_0$. For the very small effect $b_1 = 0.05$, a non-monotonic pattern
#|   may be visible at small-to-moderate $N$ (see main text). Error bars show the
#|   interquartile range across replications. Dashed lines mark $BF_{10} = 3$ and
#|   $BF_{10} = 1/3$. BFs estimated via bridge sampling."
#| fig-height: 5
#| fig-width: 7
#| eval: !expr consistency_available

dodge_n <- position_dodge(width = 8)

consistency_continuous |>
  filter(!is.na(median_log10_BF10)) |>
  mutate(true_b1 = factor(true_b1)) |>
  ggplot(aes(x = n_subjects, y = median_log10_BF10,
             colour = true_b1, group = true_b1)) +
  geom_hline(yintercept = c(log10(3), log10(1/3)), linetype = "dashed",
             colour = "grey50", linewidth = 0.5) +
  geom_linerange(aes(ymin = q25_log10_BF10, ymax = q75_log10_BF10),
                 alpha = 0.4, linewidth = 0.6,
                 position = dodge_n) +
  geom_line(linewidth = 0.8, position = dodge_n) +
  geom_point(size = 2, position = dodge_n) +
  scale_colour_viridis_d(expression(b[1]), option = "D", end = 0.85) +
  scale_x_continuous("Number of subjects") +
  scale_y_continuous(
    expression(log[10](BF[10])),
    breaks = bf_breaks,
    labels = bf_labels
  ) +
  coord_cartesian(ylim = c(-3, 15)) +
  prior_theme()
```

@fig-bf-consistency-effect shows information consistency empirically: for every sample size, the median $\log_{10}(\mathrm{BF}_{10})$ increases monotonically with the true effect size. @fig-bf-consistency-n shows the corresponding Lindley consistency results. For all substantively relevant effects ($b_1 \geq 0.10$), the BF grows monotonically with sample size, and the BF for the null ($b_1 = 0$) decreases monotonically, correctly accumulating evidence for $H_0$.

For the very small effect ($b_1 = 0.05$), however, the BF may initially support $H_0$ at small-to-moderate sample sizes before eventually reversing toward $H_1$ at sufficiently large $N$. This non-monotonic pattern is known as the Jeffreys-Lindley paradox and is a theoretically expected property of Bayes factors under fixed-scale priors, not an estimation artifact [@huisman2022; @sarafoglou2024]. @huisman2022 argued that this pattern renders Bayes factors invalid as measures of evidential strength, because they are not stochastically ordered by sample size when the true effect is small. @sarafoglou2024 showed, however, that this behavior is an inevitable consequence of probability theory that has been understood since Jeffreys's original work in the 1930s, and @rouder2009 [p.\ 233] explicitly characterised it as ideal. The mechanism is model parsimony: when the true effect is negligible relative to the prior scale, $H_1$ assigns predictive probability to larger effects that fail to materialize, so the more parsimonious $H_0$ temporarily receives more support. The non-monotonicity attenuates or disappears when the prior more tightly concentrates mass around the anticipated effect size — for example, via a power-analytic prior [@schonbrodt2018] or a narrower scale. For the Logistic(0, 0.25) prior recommended here, the anomaly is confined to effects at the boundary of detectability ($b_1 = 0.05$, corresponding to an accuracy difference of approximately 1–2 percentage points); for all effects within the practically relevant range ($b_1 \geq 0.10$), Lindley consistency holds throughout the sample sizes typical in psychological research. Together, these results confirm that matched priors satisfy the key consistency desiderata within the plausible range of effect sizes.

### Effect of Random-Effects Hyperprior

```{r}
#| label: fig-bf-violin-re
#| fig-cap: "Distribution of log10(BF10) by random-effects hyperprior for
#|   N = 30, 20 trials. All three hyperpriors produce nearly identical BF
#|   distributions. Black dots indicate medians. Dashed lines mark BF10 = 3
#|   and BF10 = 1/3."
#| fig-height: 5
#| fig-width: 8
#| eval: !expr bf_raw_re_available

bf_raw_re |>
  filter(
    true_b1 %in% c(0, 0.10, 0.20),
    n_subjects == 30,
    n_trials == 20
  ) |>
  mutate(
    sd_prior_re = factor(sd_prior_re,
                         levels = c("default", "gamma", "exponential"),
                         labels = c("Student-t(3,0,2.5)", "Gamma(2,4)",
                                    "Exponential(4)")),
    true_b1_label = factor(paste0("b[1] == ", true_b1))
  ) |>
  ggplot(aes(x = sd_prior_re, y = log10_BF10, fill = sd_prior_re)) +
  geom_hline(yintercept = c(log10(3), log10(1/3)), linetype = "dashed",
             colour = "grey50", linewidth = 0.5) +
  geom_hline(yintercept = 0, colour = "grey30", linewidth = 0.4) +
  geom_violin(alpha = 0.6, scale = "width", width = 0.7) +
  stat_summary(fun = median, geom = "point", size = 2, colour = "black") +
  facet_wrap(~ true_b1_label, nrow = 1, labeller = label_parsed) +
  scale_fill_viridis_d("RE hyperprior", option = "D", end = 0.85) +
  scale_y_continuous(
    expression(BF[10]),
    breaks = log10(c(1/100, 1/10, 1/3, 1, 3, 10, 100)),
    labels = c("1/100", "1/10", "1/3", "1", "3", "10", "100")
  ) +
  coord_cartesian(ylim = c(-2.5, 2.5)) +
  labs(x = "Random-effects hyperprior") +
  prior_theme() +
  theme(axis.text.x = element_text(angle = 30, hjust = 1))
```

The random-effects hyperprior (i.e., the prior on the standard deviation of individual differences) has no meaningful impact on Bayes Factors for the fixed effect (@fig-bf-violin-re), even though the Student-*t*(3, 0, 2.5) is far wider than Exponential(4). This nuisance-parameter robustness arises because the BF is a ratio of marginal likelihoods: the random intercepts $u_j$ are present under both $H_1$ and $H_0$ and are marginalized out in each. The common random-effects structure therefore largely cancels in the ratio, leaving the BF dominated by the fixed-effect signal-to-noise ratio. This confirms that the random-effects hyperprior choice has little influence on the Bayes Factor for the fixed effect of interest, though it may matter for other inferential goals (e.g., estimating the random-effects variance itself).

### Summary

@tbl-bf-summary-power summarises detection rates ($P(\mathrm{BF}_{10} > 3)$) and correct null-support rates ($P(\mathrm{BF}_{01} > 3)$) — the Bayesian analogue of $1 - \alpha$ (Type I error control) — for the matched and JZS/Cauchy priors across representative conditions (20 trials per subject; full results in the Supplementary Materials). Each cell shows power / null-support rate.

```{r}
#| label: tbl-bf-summary-power
#| tbl-cap: "Detection rates ($P(\\mathrm{BF}_{10} > 3)$) and correct null-support
#|   rates ($P(\\mathrm{BF}_{01} > 3)$) for the matched Logistic(0, 0.25) and
#|   JZS Cauchy(0, $\\sqrt{2}/2$) priors at representative conditions (20 trials
#|   per subject). Each cell shows power / null-support rate. Under the null
#|   ($b_1 = 0$), $P(\\mathrm{BF}_{01} > 3)$ is the Type I error analogue ---
#|   the probability of correctly supporting $H_0$ (high values desirable);
#|   under nonzero effects, it is the false-$H_0$ support rate."
#| eval: !expr jzs_available

jzs_summaries |>
  filter(n_trials == 20, n_subjects %in% c(30, 100)) |>
  mutate(
    prior_label = case_when(
      dist_b1 == "logistic" ~ "Matched",
      dist_b1 == "cauchy"   ~ "JZS"
    ),
    col_label = paste0("N=", n_subjects)
  ) |>
  select(prior_label, true_b1, col_label, P_BF10_gt3, P_BF01_gt3) |>
  mutate(
    P_BF10_gt3 = round(P_BF10_gt3, 2),
    P_BF01_gt3 = round(P_BF01_gt3, 2),
    cell = paste0(P_BF10_gt3, " / ", P_BF01_gt3)
  ) |>
  select(prior_label, true_b1, col_label, cell) |>
  pivot_wider(
    names_from  = c(col_label, prior_label),
    names_glue  = "{col_label} {prior_label}",
    values_from = cell
  ) |>
  knitr::kable(
    col.names = c("True $b_1$",
                  "N=30 Matched", "N=30 JZS",
                  "N=100 Matched", "N=100 JZS"),
    align  = c("l", rep("r", 4)),
    escape = FALSE
  )
```

Across all conditions, the most important prior-choice factor is the scale of the effect prior. The choice of the prior family, the intercept prior, and the random-effects hyperprior had negligible effects. Matched priors exhibit information and Lindley consistency, confirming that they correctly accumulate evidence as data grow. Crucially, their BF calibration is superior to the commonly used JZS/Cauchy default for small effects: the Cauchy prior's heavy tails make it overly skeptical, frequently producing misleading evidence for $H_0$ when a small genuine effect is present ($b_1 = 0.10$), while the matched Logistic prior provides substantially higher power in this range. For medium and large effects ($b_1 \geq 0.20$), both priors perform similarly. Together with the guarantee of plausible prior predictive distributions on the probability scale, the matched prior represents a strictly preferable alternative to the JZS/Cauchy default. Still, when computing Bayes Factors for drawing inferences, we recommend running a prior sensitivity analysis by computing BFs across a small set of plausible effect-prior widths (e.g., $\mathrm{sd}_{b_1} \in \{0.15, 0.25, 0.35\}$) to test the robustness of the results over varying prior scales.


## Recommendations {#sec-recommendations}

Researchers who want to apply the recommendations directly without reading the derivations can use the defaults in @tbl-prior-summary and the ready-to-use code below.

| Parameter | Link | Recommended prior | Key property |
|:----------|:-----|:------------------|:-------------|
| Intercept ($b_0$) | logit | $\mathrm{Logistic}(0, 0.75)$ | $< 10\%$ prior mass in floor/ceiling region |
| Intercept ($b_0$) | probit | $\mathrm{Normal}(0, 0.75)$ | same |
| Effect ($b_1$) | logit | $\mathrm{Logistic}(0, 0.25\text{--}0.30)$ | 90th percentile of $|\delta p| \approx 0.20$ |
| Effect ($b_1$) | probit | $\mathrm{Normal}(0, 0.25\text{--}0.30)$ | 90th percentile of $|\delta p| \approx 0.18\text{--}0.22$ |
| RE hyperprior ($\sigma_u$) | both | $\mathrm{Exponential}(4)$ | mean $= 0.25$ logit units |

: Summary of default prior recommendations for binomial GLMs. All values assume sum-to-zero $\pm 1$ coding and predictors scaled to $\mathrm{SD} = 1$; see the Scaling section below for other coding schemes. {#tbl-prior-summary}

The three workflow steps converge on a coherent set of recommendations. Priors influence Bayes Factors, making it important to consider their implications carefully. But priors also encode prior beliefs about plausible effects. Choosing overly wide priors -- as commonly done with "weakly informative" defaults -- not only biases BFs toward the null but also implies implausible prior beliefs about effect sizes on the probability scale. Our recommended priors satisfy both criteria simultaneously: They produce plausible prior predictive distributions on the probability scale (Step 2) and well-calibrated Bayes Factors across a realistic range of effects (Step 3). Importantly, the converse is not guaranteed: Some priors that happen to yield adequate BF calibration may still imply implausible effect sizes (e.g., the Cauchy).

Researchers with no domain-specific prior knowledge can use the default priors directly. If prior information about expected effect sizes is available, the prior scale can be adjusted accordingly. Furthermore, when the predictor's SD differs from 1, the default prior scale should be adapted using the scaling formulas below.

### Default Prior Specifications

@tbl-intercept-prior and @tbl-effect-prior show our recommendation for default priors for intercept and effect parameters in binomial GLMs. The hyperprior on random-effect SDs should concentrate mass below approximately 0.5 on the logit scale (corresponding to between-subject differences in success probability of roughly 10--15 percentage points around typical baselines). Our recommended default is `exponential(4)` (mean = 0.25). For larger expected between-subject variability, `exponential(2)` (mean = 0.50) provides a wider alternative while remaining well calibrated.

| Link | Recommended prior | Rationale |
|------|------------------|-----------|
| logit | $\mathrm{Logistic}(0, 0.75)$ | Matched family; sd = 0.75 gives < 10% floor/ceiling mass |
| probit | $\mathrm{Normal}(0, 0.75)$ | Matched family; same criterion |

: Recommended intercept priors for binomial GLMs. {#tbl-intercept-prior}

| Link | Recommended prior | Rationale |
|------|------------------|-----------|
| logit | $\mathrm{Logistic}(0, 0.25\mathrm{--}0.30)$ | Matched family; 90th percentile of $|\delta p| \approx 0.20\mathrm{--}0.23$ |
| probit | $\mathrm{Normal}(0, 0.25\mathrm{--}0.30)$ | Matched family; 90th percentile of $|\delta p| \approx 0.18\mathrm{--}0.22$ |

: Recommended effect priors for binomial GLMs. {#tbl-effect-prior}

#### Ready-to-use brms prior specifications

```{r}
#| label: recommended-priors
#| echo: true
#| eval: false

# Logit link (logistic regression)
logit_priors <- c(
  brms::prior(logistic(0, 0.75), class = Intercept),
  brms::prior(logistic(0, 0.25), class = b)
)

# Probit link (probit regression)
probit_priors <- c(
  brms::prior(normal(0, 0.75), class = Intercept),
  brms::prior(normal(0, 0.25), class = b)
)
```

For models with subject-level random effects:

```{r}
#| label: recommended-priors-re
#| echo: true
#| eval: false

# Logit link with random intercept
logit_priors_re <- c(
  brms::prior(logistic(0, 0.75), class = Intercept),
  brms::prior(logistic(0, 0.25), class = b),
  brms::prior(exponential(4),    class = sd)
)
```

### From Defaults to Informed Priors {#sec-scaling}

Informed priors can be generated by adjusting the defaults in two ways. First, the prior can be made wider or narrower (i.e., increasing or decreasing $\mathrm{sd}_{b_1}$) depending on whether one expects larger or smaller effects. Second, the prior width must be rescaled if the predictor's SD differs from 1. The recommended $\mathrm{sd}_{b_1} \in [0.25, 0.30]$ assumes $\mathrm{SD}(x) = 1$ (as obtained by sum-to-zero $\pm 1$ coding of a binary predictor). When the predictor's SD differs from 1, the scale of the prior should be rescaled accordingly:

$$\mathrm{sd}_{b_1}(\mathrm{predictor SD} = s) = \frac{\mathrm{sd}_{b_1}(\mathrm{SD} = 1)}{s}$$

Smaller predictor SDs require a *wider* prior on $b_1$; larger predictor SDs require a *narrower* prior. Here are some examples:

- $\pm 0.5$ coding ($s = 0.5$): use $\mathrm{sd}_{b_1} = 0.25 / 0.5 = 0.50$
- $\pm 1/\sqrt{2}$ orthonormal ($s \approx 0.707$): use $\mathrm{sd}_{b_1} = 0.25 / 0.707 \approx 0.354$
- $0/1$ treatment coding ($s = 0.5$ in a balanced design): use $\mathrm{sd}_{b_1} = 0.25 / 0.5 = 0.50$
- Standardised continuous predictor ($s = 1$): use $\mathrm{sd}_{b_1} = 0.25$ directly

For the no-intercept parameterisation (`0 + condition` in `brms`), see [Appendix B](#sec-appendix-no-intercept).

#### Dealing with multi-level factors

When a factor has more than two levels, we recommend using theoretically meaningful orthogonal contrasts that reflect specific hypotheses. These could either be a set of orthogonal contrasts that breaks down factors with more levels into a sub-set of contrasts with only two levels, or transforming factors into continuous predictors when this is reasonable.

For researchers who do not have specific directional hypotheses but want to test the omnibus question "does the factor have any effect?", orthonormal contrasts provide a principled alternative. Orthonormal contrasts [e.g., `contr.equalprior()` from the `bayestestR` package; @makowski2019; @rouder2012] equalize contrast column variances, ensuring that the prior width is the same for every pairwise comparison. In a balanced design with $J$ levels, each column has predictor $\mathrm{SD} = 1 / \sqrt{J}$, giving the rescaling formula:

$$\mathrm{sd}_{b_1}(\mathrm{orthonormal, } J \mathrm{ levels}) = 0.25 \times \sqrt{J}$$

See @tbl-orthonormal-scaling in [Appendix A](#sec-appendix-contrasts) for a look-up table of rescaled prior widths for common multi-level and factorial designs.

**In short:** use $\mathrm{Logistic}(0, 0.75)$ for the intercept, $\mathrm{Logistic}(0, 0.25)$ for the effect, and $\mathrm{Exponential}(4)$ for the random-effects SD when applying logistic regression; substitute $\mathrm{Normal}$ priors with the same scales for probit regression. Rescale the effect prior width when the predictor SD differs from 1, and report BFs across $\mathrm{sd}_{b_1} \in \{0.15, 0.25, 0.35\}$ as a sensitivity check.

```{r}
#| label: fig-orthonormal-demo
#| fig-cap: "Prior predictive distribution of a single representative
#|   pairwise $\\delta p$ under orthonormal contrasts with the rescaled prior
#|   ($\\mathrm{sd}_{b_1} = 0.25 \\times \\sqrt{J}$). The distribution is
#|   nearly identical across $J$ -- confirming that the scaling rule
#|   preserves per-comparison calibration."
#| fig-height: 4.5
#| fig-width: 6
#| message: false

set.seed(2024)
n_draws <- 50000

demo_list <- lapply(c(2, 3, 4), function(J) {
  oc <- orthonormal_contrasts(J)
  sd_b1_scaled <- 0.25 * oc$scaling_factor

  b0 <- rlogis(n_draws, location = 0, scale = 0.75)
  B <- matrix(
    rnorm(n_draws * (J - 1), mean = 0, sd = sd_b1_scaled),
    nrow = n_draws, ncol = J - 1
  )

  eta_mat <- outer(b0, rep(0, J)) + B %*% t(oc$Q)
  p_mat   <- ilogit(eta_mat)

  delta_p_pair <- p_mat[, 1] - p_mat[, 2]

  data.frame(
    J            = paste0("J = ", J),
    delta_p_pair = delta_p_pair
  )
})

demo_draws <- do.call(rbind, demo_list)

p_single <- ggplot(demo_draws, aes(x = delta_p_pair,
                                    y = J, fill = J)) +
  geom_density_ridges(alpha = 0.5, scale = 1.2) +
  geom_vline(xintercept = c(-0.20, -0.10, 0.10, 0.20),
             linetype = "dotted", linewidth = 0.4,
             colour = "grey40") +
  scale_x_continuous(expression(delta * p ~ "(single pair)"),
                     limits = c(-0.8, 0.8)) +
  scale_fill_viridis_d("Factor levels", option = "D", end = 0.85) +
  ggtitle("Single pairwise comparison") +
  prior_theme()

p_single
```


## Applying the Workflow Beyond Binomial Models {#sec-extensions}

The three-step workflow demonstrated above transfers directly to other model classes where link functions mediate between parameter and probability scales. We provide a walkthrough for ordinal regression and lognormal regression and briefly sketch the application to three other common model families.

### Ordinal regression

Ordinal regression is the method of choice when the data consist of frequencies over more than two ordered response categories, for instance, response frequencies from Likert scales. Cumulative logit and probit models use the same link functions as binomial GLMs, applied to cumulative probabilities across ordered response categories. Each threshold parameter $\tau_k$ determines the boundary between adjacent categories on the latent scale, and the inverse link converts these boundaries to cumulative probabilities.

**Step 1 (Matched priors).** The matched prior principle applies directly: a $\mathrm{Logistic}(0, 1)$ prior on each threshold under the cumulative logit link produces a uniform distribution of cumulative probabilities at that threshold. The matched prior ($\mathrm{sd} = 1$) yields approximately uniform category probabilities on average. Note that the effect of scaling differs from the binary case: scaling to $\mathrm{sd} < 1$ clusters all thresholds near the center of the latent scale, which concentrates prior mass *toward* the extreme categories (a U-shaped distribution of category probabilities); scaling to $\mathrm{sd} > 1$ spreads thresholds outward, concentrating mass in the middle categories (a bell-shaped distribution). For the effect prior on predictors, the same matched families apply as in the binomial case.

**Step 2 (Prior predictive simulation).** The key difference from the binomial case is that the probability scale is now a vector of category probabilities rather than a single success probability. Prior predictive simulation should examine the implied distribution of category probabilities and, critically, of contrasts between conditions (e.g., shifts in the most likely category, or changes in cumulative probabilities). One can draw $b_0$, threshold spacings, and effect coefficients from the joint prior, compute category probabilities via the inverse link, and inspect whether the implied distributions match substantive expectations about response patterns.

**Step 3 (BF calibration).** Bayes Factor calibration follows the same template: simulate data with known effects, fit cumulative models with the candidate priors, and evaluate BF sensitivity. The Savage-Dickey density ratio applies to the effect coefficient just as in the binomial case, because $b_1 = 0$ is a point null on the same parameter regardless of the outcome distribution, and the density ratio at zero is well-defined whenever a continuous prior is placed on $b_1$.

### Lognormal regression for response times

Response times are among the most common continuous outcomes in experimental psychology. Because they are right-skewed and strictly positive, lognormal regression (`family = lognormal()` in `brms`, with an identity link on the log scale) is a natural modelling choice. Here, the matched prior principle from Step 1 does not apply: The identity link involves no nonlinear CDF, so there is no distribution family to "match." Nevertheless, a prior-transform interaction remains, because the exponential function maps the log-scale intercept to the median response time in milliseconds. For instance, a $\mathrm{Normal}(6.5, 1)$ prior on the intercept implies a 95% range of median response times from $\exp(4) \approx 90$ ms to $\exp(8) \approx 4{,}915$ ms -- implausibly broad for most cognitive tasks. A tighter prior such as $\mathrm{Normal}(6.5, 0.5)$ concentrates the median in the range $\exp(5.5)$--$\exp(7.5) \approx 245$--$1{,}808$ ms that is more typical of a broad range of reaction-time paradigms.

For effect coefficients, the mapping is multiplicative: a coefficient $b_1$ on the log scale implies a proportional change of $\exp(b_1)$ on the response-time scale, so a $\mathrm{Normal}(0, 0.1)$ effect prior treats effects up to roughly 20% speedup or slowdown as plausible. This has the benefit that the same effect prior naturally implies larger effects for slower baseline response times and smaller effects for faster baselines, which is typically observed: Comparing groups with faster vs. slower response times, the same experimental manipulation has a larger absolute effect in the slower group; the effect size is approximately proportional to the mean [@myerson2003].

Steps 2 and 3 of the workflow apply directly: Prior predictive simulation reveals what ranges of absolute response times and proportional effects the prior treats as plausible, and Bayes Factor calibration can verify that the chosen scale yields well-behaved evidence. This illustrates that even when CDF-matching is unavailable, the simulation and calibration steps remain essential tools for principled prior specification [see also @hosack2026 for an expert-elicitation approach to lognormal GLMs].

### Other model classes

*Count data.* Poisson and negative binomial regression use a log link. The matched prior principle does not apply directly (the log link is not a CDF), but prior predictive simulation can reveal what prior scales on log-scale coefficients imply for rate ratios.

*Beta regression.* For proportional data modeled via a logit link on the mean, the matched prior principle applies identically to the intercept.

*Distributional regression.* In `brms` distributional models [@burkner2021], all parameters of the response distribution (mean, variance, shape) are linked to predictors through link functions. Each parameter faces its own prior--link interaction. The matched prior principle and simulation approach developed here provide a systematic way to address prior specification for each parameter.

In each of these cases, the three-step logic remains the same: identify whether a CDF-based link admits a matched prior, inspect what the proposed prior implies on the observable outcome scale, and calibrate against known effects to verify well-behaved Bayes factors. We discuss the scope and limitations of this framework next.

## Discussion

We have proposed and empirically validated a three-step workflow for prior specification in GLMs with nonlinear link functions: (1) Identify matched priors via the probability integral transform, (2) evaluate implied prior predictive distributions on the probability scale, and (3) calibrate priors for Bayes Factor hypothesis testing, including formal assessment of information consistency and comparison to the JZS/Cauchy default. Applying this workflow to binomial models, we recommend the matched $\mathrm{Logistic}(0, 0.75)$ intercept prior and $\mathrm{Logistic}(0, 0.25\mathrm{--}0.30)$ effect prior for logistic regression (with analogous Normal priors for probit), and $\mathrm{Exponential}(4)$ for random-effects hyperpriors as default priors.

There are four main take-home messages from our results. First, the matched prior principle provides a theoretically grounded anchor that is missing from current practice: By matching the prior family to the link function via the probability integral transform, the intercept prior becomes interpretable on the probability scale. This design-independent approach complements GLM extensions of g-priors [@bove2011; @liclyde2018], which are design-dependent because they require evaluating the Fisher information matrix at a reference prediction, and provides a practically accessible default for the applied researcher. Second, the prior predictive simulations revealed that the choice of prior *family* matters far less than the choice of prior *scale* -- a finding that held for both fixed effects and random-effects hyperpriors. Third, matched priors with fixed scale are information-consistent and Lindley-consistent, satisfying key desiderata for principled Bayesian model comparison [@bayarri2012], and their BF calibration outperforms the JZS/Cauchy default for small effects: the Cauchy prior's heavy tails cause it to frequently support $H_0$ when a small true effect is present, whereas the matched Logistic prior provides substantially higher power in that range while converging with the Cauchy for medium-to-large effects. This BF calibration advantage is in addition to the matched prior's guarantee of plausible prior predictive distributions on the probability scale. Fourth, the random-effects hyperprior has negligible influence on Bayes factors for fixed effects, providing reassurance that the recommended defaults are robust to this modelling choice.

Our demonstration is limited to the binomial case, and the specific prior scale recommendations are calibrated for the typical range of effect sizes encountered in psychological experiments (effect sizes in the range of $|b_1| \leq 0.50$ logit units). In applied terms: the recommendations are intended for paradigms where the true accuracy difference between conditions is unlikely to exceed approximately 20 percentage points — covering most detection, recognition, and choice experiments in cognitive psychology, but not paradigms with very large training effects or ceiling-limited designs. Researchers in other fields may need to adjust the defaults based on domain-specific effect size distributions. The BF calibration used the Savage-Dickey density ratio; other BF estimation methods (e.g., bridge sampling for non-nested comparisons) were not evaluated here. Future work should extend the empirical evaluation to (a) probit links — the theoretical result applies identically but empirical calibration has not been confirmed; (b) designs with random slopes — the present workflow addresses random intercepts only, and omitting random slopes when they are present can inflate false-positive rates in Bayesian hypothesis testing [@oberauer2022]; and (c) continuous predictors, where the interpretation of $\delta p$ depends on the predictor's scale and range.

The workflow itself, however, is general. Any model with a CDF-based link function admits matched priors, and prior predictive simulation is universally applicable. We hope that the proposed workflow will aid the research community in obtaining default priors for ordinal, count, and distributional regression models, and ultimately help building a library of principled default priors for hypothesis testing across the full range of GLM families used in psychological research.


## References

::: {#refs}
:::

## Appendix A: Orthonormal Contrast Scaling {#sec-appendix-contrasts .appendix}

| Design term | Predictor SD | Rescaled $\mathrm{sd}_{b_1}$ |
|:---|:---:|:---:|
| 2-level factor | $1/\sqrt{2} \approx 0.71$ | 0.35 |
| 3-level factor | $1/\sqrt{3} \approx 0.58$ | 0.43 |
| 4-level factor | $1/\sqrt{4} = 0.50$ | 0.50 |
| 5-level factor | $1/\sqrt{5} \approx 0.45$ | 0.56 |
| $2 \times 3$ interaction | $1/\sqrt{6} \approx 0.41$ | 0.61 |
| $3 \times 3$ interaction | $1/\sqrt{9} \approx 0.33$ | 0.75 |

: Rescaled effect prior widths under orthonormal contrasts for multi-level
  factors and factorial interactions. All values assume the baseline
  recommendation of $\mathrm{sd}_{b_1} = 0.25$ under $\pm 1$ coding.
  {#tbl-orthonormal-scaling}

For unbalanced designs, the predictor SD depends on cell sizes and should be computed from the contrast matrix directly. For factors with many levels ($J > 5$), the `BayesFactor` package [@rouder2012; @rouder2017] implements the full multivariate JZS prior, which accounts for the dimensionality of the effect space without requiring per-coefficient prior widths.

When using orthonormal contrasts in `brms`:

```{r}
#| label: recommended-priors-orthonormal
#| echo: true
#| eval: false

# - Single 3-level factor ---
contrasts(dat$condition) <- bayestestR::contr.equalprior(3)
# Predictor SD = 1/sqrt(3); rescaled prior: sd_b1 = 0.25 * sqrt(3) = 0.433

priors_3level <- c(
  brms::prior(logistic(0, 0.75),  class = Intercept),
  brms::prior(normal(0, 0.433),   class = b)
)

# - 2 x 3 factorial design ---
contrasts(dat$A) <- bayestestR::contr.equalprior(2)
contrasts(dat$B) <- bayestestR::contr.equalprior(3)

priors_factorial <- c(
  brms::prior(logistic(0, 0.75),  class = Intercept),
  brms::prior(normal(0, 0.354),   class = b, coef = A1),
  brms::prior(normal(0, 0.433),   class = b, coef = B1),
  brms::prior(normal(0, 0.433),   class = b, coef = B2),
  brms::prior(normal(0, 0.612),   class = b, coef = "A1:B1"),
  brms::prior(normal(0, 0.612),   class = b, coef = "A1:B2")
)
```

## Appendix B: No-Intercept Parameterisation {#sec-appendix-no-intercept .appendix}

In `brms`, specifying `0 + condition` removes the intercept and estimates separate coefficients for each condition level, so that $p_0 = g^{-1}(b_0)$, $p_1 = g^{-1}(b_1)$, and $\delta p = g^{-1}(b_1) - g^{-1}(b_0)$. Under this parameterization, the matched prior principle applies to each coefficient independently: placing $\mathrm{Logistic}(0, 1)$ on both coefficients under the logit link induces independent $\mathrm{Uniform}(0, 1)$ marginals, and the implied prior on $\delta p$ is a symmetric triangular distribution on $(-1, 1)$. Because the coefficient interpretation differs from the intercept-plus-slope parameterization, our recommended prior widths do not transfer directly -- each condition mean should receive a matched prior scaled to reflect beliefs about plausible success probabilities (e.g., $\mathrm{sd} = 0.75$), and the implied $\delta p$ should be checked via prior predictive simulation.
