---
title: "Default Priors for Binomial GLMs: Prior Predictive Analysis"
author:
  - name: Klaus Oberauer
    affiliation: University of Zurich
  - name: Joscha Dutli
    affiliation: University of Zurich
  - name: Philipp Musfeld
    affiliation: University of Zurich
  - name: Gidon Frischkorn
    affiliation: University of Lucerne
date: today
format:
  html:
    toc: true
    toc-depth: 3
    code-fold: true
    embed-resources: true
  pdf:
    toc: false
    toc-depth: 3
  docx:
    number-sections: false
execute:
  echo: false
  warning: false
  message: false
bibliography: references.bib
csl: apa.csl
abstract: |
  Bayesian inference for binomial generalized linear models (GLMs) requires
  priors on the regression coefficients, but no principled defaults exist for
  logistic or probit regression. We derive recommendations via two
  complementary approaches. First, we exploit the probability integral
  transform to identify *matched priors* — prior distributions whose family
  mirrors the link function — that induce an exactly uniform distribution on
  the success-probability scale. Specifically, a Logistic(0, 1) prior on
  the logit scale and a Normal(0, 1) prior on the probit scale both yield
  Uniform(0, 1) on the probability scale. Second, we characterise prior
  predictive distributions across 720 factorial combinations of link
  function, prior family, and prior scale via large-scale simulation.
  We recommend matched intercept priors at scale 0.75 (< 10% floor/ceiling mass)
  and Normal or Logistic effect priors at scale 0.25 (90th percentile of
  |δp| ≈ 0.20) for experiments in cognitive psychology. Cauchy priors are
  discouraged due to heavy tails that assign implausible prior probability
  to large effects. For random-effects models, we recommend Exponential(4)
  hyperpriors on between-subject SDs instead of the brms default
  half-Student-t(3, 0, 2.5), which concentrates excessive mass near the
  floor/ceiling boundary. Ready-to-use brms prior specifications are provided.
---

```{r}
#| label: setup
library(dplyr)
library(tidyr)
library(ggplot2)
library(ggridges)
library(viridis)
library(scales)
library(here)
library(knitr)

source(here("R", "link_functions.R"))
source(here("R", "prior_sampling.R"))
source(here("R", "plotting.R"))

# Load simulation results (run scripts/01 and 02 first)
summaries  <- readRDS(here("output", "prior_predictive_summaries.rds"))
draws_long <- readRDS(here("output", "prior_predictive_draws.rds"))
```

## Introduction

In cognitive psychology, the dependent variable is often accuracy: the number
of correct responses $k$ out of $n$ trials. We model these data with a
binomial generalized linear model (GLM):

$$k_i \sim \text{Binomial}(n_i,\; p_i)$$

where the success probability $p_i$ is related to a linear predictor via a
link function $g$:

$$g(p_i) = b_0 + b_1 x_i$$

Two common choices for $g$ are the **logit link** (logistic regression) and
the **probit link** (probit regression), corresponding to the CDFs of the
logistic and standard normal distributions, respectively.

Fitting these models in a Bayesian framework requires priors
on the coefficients $b_0$ and $b_1$. Inferences — especially Bayes Factor
model comparisons — are sensitive to prior specification
[@wagenmakers2018a; @schonbrodt2018], so choosing priors carefully is
essential. For normally distributed data, @rouder2012 established a widely
adopted family of default priors (the JZS prior), enabling routine Bayes
Factor computation via JASP and BayesFactor. No equivalent defaults exist
for binomial GLMs.

This gap is consequential because the nonlinear link function in GLMs creates
a **prior–link interaction**: a prior that appears reasonable on the
regression-coefficient scale may produce implausible distributions on the
probability scale, a concern raised in the estimation context by
@gelman2008 and @ghosh2018. These authors addressed the problem of
separation and weakly informative priors for logistic regression — focusing
on posterior estimation and prediction. We address a complementary problem:
choosing priors that produce well-calibrated *Bayes Factors* for hypothesis
testing about experimental effects.

More broadly, the prior–link interaction we study here is not unique to
binomial models. It arises whenever a model maps parameters through a
nonlinear transformation — including ordinal, count, and distributional
regression models [e.g., GAMLSS; @klein2015; @burkner2021]. The approach
we develop — combining analytical prior predictive results with large-scale
simulation — can serve as a **template** for establishing principled priors in
any model class where link functions mediate between the parameter and
observation scale.

### Approach

Following the Bayesian workflow outlined by @gelman2020workflow and
@schad2021, we derive prior recommendations based on two strategies:

1. Analytical derivation of what certain prior choices imply on the
probability scale, yielding exact results for matched priors — exploiting
the probability integral transform to establish a principled anchor.
2. Large-scale prior predictive simulation to characterize implied
distributions of success probabilities and effect sizes across 720
factorial combinations of prior specifications
[@gelman1996; @gabry2019; @betancourt2018].

The simulations vary:

- Link function: logit, probit
- Intercept prior family: Logistic, Normal, Cauchy
- Intercept prior SD ($\text{sd}_{b_0}$): 0.50, 0.75, 1.00, 1.50
- Effect prior family: Normal, Logistic, Cauchy
- Effect prior SD ($\text{sd}_{b_1}$): 0.10 to 2.00

All simulations use a minimal two-condition design with **sum-to-zero
contrast** $x \in \{+1, -1\}$, so that:

$$\delta p = g^{-1}(b_0 + b_1) - g^{-1}(b_0 - b_1)$$

This coding ensures the prior predictive distribution of $\delta p$ is
symmetric around zero when the prior on $b_1$ is symmetric around zero —
a desirable property for a default prior (see
[Contrast coding and prior width]).

## The Matched Prior Principle

### Exact analytical results

Prior specification for GLMs is complicated by the fact that priors are
placed on the linear predictor scale (regression coefficients) but their
implications play out on the response scale (probabilities). Previous work
on priors for logistic regression has focused on estimation problems:
@gelman2008 proposed weakly informative Cauchy(0, 2.5) priors to handle
separation, and @ghosh2018 established theoretical conditions for posterior
existence under Cauchy priors. @greenland2015 compared penalization
approaches including log-F priors. These contributions address posterior
estimation and prediction, but not the distinct requirements of Bayes
Factor hypothesis testing, where the prior density at the point null
directly enters the Savage–Dickey ratio [@ly2016; @bartos2023].

We take a different starting point: the **probability integral transform**.
For any continuous distribution $D$ with CDF $F$, if $X \sim D$, then
$F(X) \sim \text{Uniform}(0, 1)$.

Applied to binomial GLMs, the inverse link function $g^{-1}$ *is* the CDF
of some distribution. Therefore:

| Link | Inverse link $g^{-1}$ | Matched prior | Implied prior on $p$ |
|------|----------------------|---------------|----------------------|
| logit | logistic CDF | $\text{Logistic}(0, 1)$ | $\text{Uniform}(0, 1)$ — **exact** |
| probit | normal CDF $\Phi$ | $\text{Normal}(0, 1)$ | $\text{Uniform}(0, 1)$ — **exact** |

For the logit link, the inverse link $g^{-1}$ is the logistic CDF, so
placing a Logistic prior on $b_0$ produces probabilities that are exactly
uniform on $(0, 1)$. The same cancellation holds for the probit link and
the normal CDF. This is not an approximation — it follows directly from
the probability integral transform. Scaling the prior below $\text{sd} = 1$
concentrates mass away from 0 and 1, producing a regularized intercept
prior suited to experiments where extreme baseline probabilities are unlikely.

The **matched prior** at scale 1 is therefore maximally uninformative on
the probability scale. This provides a principled, interpretable anchor
that contrasts with the typical approach of specifying priors based on
intuition about regression coefficients — a strategy that
@gelman2017prior cautioned against, arguing that priors can often only be
understood in the context of the likelihood and the implied observables.

### Generality: a template for other link-based models

The matched prior principle generalises to any model that maps parameters
through a CDF. For example, in a cumulative ordinal model with a logit
link, a Logistic prior on the threshold parameters produces uniform
implied probabilities across response categories. In a log-link Poisson
model, an Exponential(1) prior on the rate produces a Uniform on the CDF
scale. More broadly, any distributional regression model
[@klein2015; @burkner2021] where the response distribution parameters are
linked to linear predictors faces the same prior–link interaction. The
approach developed here — identify the matched family, characterise
deviations via prior predictive simulation, and calibrate Bayes Factor
performance — can be applied as a template wherever such interactions
arise.

### Analytical density formulas

We can derive the implied density of $p = g^{-1}(b_0)$ analytically via
the change-of-variables formula. For the logit link with a
$\text{Logistic}(0, s)$ prior on $b_0$:

$$f(p) = \frac{f_{\text{Logis}}(\text{logit}(p),\; 0,\; s)}{p(1-p)}$$

At $s = 1$, the logistic density in the numerator and the Jacobian term
$p(1-p)$ cancel exactly, so $f(p) = 1$ — a uniform density on $(0,1)$.
For the probit link with a $\text{Normal}(0, \sigma)$ prior on $b_0$:

$$f(p) = \frac{\phi(\Phi^{-1}(p)/\sigma)/\sigma}{\phi(\Phi^{-1}(p))}
= \frac{1}{\sigma} \cdot \frac{\phi(\Phi^{-1}(p)/\sigma)}{\phi(\Phi^{-1}(p))}$$

At $\sigma = 1$, the standard normal densities in numerator and denominator
cancel, again yielding $f(p) = 1$.

We use these exact formulas to overlay analytical reference curves on the
simulation figures below.

### Why misfit matters: current software defaults

By default, brms [@burkner2017; @burkner2018] places a $\text{student\_t}(3, 0, 2.5)$ prior on the
intercept (applied to the centered design matrix, with a data-driven scale
of at least 2.5) and a flat, improper prior on all population-level effects.
Standard deviations in multilevel models receive a half-$\text{student\_t}$
prior with a data-driven scale. These defaults are designed for estimation
stability across a wide range of models, not for hypothesis testing
[@gelman2008].

Both defaults create a **prior–link misfit** for logistic regression. The
Student-$t$ intercept prior, passed through the logit link, yields a
logit-$t$ distribution on $p$ with heavier tails than the matched Logistic
prior, concentrating more mass near 0 and 1. The flat slope prior is more
problematic still: it places equal probability on arbitrarily large effect
sizes, so $|\delta p|$ near 1 is treated as equally plausible as $|\delta
p|$ near 0. For Bayes Factor testing via the Savage–Dickey density ratio,
such improper or overly diffuse priors are not merely inconvenient — they
are undefined or produce BFs that are biased toward H1
[@schonbrodt2018; @consonni2018].

## Contrast Coding and Prior Width

The choice of contrast coding directly scales the effect coefficient $b_1$
and therefore the implied prior on the probability scale.

### Sum-to-zero coding ±1

With codes $x \in \{+1, -1\}$, the effect $b_1$ represents the deviation
of each condition from the grand mean. The distribution of $\delta p$ is
**symmetric** around zero when the prior on $b_1$ is symmetric — a
desirable property for a default prior. All simulations here use ±1 coding
[see @schad2020 for a thorough treatment of contrast coding in regression].

### Treatment coding (0/1)

With treatment coding $x \in \{0, 1\}$, the baseline condition is modelled
by $b_0$ alone, whereas the treatment condition is modelled by $b_0 + b_1$.
The effect on the probability scale is
$\delta p = g^{-1}(b_0 + b_1) - g^{-1}(b_0)$.

When the prior on $b_0$ is centred at zero, $\delta p$ remains symmetric
around zero — the nonlinearity of $g^{-1}$ does not by itself introduce
skewness. However, treatment coding has two practical disadvantages compared
with sum-to-zero coding. First, the prior probability mass is distributed
asymmetrically across condition means: one condition's probability depends on
$b_0$ alone, while the other depends on $b_0 + b_1$, so the two conditions
carry different amounts of prior uncertainty. Second, if the intercept prior
mean is shifted away from zero (e.g., to reflect an expected baseline
accuracy above chance), the asymmetry in the definition of $\delta p$ leads
to a skewed prior predictive distribution of the effect — a problem that also
affects sum-to-zero coding but to a lesser degree, because both conditions
are shifted symmetrically relative to the intercept.

We therefore recommend sum-to-zero ($\pm 1$) coding, which distributes the
effect symmetrically across both conditions and produces a symmetric prior
predictive for $\delta p$ under any intercept prior centred at zero.

### Scaling conversion

When contrast codes are scaled by factor $c$ (e.g., $\pm c$ instead of
$\pm 1$), the implied effect on the link scale is
$\delta = b_1 \cdot (x_+ - x_-) = 2c \cdot b_1$.
To keep $\text{SD}(\delta)$ constant, the prior on $b_1$ must be widened
by the inverse of $c$:

$$\text{sd}_{b_1}(\text{coding } \pm c) = \frac{\text{sd}_{b_1}(\text{coding } \pm 1)}{c}$$

Smaller contrast weights require a *wider* prior on $b_1$ to induce the
same prior on the probability-scale effect.

For example, a researcher using half-unit coding ($\pm 0.5$) should specify
$\text{sd}_{b_1} = 0.50$ to obtain the same prior predictive distribution
of $\delta p$ as $\text{sd}_{b_1} = 0.25$ under $\pm 1$ coding. Orthonormal
coding ($\pm 1/\sqrt{2}$) requires $\text{sd}_{b_1} \approx 0.354$ for the
same implied effect size distribution.

### No-intercept parameterization (0 + coding in brms)

In `brms`, specifying a formula with `0 + condition` removes the intercept
and estimates separate coefficients for each condition level. In a
two-condition design, this yields coefficients $b_0$ and $b_1$ that directly
represent the condition means on the link scale, so that:

$$p_0 = g^{-1}(b_0), \qquad p_1 = g^{-1}(b_1), \qquad \delta p = g^{-1}(b_1) - g^{-1}(b_0)$$

Under this parameterization, the matched prior principle applies to each
coefficient independently: placing $\text{Logistic}(0, 1)$ on both $b_0$ and
$b_1$ under the logit link (or $\text{Normal}(0, 1)$ under the probit link)
induces independent $\text{Uniform}(0, 1)$ marginals on each condition
probability. The implied prior on $\delta p$ then follows from the
difference of two independent uniform random variables, yielding a
symmetric triangular distribution on $(-1, 1)$.

Because the interpretation of each coefficient differs from the
intercept-plus-slope parameterization, the recommended prior widths from the
present analysis do not transfer directly. Instead, each condition mean
should receive a matched prior scaled to reflect beliefs about the range of
plausible success probabilities (e.g., $\text{sd} = 0.75$ to avoid
floor/ceiling effects), and the implied prior on $\delta p$ should be
checked via prior predictive simulation.

## Results: Intercept Prior

### Figure 1: Prior predictive of $p$ by intercept prior

```{r}
#| label: fig-1-p-intercept
#| fig-cap: "Histograms of the prior predictive distribution of success
#|   probability p = g^{-1}(b0) for each intercept prior SD (rows) and link
#|   function (columns), shown separately for three prior families (Logistic,
#|   Normal, Cauchy; colour-coded). Effect prior fixed at Logistic(0, 0.25).
#|   Dashed curves show the exact analytical density for the matched prior
#|   (Logistic for logit, Normal for probit) at each sd_b0 value. At
#|   sd_b0 = 1.0, the matched prior histogram coincides with the flat
#|   (uniform) reference curve, providing visual validation of the probability
#|   integral transform result."
#| fig-height: 8
#| fig-width: 10

fig1_draws <- draws_long |>
  filter(dist_b1 == "logistic", sd_b1 == 0.25)

plot_p_intercept_ridges(fig1_draws, sd_b1_focus = 0.25)
```

As shown in @fig-1-p-intercept, at `sd_b0 = 1.0` with the matched prior
(Logistic for logit, Normal for probit), the distribution of $p$ is uniform,
consistent with the exact analytical result. At `sd_b0 = 0.75`, the prior
pulls mass away from 0 and 1, which is appropriate when floor and ceiling
effects are a concern. Misfit priors, such as Cauchy for either link,
produce heavier tails and concentrate more mass near 0 and 1.

### Figure 2: Floor/ceiling diagnostic

```{r}
#| label: fig-2-floor-ceiling
#| fig-cap: "Proportion of prior predictive mass falling in the floor/ceiling
#|   region (p < 0.05 or p > 0.95), as a function of intercept prior SD and
#|   family, for each link function."
#| fig-height: 6
#| fig-width: 10

plot_floor_ceiling_heatmap(summaries)
```

We recommend choosing the intercept prior SD so that
$P(p < 0.05 \text{ or } p > 0.95) \leq 0.10$. Under this criterion,
`Logistic(0, 0.75)` is the recommended default for the logit link and
`Normal(0, 0.75)` for the probit link (see @fig-2-floor-ceiling).

Note that the exact value of 0.10 at `sd_b0 = 1.0` for the matched priors
(Logistic under logit, Normal under probit) confirms the theoretical
prediction: when the matched prior induces a uniform distribution on
$(0, 1)$, exactly 10% of mass falls below 0.05 or above 0.95
($2 \times 0.05 = 0.10$). This serves as an internal validation of the
simulation.

## Results: Effect Prior

### Figure 3: Effect size distribution by prior scale

@fig-3-delta-p shows how the choice of $\text{sd}_{b_1}$ translates into implied
effect sizes on the probability scale ($\delta p$, signed), holding the intercept
prior fixed at the link-matched recommended prior ($\text{Logistic}(0, 0.75)$
for the logit link; $\text{Normal}(0, 0.75)$ for the probit link).

```{r}
#| label: fig-3-delta-p
#| fig-cap: "Density ridges of the prior predictive distribution of delta_p
#|   (signed effect size on probability scale) for each effect prior SD (y-axis
#|   rows), colour-coded by prior family (Normal, Logistic, Cauchy), and
#|   faceted by link function. Intercept prior fixed at the matched recommended
#|   prior per link: Logistic(0, 0.75) for logit, Normal(0, 0.75) for probit.
#|   Solid vertical line at 0; dotted lines at +/-0.10, +/-0.20, +/-0.30."
#| fig-height: 10
#| fig-width: 10

# Use the link-matched recommended intercept prior for each link:
# Logistic(0,0.75) for logit, Normal(0,0.75) for probit
fig2_draws <- draws_long |>
  filter(
    (link == "logit"  & dist_b0 == "logistic") |
    (link == "probit" & dist_b0 == "normal"),
    sd_b0 == 0.75
  )

plot_delta_p_ridges(fig2_draws, sd_b0_focus = 0.75)
```

Several patterns are immediately apparent from @fig-3-delta-p. First, the
**Cauchy prior** stands apart from Normal and Logistic at every scale: even
at $\text{sd}_{b_1} = 0.25$, its heavy tails assign substantial prior
probability to $|\delta p| > 0.50$ — effect sizes that would represent a
dramatic shift in accuracy (e.g., from 50% to above 90% correct). At
$\text{sd}_{b_1} \geq 0.50$, a Cauchy prior treats effects larger than 30
percentage points as essentially as plausible as small effects. For most
cognitive psychology experiments, such extreme effects are implausible a
priori, making the Cauchy a poor default choice.

Second, even for the better-behaved **Normal and Logistic priors**, a scale
of $\text{sd}_{b_1} = 1.0$ or above spreads mass broadly across the entire
$(-1, 1)$ range, implying near-equal plausibility of tiny and enormous
effects. At $\text{sd}_{b_1} = 0.75$, the 90th percentile of $|\delta p|$
already exceeds 0.30 — a very large effect in most experimental paradigms.
A scale below 1.0 is therefore needed to obtain a meaningful regularising
prior on the probability scale.

Within the substantively appropriate range of $\text{sd}_{b_1} \in [0.10,
0.50]$, Normal and Logistic priors behave similarly. At
$\text{sd}_{b_1} = 0.10$, almost all prior mass falls within $|\delta p| <
0.10$, suitable only when very small effects are expected. At
$\text{sd}_{b_1} = 0.25$, the distribution is centred around zero with the
90th percentile of $|\delta p|$ near 0.20 - corresponding to a moderate
effect of 20 percentage points - which aligns well with the range of
effects typically observed in cognitive psychology. Values up to
$\text{sd}_{b_1} = 0.50$ remain defensible when larger effects are
plausible, with the 90th percentile rising to roughly 0.35.

## Results: Random Effects Prior

In practice, binomial GLMs for repeated-measures cognitive psychology data
include subject-level random intercepts (and often random slopes) to account
for between-subject heterogeneity. Specifying a complete Bayesian model
therefore also requires priors on
the random-effect SDs $\sigma_u$ (intercept) and $\sigma_v$ (slope). Three
candidate hyperpriors are compared:

- **$\text{half-student\_t}(3, 0, 2.5)$**: the brms default; a half-Student-$t$ with 3
  degrees of freedom and scale 2.5, placing a heavy right tail on $\sigma_u$.
- **$\text{Gamma}(2, 4)$**: mean $= 0.5$, with approximately 71% of mass below
  $\sigma_u = 0.5$ logit units; a moderately regularising prior.
- **$\text{Exponential}(4)$**: mean $= 0.25$, with approximately 86% of mass below
  $\sigma_u = 0.5$; the strongest concentration near zero of the three.

The key question is: **what does each hyperprior imply about success
probabilities and effect heterogeneity on the probability scale?**

To answer this, we run a full hierarchical prior predictive:

$$\sigma_u \sim \text{hyperprior}, \quad u_i \sim \mathcal{N}(0, \sigma_u)$$

$$p_{ij} = g^{-1}(b_0 + u_i \pm (b_1 + v_i))$$

with $b_0 \sim \text{Logistic}(0, 0.75)$ and $b_1 \sim \mathcal{N}(0, 0.25)$
fixed at the recommended values from the fixed-effects analysis. For the
intercept + slope structure, $\sigma_v$ is drawn from the same hyperprior as
$\sigma_u$.

```{r}
#| label: setup-re
#| include: false

re_summaries <- tryCatch(
  readRDS(here("output", "prior_predictive_re_summaries.rds")),
  error = function(e) NULL
)
re_draws <- tryCatch(
  readRDS(here("output", "prior_predictive_re_draws.rds")),
  error = function(e) NULL
)

re_results_available <- !is.null(re_summaries) && !is.null(re_draws)
```

```{r}
#| label: fig-re-1-p-intercept
#| fig-cap: "Density ridges of the prior predictive distribution of
#|   subject-level success probabilities p_i = g^{-1}(b0 + u_i). Each ridge
#|   corresponds to one of three hyperpriors for the between-subject SD
#|   (student_t, gamma, exponential; y-axis), faceted by RE structure
#|   (intercept-only vs. intercept + slope). Fixed-effects priors held at
#|   Logistic(0, 0.75) for the intercept and Normal(0, 0.25) for the effect.
#|   Dashed vertical lines mark the floor/ceiling boundary (p = 0.05, 0.95)."
#| fig-height: 5
#| fig-width: 9
#| eval: !expr re_results_available

re_draws |>
  group_by(sd_prior_re, re_structure) |>
  slice_sample(n = 5000) |>
  ungroup() |>
  ggplot(aes(x = p_intercept, y = sd_prior_re,
             fill = sd_prior_re, colour = sd_prior_re)) +
  geom_density_ridges(alpha = 0.5, scale = 0.9, rel_min_height = 0.01) +
  geom_vline(xintercept = c(0.05, 0.95), linetype = "dashed",
             colour = "grey40", linewidth = 0.5) +
  facet_wrap(~re_structure) +
  scale_fill_viridis_d(begin = 0.1, end = 0.85, guide = "none") +
  scale_colour_viridis_d(begin = 0.1, end = 0.85, guide = "none") +
  scale_x_continuous("Subject-level success probability", limits = c(0, 1),
                     expand = c(0, 0)) +
  scale_y_discrete("Hyperprior") +
  labs(title = "Subject-level intercept probabilities by RE hyperprior") +
  theme_minimal(base_size = 12) +
  theme(strip.text = element_text(face = "bold"))
```

The brms default `student_t(3, 0, 2.5)` places a heavy right tail on
$\sigma_u$: occasional large SDs spread subject-level probabilities toward 0
and 1, as visible in @fig-re-1-p-intercept from the heavier tails near the
floor/ceiling boundary. The `gamma(2, 4)` and `exponential(4)` hyperpriors
concentrate $\sigma_u$ near zero, keeping subjects' probabilities away from
the boundary.

```{r}
#| label: fig-re-2-delta-p
#| fig-cap: "Prior predictive distribution of the signed effect size delta_p
#|   on the probability scale under each RE hyperprior, for both RE structures.
#|   Solid line at 0; dotted lines at +/-0.10, +/-0.20, +/-0.30."
#| fig-height: 5
#| fig-width: 9
#| eval: !expr re_results_available

re_draws |>
  group_by(sd_prior_re, re_structure) |>
  slice_sample(n = 10000) |>
  ungroup() |>
  ggplot(aes(x = delta_p, y = sd_prior_re,
             fill = sd_prior_re, colour = sd_prior_re)) +
  geom_density_ridges(alpha = 0.5, scale = 0.9, rel_min_height = 0.01) +
  geom_vline(xintercept = 0, linewidth = 0.5, colour = "grey20") +
  geom_vline(xintercept = c(-0.30, -0.20, -0.10, 0.10, 0.20, 0.30),
             linetype = "dotted", linewidth = 0.5, colour = "grey40") +
  facet_wrap(~re_structure) +
  scale_fill_viridis_d(begin = 0.1, end = 0.85, guide = "none") +
  scale_colour_viridis_d(begin = 0.1, end = 0.85, guide = "none") +
  scale_x_continuous("Effect size on probability scale (delta_p)",
                     limits = c(-1, 1), breaks = seq(-1, 1, 0.25)) +
  scale_y_discrete("Hyperprior") +
  labs(title = "Effect size distribution by RE hyperprior") +
  theme_minimal(base_size = 12) +
  theme(strip.text = element_text(face = "bold"))
```

The `student_t` hyperprior produces substantially more within-study effect
heterogeneity and sign reversals compared with the narrower alternatives
(see @fig-re-2-delta-p).
Under `student_t`, a non-trivial fraction of
simulated subjects show effects in the opposite direction to the
population-level $b_1$, even with the intercept-only RE structure — driven
purely by the wide distribution of $\sigma_u$ pulling individual subjects to
regions of the probability scale where the logistic nonlinearity reverses the
apparent direction.

## Recommendations

The approach demonstrated here — analytical derivation of matched priors
via the probability integral transform, followed by systematic prior
predictive simulation — constitutes a general workflow for establishing
principled priors in any model where link functions mediate between
parameter and observation scales. It operationalises the prior predictive
checking step of the Bayesian workflow [@gelman2020workflow; @schad2021]
into a structured, reproducible design analysis [@schonbrodt2018].

For binomial GLMs with sum-to-zero contrast coding (±1), we recommend
the following default priors:

### Intercept prior

| Link | Recommended prior | Rationale |
|------|------------------|-----------|
| logit | $\text{Logistic}(0, 0.75)$ | Matched family; sd = 0.75 gives < 10% floor/ceiling mass |
| probit | $\text{Normal}(0, 0.75)$ | Matched family; same criterion |

### Effect prior

| Prior family | Recommended SD | 90th percentile of $|\delta p|$ |
|-------------|---------------|-------------------------------|
| Logistic (default) | 0.25 | ~0.20 |
| Normal | 0.25 | ~0.18 |
| Cauchy | 0.25 | ~0.25 (heavier tails) |

Our default recommendation is $\text{Logistic}(0, 0.25)$ for logit models
and $\text{Normal}(0, 0.25)$ for probit models. At this scale, effects
larger than 20 percentage points receive decreasing prior probability,
consistent with typical effect magnitudes in cognitive psychology.

For sensitivity analyses, $\text{sd}_{b_1}$ values between 0.10 and 0.50
span a reasonable range of prior beliefs. Values above 0.50 assign
substantial probability to extreme effects and require specific theoretical
justification.

### Random effects prior

For default prior predictive behaviour consistent with the fixed-effects
recommendations:

- **Use `exponential(4)` or `gamma(2, 4)`** for both $\sigma_u$ and
$\sigma_v$. These concentrate the between-subject SD below ~0.5 logit
units, which keeps floor/ceiling mass under 10% and limits within-study
effect heterogeneity to a substantively plausible range.
- **Avoid the brms [@burkner2017] default `student_t(3, 0, 2.5)`** unless very large
between-subject variability is expected. The heavy right tail of the
half-$t$ assigns non-negligible probability to $\sigma > 2$ on the logit
scale, which is implausible for most cognitive experiments.
- Between `exponential(4)` and `gamma(2, 4)`, prefer `exponential(4)` for
its stronger concentration near zero — it matches the intercept-only
floor/ceiling criterion more reliably.

### Ready-to-use brms prior specifications

The following code blocks specify the recommended default priors directly in
`brms` [@burkner2017] syntax and can be passed to
the `prior` argument of `brm()`:

```{r}
#| label: recommended-priors
#| echo: true
#| eval: false

# Logit link (logistic regression)
# Intercept: Logistic(0, 0.75) — matched prior, < 10% floor/ceiling mass
# Effect:    Logistic(0, 0.25) — 90th percentile of |delta_p| ~ 0.20
# Note: brms accepts logistic() as a valid distribution name in prior()
logit_priors <- c(
  brms::prior(logistic(0, 0.75), class = Intercept),
  brms::prior(logistic(0, 0.25), class = b)
)

# Probit link (probit regression)
# Intercept: Normal(0, 0.75) — matched prior, < 10% floor/ceiling mass
# Effect:    Normal(0, 0.25) — 90th percentile of |delta_p| ~ 0.18
probit_priors <- c(
  brms::prior(normal(0, 0.75), class = Intercept),
  brms::prior(normal(0, 0.25), class = b)
)
```

For models with subject-level random effects, add a prior on the SD of the
random intercept (and slope, if included). Based on the prior predictive
analyses above, we recommend $\text{Exponential}(4)$ as a moderately
regularising default:

```{r}
#| label: recommended-priors-re
#| echo: true
#| eval: false

# Logit link with random intercept
logit_priors_re <- c(
  brms::prior(logistic(0, 0.75), class = Intercept),
  brms::prior(logistic(0, 0.25), class = b),
  brms::prior(exponential(4),    class = sd)
)

# Probit link with random intercept
probit_priors_re <- c(
  brms::prior(normal(0, 0.75), class = Intercept),
  brms::prior(normal(0, 0.25), class = b),
  brms::prior(exponential(4),  class = sd)
)
```

### Adjusting for contrast coding and predictor scaling

The recommended $\text{sd}_{b_1} = 0.25$ assumes that the predictor variable
has a population-level standard deviation of 1, as is the case for
sum-to-zero $\pm 1$ coding in balanced two-group designs ($\text{SD}(x) = 1$).
Whenever the predictor's SD differs from 1 — whether due to contrast coding
conventions, binary (0/1) treatment coding, or continuous predictors — the
prior on $b_1$ must be rescaled to preserve the same implied distribution of
effects on the probability scale.

The general rule is:

$$\text{sd}_{b_1}(\text{predictor SD} = s) = \frac{\text{sd}_{b_1}(\text{SD} = 1)}{s}$$

Smaller predictor SDs require a *wider* prior on $b_1$; larger predictor SDs
require a *narrower* prior. Examples:

- $\pm 0.5$ coding ($s = 0.5$): use $\text{sd}_{b_1} = 0.25 / 0.5 = 0.50$
- $\pm 1/\sqrt{2}$ orthonormal ($s \approx 0.707$): use $\text{sd}_{b_1} = 0.25 / 0.707 \approx 0.354$
- $0/1$ treatment coding ($s = 0.5$ in a balanced design): use $\text{sd}_{b_1} = 0.25 / 0.5 = 0.50$
- Standardised continuous predictor ($s = 1$): use $\text{sd}_{b_1} = 0.25$ directly

For continuous predictors that are not standardised, compute the
population-level SD of the predictor and apply the formula above. This
ensures that the prior on the predicted effect — the change in
$p$ per one-SD change in the predictor — remains at the same scale
regardless of the predictor's units.

## References

::: {#refs}
:::
