---
title: "Priors for Bayes Factors with Binomial Models"
shorttitle: "Priors for Binomial Bayes Factors"
author:
  - name: Gidon T. Frischkorn
    orcid: 0000-0002-5055-9764
    email: gidon.frischkorn@unilu.ch
    corresponding: true
    affiliations:
      - name: University of Zurich
        department: Department of Psychology
      - name: University of Lucerne
        department: Faculty of Behavioral Sciences and Psychology
  - name: Joscha Dutli
    orcid: 0009-0009-5007-1281
    affiliations:
      - name: University of Zurich
        department: Department of Psychology
  - name: Philipp Musfeld
    orcid: 0000-0002-6539-0105
    affiliations:
      - name: University of Amsterdam
        department: Department of Psychology
      - name: University of Zurich
        department: Department of Psychology
  - name: Klaus Oberauer
    orcid: 0000-0001-8815-6189
    affiliations:
      - name: University of Zurich
        department: Department of Psychology
author-note:
  disclosures:
    study-registration: "This study was not preregistered."
    data-sharing: "Simulation code and analysis scripts are available at [repository link]."
    conflict-of-interest: "The authors declare no conflicts of interest."
    financial-support: "This research was supported by a grant from the Swiss National Science Foundation (SNSF)."
    authorship-agreements: |
      Gidon T. Frischkorn: Conceptualization, Methodology, Software, Formal Analysis, Visualization, Writing -- Original Draft.
      Joscha Dutli: Conceptualization, Methodology, Writing -- Review & Editing.
      Philipp Musfeld: Conceptualization,Methodology, Writing -- Review & Editing.
      Klaus Oberauer: Conceptualization, Supervision, Methodology, Writing -- Original Draft.
floatsintext: true
output-dir: "."
format:
  apaquarto-docx: default
  apaquarto-pdf:
    keep-tex: true
  apaquarto-html: default
execute:
  echo: false
  warning: false
  message: false
bibliography: references.bib
csl: apa.csl
abstract: "Bayesian hypothesis testing with Bayes Factors in binomial generalized linear models (GLMs) requires priors on the regression coefficients, yet no principled framework exists for logistic or probit regression. The nonlinear link function creates a prior--link interaction: priors that appear reasonable on the coefficient scale can produce implausible distributions on the probability scale. We address this problem through three complementary approaches. First, we exploit the probability integral transform to identify *matched priors* - prior families that mirror the link function and induce exactly uniform distributions on the probability scale. Second, we characterise implied prior predictive distributions across 720 factorial combinations of link function, prior family, and prior scale via large-scale simulation. Third, we calibrate the recommended priors for Bayes Factor hypothesis testing, evaluating how the strength of evidence varies across prior specifications and effect magnitudes. Our recommendations provide principled defaults for researchers with no domain-specific prior knowledge, and scaling formulas that allow adaptation to any contrast coding, predictor scale, or informed prior belief. Ready-to-use brms prior specifications are provided."
---

```{r}
#| label: setup
library(dplyr)
library(tidyr)
library(ggplot2)
library(ggridges)
library(viridis)
library(scales)
library(here)
library(knitr)

source(here("R", "link_functions.R"))
source(here("R", "prior_sampling.R"))
source(here("R", "plotting.R"))
# Goal 1: Prior predictive results
summaries_pp  <- readRDS(here("output", "prior_predictive_summaries.rds"))
draws_long    <- readRDS(here("output", "prior_predictive_draws.rds"))

# Goal 2: BF calibration results
summaries_bf  <- readRDS(here("output", "bf_calibration_summaries.rds"))

re_summaries <- tryCatch(
  readRDS(here("output", "bf_calibration_re_summaries.rds")),
  error = function(e) NULL
)
re_results_available <- !is.null(re_summaries)

# Factor labels
dist_labels <- c(cauchy = "Cauchy", normal = "Normal", logistic = "Logistic")

base_theme <- function() {
  theme_minimal(base_size = 16) +
    theme(
      panel.grid.minor = element_blank(),
      strip.text       = element_text(face = "bold"),
      legend.position  = "bottom"
    )
}
```

## Introduction

In cognitive psychology, the dependent variable is often accuracy: the number of correct responses $k$ out of $n$ trials. These data follow a binomial distribution, and are therefore best modeled with a binomial generalized linear model (GLM):

$$k_i \sim \text{Binomial}(n_i, p_i)$$

where the success probability $p_i$ is related to a linear predictor via a link function $g$:

$$g(p_i) = b_0 + b_1 x_i$$
$$p_i = g^{-1}(b_0 + b_1 x_i)$$

where $g^{-1}$ is the inverse of the link function. For logistic regression, the inverse function is the logistic function; for probit regression it is the cumulative distribution function (CDF) of the standard normal distribution.

Often we want to test the hypothesis that an effect of interest - for instance, the main effect of an independent variable $x$ - exists against the null hypothesis that it does not exist. Doing this in a Bayesian framework often involves the  estimation of Bayes Factors [@wagenmakers2018a; @schonbrodt2018], which quantify the evidence for one model in comparison to another. Practically, we usually compare a model including the effect of interest (the alternative model) to one representing the null hypothesis, which is identical to the first model except that we constrain the parameter of interest - here, the effect size b1 – to zero; the Bayes factor for this comparison is usually denoted $BF_{10}$. When estimating this BF, we need to carefully set the prior for the parameter of interest, because the BF depends strongly on that prior.

For models assuming a normal distribution of the dependent variable, @rouder2012 proposed default priors for standardized effect sizes that can be used when there are no good theoretical or empirical grounds for setting priors. The rationale for default priors is that even if we know nothing about the effect of interest, we have some general knowledge about the distribution of effect sizes in our field that default priors can reflect. For binomial GLMs no such default priors exist.

This gap matters because the nonlinear link function in GLMs creates a prior-link interaction: a prior that appears reasonable on the regression-coefficient scale (i.e., the scale of the linear combination of predictors) can produce implausible distributions on the probability scale [@gelman2008; @ghosh2018]. Previous work addressed this problem for posterior estimation [@schonbrodt2018; @wagenmakers2018a]; we address the complementary problem of choosing priors that produce well-calibrated Bayes Factors for hypothesis testing.

This difference matters because priors recommended for estimation are usually highly uninformative (i.e., very broad) to minimally influence the posteriors. For instance, @gelman2008 proposed weakly informative Cauchy(0, 2.5) priors for logistic regression. The R package `brms` [@burkner2017; @burkner2018] for Bayesian GLM places a `student_t(3, 0, 2.5)` prior on the intercept and a flat, improper prior on all population-level effects. Such weakly informative priors on effects are not suited for hypothesis testing with Bayes Factors, because they bias the outcome in favor of the null hypothesis.

We provide a framework for specifying priors in binomial GLMs that covers both principled defaults - for researchers with no domain-specific knowledge - and the tools to derive informed priors from domain knowledge. Our approach proceeds in three steps:

1. Identify matched priors via the probability integral transform. We identify prior families that mirror the link function, yielding exactly uniform distributions on the probability scale. These matched priors serve as principled anchors from which informed priors can be derived by appropriate scaling.

2. Perform prior predictive simulation. We simulate priors on the regression-coefficient scale and investigate the resulting priors on the probability scale. In this way we characterise the implied distributions of baseline probability, effect size, and between-subject variability across 720 factorial combinations of link function, prior family, and prior scale (i.e. how wide the prior is).

3. Bayes Factor calibration. We evaluate how the recommended priors perform for hypothesis testing, examining how the strength of evidence expressed in the BF varies across prior specifications and effect magnitudes.

For all these steps, we model accuracy data as $k_i \sim \text{Binomial}(n_i, p_i)$ with the linear model $g(p_i) = b_0 + b_1 x_i$, where $g$ is either the logit or probit link. All simulations use a two-condition design with a single predictor variable $x$. We code $x$ with a sum-to-zero contrast $x \in \{+1, -1\}$, so that the difference on the probability scale is $\delta p = g^{-1}(b_0 + b_1) - g^{-1}(b_0 - b_1)$. This coding ensures the prior predictive distribution of $\delta p$ is symmetric around zero when the prior on $b_1$ is symmetric - a desirable property for a default prior. For researchers using other contrast codes or predictor scales, we provide scaling formulas in the [Recommendations].

## The Matched Prior Principle {#sec-matched-prior}

We first consider the prior for the intercept $b_0$ that, with sum-to-zero contrast coding, represents the grand mean. The key insight is simple: when the prior distribution on the regression-coefficient scale belongs to the same family as the link function, it produces a flat (uninformative) prior on the probability scale. This follows from the probability integral transform: applying a distribution's own CDF to a random variable drawn from that distribution always yields a uniform result. More formally, for any continuous distribution $D$ with CDF $F$, if $X \sim D$, then $F(X) \sim \text{Uniform}(0, 1)$. In a binomial GLM, the inverse link $g^{-1}$ *is* a CDF, so we can exploit this result directly.

| Link | Inverse link $g^{-1}$ | Matched prior | Implied prior on $p$ |
|------|----------------------|---------------|----------------------|
| logit | logistic CDF | $\text{Logistic}(0, 1)$ | $\text{Uniform}(0, 1)$ |
| probit | normal CDF $\Phi$ | $\text{Normal}(0, 1)$ | $\text{Uniform}(0, 1)$  |

In other words, when we transform a $\text{Logistic}(0, 1)$ prior on the regression-coefficient scale back to the probability scale, the result is a uniform distribution on $(0, 1)$; the same holds for a $\text{Normal}(0, 1)$ prior under the probit link. Conversely, scaling the prior with an $\text{sd} > 1$ concentrates probability mass toward the extreme values 0 and 1, reflecting the typically unreasonable assumption that extreme baseline probabilities are particularly likely. Scaling with $\text{sd} < 1$ concentrates mass away from 0 and 1, producing a regularized (i.e., shrunk away from extreme values) intercept prior suited to experiments where floor and ceiling performance are unlikely.

The matched prior at scale 1 is therefore maximally uninformative on the probability scale - a principled anchor. This follows recommendations by @gelman2017prior, who cautioned that priors on regression coefficients are difficult to interpret in isolation and should be understood in the context of the likelihood. The matched prior principle addresses this concern by anchoring the prior on an interpretable probability scale. The principle generalises to any model with a CDF-based link function; we return to this point in the [Recommendations].

The matched prior for the intercept primarily improves interpretability of prior predictive distributions. As the BF calibration analysis shows (@fig-bf-violin-intercept), its effect on Bayes Factor performance is negligible. Our recommendation to use them rests on methodological hygiene - it is the principled choice even when practical consequences are small. The BF estimates are much more strongly influenced by the priors on the effect sizes, to which we turn now.

## Priors for Effect Sizes {#sec-contrast-coding}

Our goal for setting priors on effect sizes is to reflect our prior knowledge about to-be-expected effect sizes on the probability scale. If we know nothing about the effect we are studying, we draw on our general knowledge about typical effect sizes in our discipline to define default priors; when we have specific knowledge about the effect sizes under investigation, we can use that knowledge to define more informed priors.

The width of the prior distribution depends on how we code the independent variable because the prediction depends on the product of the effect
parameter $b_n$ with the independent variable $x$. We recommend normalizing all independent variables to a standard deviation of 1 because then the effects have standardized effect sizes, and we can bring to bear our general knowledge about typical standardized effect sizes. For binary independent variables, this translates into coding them through sum-to-zero contrasts, which codes $x \in \{+1, -1\}$, for which the population standard deviation equals 1. With sum-to-zero contrasts, the effect $b_1$ represents the deviation of each condition from the grand mean, and the distribution of $\delta p$ is symmetric around zero when the prior on $b_1$ is symmetric. All following simulations use $\pm 1$ coding [see @schad2020 for a thorough treatment of contrast coding in regression].

With treatment coding ($x \in \{0, 1\}$), one condition depends on $b_0$ alone and the other on $b_0 + b_1$, creating asymmetric prior uncertainty across conditions. If the intercept prior mean is shifted away from zero, this asymmetry also skews the prior predictive distribution of $\delta p$. We therefore recommend sum-to-zero ($\pm 1$) coding as the default for binary predictors, unless there are substantive reasons to prefer treatment coding - in which case the resulting asymmetry may actually be desirable, because it reflects a prior belief that effects are more likely in one direction. Guidance for adapting the recommended prior widths to other contrast codes and predictor scales is provided in the *From Defaults to Informed Priors* section below.

## Prior Predictive Analysis {#sec-prior-predictive}

For the prior-predictive simulations we analyze the two popular link functions for binomial GLMs, the logit and the probit function. We consider three distribution families for priors for the intercept, $b_0$, as well as for fixed effects of predictor variables, $b_1$ to $b_n$: Normal, Logistic, and Cauchy. The first two were included because they each match one of the link functions; the Cauchy was included because @rouder2012 proposed Cauchy distributed default priors for linear models. Finally, we vary the standard deviation (SD) of the prior for the intercept, $sd_{b_0}$, from 0.5 to 1.5, and the SD of the prior for fixed effects, $sd_{b_1}$, from 0.1 to 2.0 (with an additional grid point at 0.25 to provide finer resolution in the expected recommendation range). The goal for this prior predictive simulation is to characterise the implied distributions of baseline probability, effect size, and between-subject variability across the 720 factorial combinations of link function, prior family, and prior scale.

### Intercept Prior on the Probability Scale {#sec-pp-intercept}

The intercept $b_0$ determines the baseline success probability before any experimental manipulation. A poorly chosen intercept prior can place substantial mass near $p = 0$ or $p = 1$, predicting that participants perform at floor or ceiling even before any experimental effect is applied. How do different intercept priors perform? In the figures below, we overlay exact analytical density curves (derived by transforming the prior density through the inverse link function) as reference.[^density-formulas]

[^density-formulas]: For the logit link with a $\text{Logistic}(0, s)$ prior on $b_0$: $f(p) = f_{\text{Logis}}(\text{logit}(p), 0, s) / [p(1-p)]$. At $s = 1$, the logistic density and the Jacobian $p(1-p)$ cancel, giving $f(p) = 1$. For the probit link with a $\text{Normal}(0, \sigma)$ prior: $f(p) = \phi(\Phi^{-1}(p)/\sigma) / [\sigma \cdot \phi(\Phi^{-1}(p))]$. At $\sigma = 1$, the normal densities cancel, again yielding $f(p) = 1$.

```{r}
#| label: fig-1-p-intercept
#| fig-cap: "Prior predictive distribution of success probability
#|   p = $g^{-1}(b_0)$ for each intercept prior SD (rows) and link function
#|   (columns). At $sd_{b_0} = 1.0$, the matched prior (Logistic for logit,
#|   Normal for probit) produces a perfectly uniform distribution - the
#|   visual signature of the probability integral transform. Reducing $sd_{b_0}$
#|   to 0.75 pulls mass away from floor and ceiling. Misfit priors (e.g.,
#|   Cauchy under either link) concentrate more mass near 0 and 1. Dashed
#|   curves show exact analytical densities."
#| fig-height: 8
#| fig-width: 10

fig1_draws <- draws_long |>
  filter(dist_b1 == "logistic", sd_b1 == 0.25)

plot_p_intercept_ridges(fig1_draws, sd_b1_focus = 0.25)
```

@fig-1-p-intercept confirms the analytical result: at $\text{sd}_{b_0} = 1.0$ with the matched prior, the distribution of $p$ is uniform. At $\text{sd}_{b_0} = 0.75$, the prior pulls mass away from 0 and 1, which is appropriate for experiments in which researchers avoid floor and ceiling effects. Misfit priors, such as Cauchy for either link, produce heavier tails and concentrate more mass near the boundaries.

```{r}
#| label: fig-2-floor-ceiling
#| fig-cap: "Proportion of prior predictive mass in the floor/ceiling region
#|   (p < 0.05 or p > 0.95) as a function of intercept prior SD and family.
#|   The matched prior at $sd_{b_0} = 1.0$ places exactly 10% in this region
#|   (2 x 0.05 = 0.10), confirming the uniform result. At $sd_{b_0} = 0.75$,
#|   floor/ceiling mass drops below 10% for both matched priors."
#| fig-height: 6
#| fig-width: 10

plot_floor_ceiling_heatmap(summaries_pp)
```

Experiments in cognitive psychology are typically designed to avoid floor and ceiling effects that might mask the effects of experimental conditions. This knowledge should be incorporated into the intercept prior by choosing a prior that places less probability mass at the extremes without being overly restrictive. @fig-2-floor-ceiling shows the proportion of prior predictive mass in the floor/ceiling region as a function of prior SD and family. At $\text{sd}_{b_0} = 1.0$, the matched priors produce exactly 10% floor/ceiling mass (5% in each tail), consistent with the uniform distribution and serving as an internal validation of the simulation against the analytical result. Any matched prior with $\text{sd}_{b_0} < 1$ satisfies the $< 10\%$ floor/ceiling criterion; we recommend $\text{sd}_{b_0} = 0.75$ as a practical default because it provides meaningful regularization away from the boundaries while remaining agnostic about baseline accuracy within the plausible range. This makes $\text{Logistic}(0, 0.75)$ our recommended default for the logit link and $\text{Normal}(0, 0.75)$ for the probit link.

### Effect Prior on the Probability Scale {#sec-pp-effect}

The prior on $b_1$ determines how much prior mass falls on small versus large effects on the probability scale. @fig-3-delta-p shows how the choice of $\text{sd}_{b_1}$ translates into implied effect sizes ($\delta p$), holding the intercept prior fixed at the matched recommendation ($\text{Logistic}(0, 0.75)$ for logit; $\text{Normal}(0, 0.75)$ for probit).

The Cauchy prior stands apart at every scale: even at $\text{sd}_{b_1} = 0.25$, its heavy tails assign substantial probability to $|\delta p| > 0.50$, making it a poor default for cognitive psychology experiments. Note that effects of $|\delta p| = 1$ are only possible if participants have perfect accuracy in one condition and not a single correct response in the other - a scenario that is virtually impossible in practice. Within the substantively appropriate range of $\text{sd}_{b_1} \in [0.10, 0.50]$, Normal and Logistic priors behave similarly - researchers can use either without meaningful loss.

We use the 90th percentile of $|\delta p|$ as the key summary statistic, mirroring the 10% floor/ceiling criterion from the intercept analysis: less than 10% of the prior predictive mass should fall in implausible regions, or in other words, the 90th percentile indicates the largest effect that the prior treats as common. At $\text{sd}_{b_1} = 0.25$, the 90th percentile of $|\delta p|$ is near 0.20 - already a very large effect by the standards of most psychological research, though still within the range observed in some cognitive paradigms (e.g., set-size effects in working memory). Values up to $\text{sd}_{b_1} = 0.50$ remain defensible when very large effects are plausible, with the 90th percentile rising to roughly 0.35.

For fixed-effects models, an effect prior with $\text{sd}_{b_1} \in [0.25, 0.30]$ under sum-to-zero coding provides a sensible default range. Most experiments, however, also require random effects to account for between-subject heterogeneity, which raises the question of how the hyperprior that controls the variability between subjects should be specified.


```{r}
#| label: fig-3-delta-p
#| fig-cap: "Prior predictive distribution of delta_p (signed effect size on
#|   the probability scale) for each effect prior SD, colour-coded by prior
#|   family. At sd_b1 = 0.25, Normal and Logistic priors concentrate mass on
#|   effects up to about 20 percentage points - a range consistent with
#|   typical cognitive experiments. The Cauchy stands apart even at this
#|   moderate scale, assigning substantial probability to effects exceeding
#|   50 percentage points. Dotted lines mark +/-0.10, +/-0.20, +/-0.30."
#| fig-height: 14
#| fig-width: 12

# Use the link-matched recommended intercept prior for each link:
# Logistic(0,0.75) for logit, Normal(0,0.75) for probit
fig2_draws <- draws_long |>
  filter(
    (link == "logit"  & dist_b0 == "logistic") |
    (link == "probit" & dist_b0 == "normal"),
    sd_b0 == 0.75
  )

plot_delta_p_ridges(fig2_draws, sd_b0_focus = 0.75)
```

### Random Effects Hyperprior {#sec-pp-re}

Random effects in hierarchical models are controlled by a prior on the SD parameter $\sigma$ (the standard deviation of the distribution from which individual effects are drawn). Candidate hyperpriors should therefore be defined on the positive reals. Any principled hyperprior for a SD parameter should have three properties: (a) mode at zero, expressing a preference for parsimony; (b) monotonically decreasing density, so that larger variability receives progressively less prior support; and (c) scale calibrated to domain knowledge about plausible between-subject variability [@gelman2006; @simpson2017].

Several distribution families satisfy these requirements, each with a distinct theoretical motivation. The **Gamma** family is connected to the sampling distribution of variances: if observations are normally distributed, the sample variance follows a scaled chi-squared distribution, which is a special case of the Gamma ($\chi^2(k) = \text{Gamma}(k/2, 1/2)$). The **Exponential** distribution is the maximum-entropy distribution on $[0, \infty)$ given only a mean constraint -- it encodes the least additional structure beyond a target scale. It also arises as the penalized-complexity (PC) prior [@simpson2017], which penalises deviation from the base model ($\sigma = 0$) at a constant rate. **Half-Normal** and **Half-Student-t** distributions were recommended by @gelman2006 as replacements for inverse-Gamma priors, with the Half-t providing more robust tail behaviour for small numbers of groups [@polson2012].

To disentangle the effects of distribution family and scale, we conducted a fully crossed simulation comparing four hyperprior families -- Exponential, Gamma(shape = 2), Half-Normal, and Half-Student-t(df = 3) -- at five scale levels matched by their mean: 0.15, 0.25, 0.50, 1.00, and 2.50. The Half-Student-t is included primarily as a reference point for comparison with the `brms` default `half-student_t(3, 0, 2.5)` (mean $\approx$ 2.76), which corresponds to the largest scale level. This simulation uses the logit link; by the matched-prior principle, probit results are expected to be analogous. The fixed-effects priors are held at the recommended values ($b_0 \sim \text{Logistic}(0, 0.75)$, $b_1 \sim \text{Logistic}(0, 0.25)$). We ran each configuration with random intercepts only and with random intercepts plus slopes, where $\sigma_v$ is drawn from the same hyperprior family and scale as $\sigma_u$.

```{r}
#| label: setup-re-pp
#| include: false

re_pp_summaries <- tryCatch(
  readRDS(here("output", "prior_predictive_re_summaries.rds")),
  error = function(e) NULL
)
re_pp_draws <- tryCatch(
  readRDS(here("output", "prior_predictive_re_draws.rds")),
  error = function(e) NULL
)

re_pp_available <- !is.null(re_pp_summaries) && !is.null(re_pp_draws)
```

```{r}
#| label: fig-re-1-p-intercept
#| fig-cap: "Subject-level success probabilities by hyperprior family and scale
#|   (mean of the hyperprior). Rows: hyperprior family. Columns: random-effects
#|   structure. Within each panel, facets show scale levels. At matched scale,
#|   all four families produce nearly identical distributions; the primary driver
#|   is the scale parameter. Dashed vertical lines mark p = 0.05 and p = 0.95."
#| fig-height: 10
#| fig-width: 12
#| eval: !expr re_pp_available

re_pp_draws |>
  group_by(hp_family, hp_mean, re_structure) |>
  slice_sample(n = 5000) |>
  ungroup() |>
  ggplot(aes(x = p_intercept, y = hp_family,
             fill = hp_family, colour = hp_family)) +
  geom_density_ridges(alpha = 0.5, scale = 0.9, rel_min_height = 0.01) +
  geom_vline(xintercept = c(0.05, 0.95), linetype = "dashed",
             colour = "grey40", linewidth = 0.5) +
  facet_grid(hp_mean ~ re_structure,
             labeller = labeller(
               hp_mean = function(x) lapply(paste0("mu == ", x), function(s) parse(text = s)))) +
  scale_fill_viridis_d(begin = 0.1, end = 0.85, guide = "none") +
  scale_colour_viridis_d(begin = 0.1, end = 0.85, guide = "none") +
  scale_x_continuous(
    expression("Success probability" ~ italic(p)),
    limits = c(0, 1), expand = c(0, 0)
  ) +
  scale_y_discrete("Family") +
  theme_minimal(base_size = 16) +
  theme(strip.text = element_text(face = "bold"))
```

@fig-re-1-p-intercept reveals the central finding: at matched scale (i.e., the same hyperprior mean), all four distribution families produce nearly identical prior predictive distributions of subject-level success probabilities. The primary driver is the scale of the hyperprior, not the distribution family. At a mean of 0.15--0.25, subject-level probabilities are well concentrated away from floor and ceiling across all families. At the largest scale (mean = 2.50, corresponding to the brms default territory), all families spread subject-level probabilities heavily toward 0 and 1.

```{r}
#| label: fig-re-2-delta-p
#| fig-cap: "Effect size heterogeneity by hyperprior family and scale for the
#|   intercept + slope model. At matched scale, all families produce similar
#|   effect heterogeneity. Dotted lines mark +/-0.10, +/-0.20, +/-0.30."
#| fig-height: 10
#| fig-width: 10
#| eval: !expr re_pp_available

re_pp_draws |>
  filter(re_structure == "Intercept + slope") |>
  group_by(hp_family, hp_mean) |>
  slice_sample(n = 10000) |>
  ungroup() |>
  ggplot(aes(x = delta_p, y = hp_family,
             fill = hp_family, colour = hp_family)) +
  geom_density_ridges(alpha = 0.5, scale = 0.9, rel_min_height = 0.01) +
  geom_vline(xintercept = 0, linewidth = 0.5, colour = "grey20") +
  geom_vline(xintercept = c(-0.30, -0.20, -0.10, 0.10, 0.20, 0.30),
             linetype = "dotted", linewidth = 0.5, colour = "grey40") +
  facet_wrap(~ hp_mean, ncol = 1,
             labeller = labeller(
               hp_mean = function(x) lapply(paste0("mu == ", x), function(s) parse(text = s)))) +
  scale_fill_viridis_d(begin = 0.1, end = 0.85, guide = "none") +
  scale_colour_viridis_d(begin = 0.1, end = 0.85, guide = "none") +
  scale_x_continuous(
    expression(delta * italic(p)),
    limits = c(-1, 1), breaks = seq(-1, 1, 0.25)
  ) +
  scale_y_discrete("Family") +
  theme_minimal(base_size = 16) +
  theme(strip.text = element_text(face = "bold"))
```

The same pattern holds for effect heterogeneity (@fig-re-2-delta-p): at a given scale, the four families produce very similar distributions of subject-level $\delta p$ values. What changes dramatically is the scale: at a mean of 2.50, a non-trivial fraction of simulated subjects shows effects near $|\delta p| = 1$ -- a scenario requiring perfect accuracy in one condition and zero in the other, which is virtually impossible in practice. At a mean of 0.25, between-subject variability is concentrated in a substantively plausible range.

We therefore recommend choosing the hyperprior scale such that the prior concentrates between-subject SD below approximately 0.5 logit units. At a hyperprior mean of 0.25, all four families achieve this -- confirming that the distribution family is secondary to the scale. Among the candidates that satisfy this criterion, we recommend the Exponential distribution as the default for three reasons: (1) it is the maximum-entropy distribution on $[0, \infty)$ given only a mean constraint, encoding the least additional structure beyond the target scale; (2) it corresponds to the penalized-complexity (PC) prior, which penalises deviation from the base model at a constant rate; and (3) it has a single parameter (the rate), which directly determines the mean ($\text{mean} = 1/\text{rate}$), making it straightforward to calibrate. An $\text{Exponential}(4)$ hyperprior (mean = 0.25) is our recommended default. For researchers who expect somewhat larger between-subject variability, a $\text{Gamma}(2, 4)$ (mean = 0.50) provides a slightly wider but still well-calibrated alternative. Researchers who prefer half-distribution families (e.g., Half-Normal) can achieve equivalent results by matching the mean to 0.25; the choice of family is a matter of convenience.

## Bayes Factor Calibration {#sec-bf-calibration}

The prior predictive analysis established which prior families and scales produce plausible distributions on the probability scale. We now ask whether these recommended priors also yield well-calibrated Bayes Factors for hypothesis testing. Specifically, we examine how the strength of evidence depends on prior specification -- family, scale, intercept matching, and hyperprior choice -- when applied to realistic simulated data. This analysis verifies that our prior recommendations are not only interpretable but also effective for Bayesian inference.

### Simulation Design {#sec-bf-design}

We simulated data across a range of realistic experimental scenarios. The true value of the effect coefficient $b_1$ ranged from $b_1 = 0$ (no effect) to $b_1 = 0.50$ on the regression coefficient scale, covering the span from a null effect through a large effect in cognitive experiments (on the probability scale with a typical 50% baseline, $b_1 = 0.50$ corresponds to roughly 24 percentage points difference between conditions). We crossed these data-generating factors with three distribution families for effect priors (Normal, Logistic, Cauchy) at two scale values ($\text{sd}_{b_1} = 0.25$ and $0.50$), and two intercept priors (Logistic and Normal, both at $\text{sd}_{b_0} = 0.75$; as we only use logistic models in this simulation, the former is a matched prior whereas the latter is not). To ensure results are not specific to a single sample size, we varied the number of subjects (30, 60, 100) and trials per subject (20, 50), yielding 360 conditions with 100 replications each. We present results for the most data-sparse scenario ($N = 30$, 20 trials per subject), where prior sensitivity is most visible; the pattern of results is consistent across all sample sizes.

A second simulation specifically investigated the random-effects hyperprior. It used a within-subjects design with correlated random intercepts and slopes, holding the fixed-effects priors at the recommended matched configuration (Logistic intercept and effect prior) and varying only the hyperprior on the random-effects SDs across three candidates: The brms default Student-*t*(3, 0, 2.5), Gamma(2, 4), and Exponential(4). In the main BF calibration above, the random-effects hyperprior was held constant at Exponential(4); this second simulation tests sensitivity to that choice. It produced 90 conditions with 100 replications each. We return to these results in the *Effect of Random-Effects Hyperprior* section below.

#### How We Computed Bayes Factors

All models used the logit link (the probit link is expected to produce analogous results via the matched-prior principle). To ensure that results generalize across realistic baseline accuracies, we did not fix the true intercept: In each replication, $b_0$ was sampled from $\text{Uniform}(0.4, 0.9)$ on the probability scale and transformed to the logit scale. Each simulated dataset included subject-level random intercepts ($u_i \sim \mathcal{N}(0, 0.25)$) to mimic between-subject heterogeneity.

The analysis model mirrored this structure: A binomial GLMM with one random intercept per subject, fitted with brms [@burkner2017], with a prior on the intercept mean with SD fixed at 0.75 and an $\text{Exponential}(4)$ hyperprior on the random-intercept SD. Bayes Factors were computed via the Savage--Dickey density ratio [@wagenmakers2018; @ly2016], which estimates the BF in favour of an effect as the ratio between prior and posterior density at zero: $\text{BF}_{10} = \pi(b_1 = 0) / p(b_1 = 0 \mid \text{data})$.

### How Do Priors Affect the Strength of Evidence? {#sec-bf-strength}

To examine how the choice of priors affects the strength of evidence, we focus on the scenarios where priors are most likely to matter: small to moderate effects ($b_1 \in \{0, 0.10, 0.20\}$) with $N = 30$ subjects and 20 trials per subject - the most data-sparse conditions in our simulation. We display the full distribution of $\log_{10}(\text{BF}_{10})$ as violin plots, separately for the effect prior family and width, the intercept prior, and the random-effects hyperprior.

```{r}
#| label: setup-bf-raw
#| include: false

# Load raw BF10 values from per-condition files for violin plots
bf_raw <- tryCatch({
  results_dir <- here("output", "Simulation_BFCalibration")
  cond_files <- list.files(results_dir, pattern = "^BFCalib_Cond-",
                           full.names = TRUE)
  raw_rows <- lapply(cond_files, function(f) {
    d <- readRDS(f)
    bf10_vec <- d$results[, "BF10"]
    cbind(
      as.data.frame(as.list(d$condition)),
      data.frame(BF10 = bf10_vec, log10_BF10 = log10(bf10_vec))
    )
  })
  bind_rows(raw_rows)
}, error = function(e) NULL)
bf_raw_available <- !is.null(bf_raw)

# Load raw RE BF10 values
bf_raw_re <- tryCatch({
  results_dir_re <- here("output", "Simulation_BFCalibration_RE")
  cond_files_re <- list.files(results_dir_re, pattern = "^BFCalib_RE_Cond-",
                              full.names = TRUE)
  raw_rows_re <- lapply(cond_files_re, function(f) {
    d <- readRDS(f)
    bf10_vec <- d$results[, "BF10"]
    cbind(
      as.data.frame(as.list(d$condition)),
      data.frame(BF10 = bf10_vec, log10_BF10 = log10(bf10_vec))
    )
  })
  bind_rows(raw_rows_re)
}, error = function(e) NULL)
bf_raw_re_available <- !is.null(bf_raw_re)
```

#### Effect of Prior Family and Width

```{r}
#| label: fig-bf-violin-family
#| fig-cap: "Distribution of log10(BF10) by effect prior family and width for
#|   N = 30 subjects and 20 trials per subject. Rows: true effect size. Columns:
#|   prior SD. The choice of prior family has minimal impact on the evidence
#|   distribution; prior width shifts the distribution modestly. Dashed lines
#|   mark BF10 = 3 and BF10 = 1/3."
#| fig-height: 7
#| fig-width: 10
#| eval: !expr bf_raw_available

bf_raw |>
  filter(
    dist_b0 == "logistic",
    true_b1 %in% c(0, 0.10, 0.20),
    n_subjects == 30,
    n_trials == 20
  ) |>
  mutate(
    dist_b1 = factor(dist_b1, levels = c("cauchy", "normal", "logistic"),
                     labels = c("Cauchy", "Normal", "Logistic")),
    true_b1_label = factor(paste0("b[1] == ", true_b1)),
    sd_b1_label   = factor(paste0("sd[b[1]] == ", sd_b1))
  ) |>
  ggplot(aes(x = dist_b1, y = log10_BF10, fill = dist_b1)) +
  geom_hline(yintercept = c(log10(3), log10(1/3)), linetype = "dashed",
             colour = "grey50", linewidth = 0.4) +
  geom_hline(yintercept = 0, colour = "grey30", linewidth = 0.3) +
  geom_violin(alpha = 0.6, scale = "width", width = 0.8) +
  stat_summary(fun = median, geom = "point", size = 2, colour = "black") +
  facet_grid(true_b1_label ~ sd_b1_label, labeller = label_parsed) +
  scale_fill_viridis_d("Prior family", option = "D", end = 0.85) +
  labs(
    x = "Effect prior family",
    y = expression(log[10](BF[10]))
  ) +
  base_theme()
```

@fig-bf-violin-family shows that the choice of effect prior family has minimal impact on the distribution of evidence. All three families produce nearly identical BF distributions at each effect size. The prior width ($\text{sd}_{b_1}$) has a modest effect: the narrower prior ($\text{sd} = 0.25$) yields slightly stronger evidence for H1 when the effect is present, and slightly weaker evidence for H0 under the null, compared to the wider prior ($\text{sd} = 0.50$). The Cauchy prior performs similarly to the Normal and Logistic at these scale values but should be avoided due to its implausible prior predictive distributions (see *Effect Prior on the Probability Scale* above).

#### Effect of Intercept Prior

```{r}
#| label: fig-bf-violin-intercept
#| fig-cap: "Distribution of log10(BF10) by intercept prior (matched Logistic
#|   vs. misfit Normal) for N = 30, 20 trials, Logistic effect prior with
#|   sd = 0.25. The matched and misfit intercept priors produce virtually
#|   identical BF distributions. Dashed lines mark BF10 = 3 and BF10 = 1/3."
#| fig-height: 5
#| fig-width: 8
#| eval: !expr bf_raw_available

bf_raw |>
  filter(
    dist_b1 == "logistic",
    sd_b1 == 0.25,
    true_b1 %in% c(0, 0.10, 0.20),
    n_subjects == 30,
    n_trials == 20
  ) |>
  mutate(
    dist_b0 = factor(dist_b0, levels = c("logistic", "normal"),
                     labels = c("Logistic (matched)", "Normal (misfit)")),
    true_b1_label = factor(paste0("b[1] == ", true_b1))
  ) |>
  ggplot(aes(x = dist_b0, y = log10_BF10, fill = dist_b0)) +
  geom_hline(yintercept = c(log10(3), log10(1/3)), linetype = "dashed",
             colour = "grey50", linewidth = 0.4) +
  geom_hline(yintercept = 0, colour = "grey30", linewidth = 0.3) +
  geom_violin(alpha = 0.6, scale = "width", width = 0.7) +
  stat_summary(fun = median, geom = "point", size = 2, colour = "black") +
  facet_wrap(~ true_b1_label, nrow = 1, labeller = label_parsed) +
  scale_fill_viridis_d("Intercept prior", option = "D", end = 0.7) +
  labs(
    x = "Intercept prior",
    y = expression(log[10](BF[10]))
  ) +
  base_theme()
```

@fig-bf-violin-intercept confirms that the matched-prior principle does not meaningfully affect Bayes Factor calibration. The matched (Logistic) and misfit (Normal) intercept priors produce virtually identical BF distributions across all effect sizes. The recommendation for the matched prior therefore rests on its prior predictive properties and interpretability, not on BF performance. 

#### Effect of Random-Effects Hyperprior {#sec-re-hyperprior}

```{r}
#| label: fig-bf-violin-re
#| fig-cap: "Distribution of log10(BF10) by random-effects hyperprior for
#|   N = 30, 20 trials. Within-subjects design with correlated random
#|   intercepts and slopes, matched fixed-effects priors. All three hyperpriors
#|   produce nearly identical BF distributions. Dashed lines mark BF10 = 3
#|   and BF10 = 1/3."
#| fig-height: 5
#| fig-width: 8
#| eval: !expr bf_raw_re_available

bf_raw_re |>
  filter(
    true_b1 %in% c(0, 0.10, 0.20),
    n_subjects == 30,
    n_trials == 20
  ) |>
  mutate(
    sd_prior_re = factor(sd_prior_re,
                         levels = c("default", "gamma", "exponential"),
                         labels = c("Student-t(3,0,2.5)", "Gamma(2,4)",
                                    "Exponential(4)")),
    true_b1_label = factor(paste0("b[1] == ", true_b1))
  ) |>
  ggplot(aes(x = sd_prior_re, y = log10_BF10, fill = sd_prior_re)) +
  geom_hline(yintercept = c(log10(3), log10(1/3)), linetype = "dashed",
             colour = "grey50", linewidth = 0.4) +
  geom_hline(yintercept = 0, colour = "grey30", linewidth = 0.3) +
  geom_violin(alpha = 0.6, scale = "width", width = 0.7) +
  stat_summary(fun = median, geom = "point", size = 2, colour = "black") +
  facet_wrap(~ true_b1_label, nrow = 1, labeller = label_parsed) +
  scale_fill_viridis_d("RE hyperprior", option = "D", end = 0.85) +
  labs(
    x = "Random-effects hyperprior",
    y = expression(log[10](BF[10]))
  ) +
  base_theme()
```

@fig-bf-violin-re shows that the choice of random-effects hyperprior has no meaningful impact on the distribution of Bayes Factor evidence. All three hyperpriors -- Student-*t*(3, 0, 2.5), Gamma(2, 4), and Exponential(4) -- produce nearly identical BF distributions across all effect sizes. Note that the Student-*t*(3, 0, 2.5) is included at its `brms` default scale (mean $\approx$ 2.76), which is far wider than the Gamma(2, 4) (mean = 0.50) and Exponential(4) (mean = 0.25). As the prior predictive analysis showed (see *Random Effects Hyperprior* above), at matched scale, all distribution families produce virtually identical results -- the key driver is the scale, not the family. The fact that even this large difference in scale does not meaningfully shift the BF distribution indicates that the random-effects hyperprior has little influence on the Bayes Factor for the fixed-effect of interest, which is the primary target of hypothesis testing. Our recommendation for Exponential(4) therefore rests on its desirable prior predictive properties and its simplicity (one parameter), not on BF calibration differences.

### Summary {#sec-bf-summary}

Across all comparisons, prior *scale* ($\text{sd}_{b_1}$) has the largest effect on BF calibration, followed by the amount of data. Prior family, intercept prior, and RE hyperprior have negligible impact. We recommend the matched Logistic intercept prior for interpretability and the Logistic effect prior at $\text{sd}_{b_1} \in [0.25, 0.30]$ as the default, but these choices are not critical for BF performance.

## Recommendations {#sec-recommendations}

For binomial GLMs, we recommend the following prior specifications given below. Researchers with no domain-specific prior knowledge can use the default priors directly. Those with prior information about expected effect sizes can use the scaling formulas given below to adapt the defaults to their context.

### Default Prior Specifications

#### Intercept prior

| Link | Recommended prior | Rationale |
|------|------------------|-----------|
| logit | $\text{Logistic}(0, 0.75)$ | Matched family; sd = 0.75 gives < 10% floor/ceiling mass |
| probit | $\text{Normal}(0, 0.75)$ | Matched family; same criterion |

At $\text{sd} = 0.75$, less than 10% of prior mass falls in the floor/ceiling region, while remaining agnostic about baseline accuracy within the plausible range.

#### Effect prior

| Link | Recommended prior | Rationale |
|------|------------------|-----------|
| logit | $\text{Logistic}(0, 0.25\text{--}0.30)$ | Matched family; 90th percentile of $|\delta p| \approx 0.20\text{--}0.23$ |
| probit | $\text{Normal}(0, 0.25\text{--}0.30)$ | Matched family; 90th percentile of $|\delta p| \approx 0.18\text{--}0.22$ |

Our default recommendation is the matched effect prior at $\text{sd}_{b_1} \in [0.25, 0.30]$. Within this range, the 90th percentile of $|\delta p|$ stays between approximately 0.18 and 0.23 - well-calibrated for typical cognitive experiments. The Logistic prior offers the best specificity profile due to its lowest density at zero among the candidate families. Because Bayes Factors reflect how much the prior has been updated by the data, they depend on the prior scale. We therefore recommend running a prior sensitivity analysis by computing BFs across a small set of prior widths (e.g., $\text{sd}_{b_1} \in \{0.15, 0.25, 0.35\}$) to assess how the choice of scale influences the evidence.

#### Random effects prior

The hyperprior on random-effect SDs should concentrate mass below approximately 0.5 logit units. The choice of distribution family is secondary to the choice of scale (see *Random Effects Hyperprior* above). Our recommended default is `exponential(4)` (mean = 0.25) for its simplicity (one parameter, $\text{mean} = 1/\text{rate}$). For larger expected between-subject variability, `gamma(2, 4)` (mean = 0.50) is a well-calibrated alternative. The `brms` default `student_t(3, 0, 2.5)` places far too much mass on large random-effect SDs for typical cognitive experiments.

#### Ready-to-use brms prior specifications

```{r}
#| label: recommended-priors
#| echo: true
#| eval: false

# Logit link (logistic regression)
logit_priors <- c(
  brms::prior(logistic(0, 0.75), class = Intercept),
  brms::prior(logistic(0, 0.25), class = b)
)

# Probit link (probit regression)
probit_priors <- c(
  brms::prior(normal(0, 0.75), class = Intercept),
  brms::prior(normal(0, 0.25), class = b)
)
```

For models with subject-level random effects:

```{r}
#| label: recommended-priors-re
#| echo: true
#| eval: false

# Logit link with random intercept
logit_priors_re <- c(
  brms::prior(logistic(0, 0.75), class = Intercept),
  brms::prior(logistic(0, 0.25), class = b),
  brms::prior(exponential(4),    class = sd)
)

# Probit link with random intercept
probit_priors_re <- c(
  brms::prior(normal(0, 0.75), class = Intercept),
  brms::prior(normal(0, 0.25), class = b),
  brms::prior(exponential(4),  class = sd)
)
```

### From Defaults to Informed Priors {#sec-scaling}

Informed priors can be generated easily by adjusting default priors in two ways. The first adjustment is to make the priors wider or narrower (i.e., increasing or decreasing $\text{sd}_{b_1}$) depending on whether one expects larger or smaller effects compared to the default priors. A second adjustment needs to be made if the independent variables are not normalised. The recommended $\text{sd}_{b_1} \in [0.25, 0.30]$ assumes normalised $\pm 1$ coding ($\text{SD}(x) = 1$). Whenever the predictor's SD differs from 1, the prior on $b_1$ must be rescaled, using the following scaling formula:

$$\text{sd}_{b_1}(\text{predictor SD} = s) = \frac{\text{sd}_{b_1}(\text{SD} = 1)}{s}$$

Smaller predictor SDs require a *wider* prior on $b_1$; larger predictor SDs require a *narrower* prior. Here are some examples:

- $\pm 0.5$ coding ($s = 0.5$): use $\text{sd}_{b_1} = 0.25 / 0.5 = 0.50$
- $\pm 1/\sqrt{2}$ orthonormal ($s \approx 0.707$): use $\text{sd}_{b_1} = 0.25 / 0.707 \approx 0.354$
- $0/1$ treatment coding ($s = 0.5$ in a balanced design): use $\text{sd}_{b_1} = 0.25 / 0.5 = 0.50$
- Standardised continuous predictor ($s = 1$): use $\text{sd}_{b_1} = 0.25$ directly

#### Coding multi-level factors

When a factor has more than two levels, we recommend using theoretically meaningful orthogonal contrasts that reflect specific hypotheses. These could either be a set of orthogonal contrasts that breaks down factors with more levels into a sub-set of contrasts with only two levels, or transforming factors into continuous predictors when this is reasonable.

For researchers who do not have specific directional hypotheses but want to test the omnibus question "does the factor have any effect?", orthonormal contrasts provide a principled alternative. Non-orthonormal contrast codes (e.g., Helmert or simple effects coding) produce contrast columns with unequal variances. When a single prior width is applied to all coefficients, this results in different implied prior distributions for different pairwise comparisons -- some comparisons would receive wider priors and others narrower, without substantive justification. Orthonormal contrasts equalize the column variances, ensuring that a single prior width induces the same prior distribution on every pairwise comparison. The `contr.equalprior()` function from the `bayestestR` package [@makowski2019], following the construction in @rouder2012 [p. 363], produces contrast columns with this property. In a balanced design with $J$ levels, each orthonormal contrast column has predictor $\text{SD} = 1 / \sqrt{J}$, giving the rescaled prior width:

$$\text{sd}_{b_1}(\text{orthonormal, } J \text{ levels}) = 0.25 \times \sqrt{J}$$

An overview of the resulting predictor SDs and rescaled prior widths for $J = 2$ to $5$ is given in @tbl-orthonormal-scaling.

| Design term | Predictor SD | Rescaled $\text{sd}_{b_1}$ |
|:---|:---:|:---:|
| 2-level factor | $1/\sqrt{2} \approx 0.71$ | 0.35 |
| 3-level factor | $1/\sqrt{3} \approx 0.58$ | 0.43 |
| 4-level factor | $1/\sqrt{4} = 0.50$ | 0.50 |
| 5-level factor | $1/\sqrt{5} \approx 0.45$ | 0.56 |
| $2 \times 3$ interaction | $1/\sqrt{6} \approx 0.41$ | 0.61 |
| $3 \times 3$ interaction | $1/\sqrt{9} \approx 0.33$ | 0.75 |

: Rescaled effect prior widths under orthonormal contrasts for multi-level
  factors and factorial interactions. All values assume the baseline
  recommendation of $\text{sd}_{b_1} = 0.25$ under $\pm 1$ coding.
  {#tbl-orthonormal-scaling}

@fig-orthonormal-demo shows the resulting prior predictive distributions for $J = 2$, $3$, and $4$. As can be seen there, the prior predictive distribution of any single pairwise comparison is nearly identical across $J = 2$, $3$, and $4$: the $\sqrt{J}$ scaling rule preserves the calibration established for the two-condition case.

```{r}
#| label: fig-orthonormal-demo
#| fig-cap: "Prior predictive distribution of a single representative
#|   pairwise $\\delta p$ under orthonormal contrasts with the rescaled prior
#|   ($\\text{sd}_{b_1} = 0.25 \\times \\sqrt{J}$). The distribution is
#|   nearly identical across $J$ -- confirming that the scaling rule
#|   preserves per-comparison calibration."
#| fig-height: 4.5
#| fig-width: 6

set.seed(2024)
n_draws <- 50000

demo_list <- lapply(c(2, 3, 4), function(J) {
  oc <- orthonormal_contrasts(J)
  sd_b1_scaled <- 0.25 * oc$scaling_factor

  b0 <- rlogis(n_draws, location = 0, scale = 0.75)
  B <- matrix(
    rnorm(n_draws * (J - 1), mean = 0, sd = sd_b1_scaled),
    nrow = n_draws, ncol = J - 1
  )

  eta_mat <- outer(b0, rep(0, J)) + B %*% t(oc$Q)
  p_mat   <- ilogit(eta_mat)

  delta_p_pair <- p_mat[, 1] - p_mat[, 2]

  data.frame(
    J            = paste0("J = ", J),
    delta_p_pair = delta_p_pair
  )
})

demo_draws <- do.call(rbind, demo_list)

p_single <- ggplot(demo_draws, aes(x = delta_p_pair,
                                    fill = J, colour = J)) +
  geom_density(alpha = 0.3, linewidth = 0.6) +
  geom_vline(xintercept = c(-0.20, -0.10, 0.10, 0.20),
             linetype = "dotted", linewidth = 0.4,
             colour = "grey40") +
  scale_x_continuous(expression(delta * p ~ "(single pair)"),
                     limits = c(-0.8, 0.8)) +
  scale_fill_viridis_d("Factor levels", option = "D", end = 0.85) +
  scale_colour_viridis_d("Factor levels", option = "D", end = 0.85) +
  ggtitle("Single pairwise comparison") +
  prior_theme()

p_single
```



When using orthonormal contrasts in `brms`, the recommended workflow is to set orthonormal contrasts before fitting the model and adjust the prior width accordingly:

```{r}
#| label: recommended-priors-orthonormal
#| echo: true
#| eval: false

# - Single 3-level factor ---
contrasts(dat$condition) <- bayestestR::contr.equalprior(3)
# Predictor SD = 1/sqrt(3); rescaled prior: sd_b1 = 0.25 * sqrt(3) = 0.433

priors_3level <- c(
  brms::prior(logistic(0, 0.75),  class = Intercept),
  brms::prior(normal(0, 0.433),   class = b)
)

# - 2 x 3 factorial design ---
contrasts(dat$A) <- bayestestR::contr.equalprior(2)
contrasts(dat$B) <- bayestestR::contr.equalprior(3)

priors_factorial <- c(
  brms::prior(logistic(0, 0.75),  class = Intercept),
  brms::prior(normal(0, 0.354),   class = b, coef = A1),
  brms::prior(normal(0, 0.433),   class = b, coef = B1),
  brms::prior(normal(0, 0.433),   class = b, coef = B2),
  brms::prior(normal(0, 0.612),   class = b, coef = "A1:B1"),
  brms::prior(normal(0, 0.612),   class = b, coef = "A1:B2")
)
```

## Extension to Other Model Classes {#sec-extensions}

The workflow demonstrated here - analytical matched priors, systematic prior predictive simulation, and large-scale BF calibration - transfers directly to other model classes where link functions mediate between parameter and observation scales.

*Ordinal regression.* Cumulative logit and probit models use the same link functions as binomial GLMs, so the matched prior principle applies directly to the threshold parameters. A Logistic prior on logit-scale thresholds produces uniform implied category probabilities.

*Count data.* Poisson and negative binomial regression use a log link. The matched prior principle suggests an Exponential prior on the log-scale intercept. Prior predictive simulation can reveal what prior scales on log-scale coefficients imply for rate ratios.

*Beta regression.* For proportional data, beta regression models the mean via a logit link. The matched prior principle applies identically to the intercept.

*Distributional regression.* In GAMLSS [@klein2015] and brms distributional models [@burkner2021], *all* parameters of the response distribution (mean, variance, shape) are linked to predictors through link functions. Each parameter therefore faces its own prior--link interaction. The matched prior principle and the simulation approach developed here provide a systematic way to address prior specification for each parameter.

In each case, the key steps are: (1) identify the link function and the corresponding matched prior family via the probability integral transform; (2) use prior predictive simulation to map prior scales to interpretable quantities on the observation scale; and (3) if Bayes Factor testing is intended, run a calibration simulation to verify the sensitivity of the strength of evidence for one hypothesis over the other.

## References

::: {#refs}
:::

## Appendix A: Orthonormal Contrast Scaling {#sec-appendix-contrasts .appendix}

For unbalanced designs, the predictor SD depends on cell sizes and should be computed from the contrast matrix directly. For factors with many levels ($J > 5$), the `BayesFactor` package [@rouder2012; @rouder2017] implements the full multivariate JZS prior, which accounts for the dimensionality of the effect space without requiring per-coefficient prior widths.

The full scaling produces wider priors for interactions than for main effects. In many experimental contexts, interactions are expected to be smaller than main effects, so using the *unscaled* main-effect prior for interaction terms is a reasonable alternative. @tbl-orthonormal-scaling in the main text gives the default when no such expectation exists.

A subtlety arises when priors on interaction terms are specified independently of the main-effect priors. In principle, the priors on main effects A and B already imply a distribution over the interaction A $\times$ B, because the joint prior on all coefficients determines the prior on any function of those coefficients. Manually specifying a separate prior on the interaction term can create an inconsistency: the prior beliefs implied by the main-effect priors may not be compatible with the interaction prior. In practice, this is unlikely to cause problems when the interaction has a clear substantive interpretation (e.g., a theoretically motivated moderation effect) and the priors are set to reflect genuine prior beliefs. However, when using default priors for exploratory analyses, we recommend using the scaled priors from @tbl-orthonormal-scaling rather than independently chosen interaction priors, as the scaling rule ensures internal consistency across the coefficient space.

## Appendix B: No-Intercept Parameterisation {#sec-appendix-no-intercept .appendix}

In `brms`, specifying `0 + condition` removes the intercept and estimates separate coefficients for each condition level, so that $p_0 = g^{-1}(b_0)$, $p_1 = g^{-1}(b_1)$, and $\delta p = g^{-1}(b_1) - g^{-1}(b_0)$. Under this parameterization, the matched prior principle applies to each coefficient independently: placing $\text{Logistic}(0, 1)$ on both coefficients under the logit link induces independent $\text{Uniform}(0, 1)$ marginals, and the implied prior on $\delta p$ is a symmetric triangular distribution on $(-1, 1)$. Because the coefficient interpretation differs from the intercept-plus-slope parameterization, our recommended prior widths do not transfer directly - each condition mean should receive a matched prior scaled to reflect beliefs
about plausible success probabilities (e.g., $\text{sd} = 0.75$), and the implied $\delta p$ should be checked via prior predictive simulation.

