---
title: "Default Priors for Binomial GLMs: Prior Predictive Analysis"
shorttitle: "Priors for Binomial GLMs"
author:
  - name: Klaus Oberauer
    orcid: 0000-0001-8815-6189
    affiliations:
      - name: University of Zurich
        department: Department of Psychology
  - name: Joscha Dutli
    affiliations:
      - name: University of Zurich
        department: Department of Psychology
  - name: Philipp Musfeld
    affiliations:
      - name: University of Zurich
        department: Department of Psychology
  - name: Gidon T. Frischkorn
    orcid: 0000-0002-5055-9764
    email: gidon.frischkorn@unilu.ch
    corresponding: true
    affiliations:
      - name: University of Lucerne
        department: Faculty of Health Sciences and Medicine
author-note:
  disclosures:
    study-registration: "This study was not preregistered."
    data-sharing: "Simulation code and analysis scripts are available at [repository link]."
    conflict-of-interest: "The authors declare no conflicts of interest."
    financial-support: "This research was supported by a grant from the Swiss National Science Foundation (SNSF)."
    authorship-agreements: |
      Klaus Oberauer: Conceptualization, Methodology, Writing -- Original Draft.
      Joscha Dutli: Methodology, Software, Formal Analysis.
      Philipp Musfeld: Methodology, Software, Formal Analysis.
      Gidon T. Frischkorn: Conceptualization, Methodology, Software, Formal Analysis, Visualization, Writing -- Original Draft.
format:
  apaquarto-docx: default
  apaquarto-pdf:
    keep-tex: true
  apaquarto-html: default
execute:
  echo: false
  warning: false
  message: false
bibliography: references.bib
csl: apa.csl
abstract: |
  Bayesian inference for binomial generalized linear models (GLMs) requires
  priors on the regression coefficients, but no principled defaults exist for
  logistic or probit regression. The nonlinear link function creates a
  prior--link interaction: priors that appear reasonable on the coefficient
  scale can produce implausible distributions on the probability scale. We
  address this problem via two complementary approaches. First, we exploit
  the probability integral transform to identify *matched priors* --- prior
  families that mirror the link function and induce exactly uniform
  distributions on the probability scale. Second, we characterise implied
  prior predictive distributions across 720 factorial combinations of link
  function, prior family, and prior scale via large-scale simulation. Based
  on these analyses, we recommend matched intercept priors at scale 0.75,
  effect priors at scale 0.25, and Exponential(4) hyperpriors for
  random-effects SDs. Ready-to-use brms prior specifications are provided.
---

```{r}
#| label: setup
library(dplyr)
library(tidyr)
library(ggplot2)
library(ggridges)
library(viridis)
library(scales)
library(here)
library(knitr)

source(here("R", "link_functions.R"))
source(here("R", "prior_sampling.R"))
source(here("R", "plotting.R"))

# Load simulation results (run scripts/01 and 02 first)
summaries  <- readRDS(here("output", "prior_predictive_summaries.rds"))
draws_long <- readRDS(here("output", "prior_predictive_draws.rds"))
```

## Introduction

A researcher fitting a logistic regression in `brms` faces an immediate practical question: what prior should they place on the regression coefficients? For normally distributed data, @rouder2012 established the widely adopted JZS prior, enabling routine Bayes Factor computation via JASP and BayesFactor. No equivalent defaults exist for binomial GLMs --- the workhorse model for accuracy data ($k$ correct out of $n$ trials) in cognitive psychology.

This gap matters because the nonlinear link function in GLMs creates a **prior--link interaction**: a prior that appears reasonable on the regression-coefficient scale can produce implausible distributions on the probability scale [@gelman2008; @ghosh2018]. Previous work addressed this
problem for posterior estimation; we address the complementary problem of
choosing priors that produce well-calibrated *Bayes Factors* for hypothesis
testing [@wagenmakers2018a; @schonbrodt2018].

We model accuracy data as $k_i \sim \text{Binomial}(n_i, p_i)$ with
$g(p_i) = b_0 + b_1 x_i$, where $g$ is either the **logit** or **probit**
link. We address four questions:

1. **What prior family and scale produce well-behaved baseline probabilities?**
   The intercept prior determines the distribution of $p$ before any
   experimental manipulation --- can we identify a principled default?
2. **What effect prior scale concentrates mass on plausible effect sizes?**
   The prior on $b_1$ implies a distribution of $\delta p$ --- the difference
   in success probability between conditions --- on the probability scale.
   What scale matches typical effects in cognitive experiments?
3. **Does the choice of prior family matter?** Normal, Logistic, and Cauchy
   priors differ in tail behaviour --- how much does this affect the implied
   prior predictive distribution?
4. **How should priors be specified for random-effects models?** Most
   experiments require subject-level random effects --- what hyperprior keeps
   between-subject variability in a substantively plausible range?

Following the Bayesian workflow outlined by @gelman2020workflow and
@schad2021, we combine two strategies: (1) analytical derivation of *matched
priors* via the probability integral transform, yielding exact results for
the intercept prior [@gelman1996; @gabry2019; @betancourt2018]; and (2)
large-scale prior predictive simulation across 720 factorial combinations of
link function (logit, probit), intercept prior family (Logistic, Normal,
Cauchy), intercept prior SD (0.50--1.50), effect prior family (Normal,
Logistic, Cauchy), and effect prior SD (0.10--2.00). All simulations use a
two-condition design with **sum-to-zero contrast** $x \in \{+1, -1\}$, so
that $\delta p = g^{-1}(b_0 + b_1) - g^{-1}(b_0 - b_1)$. This coding
ensures the prior predictive distribution of $\delta p$ is symmetric around
zero when the prior on $b_1$ is symmetric --- a desirable property for a
default prior (see [How does contrast coding interact with the prior?]).

## What Prior Makes the Intercept Maximally Uninformative?

The key to choosing an intercept prior is the **probability integral
transform**: for any continuous distribution $D$ with CDF $F$, if
$X \sim D$, then $F(X) \sim \text{Uniform}(0, 1)$. In a binomial GLM, the
inverse link $g^{-1}$ *is* a CDF, so we can exploit this result
directly.[^prior-lit]

[^prior-lit]: Previous work on priors for logistic regression focused on
estimation problems: @gelman2008 proposed weakly informative Cauchy(0, 2.5)
priors for separation, @ghosh2018 established conditions for posterior
existence, and @greenland2015 compared penalization approaches. These
contributions address estimation, not the distinct requirements of Bayes
Factor testing, where the prior density at the point null directly enters
the Savage--Dickey ratio [@ly2016; @bartos2023].

| Link | Inverse link $g^{-1}$ | Matched prior | Implied prior on $p$ |
|------|----------------------|---------------|----------------------|
| logit | logistic CDF | $\text{Logistic}(0, 1)$ | $\text{Uniform}(0, 1)$ --- **exact** |
| probit | normal CDF $\Phi$ | $\text{Normal}(0, 1)$ | $\text{Uniform}(0, 1)$ --- **exact** |

Placing a Logistic prior on $b_0$ under the logit link produces
probabilities that are exactly uniform on $(0, 1)$; the same holds for a
Normal prior under the probit link. This is not an approximation --- it
follows directly from the probability integral transform. Scaling the prior
below $\text{sd} = 1$ concentrates mass away from 0 and 1, producing a
regularized intercept prior suited to experiments where extreme baseline
probabilities are unlikely.

The **matched prior** at scale 1 is therefore maximally uninformative on the
probability scale --- a principled anchor that contrasts with specifying
priors based on intuition about regression coefficients, a strategy
@gelman2017prior cautioned against. The matched prior principle generalises
to any model with a CDF-based link function; we return to this point in the
[Recommendations].

We use exact analytical density formulas (derived via the change-of-variables
theorem) to overlay reference curves on the simulation figures below.[^density-formulas]

[^density-formulas]: For the logit link with a $\text{Logistic}(0, s)$ prior
on $b_0$: $f(p) = f_{\text{Logis}}(\text{logit}(p), 0, s) / [p(1-p)]$. At
$s = 1$, the logistic density and the Jacobian $p(1-p)$ cancel, giving
$f(p) = 1$. For the probit link with a $\text{Normal}(0, \sigma)$ prior:
$f(p) = \phi(\Phi^{-1}(p)/\sigma) / [\sigma \cdot \phi(\Phi^{-1}(p))]$. At
$\sigma = 1$, the normal densities cancel, again yielding $f(p) = 1$.

### Why misfit matters: current software defaults

Current `brms` [@burkner2017; @burkner2018] defaults prioritize estimation
stability, not hypothesis testing --- and they create a **prior--link
misfit** for logistic regression. By default, `brms` places a
$\text{student\_t}(3, 0, 2.5)$ prior on the intercept and a flat, improper
prior on all population-level effects. The Student-$t$ intercept prior,
passed through the logit link, concentrates more mass near $p = 0$ and
$p = 1$ than the matched Logistic prior. The flat slope prior is more
problematic still: it treats $|\delta p|$ near 1 as equally plausible as
$|\delta p|$ near 0. For Bayes Factor testing via the Savage--Dickey density
ratio, improper or overly diffuse priors are not merely inconvenient --- they
are undefined or produce BFs biased toward H1
[@schonbrodt2018; @consonni2018; @gelman2008].

Before examining the simulation results, we address one more design choice
that interacts with the prior: the contrast coding of the experimental
factor.

## How Does Contrast Coding Interact with the Prior?

The choice of contrast coding directly scales the effect coefficient $b_1$
and therefore the implied prior on the probability scale. With sum-to-zero
codes $x \in \{+1, -1\}$, the effect $b_1$ represents the deviation of each
condition from the grand mean, and the distribution of $\delta p$ is
**symmetric** around zero when the prior on $b_1$ is symmetric --- a
desirable property for a default prior. All simulations here use $\pm 1$
coding [see @schad2020 for a thorough treatment of contrast coding in
regression].

With treatment coding ($x \in \{0, 1\}$), one condition depends on $b_0$
alone and the other on $b_0 + b_1$, creating asymmetric prior uncertainty
across conditions. If the intercept prior mean is shifted away from zero,
this asymmetry also skews the prior predictive distribution of $\delta p$.
We therefore recommend sum-to-zero ($\pm 1$) coding, which distributes the
effect symmetrically. Guidance for adapting the recommended prior widths to
other contrast codes and predictor scales is provided in the
[Recommendations].

## What Does the Intercept Prior Imply on the Probability Scale?

The intercept $b_0$ determines the baseline success probability before any
experimental manipulation. A poorly chosen intercept prior can place
substantial mass near $p = 0$ or $p = 1$, predicting that participants
perform at floor or ceiling even before any experimental effect is applied.
How do different intercept priors perform?

```{r}
#| label: fig-1-p-intercept
#| fig-cap: "Prior predictive distribution of success probability
#|   p = g^{-1}(b0) for each intercept prior SD (rows) and link function
#|   (columns). At sd_b0 = 1.0, the matched prior (Logistic for logit,
#|   Normal for probit) produces a perfectly uniform distribution --- the
#|   visual signature of the probability integral transform. Reducing sd_b0
#|   to 0.75 pulls mass away from floor and ceiling. Misfit priors (e.g.,
#|   Cauchy under either link) concentrate more mass near 0 and 1. Dashed
#|   curves show exact analytical densities."
#| fig-height: 8
#| fig-width: 10

fig1_draws <- draws_long |>
  filter(dist_b1 == "logistic", sd_b1 == 0.25)

plot_p_intercept_ridges(fig1_draws, sd_b1_focus = 0.25)
```

@fig-1-p-intercept confirms the analytical result: at $\text{sd}_{b_0} = 1.0$
with the matched prior, the distribution of $p$ is uniform. At
$\text{sd}_{b_0} = 0.75$, the prior pulls mass away from 0 and 1, which is
appropriate when floor and ceiling effects are a concern. Misfit priors,
such as Cauchy for either link, produce heavier tails and concentrate more
mass near the boundaries.

```{r}
#| label: fig-2-floor-ceiling
#| fig-cap: "Proportion of prior predictive mass in the floor/ceiling region
#|   (p < 0.05 or p > 0.95) as a function of intercept prior SD and family.
#|   The matched prior at sd_b0 = 1.0 places exactly 10% in this region
#|   (2 x 0.05 = 0.10), confirming the uniform result. At sd_b0 = 0.75,
#|   floor/ceiling mass drops below 10% for both matched priors."
#| fig-height: 6
#| fig-width: 10

plot_floor_ceiling_heatmap(summaries)
```

We recommend choosing the intercept prior SD so that
$P(p < 0.05 \text{ or } p > 0.95) \leq 0.10$. Under this criterion,
$\text{Logistic}(0, 0.75)$ is the recommended default for the logit link and
$\text{Normal}(0, 0.75)$ for the probit link (@fig-2-floor-ceiling). The
exact 10% value at $\text{sd}_{b_0} = 1.0$ for the matched priors serves as
an internal validation of the simulation against the analytical result.

With the intercept prior established, we turn to the effect prior.

## How Wide Should the Effect Prior Be?

The prior on $b_1$ determines how much prior mass falls on small versus
large effects on the probability scale. @fig-3-delta-p shows how the choice
of $\text{sd}_{b_1}$ translates into implied effect sizes ($\delta p$),
holding the intercept prior fixed at the matched recommendation
($\text{Logistic}(0, 0.75)$ for logit; $\text{Normal}(0, 0.75)$ for probit).

```{r}
#| label: fig-3-delta-p
#| fig-cap: "Prior predictive distribution of delta_p (signed effect size on
#|   the probability scale) for each effect prior SD, colour-coded by prior
#|   family. At sd_b1 = 0.25, Normal and Logistic priors concentrate mass on
#|   effects up to about 20 percentage points --- a range consistent with
#|   typical cognitive experiments. The Cauchy stands apart even at this
#|   moderate scale, assigning substantial probability to effects exceeding
#|   50 percentage points. Dotted lines mark +/-0.10, +/-0.20, +/-0.30."
#| fig-height: 10
#| fig-width: 10

# Use the link-matched recommended intercept prior for each link:
# Logistic(0,0.75) for logit, Normal(0,0.75) for probit
fig2_draws <- draws_long |>
  filter(
    (link == "logit"  & dist_b0 == "logistic") |
    (link == "probit" & dist_b0 == "normal"),
    sd_b0 == 0.75
  )

plot_delta_p_ridges(fig2_draws, sd_b0_focus = 0.75)
```

@fig-3-delta-p shows several patterns. The **Cauchy prior** stands apart at
every scale: even at $\text{sd}_{b_1} = 0.25$, its heavy tails assign
substantial probability to $|\delta p| > 0.50$, representing a shift in
accuracy from 50% to above 90%. At $\text{sd}_{b_1} \geq 0.50$, effects
larger than 30 percentage points are treated as roughly as plausible as small
effects, making the Cauchy a poor default for cognitive psychology.

Even for the **Normal and Logistic priors**, a scale of
$\text{sd}_{b_1} = 1.0$ or above spreads mass broadly across the entire
$(-1, 1)$ range. At $\text{sd}_{b_1} = 0.75$, the 90th percentile of
$|\delta p|$ already exceeds 0.30. A scale below 1.0 is needed for
meaningful regularisation.

Within the range $\text{sd}_{b_1} \in [0.10, 0.50]$, Normal and Logistic
priors behave similarly. At $\text{sd}_{b_1} = 0.10$, almost all mass falls
within $|\delta p| < 0.10$, suitable only when very small effects are
expected. At $\text{sd}_{b_1} = 0.25$, the 90th percentile of $|\delta p|$
is near 0.20 --- a moderate effect of 20 percentage points that aligns well
with typical cognitive experiments. Values up to $\text{sd}_{b_1} = 0.50$
remain defensible when larger effects are plausible, with the 90th percentile
rising to roughly 0.35.

For fixed-effects models, an effect prior with $\text{sd}_{b_1} = 0.25$
under sum-to-zero coding provides a sensible default. Most experiments,
however, also require random effects to account for between-subject
heterogeneity, which raises the question of how the hyperprior should be
specified.

## How Should Priors Be Specified for Random Effects?

The `brms` default hyperprior on random-effect SDs ---
$\text{half-student\_t}(3, 0, 2.5)$ --- places a heavy right tail on
between-subject variability. Does this matter on the probability scale? To
find out, we compare three candidate hyperpriors for the random-effect SDs
$\sigma_u$ (intercept) and $\sigma_v$ (slope):

- **$\text{half-student\_t}(3, 0, 2.5)$**: the `brms` default, with a heavy
  right tail on $\sigma_u$.
- **$\text{Gamma}(2, 4)$**: mean $= 0.5$, with approximately 71% of mass
  below $\sigma_u = 0.5$ logit units.
- **$\text{Exponential}(4)$**: mean $= 0.25$, with approximately 86% of mass
  below $\sigma_u = 0.5$; the strongest concentration near zero.

We run a full hierarchical prior predictive with
$\sigma_u \sim \text{hyperprior}$, subject-level random intercepts
$u_i \sim \mathcal{N}(0, \sigma_u)$, and
$p_{ij} = g^{-1}(b_0 + u_i \pm (b_1 + v_i))$, where $v_i$ is the
subject-level random slope. The fixed-effects priors are held at the
recommended values ($b_0 \sim \text{Logistic}(0, 0.75)$,
$b_1 \sim \mathcal{N}(0, 0.25)$). For the intercept + slope structure,
the random-slope SD $\sigma_v$ is drawn from the same hyperprior as
$\sigma_u$.

```{r}
#| label: setup-re
#| include: false

re_summaries <- tryCatch(
  readRDS(here("output", "prior_predictive_re_summaries.rds")),
  error = function(e) NULL
)
re_draws <- tryCatch(
  readRDS(here("output", "prior_predictive_re_draws.rds")),
  error = function(e) NULL
)

re_results_available <- !is.null(re_summaries) && !is.null(re_draws)
```

```{r}
#| label: fig-re-1-p-intercept
#| fig-cap: "The brms default student_t(3, 0, 2.5) spreads subject-level
#|   probabilities toward floor and ceiling due to its heavy right tail on
#|   between-subject SDs. The exponential(4) and gamma(2, 4) hyperpriors keep
#|   subjects' probabilities well away from the boundary. Dashed vertical
#|   lines mark p = 0.05 and p = 0.95."
#| fig-height: 5
#| fig-width: 9
#| eval: !expr re_results_available

re_draws |>
  group_by(sd_prior_re, re_structure) |>
  slice_sample(n = 5000) |>
  ungroup() |>
  ggplot(aes(x = p_intercept, y = sd_prior_re,
             fill = sd_prior_re, colour = sd_prior_re)) +
  geom_density_ridges(alpha = 0.5, scale = 0.9, rel_min_height = 0.01) +
  geom_vline(xintercept = c(0.05, 0.95), linetype = "dashed",
             colour = "grey40", linewidth = 0.5) +
  facet_wrap(~re_structure) +
  scale_fill_viridis_d(begin = 0.1, end = 0.85, guide = "none") +
  scale_colour_viridis_d(begin = 0.1, end = 0.85, guide = "none") +
  scale_x_continuous("Subject-level success probability", limits = c(0, 1),
                     expand = c(0, 0)) +
  scale_y_discrete("Hyperprior") +
  labs(title = "Subject-level intercept probabilities by RE hyperprior") +
  theme_minimal(base_size = 12) +
  theme(strip.text = element_text(face = "bold"))
```

@fig-re-1-p-intercept shows the consequences: the `brms` default
`student_t(3, 0, 2.5)` spreads subject-level probabilities toward 0 and 1,
with heavier tails near the floor/ceiling boundary. The `gamma(2, 4)` and
`exponential(4)` hyperpriors concentrate $\sigma_u$ near zero, keeping
subjects' probabilities away from the boundary.

```{r}
#| label: fig-re-2-delta-p
#| fig-cap: "Under the student_t default, a non-trivial fraction of simulated
#|   subjects show effects in the opposite direction to the population-level
#|   b1. The more concentrated hyperpriors (gamma, exponential) limit this
#|   effect heterogeneity to a substantively plausible range. Dotted lines
#|   mark +/-0.10, +/-0.20, +/-0.30."
#| fig-height: 5
#| fig-width: 9
#| eval: !expr re_results_available

re_draws |>
  group_by(sd_prior_re, re_structure) |>
  slice_sample(n = 10000) |>
  ungroup() |>
  ggplot(aes(x = delta_p, y = sd_prior_re,
             fill = sd_prior_re, colour = sd_prior_re)) +
  geom_density_ridges(alpha = 0.5, scale = 0.9, rel_min_height = 0.01) +
  geom_vline(xintercept = 0, linewidth = 0.5, colour = "grey20") +
  geom_vline(xintercept = c(-0.30, -0.20, -0.10, 0.10, 0.20, 0.30),
             linetype = "dotted", linewidth = 0.5, colour = "grey40") +
  facet_wrap(~re_structure) +
  scale_fill_viridis_d(begin = 0.1, end = 0.85, guide = "none") +
  scale_colour_viridis_d(begin = 0.1, end = 0.85, guide = "none") +
  scale_x_continuous("Effect size on probability scale (delta_p)",
                     limits = c(-1, 1), breaks = seq(-1, 1, 0.25)) +
  scale_y_discrete("Hyperprior") +
  labs(title = "Effect size distribution by RE hyperprior") +
  theme_minimal(base_size = 12) +
  theme(strip.text = element_text(face = "bold"))
```

The `student_t` hyperprior also produces more within-study effect
heterogeneity (@fig-re-2-delta-p): a non-trivial fraction of simulated
subjects show effects in the opposite direction to the population-level
$b_1$. The wide distribution of $\sigma_u$ pulls individual subjects into
regions where the logistic nonlinearity reverses the apparent effect
direction.

We therefore recommend `exponential(4)` for random-effect SD hyperpriors.
It concentrates between-subject variability in a plausible range and avoids
excessive floor/ceiling effects.

## Recommendations

For binomial GLMs with sum-to-zero contrast coding ($\pm 1$), we recommend
the following default priors:

### Intercept prior

| Link | Recommended prior | Rationale |
|------|------------------|-----------|
| logit | $\text{Logistic}(0, 0.75)$ | Matched family; sd = 0.75 gives < 10% floor/ceiling mass |
| probit | $\text{Normal}(0, 0.75)$ | Matched family; same criterion |

At $\text{sd} = 0.75$, less than 10% of prior mass falls in the
floor/ceiling region, while remaining agnostic about baseline accuracy
within the plausible range.

### Effect prior

| Prior family | Recommended SD | 90th percentile of $|\delta p|$ |
|-------------|---------------|-------------------------------|
| Logistic (default) | 0.25 | ~0.20 |
| Normal | 0.25 | ~0.18 |
| Cauchy | 0.25 | ~0.25 (heavier tails) |

At $\text{sd}_{b_1} = 0.25$, effects larger than 20 percentage points
receive decreasing prior probability, while effects above 30 percentage
points become increasingly implausible.

Our default recommendation is $\text{Logistic}(0, 0.25)$ for logit models
and $\text{Normal}(0, 0.25)$ for probit models, consistent with typical
effect magnitudes in cognitive psychology.

For sensitivity analyses, $\text{sd}_{b_1}$ values between 0.10 and 0.50
span a reasonable range of prior beliefs. Values above 0.50 assign
substantial probability to extreme effects and require specific theoretical
justification.

### Random effects prior

For prior predictive behaviour consistent with the fixed-effects
recommendations:

- **Use `exponential(4)` or `gamma(2, 4)`** for both $\sigma_u$ and
$\sigma_v$. These concentrate the between-subject SD below ~0.5 logit
units, which keeps floor/ceiling mass under 10% and limits within-study
effect heterogeneity to a substantively plausible range.
- **Avoid the brms [@burkner2017] default `student_t(3, 0, 2.5)`** unless very large
between-subject variability is expected. The heavy right tail of the
half-$t$ assigns non-negligible probability to $\sigma > 2$ on the logit
scale, which is implausible for most cognitive experiments.
- Between `exponential(4)` and `gamma(2, 4)`, prefer `exponential(4)` for
its stronger concentration near zero --- it matches the intercept-only
floor/ceiling criterion more reliably.

### Ready-to-use brms prior specifications

The following code blocks specify the recommended default priors directly in
`brms` [@burkner2017] syntax and can be passed to
the `prior` argument of `brm()`:

```{r}
#| label: recommended-priors
#| echo: true
#| eval: false

# Logit link (logistic regression)
# Intercept: Logistic(0, 0.75) — matched prior, < 10% floor/ceiling mass
# Effect:    Logistic(0, 0.25) — 90th percentile of |delta_p| ~ 0.20
# Note: brms accepts logistic() as a valid distribution name in prior()
logit_priors <- c(
  brms::prior(logistic(0, 0.75), class = Intercept),
  brms::prior(logistic(0, 0.25), class = b)
)

# Probit link (probit regression)
# Intercept: Normal(0, 0.75) — matched prior, < 10% floor/ceiling mass
# Effect:    Normal(0, 0.25) — 90th percentile of |delta_p| ~ 0.18
probit_priors <- c(
  brms::prior(normal(0, 0.75), class = Intercept),
  brms::prior(normal(0, 0.25), class = b)
)
```

For models with subject-level random effects, add a prior on the SD of the
random intercept (and slope, if included). Based on the prior predictive
analyses above, we recommend $\text{Exponential}(4)$ as a moderately
regularising default:

```{r}
#| label: recommended-priors-re
#| echo: true
#| eval: false

# Logit link with random intercept
logit_priors_re <- c(
  brms::prior(logistic(0, 0.75), class = Intercept),
  brms::prior(logistic(0, 0.25), class = b),
  brms::prior(exponential(4),    class = sd)
)

# Probit link with random intercept
probit_priors_re <- c(
  brms::prior(normal(0, 0.75), class = Intercept),
  brms::prior(normal(0, 0.25), class = b),
  brms::prior(exponential(4),  class = sd)
)
```

### Adapting to other contrast codes and predictor scales

The recommended $\text{sd}_{b_1} = 0.25$ assumes sum-to-zero $\pm 1$ coding
($\text{SD}(x) = 1$). Whenever the predictor's SD differs from 1, the prior
on $b_1$ must be rescaled:

$$\text{sd}_{b_1}(\text{predictor SD} = s) = \frac{\text{sd}_{b_1}(\text{SD} = 1)}{s}$$

Smaller predictor SDs require a *wider* prior on $b_1$; larger predictor SDs
require a *narrower* prior. Examples:

- $\pm 0.5$ coding ($s = 0.5$): use $\text{sd}_{b_1} = 0.25 / 0.5 = 0.50$
- $\pm 1/\sqrt{2}$ orthonormal ($s \approx 0.707$): use $\text{sd}_{b_1} = 0.25 / 0.707 \approx 0.354$
- $0/1$ treatment coding ($s = 0.5$ in a balanced design): use $\text{sd}_{b_1} = 0.25 / 0.5 = 0.50$
- Standardised continuous predictor ($s = 1$): use $\text{sd}_{b_1} = 0.25$ directly

For continuous predictors that are not standardised, compute the
population-level SD and apply the formula above.

#### Generalising to multi-level and factorial designs

When a factor has more than two levels, the choice of contrast coding affects
not only the interpretation of coefficients but also the *prior* on each
factor-level mean. With standard coding schemes (treatment, effects), placing
the same prior on all coefficients produces unequal priors across factor
levels --- the reference level is treated differently from the remaining
levels. This asymmetry propagates to the Bayes factor, biasing evidence for
some comparisons relative to others [@rouder2012; @schad2020].

Orthonormal contrasts avoid this problem. Each contrast column sums to
zero and has unit length, so placing the same prior on each coefficient
produces equal marginal priors on all factor levels. The
`contr.equalprior()` function from the `bayestestR` package
[@makowski2019] provides such contrasts, following the construction in
@rouder2012 [p. 363]. Because the prior is symmetric across all directions
in the effect space, the Bayes factor does not depend on which specific
orthonormal basis is chosen (e.g., Helmert vs. polynomial) --- only the
orthonormality property matters [@rouder2017].

In a balanced design with $J$ levels, each orthonormal contrast column has
predictor $\text{SD} = 1 / \sqrt{J}$. Applying the scaling formula from
above gives:

$$\text{sd}_{b_1}(\text{orthonormal, } J \text{ levels}) = 0.25 \times \sqrt{J}$$

This extends to factorial designs. In a balanced $J \times K$ factorial with
orthonormal contrasts on both factors, the interaction columns are
element-wise products of the main-effect columns, yielding predictor
$\text{SD} = 1 / \sqrt{JK}$ [@rouder2017]. @tbl-orthonormal-scaling gives
the rescaled prior widths for common cases.

| Design term | Predictor SD | Rescaled $\text{sd}_{b_1}$ |
|:---|:---:|:---:|
| 2-level factor | $1/\sqrt{2} \approx 0.71$ | 0.35 |
| 3-level factor | $1/\sqrt{3} \approx 0.58$ | 0.43 |
| 4-level factor | $1/\sqrt{4} = 0.50$ | 0.50 |
| 5-level factor | $1/\sqrt{5} \approx 0.45$ | 0.56 |
| $2 \times 3$ interaction | $1/\sqrt{6} \approx 0.41$ | 0.61 |
| $3 \times 3$ interaction | $1/\sqrt{9} \approx 0.33$ | 0.75 |

: Rescaled effect prior widths under orthonormal contrasts for multi-level
  factors and factorial interactions. All values assume the baseline
  recommendation of $\text{sd}_{b_1} = 0.25$ under $\pm 1$ coding.
  {#tbl-orthonormal-scaling}

The full scaling produces wider priors for interactions than for main
effects. In many experimental contexts, interactions are expected to be
smaller than main effects, so using the *unscaled* main-effect prior for
interaction terms is a reasonable alternative. The table above gives the
default when no such expectation exists.

@fig-orthonormal-demo shows the resulting prior predictive distributions
for $J = 2$, $3$, and $4$.

```{r}
#| label: fig-orthonormal-demo
#| fig-cap: "Prior predictive distribution of pairwise effect sizes under
#|   orthonormal contrasts with the rescaled prior
#|   ($\\text{sd}_{b_1} = 0.25 \\times \\sqrt{J}$). Left: distribution of
#|   a single representative pairwise $\\delta p$, which is nearly identical
#|   across $J$ --- confirming that the scaling rule preserves per-comparison
#|   calibration. Right: distribution of the maximum pairwise
#|   $|\\delta p|$, which shifts right with $J$ as expected
#|   (more pairwise comparisons produce a larger maximum)."
#| fig-height: 4.5
#| fig-width: 10

source(here("R", "link_functions.R"))
source(here("R", "plotting.R"))

set.seed(2024)
n_draws <- 50000

demo_list <- lapply(c(2, 3, 4), function(J) {
  oc <- orthonormal_contrasts(J)
  sd_b1_scaled <- 0.25 * oc$scaling_factor

  b0 <- rlogis(n_draws, location = 0, scale = 0.75)
  # J-1 contrast coefficients from the rescaled prior
  B <- matrix(
    rnorm(n_draws * (J - 1), mean = 0, sd = sd_b1_scaled),
    nrow = n_draws, ncol = J - 1
  )

  # Cell means on the linear predictor: eta = b0 + Q %*% beta
  eta_mat <- outer(b0, rep(0, J)) + B %*% t(oc$Q)
  p_mat   <- ilogit(eta_mat)

  # Single representative pairwise delta_p (levels 1 vs 2)
  delta_p_pair <- p_mat[, 1] - p_mat[, 2]

  # Maximum absolute pairwise delta_p
  max_abs_dp <- apply(p_mat, 1, function(row) {
    diffs <- outer(row, row, "-")
    max(abs(diffs[upper.tri(diffs)]))
  })

  data.frame(
    J            = paste0("J = ", J),
    delta_p_pair = delta_p_pair,
    max_abs_dp   = max_abs_dp
  )
})

demo_draws <- do.call(rbind, demo_list)

p_single <- ggplot(demo_draws, aes(x = delta_p_pair,
                                    fill = J, colour = J)) +
  geom_density(alpha = 0.3, linewidth = 0.6) +
  geom_vline(xintercept = c(-0.20, -0.10, 0.10, 0.20),
             linetype = "dotted", linewidth = 0.4,
             colour = "grey40") +
  scale_x_continuous(expression(delta * p ~ "(single pair)"),
                     limits = c(-0.8, 0.8)) +
  scale_fill_viridis_d("Factor levels", option = "D", end = 0.85) +
  scale_colour_viridis_d("Factor levels", option = "D", end = 0.85) +
  ggtitle("Single pairwise comparison") +
  prior_theme()

p_max <- ggplot(demo_draws, aes(x = max_abs_dp,
                                 fill = J, colour = J)) +
  geom_density(alpha = 0.3, linewidth = 0.6) +
  geom_vline(xintercept = c(0.10, 0.20, 0.30),
             linetype = "dotted", linewidth = 0.4,
             colour = "grey40") +
  scale_x_continuous(expression(max ~ "|" * delta * p * "|"),
                     limits = c(0, 1)) +
  scale_fill_viridis_d("Factor levels", option = "D", end = 0.85) +
  scale_colour_viridis_d("Factor levels", option = "D", end = 0.85) +
  ggtitle("Maximum pairwise difference") +
  prior_theme()

patchwork::wrap_plots(p_single, p_max, guides = "collect") &
  theme(legend.position = "bottom")
```

The left panel of @fig-orthonormal-demo shows that the prior predictive
distribution of any single pairwise comparison is nearly identical across
$J = 2$, $3$, and $4$: the $\sqrt{J}$ scaling rule preserves the
calibration established for the two-condition case. The right panel shows
that the maximum pairwise $|\delta p|$ shifts right with increasing $J$.
This shift reflects the number of comparisons, not prior misspecification
--- with more pairs, the largest difference naturally grows.

In `brms`, the recommended workflow is to set orthonormal contrasts before
fitting the model and adjust the prior width accordingly:

```{r}
#| label: recommended-priors-orthonormal
#| echo: true
#| eval: false

# --- Single 3-level factor ---
contrasts(dat$condition) <- bayestestR::contr.equalprior(3)
# Predictor SD = 1/sqrt(3); rescaled prior: sd_b1 = 0.25 * sqrt(3) = 0.433

priors_3level <- c(
  brms::prior(logistic(0, 0.75),  class = Intercept),
  brms::prior(normal(0, 0.433),   class = b)
)

# --- 2 x 3 factorial design ---
contrasts(dat$A) <- bayestestR::contr.equalprior(2)
contrasts(dat$B) <- bayestestR::contr.equalprior(3)
# Main effect A: sd = 0.25 * sqrt(2) = 0.354
# Main effect B: sd = 0.25 * sqrt(3) = 0.433
# Interaction:   sd = 0.25 * sqrt(6) = 0.612
#   (or use the main-effect prior if interactions are expected to be small)

priors_factorial <- c(
  brms::prior(logistic(0, 0.75),  class = Intercept),
  brms::prior(normal(0, 0.354),   class = b, coef = A1),
  brms::prior(normal(0, 0.433),   class = b, coef = B1),
  brms::prior(normal(0, 0.433),   class = b, coef = B2),
  brms::prior(normal(0, 0.612),   class = b, coef = "A1:B1"),
  brms::prior(normal(0, 0.612),   class = b, coef = "A1:B2")
)
```

For unbalanced designs, the predictor SD depends on cell sizes and should
be computed from the contrast matrix directly. For factors with many levels
($J > 5$), the `BayesFactor` package [@rouder2012; @rouder2017] implements
the full multivariate JZS prior, which accounts for the dimensionality of
the effect space without requiring per-coefficient prior widths.

#### No-intercept parameterization (0 + coding in brms)

In `brms`, specifying `0 + condition` removes the intercept and estimates
separate coefficients for each condition level, so that
$p_0 = g^{-1}(b_0)$, $p_1 = g^{-1}(b_1)$, and
$\delta p = g^{-1}(b_1) - g^{-1}(b_0)$. Under this parameterization, the
matched prior principle applies to each coefficient independently: placing
$\text{Logistic}(0, 1)$ on both coefficients under the logit link induces
independent $\text{Uniform}(0, 1)$ marginals, and the implied prior on
$\delta p$ is a symmetric triangular distribution on $(-1, 1)$. Because the
coefficient interpretation differs from the intercept-plus-slope
parameterization, our recommended prior widths do not transfer directly ---
each condition mean should receive a matched prior scaled to reflect beliefs
about plausible success probabilities (e.g., $\text{sd} = 0.75$), and the
implied $\delta p$ should be checked via prior predictive simulation.

### Beyond binomial models

The approach here --- analytical derivation of matched priors via the
probability integral transform, followed by prior predictive simulation ---
applies to any model where link functions mediate between parameter and
observation scales [@gelman2020workflow; @schad2021; @schonbrodt2018]. The
matched prior principle generalises to any model that maps parameters through
a CDF. In a cumulative ordinal model with a logit link, a Logistic prior on
the threshold parameters produces uniform implied probabilities across
response categories. In a log-link Poisson model, an Exponential(1) prior on
the rate produces a Uniform on the CDF scale. More broadly, any
distributional regression model [@klein2015; @burkner2021] where response
parameters are linked to predictors faces the same prior--link interaction.
The same workflow --- identify the matched family, characterise deviations
via prior predictive simulation, calibrate Bayes Factor performance --- can
be applied wherever such interactions arise.

The simulation code, ready-to-use `brms` prior specifications, and all
output data are available in the project repository. A companion report
examines how these priors perform in Bayes Factor hypothesis testing.

## References

::: {#refs}
:::
